### 2016-08-05 paper.tex
Initial version.

\documentclass{sigchi}

%% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is      granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from permissions@acm.org. \\
% {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
% Copyright \copyright~2014 ACM ISBN/14/04...\$15.00. \\
% DOI string from ACM form confirmation}
%% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)

\toappear{}

% Arabic page numbers for submission.  Remove this line to eliminate
% page numbers for the camera ready copy

%\pagenumbering{arabic}

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage[pdftex]{hyperref}
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
%\hypersetup{%
%  pdftitle={SIGCHI Conference Proceedings Format},
%  pdfauthor={LaTeX},
%  pdfkeywords={SIGCHI, proceedings, archival format},
%  bookmarksnumbered,
%  pdfstartview={FitH},
%  colorlinks,
%  citecolor=black,
%  filecolor=black,
%  linkcolor=black,
%  urlcolor=linkColor,
%  breaklinks=true,
%}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{PeerMind}

%\numberofauthors{3}
%\author{%
%  \alignauthor{1st Author Name\\
%    \affaddr{Affiliation}\\
%    \affaddr{City, Country}\\
%    \email{e-mail address}}\\
%  \alignauthor{2nd Author Name\\
%    \affaddr{Affiliation}\\
%    \affaddr{City, Country}\\
%    \email{e-mail address}}\\
%  \alignauthor{3rd Author Name\\
%    \affaddr{Affiliation}\\
%    \affaddr{City, Country}\\
%    \email{e-mail address}}\\
%}

\maketitle

\begin{abstract}

\todo[inline]{TODO}

\end{abstract}

%\keywords{Authors' choice; of terms; separated; by semi\-colons;
%  commas, within terms only; this section is required.}

%\category{H.5.m.}{Information Interfaces and Presentation
%  (e.g. HCI)}{Miscellaneous} \category{See
%  \url{http://acm.org/about/class/1998/} for the full list of ACM
%  classifiers. This section is required.}{}{}

\section{Introduction}

\bibliographystyle{sigchi}
\bibliography{refs}

\end{document}

### 2016-08-06 paper.tex
Citations.
Cleaning.
Delegation.

@@
 \usepackage{todonotes}
+\usepackage{pgf}
+\usepackage{tikz}
+\usepackage{tkz-graph}
+\usepackage{mathtools}
 
@@
 
-% End of preamble. Here it comes the document.
+\GraphInit[vstyle=Dijkstra]
+\tikzset{
+  EdgeStyle/.append style={->,bend left}
+}
+
+\hyphenation{PeerMind}
+
 \begin{document}
@@
 
-\todo[inline]{TODO}
+%\todo[inline]{TODO}
 
@@
 
+Terminology:
+\begin{itemize}
+\item community -- group which is making a decision using PeerMind
+\item member -- user of the app who is participating in decision making
+\end{itemize}
+
+% decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)
+
+% motion
+
+\section{Related work}
+
+\cite{andersen2008trust, dirnstorfer2010voting, ford2002delegative, rodriguez2007smartocracy, yamakawa2007toward}
+
+\section{Features}
+
+\subsection{Dynamic quroum}
+
+\subsection{Score voting}
+
+\subsection{Delegation}
+
+Because not every member of the community has time or knowledge to vote all the time, each member can delegate
+decisions to one or more other members of the community.
+When a member does not vote themselves, information how those delegates voted is combined into a vote of
+the non-voting member.
+If a delegate has not voted themselves either, their vote is computed through delegation as well, recursivelly.
+
+Each member $M$ can declare their delegates and ratios $r$ in which delegates' votes are combined. For example:
+
+\begin{itemize}
+\item Member $M_0$ decides not to delegate at all.
+\item Member $M_1$ delegates to $M_0$ and $M_3$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectivelly.
+\item Member $M_2$ delegates to $M_3$ with ratio $1.00$ ($r_{2,3}$).
+\item Member $M_3$ delegates to $M_1$ and $M_2$ with ratios $0.40$ ($r_{3,1}$) and $0.40$ ($r_{3,2}$).
+\end{itemize}
+
+Observe that member $M_3$ decided to use only $0.8$ of their vote in the case they do not vote themselves.
+The rest ($0.2$) counts towards non-voting population.
+In this way a member can decide to participate in voting through delegation, but not fully.
+Member $M_0$ deciced to not participate at all and if they do not vote they are counted towards non-voting population.
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \SetGraphUnit{2.5}
+
+    \Vertex[Math,L=M_0]{M0}
+    \EA[Math,L=M_1](M0){M1}
+    \SO[Math,L=M_3](M1){M3}
+    \WE[Math,L=M_2](M3){M2}
+
+    \Edge[label=$0.40$](M1)(M3)
+    \Edge[label=$0.75$](M0)(M1)
+    \Edge[label=$0.25$](M3)(M1)
+
+    \tikzset{EdgeStyle/.append style={bend right}}
+    \Edge[label=$1.00$](M3)(M2)
+    \Edge[label=$0.40$](M2)(M3)
+  \end{tikzpicture}
+  \caption{Example delegation network. Edges are directed towards a member who is delegating, representing how their
+  vote is computed from delegates' votes. E.g., $V_{M_1} = 0.75 V_{M_0} + 0.25 V_{M_3}$}~\label{fig:delegation-network}
+\end{figure}
+
+We can represent these delegations as a graph we call \emph{delegation network} as seen in
+Figure~\ref{fig:delegation-network}.
+We direct edges towards a member who is delegating, to represent how their vote is computed from delegates' votes.
+For example, member $M_1$ delegates the decision to members $M_0$ and $M_3$.
+$M_0$ and $M_3$ become $M_1$'s delegates.
+As a consequence, $M_1$'s vote ($V_{M_1}$) is computed from $M_0$'s and $M_3$'s votes ($V_{M_0}$, $V_{M_3}$).
+
+A delegation network can be described as a system of equations.
+
+\begin{displaymath}
+\begin{array}{rcl}
+V_{M_0} & = & 0 \\
+V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
+V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
+V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} = 0.40 V_{M_1} + 0.40 V_{M_2} \\
+\end{array}
+\end{displaymath}
+
+When only some members vote ($\mathbf{v}$) we can use a delegation network to compute votes
+of all members ($\mathbf{x}$).
+We convert a delegation network to a matrix $\mathrm{D}$:
+
+\begin{displaymath}
+\mathrm{D} = \left[d_{ij}\right] = \begin{dcases*}
+ 1 & if $i = j$ \\
+ -r_{i,j} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with ratio $r_{i,j}$} \\
+ 0 & otherwise \\
+\end{dcases*}
+\end{displaymath}
+
+We solve the matrix equation:
+
+\begin{displaymath}
+\mathrm{D} \mathbf{x_v} = \mathbf{v}
+\end{displaymath}
+
+To continue with our example, let's say that $M_0$ and $M_2$ voted for a decision with $0.7$ and against with $0.4$,
+respectivelly.
+We are using score voting on $[-1, 1]$ range, so those votes are represented with $V_{M_0} = 0.7$ and
+$V_{M_2} = -0.4$.
+
+\begin{displaymath}
+\mathrm{D} = \left[ \begin{array}{cccc}
+1 & 0 & 0 & 0 \\
+-0.75 & 1 & -0.25 & 0 \\
+0 & 0 & 1 & 0 \\
+0 & -0.40 & -0.40 & 1 \\
+\end{array} \right]
+\end{displaymath}
+
+\begin{displaymath}
+\mathbf{v} = \left[ \begin{array}{c}
+0.7 \\
+0 \\
+-0.4 \\
+0 \\
+\end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
+0.7 \\
+0.54 \\
+-0.4 \\
+0.056 \\
+\end{array} \right]
+\end{displaymath}
+
+The result (average of scores) without delegation is $\overline{v} = 0.15$ and
+with delegation is $\overline{x} = 0.22$.
+Delegation pushes decisision slightly in favor.
+
+There are few details we have to look into.
+Despite delegation some votes or parts of votes can still happen to be discarded and should count
+towards non-voting population.
+Members might not declare their delegates, or they might declare them without using their vote fully.
+Other members can delegate to them as well, potentially propagating lack of a vote further.
+For example, if member $M_0$ would not vote, then their vote would be discarded.
+Moreover, we could not use $M_0$'s vote to compute $M_1$'s vote fully, which would influence $M_3$'s vote as well.
+
+One option is that we would normalize ratio in such case, so if $M_0$ does not vote, then $r_{1,3}$ is normalized
+from $0.25$ to $1.0$.
+We could do this automatically for all members, or we could even leave to members to configure
+what to do in such case for them.
+In current version of PeerMind we decided to not handle this situation in any special way and we leave
+discarded votes to propagate.
+
+As a consequence we have to determine how many votes have been discarded to know the size of the non-voting
+population.
+Another issue is also a question of abstentions.
+In PeerMind members can abstain from voting which is different from simply not participating.
+Abstentions are counted towards quorum, but they do not contribute to the result.
+To address this, we define two vectors:
+
+\begin{displaymath}
+\mathrm{w} = \left[w_{i}\right] = \begin{dcases*}
+ 1 & if $M_i$ voted \\
+ 0 & otherwise \\
+\end{dcases*},\quad \mathrm{a} = \left[a_{i}\right] = \begin{dcases*}
+ 1 & if $M_i$ abstained \\
+ 0 & otherwise \\
+\end{dcases*}
+\end{displaymath}
+
+Solving $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ gives
+us number of votes and number of abstentions, respectivelly, by summing together all values in the
+resulting vectors: $w = \sum \mathbf{x_w}$, $a = \sum \mathbf{x_a}$.
+
+\todo[inline]{Is the result then $\frac{\sum \mathbf{x_v}}{w}$ or $\frac{\sum \mathbf{x_v}}{\textrm{number of entries
+in}~\mathbf{x_w}~\textrm{which are non-zero}}$?}
+
+We pass computed votes and number of votes and abstaintions into computation of dynamic quorum to determine
+a quorum and confidence, in the same manner as if members voted themselves.
+
+% In general, there are various ways to use this delegation scheme.
+To whom to delegate could be done for each vote separately, or decisions could be categorized into various
+categories and then delegations for that category would be used.
+But for simplicity, PeerMind currently supports only defining delegates separately from a decision making process,
+and those delegates are then in effect for all votes where a member does not vote themselves.
+A member is invited to define initial delegations at a registration time, and can change them later on at will,
+but independently from a particular decision being made.
+
+\subsection{Integration with offline meetings}
+
 \bibliographystyle{sigchi}

### 2016-08-06 paper.tex
TODOs.

@@
 
+\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
+we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
+also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
+
 \subsection{Delegation}
@@
 
+\todo[inline]{Or should we be normalizing? In a way we should not normalize if votes are discarded because
+member explicitly delegated so ($M_3$ above), but if votes are discarded because we could not compute
+all delegations, we might want to automatically normalize.}
+
 As a consequence we have to determine how many votes have been discarded to know the size of the non-voting

### 2016-08-06 paper.tex
Adding various notes.

@@
 
+\todo[inline]{This section is mostly notes. Organize and improve.}
+
 Terminology:
@@
 
-% decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)
+\todo[inline]{Terminology: decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)}
+\todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}
 
-% motion
+The main question is: can we increase legitimacy and decrease attention required at the same time.
+Many others are trying to increase attention (participation) because they want to increase legitimacy,
+assuming that this is the only way.
+
+A list of potential other points:
+\begin{itemize}
+\item legitimacy (quorum, involvement of more people through delegation)
+\item scaling (delegation helps, less attention helps)
+\item attention (at same or better decisions)
+\item bad assumptions for democracy - time, knowledge, it s rational to know that you do not know, common practice is to ask friends
+\item better decisions (more people consulted, attention directed where it should be)
+\end{itemize}
+
+Another question: should feedback be visible (to see how others voted before you voted)?
+People go with majority.
+Better for wisdom of crowds is an independent sample.
+Maybe we should only allow how other voted to be seen after you voted ()but you can still change the vote),
+but we could then observe that and learn from this and see if this is good or bad.
+Or we could just store information if you observed existing votes before you voted (you have to click a button to see them).
+Maybe going with majority is also good when you want more social cohesion in the community and not so much "optimal"
+wisdom of crowds result -  you would want that people relax their positions based on how other people care.
+Cohesion of the group vs. accuracy.
+Maybe display only number of votes and confidence, maybe confidence direction or graph, but not average votes.
+
+A list of potential questions:
+\begin{itemize}
+\item Delegation works?
+\item Statistical quorum works?
+\item Real-time feedback influences?
+\item Private vs. public delegation (do you know how member to whom you delegated is voting)?
+\item Knowing how much power you have vs.
+not knowing -- does it change how people behave?
+(Knowing might corrupt you, but also motivate you to do more thorough work voting because you know other depend on you.)
+\end{itemize}
+
+Performance/theory/scaling in terms in computational (how big could effectively community can be to be able to compute it).
+Some other theoretical properties?
 
@@
 
-\subsection{Dynamic quroum}
+\subsection{Dynamic quorum}
 
@@
 the non-voting member.
-If a delegate has not voted themselves either, their vote is computed through delegation as well, recursivelly.
+If a delegate has not voted themselves either, their vote is computed through delegation as well, recursively.
 
@@
 \item Member $M_0$ decides not to delegate at all.
-\item Member $M_1$ delegates to $M_0$ and $M_3$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectivelly.
+\item Member $M_1$ delegates to $M_0$ and $M_3$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectively.
 \item Member $M_2$ delegates to $M_3$ with ratio $1.00$ ($r_{2,3}$).
@@
 In this way a member can decide to participate in voting through delegation, but not fully.
-Member $M_0$ deciced to not participate at all and if they do not vote they are counted towards non-voting population.
+Member $M_0$ decided to not participate at all and if they do not vote they are counted towards non-voting population.
 
@@
 To continue with our example, let's say that $M_0$ and $M_2$ voted for a decision with $0.7$ and against with $0.4$,
-respectivelly.
+respectively.
 We are using score voting on $[-1, 1]$ range, so those votes are represented with $V_{M_0} = 0.7$ and
@@
 with delegation is $\overline{x} = 0.22$.
-Delegation pushes decisision slightly in favor.
+Delegation pushes decision slightly in favor.
 
@@
 Solving $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ gives
-us number of votes and number of abstentions, respectivelly, by summing together all values in the
+us number of votes and number of abstentions, respectively, by summing together all values in the
 resulting vectors: $w = \sum \mathbf{x_w}$, $a = \sum \mathbf{x_a}$.
@@
 
-We pass computed votes and number of votes and abstaintions into computation of dynamic quorum to determine
+We pass computed votes and number of votes and abstentions into computation of dynamic quorum to determine
 a quorum and confidence, in the same manner as if members voted themselves.
@@
 
+\subsection{Three step process}
+
+\subsection{Real-time feedback}
+
+% Real-time communication in general.
+
 \subsection{Integration with offline meetings}

### 2016-08-06 paper.tex
Improving math symbols.

@@
 $M_0$ and $M_3$ become $M_1$'s delegates.
-As a consequence, $M_1$'s vote ($V_{M_1}$) is computed from $M_0$'s and $M_3$'s votes ($V_{M_0}$, $V_{M_3}$).
+As a consequence, $M_1$'s vote $V_{M_1}$ is computed from $M_0$'s and $M_3$'s votes ($V_{M_0}$ and $V_{M_3}$).
 
@@
 \begin{displaymath}
-\mathrm{w} = \left[w_{i}\right] = \begin{dcases*}
+\mathbf{w} = \left[w_{i}\right] = \begin{dcases*}
  1 & if $M_i$ voted \\
  0 & otherwise \\
-\end{dcases*},\quad \mathrm{a} = \left[a_{i}\right] = \begin{dcases*}
+\end{dcases*},\quad \mathbf{a} = \left[a_{i}\right] = \begin{dcases*}
  1 & if $M_i$ abstained \\
@@
 Solving $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ gives
-us number of votes and number of abstentions, respectively, by summing together all values in the
-resulting vectors: $w = \sum \mathbf{x_w}$, $a = \sum \mathbf{x_a}$.
+us number of votes $W$ and number of abstentions $A$, respectively, by summing together all values in the
+resulting vectors: $W = \sum \mathbf{x_w}$, $A = \sum \mathbf{x_a}$.
 
-\todo[inline]{Is the result then $\frac{\sum \mathbf{x_v}}{w}$ or $\frac{\sum \mathbf{x_v}}{\textrm{number of entries
+\todo[inline]{Is the result then $\frac{\sum \mathbf{x_v}}{W}$ or $\frac{\sum \mathbf{x_v}}{\textrm{number of entries
 in}~\mathbf{x_w}~\textrm{which are non-zero}}$?}

### 2016-08-06 paper.tex
More notes.

@@
 
-\todo[inline]{This section is mostly notes. Organize and improve.}
+The broad idea is that instead of just trying to move the voting process online -- something many others do -- we in
+fact add additional value by voting online.
+Specifically, we can compute results faster and more accurately, we can use sophisticated statistics to estimate
+quantities for the entire voting population, and we can promote clarity and full understanding of the voting items for all.
+
+\todo[inline]{This rest of the section is mostly notes. Organize and improve.}
 
@@
 Better for wisdom of crowds is an independent sample.
-Maybe we should only allow how other voted to be seen after you voted ()but you can still change the vote),
+Maybe we should only allow how other voted to be seen after you voted (but you can still change the vote),
 but we could then observe that and learn from this and see if this is good or bad.
 Or we could just store information if you observed existing votes before you voted (you have to click a button to see them).
-Maybe going with majority is also good when you want more social cohesion in the community and not so much "optimal"
+Maybe going with majority is also good when you want more social cohesion in the community and not so much ``optimal''
 wisdom of crowds result -  you would want that people relax their positions based on how other people care.
@@
 \item Private vs. public delegation (do you know how member to whom you delegated is voting)?
+\item When delegating, how to assure that you cannot learn how somebody else voted, but that you can still somehow
+learn how your vote is used?
 \item Knowing how much power you have vs.
@@
 
+Out of scope of this work:
+
+\begin{itemize}
+\item How to assure privacy of the delegation network (it is highly sensitive data, probably even more than votes)?
+\item How to compute delegation results in a way that it does not require access to the whole network of delegation at once?
+\item How can we make this system distributed (and still assure privacy)?
+\end{itemize}
+
 \section{Related work}
@@
 \subsection{Dynamic quorum}
+
+Using statistics to compute dynamic quorum based on how people vote using hypothesis testing.
+If population all votes for one position, then quorum can be much lower than if population is completely split on an
+issue, then in fact the whole population should vote because this is the only way we know what the population stands for.
+So we can base quorum on the confidentiality.
+The goal is to always minimize the amount of information needed from
+the population in order to confidently determine what actions the population as a whole wants to take.
+This allows faster decision making for non-contentious issues and prevents strategizing with quorum in
+referendums, where if there is a quorum a possible best action for you is not to attend.
+
+\todo[inline]{When we have multiple options?}
+\todo[inline]{Generalization so that majority and super-majority and consensus is just a parameter.}
 
@@
 
+Decoupling the information presentation and voting processes from the debating process in the full decision-making
+pipeline (with the idea being that debating is best done in person, and presenting information/voting is best done online).
+
+\section{Evaluation}
+
+\subsection{Design}
+
+User study/evaluation design.
+To create a controlled evaluation of our solutions, to not depend on general usage of our system among communities,
+we can evaluate the design with the following user study:
+\begin{itemize}
+\item we find a tight community of around 100 people willing to participate in the user study
+\item they should know each other well (to know how to delegate)
+\item we offer them to pay some Amazon credits or something for participation
+\item every member of the community should define delegations
+\item we ask them to split into two groups, group A and group B
+\item group A (target size is around 20% of the community) will meet together or online and discuss various topics through our system
+\item discussed topics some can be completely random, to get people to be familiar with they system, and also have some
+easy topics to decide on, like, what ``should be the color of the sky if we could change it'', and some topics which are relevant to the community
+\item those discussions would be just for evaluation, but community can use the results in any way they find useful
+\item after this discussions and decisions were done we ask everyone from group B to evaluate the results, we present them with two results (blinded, them not knowing which was made in which way)
+\begin{itemize}
+\item results which would be obtained using traditional methods (majority voting without any our improvements)
+\item our results using our system
+\end{itemize}
+\item we ask them to evaluate which one (or both, or none) of decisions they agree with
+\item prediction is that members who have not participated in the decision making process will
+find decisions made through our system more often agreeable
+\item hypothesis here is that what a decision making system is trying to achieve is to find a decision most of the
+whole community can agree with, or like it, even when not everyone participates and not all information is available
+on their positions on a topic
+\end{itemize}
+
 \bibliographystyle{sigchi}
@@
 
+More:
+\begin{itemize}
+\item https://mason.gmu.edu/~rhanson/futarchy.html
+\item http://merkle.com/papers/DAOdemocracyDraft.pdf
+\end{itemize}
+
 \end{document}

### 2016-08-07 paper.tex
TODOs.
Order.
More notes.

@@
 \todo[inline]{This rest of the section is mostly notes. Organize and improve.}
+
+We can often hear that the issue of our democracy is low engagement.
+But this is a fallacy, because we should be instead trying to decrease it.
+The best governance system is one where you do not have to be involved yourself, but it is still doing things
+like you would like to do them.
+% Or even, doing things you might not even know that you would like to be done, especially in a long term.
+We want to outsource the job of coordination between people.
+Sometimes you want to coordinate people yourself (in your company, team), you might enjoy it, but in most cases
+you cannot or do not want to, so you outsource this service, especially at a large scale.
+
+One reason why increasing engagement sounds like a good idea is because current democratic systems have two
+questionable assumptions: participants have time to participate, and knowledge to participate.
+If you do not participate it is your fault.
+Moreover, the legitimacy of decisions is based on participation as well.
+This is why those participating in the these democratic systems have motivation to try to increase participation
+of others (especially if they believe they will still be in majority), or claim they are at fault to not participate.
+
+engagement is then increased only to assure that people like the results
+
+so if we try to minimize engagement, but still make sure that people like the results, then this is the perfect system
+
+and only in the case when they are on the loosing side, then we have to make sure that they understand the reasons why
+they lost (eg. the middle panel in our design) or at least to feel that they participated, in at least some way
+(voting, or just delegating)
+
+but the participation is the last on this list of important factors
+
+the more important thing is that they understand why others were thinking differently
+or that they do not even have to be engaged
+
+we want that people do not have to pay attention to what government is doing
+
+currently, we are in a loop where we have a useless feedback loop, where government in fact would like for people
+to pay some type of attention because they need to be elected, so they have to show like they are doing something,
+but on the other hand they do not want too much engagement, because they do not have things in place to really
+engage with everyone
+
+government should be like infrastructure, you should not have to care about it while it works, and be there to fix
+things if it starts failing, but not even you personally, but have mechanisms and processes in place to find this
 
@@
 
+\todo[inline]{We also have to explain that with non-zero values of $\mathbf{x_w}$ we also get which zero values in
+$\mathbf{x_v}$ are real results and which are zero because delegation could not be computed. One has to use that
+to compute delegated votes as well.}
+
 We pass computed votes and number of votes and abstentions into computation of dynamic quorum to determine
@@
 we can evaluate the design with the following user study:
-\begin{itemize}
+\begin{enumerate}
 \item we find a tight community of around 100 people willing to participate in the user study
@@
 on their positions on a topic
-\end{itemize}
+\end{enumerate}
 

### 2016-08-11 paper.tex
Some questions and answers.

@@
 quantities for the entire voting population, and we can promote clarity and full understanding of the voting items for all.
+
+\todo[inline]{Main questions and answers follows.}
+
+What problem are we solving? Why is it important to solve?
+
+Finding a way to lower attention required but increase legitimacy of democratic decisions.
+Traditionally legitimacy is increased by increasing participation/attention required.
+It is an everlasting problem.
+How to outsource governance and at the same time assure it does not get corrupt
+or tyrannic.
+
+Has anyone else tried to solve it or related problems? How did they try
+and what were the results (subjective or objective)?
+
+Described here: \url{http://mitar.tnode.com/post/73983101095/peer-to-peer-voting-scheme}.
+Some tried delegations, but mostly only to one person.
+Getting issues that you have tree-structure with possible cycles you cannot resolve.
+Often there were reported UI problems as well.
+So just technically improving is not enough, you also have to have UI for the technical solution so that people can trust.
+
+How are you solving the problem?
+Using dynamic quorum, score voting, delegation, and nice integration with UI/UX.
+
+How will you know how well you solved the problem?
+If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
+So currently we have issues with reaching the quorum and many people do not participate.
+I would like for them to have easier time participating (less time required) and as a consequence more people to
+participate collectively.
+If I observe that I think I will know I solved the problem.
 

### 2016-08-11 paper.tex
Paragraphs.
Normalize delegations.

@@
 
-%\pagenumbering{arabic}
+\pagenumbering{arabic}
 
@@
 How will you know how well you solved the problem?
+
 If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
@@
 \todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}
+\todo[inline]{Terminology: discarding a vote if you didn't vote and didn't delegate. Some other term?}
 
@@
 \item Member $M_2$ delegates to $M_3$ with ratio $1.00$ ($r_{2,3}$).
-\item Member $M_3$ delegates to $M_1$ and $M_2$ with ratios $0.40$ ($r_{3,1}$) and $0.40$ ($r_{3,2}$).
+\item Member $M_3$ delegates to $M_1$, $M_2$, and $M_4$ with ratios $0.40$ ($r_{3,1}$), $0.40$ ($r_{3,2}$), and $0.20$ ($r_{3,4}$), respectively.
+\item Member $M_4$ decides not to delegate.
 \end{itemize}
 
-Observe that member $M_3$ decided to use only $0.8$ of their vote in the case they do not vote themselves.
-The rest ($0.2$) counts towards non-voting population.
-In this way a member can decide to participate in voting through delegation, but not fully.
-Member $M_0$ decided to not participate at all and if they do not vote they are counted towards non-voting population.
+The sum of ratios of all delegations of a member has to be $1.0$.
+Members $M_0$ and $M_4$ decided to not participate in delegation and if they do not vote they are
+counted towards non-voting population.
 
@@
     \WE[Math,L=M_2](M3){M2}
+    \EA[Math,L=M_4](M3){M4}
 
@@
     \Edge[label=$0.25$](M3)(M1)
+    \Edge[label=$0.20$](M4)(M3)
 
@@
 V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
-V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} = 0.40 V_{M_1} + 0.40 V_{M_2} \\
+V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} + r_{3,4} V_{M_4} = \\
+& = & 0.40 V_{M_1} + 0.40 V_{M_2} + 0.20 V_{M_4} \\
 \end{array}
@@
 \begin{displaymath}
-\mathrm{D} = \left[ \begin{array}{cccc}
-1 & 0 & 0 & 0 \\
--0.75 & 1 & -0.25 & 0 \\
-0 & 0 & 1 & 0 \\
-0 & -0.40 & -0.40 & 1 \\
+\mathrm{D} = \left[ \begin{array}{ccccc}
+1 & 0 & 0 & 0 & 0 \\
+-0.75 & 1 & -0.25 & 0 & 0 \\
+0 & 0 & 1 & 0 & 0 \\
+0 & -0.40 & -0.40 & 1 & -0.20 \\
+0 & 0 & 0 & 0 & 1 \\
 \end{array} \right]
@@
 0 \\
+0 \\
 \end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
@@
 0.056 \\
+0.0 \\
 \end{array} \right]
 \end{displaymath}
-
-The result (average of scores) without delegation is $\overline{v} = 0.15$ and
-with delegation is $\overline{x} = 0.22$.
-Delegation pushes decision slightly in favor.
 
@@
 towards non-voting population.
-Members might not declare their delegates, or they might declare them without using their vote fully.
+Members might not declare their delegates.
 Other members can delegate to them as well, potentially propagating lack of a vote further.
-For example, if member $M_0$ would not vote, then their vote would be discarded.
-Moreover, we could not use $M_0$'s vote to compute $M_1$'s vote fully, which would influence $M_3$'s vote as well.
+In our example, member $M_4$ has not voted so their vote is discarded.
+As a consequence $M_3$'s vote cannot be computed fully, which influences $M_1$'s vote as well.
 
-One option is that we would normalize ratio in such case, so if $M_0$ does not vote, then $r_{1,3}$ is normalized
-from $0.25$ to $1.0$.
-We could do this automatically for all members, or we could even leave to members to configure
-what to do in such case for them.
-In current version of PeerMind we decided to not handle this situation in any special way and we leave
-discarded votes to propagate.
+In such case we normalize ratios of other delegates, those who voted or for which we can
+compute their votes recursively.
+So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.
 
-\todo[inline]{Or should we be normalizing? In a way we should not normalize if votes are discarded because
-member explicitly delegated so ($M_3$ above), but if votes are discarded because we could not compute
-all delegations, we might want to automatically normalize.}
-
-As a consequence we have to determine how many votes have been discarded to know the size of the non-voting
-population.
-Another issue is also a question of abstentions.
+Another issue is a question of abstentions.
 In PeerMind members can abstain from voting which is different from simply not participating.
@@
 
-Solving $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ gives
-us number of votes $W$ and number of abstentions $A$, respectively, by summing together all values in the
-resulting vectors: $W = \sum \mathbf{x_w}$, $A = \sum \mathbf{x_a}$.
+We solve $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$.
+To normalize votes we compute $x = x_v / x_w$.
+A consequence is that those elements of $x_w$ equal to $0$ make corresponding elements of $x$ undefined.
+Those undefined elements of $x$ we count towards non-voting population.
+We just have to make sure that they are not in fact abstentions.
+We check corresponding elements of $x_a$ and if they are non-zero we count them as abstentions,
+otherwise as non-votes.
+In our example:
 
-\todo[inline]{Is the result then $\frac{\sum \mathbf{x_v}}{W}$ or $\frac{\sum \mathbf{x_v}}{\textrm{number of entries
-in}~\mathbf{x_w}~\textrm{which are non-zero}}$?}
+\begin{displaymath}
+\mathbf{x} = \left[ \begin{array}{c}
+0.7 \\
+0.57 \\
+-0.4 \\
+0.07 \\
+\mathrm{undefined} \\
+\end{array} \right]
+\end{displaymath}
 
-\todo[inline]{We also have to explain that with non-zero values of $\mathbf{x_w}$ we also get which zero values in
-$\mathbf{x_v}$ are real results and which are zero because delegation could not be computed. One has to use that
-to compute delegated votes as well.}
+\todo[inline]{Are we discarding abstentions too much here? If we have 5 people and each of them is computed 0.2 of abstention
+through delegation, while the rest 0.8 is a vote, then we normalize 0.8 to 1.0 for all of them and simply discard 0.2
+of abstention. But otherwise those 0.2 of abstentions would count together towards 1.0. But probably this is not a
+problem because the quorum is still the same: those 5 people have participated through delegation in voting. Only
+if somebody has not participated through delegation in voting we care if they might participated in abstentions
+through voting. To maximize delegated participation.}
 
-We pass computed votes and number of votes and abstentions into computation of dynamic quorum to determine
+The result (average of scores) without delegation is $\overline{v} = 0.15$ and
+with delegation is $\overline{x} = 0.24$.
+Delegation pushes decision slightly in favor.
+
+\todo[inline]{$x = x_v / x_w$ does not really produce the same result as computing with 0.50 as ratio to begin
+with. Why?}
+
+We pass computed votes and number of abstentions into computation of dynamic quorum to determine
 a quorum and confidence, in the same manner as if members voted themselves.

### 2016-08-12 paper.tex
Max: comments + additions

@@
 Some other theoretical properties?
+Computing statistical quorum is an ``embarrassingly parallel'' problem, and so its use would not suffer in large populations.
+Sparsity of delegation matrix also lends to using recently developed efficient matrix computation algorithms.
 
@@
 issue, then in fact the whole population should vote because this is the only way we know what the population stands for.
-So we can base quorum on the confidentiality.
+So we can base quorum on the credibility.
 The goal is to always minimize the amount of information needed from
@@
 \subsection{Delegation}
-
+\todo[inline]{Max: I think delegation graph edges should point in the direction towards the person that the vote is being given to (more intuitive).}
 Because not every member of the community has time or knowledge to vote all the time, each member can delegate
@@
 A delegation network can be described as a system of equations.
+
+\todo[inline]{Max: Double check accuracy of this method, Improve explanation here and perhaps use nicer numbers so readers can follow along certain parts in their heads easily.}
 
@@
 \end{displaymath}
-
+\todo[inline]{Max: FYI, computing who to discard from the population as ``non-voting'' is a separate computation in itself.
+You need to figure out what votes can be calculated partially as a consequence of delegation, and which cannot.
+Use indicator vectors denoting 1 if member has voted, else 0.}
 There are few details we have to look into.
@@
 
+One additional part of the experimental delegation study:
+
+Compute statistical quorum with and without delegations.
+This semester, do not actually allow delegation to be used for passing statistical quorum, but collect data so that we can see how much earlier delegation reached quorum before official statistical quorum.
+(Only count how much earlier at the point where delegation passed quorum and never went below thereafter, thus showing the point where the decision truly was clear.)
+
+Evaluation of statistical quorum (given that we don't officially decide by including delegation this semester): From email: ``Do an internal Cloyne comparison that is more mathematical, showing how many agenda items were decided using fewer actual votes than old 33\% quorum, as a result of statistical quorum.
+This could be measured by average \# votes needed to pass statistical quorum vs.
+\% of those items that get formally passed at council via consent agenda.
+(If it's 100\%, then you can say that we saved X\% of votes on average by using our online voting).
+Then you could also show that some items needed more than 33\% of votes, implying that a hard-line 33\% is also bad.''
+
+Subjective evaluation of entire platform: Council Surveys comparing past two semesters to this semester for other houses vs.
+Cloyne.
+``use the previous 2 semesters' worth of council surveys as control, and a council survey for Cloyne at the end of this semester as results.
+These results would speak to those specific subjective questions and would be one useful thing to look at.''
+
 \bibliographystyle{sigchi}

### 2016-08-12 paper.tex
Fixing delegations.

--- 2016-08-11 paper.tex
@@
 \begin{array}{rcl}
-V_{M_0} & = & 0 \\
+V_{M_0} & = & V_{M_0} \\
 V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
@@
 & = & 0.40 V_{M_1} + 0.40 V_{M_2} + 0.20 V_{M_4} \\
+V_{M_4} & = & V_{M_4} \\
 \end{array}
@@
 
-We solve the matrix equation:
+We solve the matrix equation for $\mathbf{x_v}$:
 
 \begin{displaymath}
-\mathrm{D} \mathbf{x_v} = \mathbf{v}
+\mathrm{D} \mathbf{x_v} = \mathbf{v},\quad \mathbf{v} = \left[v_{i}\right] = \begin{dcases*}
+ V_{M_i} & if $M_i$ voted \\
+ 0 & otherwise \\
+\end{dcases*}
 \end{displaymath}
@@
 \mathbf{v} = \left[ \begin{array}{c}
-0.7 \\
+0.70 \\
 0 \\
@@
 \end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
-0.7 \\
+0.70 \\
 0.54 \\
--0.4 \\
+-0.40 \\
 0.056 \\
-0.0 \\
+0 \\
 \end{array} \right]
@@
 There are few details we have to look into.
-Despite delegation some votes or parts of votes can still happen to be discarded and should count
-towards non-voting population.
-Members might not declare their delegates.
-Other members can delegate to them as well, potentially propagating lack of a vote further.
+Despite delegation some votes or parts of votes can still happen to be discarded.
+Members who did not vote might have not declared their delegates.
+Other members could delegate to them as well, potentially propagating lack of a vote further.
 In our example, member $M_4$ has not voted so their vote is discarded.
@@
 
-In such case we normalize ratios of other delegates, those who voted or for which we can
+In such case we want to normalize ratios of other delegates, those who voted or for which we can
 compute their votes recursively.
@@
 Abstentions are counted towards quorum, but they do not contribute to the result.
-To address this, we define two vectors:
+
+Moreover, from $\mathbf{x_v}$ alone we cannot see which elements represent vote $0$ and which
+represent that it was not possible to compute a vote for a member.
+
+To address all this, we define two vectors:
 
@@
 
-We solve $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$.
-To normalize votes we compute $x = x_v / x_w$.
-A consequence is that those elements of $x_w$ equal to $0$ make corresponding elements of $x$ undefined.
-Those undefined elements of $x$ we count towards non-voting population.
-We just have to make sure that they are not in fact abstentions.
-We check corresponding elements of $x_a$ and if they are non-zero we count them as abstentions,
-otherwise as non-votes.
-In our example:
+We solve $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ for
+$\mathbf{x_w}$ and $\mathbf{x_a}$, respectively.
+Non-zero elements of $\mathbf{x_w}$ correspond to members who voted or for which we can compute
+a vote.
+We define a normalized matrix:
+
+\begin{displaymath}
+\mathrm{D'} = \left[d'_{ij}\right] = \begin{dcases*}
+ 1 & if $i = j$ \\
+ 0 & if $x_{w_j} = 0$ ($V_{M_j}$ cannot be computed) \\
+ -\frac{r_{i,j}}{R_i} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with normalized ratio $\frac{r_{i,j}}{R_i}$, $R_i = \sum_{x_{w_k} \ne 0} r_{i,k}$} \\
+ 0 & otherwise \\
+\end{dcases*}
+\end{displaymath}
+
+To compute votes with normalized ratios for only those delegates who voted or for which we can compute votes recursively,
+we solve $\mathrm{D'} \mathbf{x} = \mathbf{v}$ for $\mathbf{x}$:
 
@@
 \mathbf{x} = \left[ \begin{array}{c}
-0.7 \\
-0.57 \\
--0.4 \\
+0.70 \\
+0.54 \\
+-0.40 \\
 0.07 \\
-\mathrm{undefined} \\
+0 \\
 \end{array} \right]
@@
 
-\todo[inline]{Are we discarding abstentions too much here? If we have 5 people and each of them is computed 0.2 of abstention
-through delegation, while the rest 0.8 is a vote, then we normalize 0.8 to 1.0 for all of them and simply discard 0.2
-of abstention. But otherwise those 0.2 of abstentions would count together towards 1.0. But probably this is not a
-problem because the quorum is still the same: those 5 people have participated through delegation in voting. Only
-if somebody has not participated through delegation in voting we care if they might participated in abstentions
-through voting. To maximize delegated participation.}
+If an element of $\mathbf{x_w}$ is equal to $0$, then we could not compute a vote for that member and we count
+it towards non-voting population.
+We just have to make sure that they are not in fact abstentions.
+We check corresponding elements of $\mathbf{x_a}$ and if they are non-zero we count them as abstentions,
+otherwise as non-votes.
 
 The result (average of scores) without delegation is $\overline{v} = 0.15$ and
-with delegation is $\overline{x} = 0.24$.
+with delegation is $\overline{x} = 0.23$.
 Delegation pushes decision slightly in favor.
-
-\todo[inline]{$x = x_v / x_w$ does not really produce the same result as computing with 0.50 as ratio to begin
-with. Why?}
 

### 2016-08-12 paper.tex
Added answers.

@@
 Cohesion of the group vs. accuracy.
-Maybe display only number of votes and confidence, maybe confidence direction or graph, but not average votes.
+Maybe display only number of votes and credibility, maybe credibility direction or graph, but not average votes.
 
@@
 learn how your vote is used?
-\item Knowing how much power you have vs.
-not knowing -- does it change how people behave?
-(Knowing might corrupt you, but also motivate you to do more thorough work voting because you know other depend on you.)
+\item Knowing how much power you have vs. not knowing -- does it change how people behave? (Knowing might corrupt you,
+but also motivate you to do more thorough work voting because you know other depend on you.)
 \end{itemize}
@@
 Some other theoretical properties?
-Computing statistical quorum is an ``embarrassingly parallel'' problem, and so its use would not suffer in large populations.
-Sparsity of delegation matrix also lends to using recently developed efficient matrix computation algorithms.
+Computing statistical quorum is an ``embarrassingly parallel'' problem, and so its
+use would not suffer in large populations.
+Sparsity of delegation matrix also lends to using recently developed
+efficient matrix computation algorithms.
 
@@
 \subsection{Delegation}
-\todo[inline]{Max: I think delegation graph edges should point in the direction towards the person that the vote is being given to (more intuitive).}
+\todo[inline]{Max: I think delegation graph edges should point in the direction towards the person that the vote is
+being given to (more intuitive).
+\\
+Mitar: We are not giving votes to people, we are delegating a decision to people, but votes are computed for each person
+separately based on how delegates voted themselves. The idea is that you ask a delegate how the voted and then compute
+your vote. The do not get your vote to do anything with it. This is why arrows are like that. Maybe the word
+"delegation" is confusing and we should maybe use surrogate or some other word.}
 Because not every member of the community has time or knowledge to vote all the time, each member can delegate
@@
 
-\todo[inline]{Max: Double check accuracy of this method, Improve explanation here and perhaps use nicer numbers so readers can follow along certain parts in their heads easily.}
+\todo[inline]{Max: Double check accuracy of this method, Improve explanation here and perhaps use nicer numbers so
+readers can follow along certain parts in their heads easily.}
 
@@
 \end{displaymath}
+
 \todo[inline]{Max: FYI, computing who to discard from the population as ``non-voting'' is a separate computation in itself.
 You need to figure out what votes can be calculated partially as a consequence of delegation, and which cannot.
-Use indicator vectors denoting 1 if member has voted, else 0.}
+Use indicator vectors denoting 1 if member has voted, else 0.
+\\
+Mitar: This is exactly what $\mathbf{w}$ is later on.}
+
 There are few details we have to look into.
@@
 We pass computed votes and number of abstentions into computation of dynamic quorum to determine
-a quorum and confidence, in the same manner as if members voted themselves.
+a quorum and credibility, in the same manner as if members voted themselves.
 
@@
 Compute statistical quorum with and without delegations.
-This semester, do not actually allow delegation to be used for passing statistical quorum, but collect data so that we can see how much earlier delegation reached quorum before official statistical quorum.
-(Only count how much earlier at the point where delegation passed quorum and never went below thereafter, thus showing the point where the decision truly was clear.)
+This semester, do not actually allow delegation to be used for passing statistical quorum, but collect data so that
+we can see how much earlier delegation reached quorum before official statistical quorum.
+(Only count how much earlier at the point where delegation passed quorum and never went below thereafter, thus
+showing the point where the decision truly was clear.)
 
-Evaluation of statistical quorum (given that we don't officially decide by including delegation this semester): From email: ``Do an internal Cloyne comparison that is more mathematical, showing how many agenda items were decided using fewer actual votes than old 33\% quorum, as a result of statistical quorum.
-This could be measured by average \# votes needed to pass statistical quorum vs.
-\% of those items that get formally passed at council via consent agenda.
-(If it's 100\%, then you can say that we saved X\% of votes on average by using our online voting).
+Evaluation of statistical quorum (given that we don't officially decide by including delegation this semester):
+From email: ``Do an internal Cloyne comparison that is more mathematical, showing how many agenda items were decided
+using fewer actual votes than old 33\% quorum, as a result of statistical quorum.
+This could be measured by average \# votes needed to pass statistical quorum vs. \% of those items that get
+formally passed at council via consent agenda.
+(If it's 100\%, then you can say that we saved X\% of votes on
+average by using our online voting).
 Then you could also show that some items needed more than 33\% of votes, implying that a hard-line 33\% is also bad.''
 
-Subjective evaluation of entire platform: Council Surveys comparing past two semesters to this semester for other houses vs.
-Cloyne.
-``use the previous 2 semesters' worth of council surveys as control, and a council survey for Cloyne at the end of this semester as results.
-These results would speak to those specific subjective questions and would be one useful thing to look at.''
+Subjective evaluation of entire platform: Council Surveys comparing past two semesters to this semester for other
+houses vs. Cloyne.
+``use the previous 2 semesters' worth of council surveys as control, and a council survey for
+Cloyne at the end of this semester as results.
+These results would speak to those specific subjective questions
+and would be one useful thing to look at.''
 

### 2016-08-16 paper.tex
Improved description of delegation.

@@
 \usepackage{mathtools}
+\usepackage{amsmath}
 
@@
 \begin{itemize}
-\item Member $M_0$ decides not to delegate at all.
-\item Member $M_1$ delegates to $M_0$ and $M_3$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectively.
-\item Member $M_2$ delegates to $M_3$ with ratio $1.00$ ($r_{2,3}$).
-\item Member $M_3$ delegates to $M_1$, $M_2$, and $M_4$ with ratios $0.40$ ($r_{3,1}$), $0.40$ ($r_{3,2}$), and $0.20$ ($r_{3,4}$), respectively.
-\item Member $M_4$ decides not to delegate.
+\item Member $M_1$ decides not to delegate at all.
+\item Member $M_2$ delegates to $M_1$ and $M_4$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectively.
+\item Member $M_3$ delegates to $M_4$ with ratio $1.00$ ($r_{2,3}$).
+\item Member $M_4$ delegates to $M_2$, $M_3$, and $M_5$ with ratios $0.40$ ($r_{3,1}$), $0.40$ ($r_{3,2}$), and $0.20$ ($r_{3,4}$), respectively.
+\item Member $M_5$ decides not to delegate.
 \end{itemize}
@@
 The sum of ratios of all delegations of a member has to be $1.0$.
-Members $M_0$ and $M_4$ decided to not participate in delegation and if they do not vote they are
+Members $M_1$ and $M_5$ decided not to participate in delegation and if they do not vote they are
 counted towards non-voting population.
@@
 
-    \Vertex[Math,L=M_0]{M0}
-    \EA[Math,L=M_1](M0){M1}
-    \SO[Math,L=M_3](M1){M3}
-    \WE[Math,L=M_2](M3){M2}
-    \EA[Math,L=M_4](M3){M4}
+    \Vertex[Math,L=M_1]{M1}
+    \EA[Math,L=M_2](M1){M2}
+    \SO[Math,L=M_4](M2){M4}
+    \WE[Math,L=M_3](M4){M3}
+    \EA[Math,L=M_5](M4){M5}
 
-    \Edge[label=$0.40$](M1)(M3)
-    \Edge[label=$0.75$](M0)(M1)
-    \Edge[label=$0.25$](M3)(M1)
-    \Edge[label=$0.20$](M4)(M3)
+    \Edge[label=$0.40$](M2)(M4)
+    \Edge[label=$0.75$](M1)(M2)
+    \Edge[label=$0.25$](M4)(M2)
+    \Edge[label=$0.20$](M5)(M4)
 
     \tikzset{EdgeStyle/.append style={bend right}}
-    \Edge[label=$1.00$](M3)(M2)
-    \Edge[label=$0.40$](M2)(M3)
+    \Edge[label=$1.00$](M4)(M3)
+    \Edge[label=$0.40$](M3)(M4)
   \end{tikzpicture}
   \caption{Example delegation network. Edges are directed towards a member who is delegating, representing how their
-  vote is computed from delegates' votes. E.g., $V_{M_1} = 0.75 V_{M_0} + 0.25 V_{M_3}$}~\label{fig:delegation-network}
+  vote is computed from delegates' votes. E.g., $V_2 = 0.75 V_1 + 0.25 V_4$}~\label{fig:delegation-network}
 \end{figure}
@@
 We direct edges towards a member who is delegating, to represent how their vote is computed from delegates' votes.
-For example, member $M_1$ delegates the decision to members $M_0$ and $M_3$.
-$M_0$ and $M_3$ become $M_1$'s delegates.
-As a consequence, $M_1$'s vote $V_{M_1}$ is computed from $M_0$'s and $M_3$'s votes ($V_{M_0}$ and $V_{M_3}$).
+For example, member $M_2$ delegates the decision to members $M_1$ and $M_4$.
+$M_1$ and $M_4$ become $M_2$'s delegates.
+As a consequence, $M_2$'s vote $V_2$ is computed from $M_1$'s and $M_4$'s computed votes ($V_1$ and $V_4$)
+if $M_2$ does not vote themselves.
+
+With $v_i$ we represent a vote of a member $M_i$ if they voted, otherwise it is $0$.
+$V_i$ represents their vote, if they have voted, or a computed vote based on their delegations.
 
@@
 \begin{array}{rcl}
-V_{M_0} & = & V_{M_0} \\
-V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
-V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
-V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} + r_{3,4} V_{M_4} = \\
-& = & 0.40 V_{M_1} + 0.40 V_{M_2} + 0.20 V_{M_4} \\
-V_{M_4} & = & V_{M_4} \\
+V_1 & = & \begin{dcases*}
+ v_1 & if $M_1$ votes \\
+ 0 & otherwise \\
+\end{dcases*} \\
+V_2 & = & \begin{dcases*}
+ v_2 & if $M_2$ votes \\
+ r_{1,0} V_1 + r_{1,3} V_4 = 0.75 V_1 + 0.25 V_4 & otherwise \\
+\end{dcases*} \\
+V_3 & = & \begin{dcases*}
+ v_3 & if $M_3$ votes \\
+ r_{2,3} V_4 = 1.00 V_4 & otherwise \\
+\end{dcases*} \\
+V_4 & = & \begin{dcases*}
+ v_4 & if $M_4$ votes \\
+ r_{3,1} V_2+ r_{3,2} V_3 + r_{3,4} V_5 = & otherwise \\
+ = 0.40 V_2 + 0.40 V_3 + 0.20 V_5 & \\
+\end{dcases*} \\
+V_5 & = & \begin{dcases*}
+ v_5 & if $M_5$ votes \\
+ 0 & otherwise \\
+\end{dcases*} \\
 \end{array}
@@
 
-When only some members vote ($\mathbf{v}$) we can use a delegation network to compute votes
-of all members ($\mathbf{x}$).
-We convert a delegation network to a matrix $\mathrm{D}$:
+When only some members vote we can use a delegation network to compute votes
+of all members.
+We convert a delegation network to a matrix $D$:
 
 \begin{displaymath}
-\mathrm{D} = \left[d_{ij}\right] = \begin{dcases*}
+D = \left[d_{ij}\right] = \begin{dcases*}
  1 & if $i = j$ \\
@@
 
-We solve the matrix equation for $\mathbf{x_v}$:
+We solve the matrix equation for $\boldsymbol{V}$:
 
 \begin{displaymath}
-\mathrm{D} \mathbf{x_v} = \mathbf{v},\quad \mathbf{v} = \left[v_{i}\right] = \begin{dcases*}
- V_{M_i} & if $M_i$ voted \\
- 0 & otherwise \\
-\end{dcases*}
+\mathrm{D} \boldsymbol{V} = \boldsymbol{v},\quad \boldsymbol{v} = \left[v_{i}\right],\quad \boldsymbol{V} = \left[V_{i}\right]
 \end{displaymath}
 
-To continue with our example, let's say that $M_0$ and $M_2$ voted for a decision with $0.7$ and against with $0.4$,
+To continue with our example, let's say that $M_1$ and $M_3$ voted for a decision with $0.7$ and against with $0.4$,
 respectively.
-We are using score voting on $[-1, 1]$ range, so those votes are represented with $V_{M_0} = 0.7$ and
-$V_{M_2} = -0.4$.
+We are using score voting on $[-1, 1]$ range, so those votes are represented with $v_1 = V_1 = 0.7$ and
+$v_3 = V_3 = -0.4$.
 
@@
 \begin{displaymath}
-\mathbf{v} = \left[ \begin{array}{c}
-0.70 \\
+\boldsymbol{v} = \left[ \begin{array}{c}
+0.7 \\
 0 \\
@@
 0 \\
-\end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
+\end{array} \right],\quad \boldsymbol{V} = \left[ \begin{array}{c}
 0.70 \\
@@
 \\
-Mitar: This is exactly what $\mathbf{w}$ is later on.}
+Mitar: This is exactly what $\boldsymbol{w}$ is later on.}
 
@@
 Other members could delegate to them as well, potentially propagating lack of a vote further.
-In our example, member $M_4$ has not voted so their vote is discarded.
-As a consequence $M_3$'s vote cannot be computed fully, which influences $M_1$'s vote as well.
-
-In such case we want to normalize ratios of other delegates, those who voted or for which we can
-compute their votes recursively.
-So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.
+In our example, member $M_5$ has not voted so their vote is discarded.
+As a consequence $M_4$'s vote cannot be computed fully, which influences $M_2$'s vote as well.
 
@@
 
-Moreover, from $\mathbf{x_v}$ alone we cannot see which elements represent vote $0$ and which
+Moreover, from $\boldsymbol{V}$ alone we cannot see which elements represent vote $0$ and which
 represent that it was not possible to compute a vote for a member.
@@
 \begin{displaymath}
-\mathbf{w} = \left[w_{i}\right] = \begin{dcases*}
+\boldsymbol{w} = \left[w_i\right] = \begin{dcases*}
  1 & if $M_i$ voted \\
  0 & otherwise \\
-\end{dcases*},\quad \mathbf{a} = \left[a_{i}\right] = \begin{dcases*}
+\end{dcases*},\quad \boldsymbol{a} = \left[a_i\right] = \begin{dcases*}
  1 & if $M_i$ abstained \\
@@
 
-We solve $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ for
-$\mathbf{x_w}$ and $\mathbf{x_a}$, respectively.
-Non-zero elements of $\mathbf{x_w}$ correspond to members who voted or for which we can compute
+We solve $D \boldsymbol{W} = \boldsymbol{w}$ and $\mathrm{D} \boldsymbol{A} = \boldsymbol{a}$ for
+$\boldsymbol{W} = \left[W_{i}\right]$ and $\boldsymbol{A} = \left[A_{i}\right]$, respectively.
+
+Non-zero elements of $\boldsymbol{W}$ correspond to members who voted or for which we can compute
 a vote.
+
+\todo[inline]{If there is a clique in delegation, without any vote inside a clique, the matrix becomes singular.
+We should find a way to detect such rows and set corresponding elements of $\boldsymbol{W}$ to $0$ (or the row
+to identity row before we try to compute $\boldsymbol{W}$).
+This is exactly what we want in fact, that those in such a clique are seen as non-computable.
+Then the rest works.}
+
 We define a normalized matrix:
@@
 \begin{displaymath}
-\mathrm{D'} = \left[d'_{ij}\right] = \begin{dcases*}
+D' = \left[d'_{ij}\right] = \begin{dcases*}
  1 & if $i = j$ \\
- 0 & if $x_{w_j} = 0$ ($V_{M_j}$ cannot be computed) \\
- -\frac{r_{i,j}}{R_i} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with normalized ratio $\frac{r_{i,j}}{R_i}$, $R_i = \sum_{x_{w_k} \ne 0} r_{i,k}$} \\
+ 0 & if $W_j = 0$ ($V_j$ cannot be computed) \\
+ -r'_{i,j} & if $M_i$ did not vote \\
  0 & otherwise \\
@@
 
-To compute votes with normalized ratios for only those delegates who voted or for which we can compute votes recursively,
-we solve $\mathrm{D'} \mathbf{x} = \mathbf{v}$ for $\mathbf{x}$:
+Where:
 
 \begin{displaymath}
-\mathbf{x} = \left[ \begin{array}{c}
+r'_{i,j} = \frac{r_{i,j}}{R_i},\quad R_i = \sum_{W_k \ne 0} r_{i,k}
+\end{displaymath}
+
+We normalize ratios to only those delegates who voted or for which we can
+compute their votes recursively.
+So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.
+
+To compute votes with normalized ratios for only those delegates who voted or for which we can compute votes recursively,
+we solve $D' \boldsymbol{V} = \boldsymbol{v}$ for $\boldsymbol{V}$:
+
+\begin{displaymath}
+\mathbf{V} = \left[ \begin{array}{c}
 0.70 \\
@@
 
-If an element of $\mathbf{x_w}$ is equal to $0$, then we could not compute a vote for that member and we count
-it towards non-voting population.
+If an element of $\boldsymbol{W}$ is equal to $0$, then we could not compute a vote for the corresponding
+member and we count it towards the non-voting population.
 We just have to make sure that they are not in fact abstentions.
-We check corresponding elements of $\mathbf{x_a}$ and if they are non-zero we count them as abstentions,
-otherwise as non-votes.
+For elements $\boldsymbol{W} = 0$ we check their corresponding elements of $\boldsymbol{A}$ and if they are
+non-zero we count them as abstentions, otherwise as non-votes.
 
-The result (average of scores) without delegation is $\overline{v} = 0.15$ and
-with delegation is $\overline{x} = 0.23$.
-Delegation pushes decision slightly in favor.
+\begin{itemize}
+\item Number of directly voting members: $\left| v \right|$
+\item Number of voting members: $\left| V \right| = \left| W_k \ne 0 \right|$
+\item Number of abstentions: $\left| A \right| = \left| W_k = 0 \land A_k \ne 0 \right|$
+\item Number of non-voting members: $\left| W_k = 0 \land A_k = 0 \right|$
+\end{itemize}
 
-We pass computed votes and number of abstentions into computation of dynamic quorum to determine
+The result (average of scores) without delegation is:
+
+\begin{displaymath}
+\overline{v} = \frac{\sum_{M_k voted} v_k}{\left| v \right|} = 0.15
+\end{displaymath}
+
+With delegation:
+
+\begin{displaymath}
+\overline{V} = \frac{\sum_{W_k \ne 0} V_k}{\left| V \right|} = 0.23
+\end{displaymath}
+
+In our example, delegation slightly pushes decision in favor.
+
+We pass computed votes $V$ and number of abstentions $\left| A \right|$ into computation of dynamic quorum to determine
 a quorum and credibility, in the same manner as if members voted themselves.
+
+\todo[inline]{Do we want really to compute quorum using delegation, or should quorum be computed
+using direct votes only?}
 

### 2016-08-17 paper.tex
More questions.
Small improvements.

@@
 
-\todo[inline]{Main questions and answers follows.}
+\todo[inline]{Main questions and answers follow.}
 
-What problem are we solving? Why is it important to solve?
+What problem are we solving? Why is it important to solve? What is the story?
 
@@
 
-How are you solving the problem?
+How are you solving the problem? What technologies are you using?
+
 Using dynamic quorum, score voting, delegation, and nice integration with UI/UX.
 
-How will you know how well you solved the problem?
+How will you know how well you solved the problem? How will those technologies address the issues?
 
@@
 We can often hear that the issue of our democracy is low engagement.
-But this is a fallacy, because we should be instead trying to decrease it.
+But this is a fallacy, because we should be instead trying to decrease engagement.
 The best governance system is one where you do not have to be involved yourself, but it is still doing things
@@
 and only in the case when they are on the loosing side, then we have to make sure that they understand the reasons why
-they lost (eg. the middle panel in our design) or at least to feel that they participated, in at least some way
+they lost (e.g.,\ the middle panel in our design) or at least to feel that they participated, in at least some way
 (voting, or just delegating)
@@
 \item attention (at same or better decisions)
-\item bad assumptions for democracy - time, knowledge, it s rational to know that you do not know, common practice is to ask friends
+\item bad assumptions for democracy - time, knowledge, it is rational to know that you do not know, common practice is to ask friends
 \item better decisions (more people consulted, attention directed where it should be)
@@
 
-Performance/theory/scaling in terms in computational (how big could effectively community can be to be able to compute it).
+Performance/theory/scaling in computational terms (how big could effectively community can be to be able to compute it).
 Some other theoretical properties?
@@
 use would not suffer in large populations.
-Sparsity of delegation matrix also lends to using recently developed
+Sparsity of delegation matrix also enables use of (recently developed?)
 efficient matrix computation algorithms.
@@
 \item Member $M_1$ decides not to delegate at all.
-\item Member $M_2$ delegates to $M_1$ and $M_4$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectively.
-\item Member $M_3$ delegates to $M_4$ with ratio $1.00$ ($r_{2,3}$).
-\item Member $M_4$ delegates to $M_2$, $M_3$, and $M_5$ with ratios $0.40$ ($r_{3,1}$), $0.40$ ($r_{3,2}$), and $0.20$ ($r_{3,4}$), respectively.
+\item Member $M_2$ delegates to $M_1$ and $M_4$ with ratios $0.75$ ($r_{2,1}$) and $0.25$ ($r_{2,4}$), respectively.
+\item Member $M_3$ delegates to $M_4$ with ratio $1.00$ ($r_{3,4}$).
+\item Member $M_4$ delegates to $M_2$, $M_3$, and $M_5$ with ratios $0.40$ ($r_{4,2}$), $0.40$ ($r_{4,3}$), and $0.20$ ($r_{4,5}$), respectively.
 \item Member $M_5$ decides not to delegate.
@@
 
-With $v_i$ we represent a vote of a member $M_i$ if they voted, otherwise it is $0$.
-$V_i$ represents their vote, if they have voted, or a computed vote based on their delegations.
+With $v_i$ we denote a vote of a member $M_i$ if they voted, otherwise it is $0$.
+$V_i$ denotes their vote, if they have voted, or a computed vote based on their delegations, if they have not voted.
 
@@
  v_2 & if $M_2$ votes \\
- r_{1,0} V_1 + r_{1,3} V_4 = 0.75 V_1 + 0.25 V_4 & otherwise \\
+ r_{2,1} V_1 + r_{2,4} V_4 = 0.75 V_1 + 0.25 V_4 & otherwise \\
 \end{dcases*} \\
@@
  v_3 & if $M_3$ votes \\
- r_{2,3} V_4 = 1.00 V_4 & otherwise \\
+ r_{3,4} V_4 = 1.00 V_4 & otherwise \\
 \end{dcases*} \\
@@
  v_4 & if $M_4$ votes \\
- r_{3,1} V_2+ r_{3,2} V_3 + r_{3,4} V_5 = & otherwise \\
+ r_{4,2} V_2+ r_{4,3} V_3 + r_{4,5} V_5 = & otherwise \\
  = 0.40 V_2 + 0.40 V_3 + 0.20 V_5 & \\
@@
 \begin{displaymath}
-\mathrm{D} \boldsymbol{V} = \boldsymbol{v},\quad \boldsymbol{v} = \left[v_{i}\right],\quad \boldsymbol{V} = \left[V_{i}\right]
+D \boldsymbol{V} = \boldsymbol{v},\quad \boldsymbol{v} = \left[v_{i}\right],\quad \boldsymbol{V} = \left[V_{i}\right]
 \end{displaymath}
@@
 respectively.
-We are using score voting on $[-1, 1]$ range, so those votes are represented with $v_1 = V_1 = 0.7$ and
+We are using score voting on $[-1, 1]$ range, so those votes are $v_1 = V_1 = 0.7$ and
 $v_3 = V_3 = -0.4$.
@@
 \begin{displaymath}
-\mathrm{D} = \left[ \begin{array}{ccccc}
+D = \left[ \begin{array}{ccccc}
 1 & 0 & 0 & 0 & 0 \\
@@
 
-We solve $D \boldsymbol{W} = \boldsymbol{w}$ and $\mathrm{D} \boldsymbol{A} = \boldsymbol{a}$ for
+We solve $D \boldsymbol{W} = \boldsymbol{w}$ and $D \boldsymbol{A} = \boldsymbol{a}$ for
 $\boldsymbol{W} = \left[W_{i}\right]$ and $\boldsymbol{A} = \left[A_{i}\right]$, respectively.
@@
 
-\todo[inline]{If there is a clique in delegation, without any vote inside a clique, the matrix becomes singular.
+\todo[inline]{If there is a clique/cycle/closed subgroup in delegation, without any vote inside a clique/cycle/closed subgroup, the matrix becomes singular.
 We should find a way to detect such rows and set corresponding elements of $\boldsymbol{W}$ to $0$ (or the row
 to identity row before we try to compute $\boldsymbol{W}$).
-This is exactly what we want in fact, that those in such a clique are seen as non-computable.
-Then the rest works.}
+This is exactly what we want in fact, that those in such a clique/cycle/closed subgroup are seen as non-computable.
+Then the rest works.
+We should describe a problem, maybe with an example.}
 
@@
 compute their votes recursively.
-So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.
+So $r_{4,2}$ and $r_{4,3}$ are normalized from $0.40$ to $0.50$.
 
@@
 \begin{itemize}
-\item Number of directly voting members: $\left| v \right|$
-\item Number of voting members: $\left| V \right| = \left| W_k \ne 0 \right|$
-\item Number of abstentions: $\left| A \right| = \left| W_k = 0 \land A_k \ne 0 \right|$
-\item Number of non-voting members: $\left| W_k = 0 \land A_k = 0 \right|$
+\item Number of directly voting members: $N_v = \left| \{ k: M_k~\mathrm{voted} \} \right|$
+\item Number of voting members: $N_V = \left| \{ k: W_k \ne 0 \} \right|$
+\item Number of abstentions: $N_A = \left| \{ k:  W_k = 0 \land A_k \ne 0 \} \right|$
+\item Number of non-voting members: $\left| \{ k: W_k = 0 \land A_k = 0 \} \right|$
 \end{itemize}
@@
 \begin{displaymath}
-\overline{v} = \frac{\sum_{M_k voted} v_k}{\left| v \right|} = 0.15
+\overline{v} = \frac{\sum_{M_k~\mathrm{voted}} v_k}{N_v} = 0.15
 \end{displaymath}
@@
 \begin{displaymath}
-\overline{V} = \frac{\sum_{W_k \ne 0} V_k}{\left| V \right|} = 0.23
+\overline{V} = \frac{\sum_{W_k \ne 0} V_k}{N_V} = 0.23
 \end{displaymath}
@@
 
-We pass computed votes $V$ and number of abstentions $\left| A \right|$ into computation of dynamic quorum to determine
+We pass computed votes $V$ and number of abstentions $N_A$ into computation of dynamic quorum to determine
 a quorum and credibility, in the same manner as if members voted themselves.

### 2016-10-13 paper.tex
statistical quorum documentation

@@
 
-\subsection{Dynamic quorum}
+\subsection{Statistical Quorum}
 
-Using statistics to compute dynamic quorum based on how people vote using hypothesis testing.
-If population all votes for one position, then quorum can be much lower than if population is completely split on an
-issue, then in fact the whole population should vote because this is the only way we know what the population stands for.
-So we can base quorum on the credibility.
-The goal is to always minimize the amount of information needed from
-the population in order to confidently determine what actions the population as a whole wants to take.
-This allows faster decision making for non-contentious issues and prevents strategizing with quorum in
-referendums, where if there is a quorum a possible best action for you is not to attend.
+\subsubsection*{Background}
+A population of voters needs to collectively decide which item (if any) to take action on out of a set of competing items.
+In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a winning item, while also demonstrating that the winning item is preferred over all other competing items.
+We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that the population does not take undesired actions, which in this case encompasses taking action on an item that is not supported by a majority of the population and/or taking action on an item that is not clearly preferred over all other competing alternatives.
 
-\todo[inline]{When we have multiple options?}
-\todo[inline]{Generalization so that majority and super-majority and consensus is just a parameter.}
+\subsubsection*{Problem Statement}
+We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority, $90\%$ for a consensus).
+In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater than the bias towards voting ``yes'' on all other competing items.
 
-\subsection{Score voting}
+\subsubsection*{Assumptions}
+We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
+
+Without any votes, we assume that this bias may be any value between $0\%$ and $100\%$ with equal probability.
+
+\subsubsection*{Voting}
+We allow for score voting where each member of the population may choose any real number in the range $[-1,+1]$ to represent their vote, with $-1$ representing opposition and $+1$ representing support.
+In the theory subsection below, we rescale votes to be in the range $[0,1]$ via the notation $\frac{votes + 1}{2}$.
+
+We allow voters to ``abstain'' from voting.
+We treat abstentions as indicating that a voter is removing themselves from the population of voters for an item.
+It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$ (indicating indifference).
+
+\subsubsection*{Bayesian Theory}
+To properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial: \url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
+
+The above tutorial is very similar to the problem we have at hand.
+The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises naturally when considering samples from a finite population.
+
+After extracting the sufficient statistics from raw votes --- namely the average score and the number of abstentions --- we can adjust the effective population size and corresponding threshold.
+In addition, we can calculate the effective number of ``yes'' and ``no'' votes, which one can think of as the number of Heads and Tails in the aforementioned tutorial.
+
+Deriving our expression of interest now immediately follows:
+
+Let
+\begin{description}
+\begin{itemize}
+\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing motions,
+\item[$N = $] population size,
+\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the raw votes for motion $X_i$, excluding ``abstentions'',
+\item[$N_i = $] $(N - $ number of abstentions on $X_i)$, the effective population voting on motion $X_i$,
+\item[$\alpha_i = $] $\sum\limits_{v \in X_i} (\frac{v+1}{2})$, the effective number of ``yes'' votes on motion $X_i$,
+\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
+\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
+\item[$p_i = $] true latent population bias towards voting ``yes'' on $X_i$.
+\end{itemize}
+\end{description}
+
+Often, we have a single motion open for voting.  In this scenario, $C=\{X_1\}$ and we simply wish to compute:
+\begin{align}\label{eq1}
+\Pr(\alpha^*_1 > T_1 \mid X_1),
+\end{align}
+where
+\begin{description}
+\begin{itemize}
+\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
+\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote (e.g. $f=\frac{1}{2}$ for a simple majority).
+\end{itemize}
+\end{description}
+Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
+\begin{align*}
+\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \Pr(\alpha^*_1 \leq T_1 \mid X_1),\\
+\Pr(\alpha^*_1 \leq T_1 \mid X_1) &= \int_0^1 \Pr(\alpha^*_1 \leq T_1 \mid p_1)\cdot \Pr(p_1 \mid X_1) \mathrm{d}p_1\\
+&= \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{K-i} {(1-p_1)}^{i} \\
+& \qquad\qquad\quad \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1,
+\end{align*}
+where
+\begin{description}
+\begin{itemize}
+\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that $X_i$ would remain below threshold.
+\end{itemize}
+\end{description}
+
+Hence, we have
+\begin{align}\label{eq2}
+& \Pr(\alpha^*_1 > T_1 \mid X_1) = \nonumber\\
+& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{K_1-i} {(1-p_1)}^{i} \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
+\end{align}
+
+In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to consider.
+Without loss of generality, assume that $X_1$ is the most popular motion within $C$.
+We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other $X_i  \in C, i \neq 1$:
+
+\begin{align}\label{eq3}
+\Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)
+\end{align}
+
+Using a safe assumption that votes on motions are independent across motions, we can decompose \eqref{eq3} as follows:
+\begin{align*}
+& \Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)\\
+=& \prod_{i=2}^n \Pr(\alpha^*_1 > \alpha^*_i \mid C)\\
+=& \prod_{i=2}^n \iint\limits_{p_1, p_i} \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \Pr(p_1 \mid X_1) \cdot \Pr(p_i \mid X_i) \mathrm{d}p_i \mathrm{d}p_1\\
+=& \prod_{i=2}^n \int_0^1 \int_0^1 \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \\
+& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1\\
+\end{align*}
+Deriving an expression for $\Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)$, we get:
+\begin{align*}
+& \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)\\
+=& 1 - \Pr(\alpha^*_1 \leq \alpha^*_i \mid p_1, p_i)\\
+=& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
+\end{align*}
+where
+\begin{description}
+\begin{itemize}
+\item[$v_{i,j} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
+\end{itemize}
+\end{description}
+
+Denoting
+\begin{align}\label{eq4}
+& \Phi_i(p_1,p_i) {\overset {\underset {\mathrm {def} }{}}{=}} \nonumber\\
+& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
+\end{align}
+we can simplify \eqref{eq3} to:
+\begin{align}\label{eq5}
+& \Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)\nonumber\\
+&= \prod_{i=2}^n \int_0^1 \int_0^1 \Phi_i(p_1,p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \nonumber\\
+& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1
+\end{align}
+
+We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.  In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/simple_majority}
+\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
+\label{simple_majority}
+\end{figure}
+%Data:
+%Curve 1: {{0,0.000299648},{1,0.00213083},{2,0.00817989},{3,0.0224777},{4,0.049512},{5,0.0928927},{6,0.154136},{7,0.232044},{8,0.322859},{9,0.421057},{10,0.520467},{11,0.615358},{12,0.701241},{13,0.775267},{14,0.836245},{15,0.884385},{16,0.920892},{17,0.947532},{18,0.966265},{19,0.978973},{20,0.987296},{21,0.992561},{22,0.99578},{23,0.997682},{24,0.998767},{25,0.999366}}
+%Curve 2: {x, 0.99} for all x
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/super_majority}
+\caption{Super Majority: $\alpha$ vs. $\Pr(\alpha > \frac{2}{3}N)$ for $N=100$, $\beta = 10$.}
+\label{super_majority}
+\end{figure}
+%Data:
+%Curve 1: {{10,0.0395173},{11,0.0591278},{12,0.0846131},{13,0.116407},{14,0.15463},{15,0.199057},{16,0.24911},{17,0.303899},{18,0.362273},{19,0.422909},{20,0.484399},{21,0.545345},{22,0.604445},{23,0.660559},{24,0.712764},{25,0.760382},{26,0.802985},{27,0.840385},{28,0.872612},{29,0.899875},{30,0.922519},{31,0.940987},{32,0.955777},{33,0.967407},{34,0.976384},{35,0.983186},{36,0.988243},{37,0.99193},{38,0.994567},{39,0.996414},{40,0.997682},{41,0.998533},{42,0.999093},{43,0.999452},{44,0.999677},{45,0.999815},{46,0.999896},{47,0.999944},{48,0.99997},{49,0.999985},{50,0.999993}}
+%Curve 2: {x, 0.99} for all x
+
+\subsection{Score Voting}
 

### 2016-11-29 paper.tex
Style.

@@
 
-\subsubsection*{Background}
+\subsubsection{Background}
 A population of voters needs to collectively decide which item (if any) to take action on out of a set of competing items.
-In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a winning item, while also demonstrating that the winning item is preferred over all other competing items.
-We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that the population does not take undesired actions, which in this case encompasses taking action on an item that is not supported by a majority of the population and/or taking action on an item that is not clearly preferred over all other competing alternatives.
+In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a
+winning item, while also demonstrating that the winning item is preferred over all other competing items.
+We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that
+the population does not take undesired actions, which in this case encompasses taking action on an item that is not
+supported by a majority of the population and/or taking action on an item that is not clearly preferred over all
+other competing alternatives.
 
-\subsubsection*{Problem Statement}
-We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority, $90\%$ for a consensus).
-In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater than the bias towards voting ``yes'' on all other competing items.
+\subsubsection{Problem Statement}
+We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some
+fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority, $90\%$ for a consensus).
+In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater
+than the bias towards voting ``yes'' on all other competing items.
 
-\subsubsection*{Assumptions}
+\subsubsection{Assumptions}
 We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
@@
 
-\subsubsection*{Voting}
-We allow for score voting where each member of the population may choose any real number in the range $[-1,+1]$ to represent their vote, with $-1$ representing opposition and $+1$ representing support.
+\subsubsection{Voting}
+We allow for score voting where each member of the population may choose any real number in the range $[-1,+1]$ to
+represent their vote, with $-1$ representing opposition and $+1$ representing support.
 In the theory subsection below, we rescale votes to be in the range $[0,1]$ via the notation $\frac{votes + 1}{2}$.
@@
 We treat abstentions as indicating that a voter is removing themselves from the population of voters for an item.
-It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$ (indicating indifference).
+It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$
+(indicating indifference).
 
-\subsubsection*{Bayesian Theory}
-To properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial: \url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
+\subsubsection{Bayesian Theory}
+To properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial:
+\url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
 
 The above tutorial is very similar to the problem we have at hand.
-The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises naturally when considering samples from a finite population.
+The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a
+consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises
+naturally when considering samples from a finite population.
 
-After extracting the sufficient statistics from raw votes --- namely the average score and the number of abstentions --- we can adjust the effective population size and corresponding threshold.
-In addition, we can calculate the effective number of ``yes'' and ``no'' votes, which one can think of as the number of Heads and Tails in the aforementioned tutorial.
+After extracting the sufficient statistics from raw votes --- namely the average score and the number of
+abstentions --- we can adjust the effective population size and corresponding threshold.
+In addition, we can calculate the effective number of ``yes'' and ``no'' votes, which one can think of as the number
+of Heads and Tails in the aforementioned tutorial.
 
@@
 \item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
-\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote (e.g. $f=\frac{1}{2}$ for a simple majority).
+\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
+(e.g. $f=\frac{1}{2}$ for a simple majority).
 \end{itemize}
 \end{description}
-Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
+Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when
+no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
 \begin{align*}
@@
 \begin{itemize}
-\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that $X_i$ would remain below threshold.
+\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that
+$X_i$ would remain below threshold.
 \end{itemize}
@@
 
-In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to consider.
+In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to
+consider.
 Without loss of generality, assume that $X_1$ is the most popular motion within $C$.
-We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other $X_i  \in C, i \neq 1$:
+We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other
+$X_i  \in C, i \neq 1$:
 
@@
 
-We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.  In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
+We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.
+In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
 

### 2016-11-29 paper.tex
Short introduction.

@@
 
-\title{PeerMind}
+\title{PeerMind: Towards Collective Decision-Making}
 
@@
 
-The broad idea is that instead of just trying to move the voting process online -- something many others do -- we in
-fact add additional value by voting online.
-Specifically, we can compute results faster and more accurately, we can use sophisticated statistics to estimate
-quantities for the entire voting population, and we can promote clarity and full understanding of the voting items for all.
+Two of the most shocking democratic votes in recent history occurred in 2016:
+The referendum vote for Britain to leave the European Union (``Brexit'') and the election of Donald Trump as
+President of the United States.
+Both votes were decided by the slimmest of margins, but more significantly, both votes had poor turnout.
+We often hear that the Achilles heel of a democracy is low engagement, but this is a fallacy~---~in fact, our goal
+should be to decrease engagement.
+The ideal governance system is one where citizens do not have to be constantly involved, but the system is still
+acting as the citizens would like it to. Can we increase legitimacy within a democracy while also decreasing the
+attention required at the same time?
 
-\todo[inline]{Main questions and answers follow.}
+In this paper we present our work on improving the collective decision-making through use of a web application
+to facilitate the process.
+In contrast with existing similar solutions, our web application integrates new ideas in how to improve this process:
+use of range voting, dynamic quorum, and vote delegation.
+The broad idea is that instead of just trying to move the voting process online~---~something many others have done
+previously~---~we in fact wish to add value by voting online.
+Specifically, we can compute results faster, more accurately,
+%and securely,
+we can use sophisticated statistics to
+estimate important quantities for the entire voting population, we can allow for members to entrust others who have
+more expertise on topics than they do themselves, and we can promote clarity and full understanding of the voting
+items for all.
 
-What problem are we solving? Why is it important to solve? What is the story?
-
-Finding a way to lower attention required but increase legitimacy of democratic decisions.
-Traditionally legitimacy is increased by increasing participation/attention required.
-It is an everlasting problem.
-How to outsource governance and at the same time assure it does not get corrupt
-or tyrannic.
-
-Has anyone else tried to solve it or related problems? How did they try
-and what were the results (subjective or objective)?
-
-Described here: \url{http://mitar.tnode.com/post/73983101095/peer-to-peer-voting-scheme}.
-Some tried delegations, but mostly only to one person.
-Getting issues that you have tree-structure with possible cycles you cannot resolve.
-Often there were reported UI problems as well.
-So just technically improving is not enough, you also have to have UI for the technical solution so that people can trust.
-
-How are you solving the problem? What technologies are you using?
-
-Using dynamic quorum, score voting, delegation, and nice integration with UI/UX.
-
-How will you know how well you solved the problem? How will those technologies address the issues?
-
-If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
-So currently we have issues with reaching the quorum and many people do not participate.
-I would like for them to have easier time participating (less time required) and as a consequence more people to
-participate collectively.
-If I observe that I think I will know I solved the problem.
-
-\todo[inline]{This rest of the section is mostly notes. Organize and improve.}
-
-We can often hear that the issue of our democracy is low engagement.
-But this is a fallacy, because we should be instead trying to decrease engagement.
-The best governance system is one where you do not have to be involved yourself, but it is still doing things
-like you would like to do them.
-% Or even, doing things you might not even know that you would like to be done, especially in a long term.
-We want to outsource the job of coordination between people.
-Sometimes you want to coordinate people yourself (in your company, team), you might enjoy it, but in most cases
-you cannot or do not want to, so you outsource this service, especially at a large scale.
-
-One reason why increasing engagement sounds like a good idea is because current democratic systems have two
-questionable assumptions: participants have time to participate, and knowledge to participate.
-If you do not participate it is your fault.
-Moreover, the legitimacy of decisions is based on participation as well.
-This is why those participating in the these democratic systems have motivation to try to increase participation
-of others (especially if they believe they will still be in majority), or claim they are at fault to not participate.
-
-engagement is then increased only to assure that people like the results
-
-so if we try to minimize engagement, but still make sure that people like the results, then this is the perfect system
-
-and only in the case when they are on the loosing side, then we have to make sure that they understand the reasons why
-they lost (e.g.,\ the middle panel in our design) or at least to feel that they participated, in at least some way
-(voting, or just delegating)
-
-but the participation is the last on this list of important factors
-
-the more important thing is that they understand why others were thinking differently
-or that they do not even have to be engaged
-
-we want that people do not have to pay attention to what government is doing
-
-currently, we are in a loop where we have a useless feedback loop, where government in fact would like for people
-to pay some type of attention because they need to be elected, so they have to show like they are doing something,
-but on the other hand they do not want too much engagement, because they do not have things in place to really
-engage with everyone
-
-government should be like infrastructure, you should not have to care about it while it works, and be there to fix
-things if it starts failing, but not even you personally, but have mechanisms and processes in place to find this
-
-Terminology:
-\begin{itemize}
-\item community -- group which is making a decision using PeerMind
-\item member -- user of the app who is participating in decision making
-\end{itemize}
-
-\todo[inline]{Terminology: decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)}
-\todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}
-\todo[inline]{Terminology: discarding a vote if you didn't vote and didn't delegate. Some other term?}
-
-The main question is: can we increase legitimacy and decrease attention required at the same time.
-Many others are trying to increase attention (participation) because they want to increase legitimacy,
-assuming that this is the only way.
-
-A list of potential other points:
-\begin{itemize}
-\item legitimacy (quorum, involvement of more people through delegation)
-\item scaling (delegation helps, less attention helps)
-\item attention (at same or better decisions)
-\item bad assumptions for democracy - time, knowledge, it is rational to know that you do not know, common practice is to ask friends
-\item better decisions (more people consulted, attention directed where it should be)
-\end{itemize}
-
-Another question: should feedback be visible (to see how others voted before you voted)?
-People go with majority.
-Better for wisdom of crowds is an independent sample.
-Maybe we should only allow how other voted to be seen after you voted (but you can still change the vote),
-but we could then observe that and learn from this and see if this is good or bad.
-Or we could just store information if you observed existing votes before you voted (you have to click a button to see them).
-Maybe going with majority is also good when you want more social cohesion in the community and not so much ``optimal''
-wisdom of crowds result -  you would want that people relax their positions based on how other people care.
-Cohesion of the group vs. accuracy.
-Maybe display only number of votes and credibility, maybe credibility direction or graph, but not average votes.
-
-A list of potential questions:
-\begin{itemize}
-\item Delegation works?
-\item Statistical quorum works?
-\item Real-time feedback influences?
-\item Private vs. public delegation (do you know how member to whom you delegated is voting)?
-\item When delegating, how to assure that you cannot learn how somebody else voted, but that you can still somehow
-learn how your vote is used?
-\item Knowing how much power you have vs. not knowing -- does it change how people behave? (Knowing might corrupt you,
-but also motivate you to do more thorough work voting because you know other depend on you.)
-\end{itemize}
-
-Performance/theory/scaling in computational terms (how big could effectively community can be to be able to compute it).
-Some other theoretical properties?
-Computing statistical quorum is an ``embarrassingly parallel'' problem, and so its
-use would not suffer in large populations.
-Sparsity of delegation matrix also enables use of (recently developed?)
-efficient matrix computation algorithms.
-
-Out of scope of this work:
-
-\begin{itemize}
-\item How to assure privacy of the delegation network (it is highly sensitive data, probably even more than votes)?
-\item How to compute delegation results in a way that it does not require access to the whole network of delegation at once?
-\item How can we make this system distributed (and still assure privacy)?
-\end{itemize}
-
-\section{Related work}
-
-\cite{andersen2008trust, dirnstorfer2010voting, ford2002delegative, rodriguez2007smartocracy, yamakawa2007toward}
+%The broad idea is that instead of just trying to move the voting process online -- something many others do -- we in
+%fact add additional value by voting online.
+%Specifically, we can compute results faster and more accurately, we can use sophisticated statistics to estimate
+%quantities for the entire voting population, and we can promote clarity and full understanding of the voting items for all.
+%
+%\todo[inline]{Main questions and answers follow.}
+%
+%What problem are we solving? Why is it important to solve? What is the story?
+%
+%Finding a way to lower attention required but increase legitimacy of democratic decisions.
+%Traditionally legitimacy is increased by increasing participation/attention required.
+%It is an everlasting problem.
+%How to outsource governance and at the same time assure it does not get corrupt
+%or tyrannic.
+%
+%Has anyone else tried to solve it or related problems? How did they try
+%and what were the results (subjective or objective)?
+%
+%Described here: \url{http://mitar.tnode.com/post/73983101095/peer-to-peer-voting-scheme}.
+%Some tried delegations, but mostly only to one person.
+%Getting issues that you have tree-structure with possible cycles you cannot resolve.
+%Often there were reported UI problems as well.
+%So just technically improving is not enough, you also have to have UI for the technical solution so that people can trust.
+%
+%How are you solving the problem? What technologies are you using?
+%
+%Using dynamic quorum, score voting, delegation, and nice integration with UI/UX.
+%
+%How will you know how well you solved the problem? How will those technologies address the issues?
+%
+%If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
+%So currently we have issues with reaching the quorum and many people do not participate.
+%I would like for them to have easier time participating (less time required) and as a consequence more people to
+%participate collectively.
+%If I observe that I think I will know I solved the problem.
+%
+%\todo[inline]{This rest of the section is mostly notes. Organize and improve.}
+%
+%We can often hear that the issue of our democracy is low engagement.
+%But this is a fallacy, because we should be instead trying to decrease engagement.
+%The best governance system is one where you do not have to be involved yourself, but it is still doing things
+%like you would like to do them.
+%% Or even, doing things you might not even know that you would like to be done, especially in a long term.
+%We want to outsource the job of coordination between people.
+%Sometimes you want to coordinate people yourself (in your company, team), you might enjoy it, but in most cases
+%you cannot or do not want to, so you outsource this service, especially at a large scale.
+%
+%One reason why increasing engagement sounds like a good idea is because current democratic systems have two
+%questionable assumptions: participants have time to participate, and knowledge to participate.
+%If you do not participate it is your fault.
+%Moreover, the legitimacy of decisions is based on participation as well.
+%This is why those participating in the these democratic systems have motivation to try to increase participation
+%of others (especially if they believe they will still be in majority), or claim they are at fault to not participate.
+%
+%engagement is then increased only to assure that people like the results
+%
+%so if we try to minimize engagement, but still make sure that people like the results, then this is the perfect system
+%
+%and only in the case when they are on the loosing side, then we have to make sure that they understand the reasons why
+%they lost (e.g.,\ the middle panel in our design) or at least to feel that they participated, in at least some way
+%(voting, or just delegating)
+%
+%but the participation is the last on this list of important factors
+%
+%the more important thing is that they understand why others were thinking differently
+%or that they do not even have to be engaged
+%
+%we want that people do not have to pay attention to what government is doing
+%
+%currently, we are in a loop where we have a useless feedback loop, where government in fact would like for people
+%to pay some type of attention because they need to be elected, so they have to show like they are doing something,
+%but on the other hand they do not want too much engagement, because they do not have things in place to really
+%engage with everyone
+%
+%government should be like infrastructure, you should not have to care about it while it works, and be there to fix
+%things if it starts failing, but not even you personally, but have mechanisms and processes in place to find this
+%
+%Terminology:
+%\begin{itemize}
+%\item community -- group which is making a decision using PeerMind
+%\item member -- user of the app who is participating in decision making
+%\end{itemize}
+%
+%\todo[inline]{Terminology: decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)}
+%\todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}
+%\todo[inline]{Terminology: discarding a vote if you didn't vote and didn't delegate. Some other term?}
+%
+%The main question is: can we increase legitimacy and decrease attention required at the same time.
+%Many others are trying to increase attention (participation) because they want to increase legitimacy,
+%assuming that this is the only way.
+%
+%A list of potential other points:
+%\begin{itemize}
+%\item legitimacy (quorum, involvement of more people through delegation)
+%\item scaling (delegation helps, less attention helps)
+%\item attention (at same or better decisions)
+%\item bad assumptions for democracy - time, knowledge, it is rational to know that you do not know, common practice is to ask friends
+%\item better decisions (more people consulted, attention directed where it should be)
+%\end{itemize}
+%
+%Another question: should feedback be visible (to see how others voted before you voted)?
+%People go with majority.
+%Better for wisdom of crowds is an independent sample.
+%Maybe we should only allow how other voted to be seen after you voted (but you can still change the vote),
+%but we could then observe that and learn from this and see if this is good or bad.
+%Or we could just store information if you observed existing votes before you voted (you have to click a button to see them).
+%Maybe going with majority is also good when you want more social cohesion in the community and not so much ``optimal''
+%wisdom of crowds result -  you would want that people relax their positions based on how other people care.
+%Cohesion of the group vs. accuracy.
+%Maybe display only number of votes and credibility, maybe credibility direction or graph, but not average votes.
+%
+%A list of potential questions:
+%\begin{itemize}
+%\item Delegation works?
+%\item Statistical quorum works?
+%\item Real-time feedback influences?
+%\item Private vs. public delegation (do you know how member to whom you delegated is voting)?
+%\item When delegating, how to assure that you cannot learn how somebody else voted, but that you can still somehow
+%learn how your vote is used?
+%\item Knowing how much power you have vs. not knowing -- does it change how people behave? (Knowing might corrupt you,
+%but also motivate you to do more thorough work voting because you know other depend on you.)
+%\end{itemize}
+%
+%Performance/theory/scaling in computational terms (how big could effectively community can be to be able to compute it).
+%Some other theoretical properties?
+%Computing statistical quorum is an ``embarrassingly parallel'' problem, and so its
+%use would not suffer in large populations.
+%Sparsity of delegation matrix also enables use of (recently developed?)
+%efficient matrix computation algorithms.
+%
+%Out of scope of this work:
+%
+%\begin{itemize}
+%\item How to assure privacy of the delegation network (it is highly sensitive data, probably even more than votes)?
+%\item How to compute delegation results in a way that it does not require access to the whole network of delegation at once?
+%\item How can we make this system distributed (and still assure privacy)?
+%\end{itemize}
+%
+%\section{Related work}
+%
+%\cite{andersen2008trust, dirnstorfer2010voting, ford2002delegative, rodriguez2007smartocracy, yamakawa2007toward}
 
@@
 
-\subsection{Three step process}
-
-\subsection{Real-time feedback}
-
-% Real-time communication in general.
-
-\subsection{Integration with offline meetings}
-
-Decoupling the information presentation and voting processes from the debating process in the full decision-making
-pipeline (with the idea being that debating is best done in person, and presenting information/voting is best done online).
-
-\section{Evaluation}
-
-\subsection{Design}
-
-User study/evaluation design.
-To create a controlled evaluation of our solutions, to not depend on general usage of our system among communities,
-we can evaluate the design with the following user study:
-\begin{enumerate}
-\item we find a tight community of around 100 people willing to participate in the user study
-\item they should know each other well (to know how to delegate)
-\item we offer them to pay some Amazon credits or something for participation
-\item every member of the community should define delegations
-\item we ask them to split into two groups, group A and group B
-\item group A (target size is around 20% of the community) will meet together or online and discuss various topics through our system
-\item discussed topics some can be completely random, to get people to be familiar with they system, and also have some
-easy topics to decide on, like, what ``should be the color of the sky if we could change it'', and some topics which are relevant to the community
-\item those discussions would be just for evaluation, but community can use the results in any way they find useful
-\item after this discussions and decisions were done we ask everyone from group B to evaluate the results, we present them with two results (blinded, them not knowing which was made in which way)
-\begin{itemize}
-\item results which would be obtained using traditional methods (majority voting without any our improvements)
-\item our results using our system
-\end{itemize}
-\item we ask them to evaluate which one (or both, or none) of decisions they agree with
-\item prediction is that members who have not participated in the decision making process will
-find decisions made through our system more often agreeable
-\item hypothesis here is that what a decision making system is trying to achieve is to find a decision most of the
-whole community can agree with, or like it, even when not everyone participates and not all information is available
-on their positions on a topic
-\end{enumerate}
-
-One additional part of the experimental delegation study:
-
-Compute statistical quorum with and without delegations.
-This semester, do not actually allow delegation to be used for passing statistical quorum, but collect data so that
-we can see how much earlier delegation reached quorum before official statistical quorum.
-(Only count how much earlier at the point where delegation passed quorum and never went below thereafter, thus
-showing the point where the decision truly was clear.)
-
-Evaluation of statistical quorum (given that we don't officially decide by including delegation this semester):
-From email: ``Do an internal Cloyne comparison that is more mathematical, showing how many agenda items were decided
-using fewer actual votes than old 33\% quorum, as a result of statistical quorum.
-This could be measured by average \# votes needed to pass statistical quorum vs. \% of those items that get
-formally passed at council via consent agenda.
-(If it's 100\%, then you can say that we saved X\% of votes on
-average by using our online voting).
-Then you could also show that some items needed more than 33\% of votes, implying that a hard-line 33\% is also bad.''
-
-Subjective evaluation of entire platform: Council Surveys comparing past two semesters to this semester for other
-houses vs. Cloyne.
-``use the previous 2 semesters' worth of council surveys as control, and a council survey for
-Cloyne at the end of this semester as results.
-These results would speak to those specific subjective questions
-and would be one useful thing to look at.''
-
-\bibliographystyle{sigchi}
-\bibliography{refs}
-
-More:
-\begin{itemize}
-\item https://mason.gmu.edu/~rhanson/futarchy.html
-\item http://merkle.com/papers/DAOdemocracyDraft.pdf
-\end{itemize}
+%\subsection{Three step process}
+%
+%\subsection{Real-time feedback}
+%
+%
+%% Real-time communication in general.
+%
+%\subsection{Integration with offline meetings}
+%
+%Decoupling the information presentation and voting processes from the debating process in the full decision-making
+%pipeline (with the idea being that debating is best done in person, and presenting information/voting is best done online).
+%
+%\section{Evaluation}
+%
+%\subsection{Design}
+%
+%User study/evaluation design.
+%To create a controlled evaluation of our solutions, to not depend on general usage of our system among communities,
+%we can evaluate the design with the following user study:
+%\begin{enumerate}
+%\item we find a tight community of around 100 people willing to participate in the user study
+%\item they should know each other well (to know how to delegate)
+%\item we offer them to pay some Amazon credits or something for participation
+%\item every member of the community should define delegations
+%\item we ask them to split into two groups, group A and group B
+%\item group A (target size is around 20% of the community) will meet together or online and discuss various topics through our system
+%\item discussed topics some can be completely random, to get people to be familiar with they system, and also have some
+%easy topics to decide on, like, what ``should be the color of the sky if we could change it'', and some topics which are relevant to the community
+%\item those discussions would be just for evaluation, but community can use the results in any way they find useful
+%\item after this discussions and decisions were done we ask everyone from group B to evaluate the results, we present them with two results (blinded, them not knowing which was made in which way)
+%\begin{itemize}
+%\item results which would be obtained using traditional methods (majority voting without any our improvements)
+%\item our results using our system
+%\end{itemize}
+%\item we ask them to evaluate which one (or both, or none) of decisions they agree with
+%\item prediction is that members who have not participated in the decision making process will
+%find decisions made through our system more often agreeable
+%\item hypothesis here is that what a decision making system is trying to achieve is to find a decision most of the
+%whole community can agree with, or like it, even when not everyone participates and not all information is available
+%on their positions on a topic
+%\end{enumerate}
+%
+%One additional part of the experimental delegation study:
+%
+%Compute statistical quorum with and without delegations.
+%This semester, do not actually allow delegation to be used for passing statistical quorum, but collect data so that
+%we can see how much earlier delegation reached quorum before official statistical quorum.
+%(Only count how much earlier at the point where delegation passed quorum and never went below thereafter, thus
+%showing the point where the decision truly was clear.)
+%
+%Evaluation of statistical quorum (given that we don't officially decide by including delegation this semester):
+%From email: ``Do an internal Cloyne comparison that is more mathematical, showing how many agenda items were decided
+%using fewer actual votes than old 33\% quorum, as a result of statistical quorum.
+%This could be measured by average \# votes needed to pass statistical quorum vs. \% of those items that get
+%formally passed at council via consent agenda.
+%(If it's 100\%, then you can say that we saved X\% of votes on
+%average by using our online voting).
+%Then you could also show that some items needed more than 33\% of votes, implying that a hard-line 33\% is also bad.'' 
+%
+%Subjective evaluation of entire platform: Council Surveys comparing past two semesters to this semester for other
+%houses vs. Cloyne.
+%``use the previous 2 semesters' worth of council surveys as control, and a council survey for
+%Cloyne at the end of this semester as results.
+%These results would speak to those specific subjective questions
+%and would be one useful thing to look at.''
+%
+%\bibliographystyle{sigchi}
+%\bibliography{refs}
+%
+%More:
+%\begin{itemize}
+%\item https://mason.gmu.edu/~rhanson/futarchy.html
+%\item http://merkle.com/papers/DAOdemocracyDraft.pdf
+%\end{itemize}
 

### 2016-11-29 paper.tex
Range voting.

@@
 
-\subsection{Statistical Quorum}
+\subsection{Range Voting}
+
+Range voting (score voting) is voting where members of a community express their positions on open proposals by assigning
+each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
+Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
+positions in a highly informative and precise manner.
+We find it especially suitable for a web application because many web users are used to use it when ranking products
+and giving other ``five-star'' feedback.
+Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
+new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
+each proposal would need reassessing (assuming members use a consistent personal scale when voting).
+
+\subsection{Dynamic Quorum}
 
 \subsubsection{Background}
-A population of voters needs to collectively decide which item (if any) to take action on out of a set of competing items.
+A population of voters (a community) needs to collectively decide which item (if any) to take action on out of a set of competing items.
 In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a
@@
 
-\subsection{Score Voting}
-
-\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
-we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
-also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
+%\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
+%we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
+%also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
 

### 2016-11-29 paper.tex
Style.

@@
 %
-%Using dynamic quorum, score voting, delegation, and nice integration with UI/UX.
+%Using dynamic quorum, range voting, delegation, and nice integration with UI/UX.
 %
@@
 
-Range voting (score voting) is voting where members of a community express their positions on open proposals by assigning
+Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
 each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
+A proposal passes if the average score of all votes on a position is positive.
+When multiple competing proposals are being voted on, the one with highest average score wins.
+
 Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
@@
 \subsection{Dynamic Quorum}
+
+%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal}
 
@@
 We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some
-fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority, $90\%$ for a consensus).
+fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority).
 In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater
@@
 \subsubsection{Voting}
-We allow for score voting where each member of the population may choose any real number in the range $[-1,+1]$ to
+We allow for range voting where each member of the population may choose any real number in the range $[-1,+1]$ to
 represent their vote, with $-1$ representing opposition and $+1$ representing support.
@@
 \subsubsection{Bayesian Theory}
-To properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial:
-\url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
 
-The above tutorial is very similar to the problem we have at hand.
 The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a
 consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises
-naturally when considering samples from a finite population.
+naturally when considering samples from a finite population.\footnote{To
+properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial:
+\url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
+The tutorial is very similar to the problem we have at hand.}
 
-After extracting the sufficient statistics from raw votes --- namely the average score and the number of
-abstentions --- we can adjust the effective population size and corresponding threshold.
-In addition, we can calculate the effective number of ``yes'' and ``no'' votes, which one can think of as the number
-of Heads and Tails in the aforementioned tutorial.
+After extracting the sufficient statistics from raw votes~---~namely the average score and the number of
+abstentions~---~we can adjust the effective population size and corresponding threshold.
+In addition, we can calculate the effective number of ``yes'' and ``no'' votes.\footnote{Which one can think
+of as the number of Heads and Tails in the aforementioned tutorial.}
 
@@
 respectively.
-We are using score voting on $[-1, 1]$ range, so those votes are $v_1 = V_1 = 0.7$ and
+We are using range voting on $[-1, 1]$ range, so those votes are $v_1 = V_1 = 0.7$ and
 $v_3 = V_3 = -0.4$.

### 2016-11-29 paper.tex
Cleaned version.

@@
 \usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 
-%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal}
+%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal, motion -> proposal}
+
+Dynamic quorum uses statistical hypothesis testing to compute the minimum number of required votes needed
+to achieve a desired level of confidence before a community adopts an outcome as being truly representative of
+its members' positions.
+For example, in a community of 100 members, if 20 members vote in full support of a proposal, then we can conclude
+with very high confidence that, had all 100 members actually voted, a majority them would support the proposal,
+despite in actuality 80 members not having voted.
+On the other hand, if 80 members vote but 41 vote in support of the proposal and 39 vote in opposition,
+the community cannot yet conclude anything with high confidence.
+Hence, the suggested action would be for the community to gather more votes and continue discussion around the proposal.
+By utilizing a dynamic quorum, we can direct attention and participation of the community towards contentious proposals,
+while reducing attention towards proposals which already have overwhelming support.
 
@@
 
-Deriving our expression of interest now immediately follows:
-
 Let
@@
 \item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the raw votes for motion $X_i$, excluding ``abstentions'',
-\item[$N_i = $] $(N - $ number of abstentions on $X_i)$, the effective population voting on motion $X_i$,
-\item[$\alpha_i = $] $\sum\limits_{v \in X_i} (\frac{v+1}{2})$, the effective number of ``yes'' votes on motion $X_i$,
+\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on motion $X_i$,
+\item[$\alpha_i = $] $\sum\limits_{v \in X_i} \left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
 \item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
@@
 
-Often, we have a single motion open for voting.  In this scenario, $C=\{X_1\}$ and we simply wish to compute:
+When we have a single motion open for voting, $C=\{X_1\}$, we simply wish to compute:
 \begin{align}\label{eq1}
@@
 \begin{align}\label{eq3}
-\Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)
+\Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)
 \end{align}
 
-Using a safe assumption that votes on motions are independent across motions, we can decompose \eqref{eq3} as follows:
+Using the assumption
+that votes on motions are independent across motions, which, in the common
+scenario where there is a negative covariance between votes on competing
+motions, will provide a lower bound on the probability, we can decompose \eqref{eq3} as follows:
 \begin{align*}
-& \Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)\\
+& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\\
 =& \prod_{i=2}^n \Pr(\alpha^*_1 > \alpha^*_i \mid C)\\
@@
 \begin{align}\label{eq5}
-& \Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)\nonumber\\
+& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\nonumber\\
 &= \prod_{i=2}^n \int_0^1 \int_0^1 \Phi_i(p_1,p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \nonumber\\
@@
 
-\begin{figure}[!ht]
+\begin{figure}[ht]
 \centering
@@
 \subsection{Delegation}
-\todo[inline]{Max: I think delegation graph edges should point in the direction towards the person that the vote is
-being given to (more intuitive).
-\\
-Mitar: We are not giving votes to people, we are delegating a decision to people, but votes are computed for each person
-separately based on how delegates voted themselves. The idea is that you ask a delegate how the voted and then compute
-your vote. The do not get your vote to do anything with it. This is why arrows are like that. Maybe the word
-"delegation" is confusing and we should maybe use surrogate or some other word.}
+%\todo[inline]{Max: I think delegation graph edges should point in the direction towards the person that the vote is
+%being given to (more intuitive).
+%\\
+%Mitar: We are not giving votes to people, we are delegating a decision to people, but votes are computed for each person
+%separately based on how delegates voted themselves. The idea is that you ask a delegate how the voted and then compute
+%your vote. The do not get your vote to do anything with it. This is why arrows are like that. Maybe the word
+%"delegation" is confusing and we should maybe use surrogate or some other word.}
 Because not every member of the community has time or knowledge to vote all the time, each member can delegate
@@
   \caption{Example delegation network. Edges are directed towards a member who is delegating, representing how their
-  vote is computed from delegates' votes. E.g., $V_2 = 0.75 V_1 + 0.25 V_4$}~\label{fig:delegation-network}
+  vote is computed from delegates' votes. E.g., $V_2 = 0.75 V_1 + 0.25 V_4$}\label{fig:delegation-network}
 \end{figure}
@@
 
-\todo[inline]{Max: Double check accuracy of this method, Improve explanation here and perhaps use nicer numbers so
-readers can follow along certain parts in their heads easily.}
+%\todo[inline]{Max: Double check accuracy of this method, Improve explanation here and perhaps use nicer numbers so
+%readers can follow along certain parts in their heads easily.}
 
@@
 
-\todo[inline]{Max: FYI, computing who to discard from the population as ``non-voting'' is a separate computation in itself.
-You need to figure out what votes can be calculated partially as a consequence of delegation, and which cannot.
-Use indicator vectors denoting 1 if member has voted, else 0.
-\\
-Mitar: This is exactly what $\boldsymbol{w}$ is later on.}
+There are few details we have to look into.
 
-There are few details we have to look into.
-Despite delegation some votes or parts of votes can still happen to be discarded.
+It happen that matrix $D$ is singular which prevents $\boldsymbol{V}$ to be directly computed.
+For example, for the following system:
+
+\begin{displaymath}
+\left[ \begin{array}{ccccc}
+1 & 0 & -1 & 0 \\
+0 & 1 & -0.3 & -0.7 \\
+-1 & 0 & 1 & 0 \\
+0 & 0 & 0 & 1\\
+\end{array} \right] \boldsymbol{V} = \left[ \begin{array}{c}
+0 \\
+0 \\
+0 \\
+1 \\
+\end{array} \right]
+\end{displaymath}
+
+we would like that for $\boldsymbol{V}$ the result is:
+
+\begin{displaymath}
+\left[ \begin{array}{c}
+0 \\
+0.7 \\
+0 \\
+1 \\
+\end{array} \right]
+\end{displaymath}
+
+Such matrix $D$ can happen when there is a cycle of delegations and nobody in a cycle
+voted.
+We can use the algorithm from Algorithm~\ref{alg:decycle} to modify matrix $D$ into
+matrix $D'$ without such cycles.
+It is worth noting that matrix $D$ has some useful properties:
+there is always a diagonal with elements $1$, other elements are from interval $[−1,0]$,
+sum of each row is always or $1$ or $0$.
+
+\begin{algorithm}[h]
+  \caption{Removing cycles in delegation matrix $D$}
+  \begin{algorithmic}[1]
+    \State Define a set of restricted indexes $Z$.
+    \State Populate $Z$ with indexes of all rows which have $1$ only on a diagonal.\footnotemark
+    \State Remove all rows and columns corresponding to those indexes.
+    \State Add to $Z$ (original) index of every row for which the sum (of its elements in this reduced matrix) is $\neq 0$.
+    \State If added any such index, go to step 3.
+    \State Otherwise rows which are making the matrix ill-defined are left. Set them to having only $1$ on a diagonal in the original matrix $D$.
+  \end{algorithmic}
+  \label{alg:decycle}
+\end{algorithm}
+\footnotetext{Equal to: all rows for which the sum of its elements is $1$; or, for which all non-diagonal elements are $0$.}
+
+For the last example above, the resulting $D'$ matrix would be:
+
+\begin{displaymath}
+\left[ \begin{array}{cccc}
+1 & 0 & 0 & 0 \\
+0 & 1 & -0.3 & -0.7 \\
+0 & 0 & 1 & 0 \\
+0 & 0 & 0 & 1\\
+\end{array} \right]
+\end{displaymath}
+
+For our initial example, no change is necessary, so $D' = D$.
+
+Despite delegation some votes or parts of votes can still happen to stay missing.
 Members who did not vote might have not declared their delegates.
 Other members could delegate to them as well, potentially propagating lack of a vote further.
-In our example, member $M_5$ has not voted so their vote is discarded.
+In our example, member $M_5$ has not voted so their vote will stay missing.
 As a consequence $M_4$'s vote cannot be computed fully, which influences $M_2$'s vote as well.
@@
 
-We solve $D \boldsymbol{W} = \boldsymbol{w}$ and $D \boldsymbol{A} = \boldsymbol{a}$ for
+We solve $D' \boldsymbol{W} = \boldsymbol{w}$ and $D' \boldsymbol{A} = \boldsymbol{a}$ for
 $\boldsymbol{W} = \left[W_{i}\right]$ and $\boldsymbol{A} = \left[A_{i}\right]$, respectively.
@@
 
-\todo[inline]{If there is a clique/cycle/closed subgroup in delegation, without any vote inside a clique/cycle/closed subgroup, the matrix becomes singular.
-We should find a way to detect such rows and set corresponding elements of $\boldsymbol{W}$ to $0$ (or the row
-to identity row before we try to compute $\boldsymbol{W}$).
-This is exactly what we want in fact, that those in such a clique/cycle/closed subgroup are seen as non-computable.
-Then the rest works.
-We should describe a problem, maybe with an example.}
-
 We define a normalized matrix:
@@
 \begin{displaymath}
-D' = \left[d'_{ij}\right] = \begin{dcases*}
+D'' = \left[D''_{ij}\right] = \begin{dcases*}
  1 & if $i = j$ \\
@@
 To compute votes with normalized ratios for only those delegates who voted or for which we can compute votes recursively,
-we solve $D' \boldsymbol{V} = \boldsymbol{v}$ for $\boldsymbol{V}$:
+we solve $D'' \boldsymbol{V} = \boldsymbol{v}$ for $\boldsymbol{V}$:
 
@@
 
-\todo[inline]{Do we want really to compute quorum using delegation, or should quorum be computed
-using direct votes only?}
+%\todo[inline]{Do we want really to compute quorum using delegation, or should quorum be computed
+%using direct votes only?}
 

### 2016-12-03 paper.tex
stash local changes

--- 2016-11-29 paper.tex
@@
 
-\title{PeerMind: Towards Collective Decision-Making}
+\title{PeerMind}
 
@@
 
-Two of the most shocking democratic votes in recent history occurred in 2016:
-The referendum vote for Britain to leave the European Union (``Brexit'') and the election of Donald Trump as
-President of the United States.
-Both votes were decided by the slimmest of margins, but more significantly, both votes had poor turnout.
-We often hear that the Achilles heel of a democracy is low engagement, but this is a fallacy~---~in fact, our goal
-should be to decrease engagement.
-The ideal governance system is one where citizens do not have to be constantly involved, but the system is still
-acting as the citizens would like it to. Can we increase legitimacy within a democracy while also decreasing the
-attention required at the same time?
+The broad idea is that instead of just trying to move the voting process online -- something many others do -- we in
+fact add additional value by voting online.
+Specifically, we can compute results faster and more accurately, we can use sophisticated statistics to estimate
+quantities for the entire voting population, and we can promote clarity and full understanding of the voting items for all.
 
-In this paper we present our work on improving the collective decision-making through use of a web application
-to facilitate the process.
-In contrast with existing similar solutions, our web application integrates new ideas in how to improve this process:
-use of range voting, dynamic quorum, and vote delegation.
-The broad idea is that instead of just trying to move the voting process online~---~something many others have done
-previously~---~we in fact wish to add value by voting online.
-Specifically, we can compute results faster, more accurately,
-%and securely,
-we can use sophisticated statistics to
-estimate important quantities for the entire voting population, we can allow for members to entrust others who have
-more expertise on topics than they do themselves, and we can promote clarity and full understanding of the voting
-items for all.
+\todo[inline]{Main questions and answers follow.}
 
-%The broad idea is that instead of just trying to move the voting process online -- something many others do -- we in
-%fact add additional value by voting online.
-%Specifically, we can compute results faster and more accurately, we can use sophisticated statistics to estimate
-%quantities for the entire voting population, and we can promote clarity and full understanding of the voting items for all.
-%
-%\todo[inline]{Main questions and answers follow.}
-%
-%What problem are we solving? Why is it important to solve? What is the story?
-%
-%Finding a way to lower attention required but increase legitimacy of democratic decisions.
-%Traditionally legitimacy is increased by increasing participation/attention required.
-%It is an everlasting problem.
-%How to outsource governance and at the same time assure it does not get corrupt
-%or tyrannic.
-%
-%Has anyone else tried to solve it or related problems? How did they try
-%and what were the results (subjective or objective)?
-%
-%Described here: \url{http://mitar.tnode.com/post/73983101095/peer-to-peer-voting-scheme}.
-%Some tried delegations, but mostly only to one person.
-%Getting issues that you have tree-structure with possible cycles you cannot resolve.
-%Often there were reported UI problems as well.
-%So just technically improving is not enough, you also have to have UI for the technical solution so that people can trust.
-%
-%How are you solving the problem? What technologies are you using?
-%
-%Using dynamic quorum, score voting, delegation, and nice integration with UI/UX.
-%
-%How will you know how well you solved the problem? How will those technologies address the issues?
-%
-%If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
-%So currently we have issues with reaching the quorum and many people do not participate.
-%I would like for them to have easier time participating (less time required) and as a consequence more people to
-%participate collectively.
-%If I observe that I think I will know I solved the problem.
-%
-%\todo[inline]{This rest of the section is mostly notes. Organize and improve.}
-%
-%We can often hear that the issue of our democracy is low engagement.
-%But this is a fallacy, because we should be instead trying to decrease engagement.
-%The best governance system is one where you do not have to be involved yourself, but it is still doing things
-%like you would like to do them.
-%% Or even, doing things you might not even know that you would like to be done, especially in a long term.
-%We want to outsource the job of coordination between people.
-%Sometimes you want to coordinate people yourself (in your company, team), you might enjoy it, but in most cases
-%you cannot or do not want to, so you outsource this service, especially at a large scale.
-%
-%One reason why increasing engagement sounds like a good idea is because current democratic systems have two
-%questionable assumptions: participants have time to participate, and knowledge to participate.
-%If you do not participate it is your fault.
-%Moreover, the legitimacy of decisions is based on participation as well.
-%This is why those participating in the these democratic systems have motivation to try to increase participation
-%of others (especially if they believe they will still be in majority), or claim they are at fault to not participate.
-%
-%engagement is then increased only to assure that people like the results
-%
-%so if we try to minimize engagement, but still make sure that people like the results, then this is the perfect system
-%
-%and only in the case when they are on the loosing side, then we have to make sure that they understand the reasons why
-%they lost (e.g.,\ the middle panel in our design) or at least to feel that they participated, in at least some way
-%(voting, or just delegating)
-%
-%but the participation is the last on this list of important factors
-%
-%the more important thing is that they understand why others were thinking differently
-%or that they do not even have to be engaged
-%
-%we want that people do not have to pay attention to what government is doing
-%
-%currently, we are in a loop where we have a useless feedback loop, where government in fact would like for people
-%to pay some type of attention because they need to be elected, so they have to show like they are doing something,
-%but on the other hand they do not want too much engagement, because they do not have things in place to really
-%engage with everyone
-%
-%government should be like infrastructure, you should not have to care about it while it works, and be there to fix
-%things if it starts failing, but not even you personally, but have mechanisms and processes in place to find this
-%
-%Terminology:
-%\begin{itemize}
-%\item community -- group which is making a decision using PeerMind
-%\item member -- user of the app who is participating in decision making
-%\end{itemize}
-%
-%\todo[inline]{Terminology: decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)}
-%\todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}
-%\todo[inline]{Terminology: discarding a vote if you didn't vote and didn't delegate. Some other term?}
-%
-%The main question is: can we increase legitimacy and decrease attention required at the same time.
-%Many others are trying to increase attention (participation) because they want to increase legitimacy,
-%assuming that this is the only way.
-%
-%A list of potential other points:
-%\begin{itemize}
-%\item legitimacy (quorum, involvement of more people through delegation)
-%\item scaling (delegation helps, less attention helps)
-%\item attention (at same or better decisions)
-%\item bad assumptions for democracy - time, knowledge, it is rational to know that you do not know, common practice is to ask friends
-%\item better decisions (more people consulted, attention directed where it should be)
-%\end{itemize}
-%
-%Another question: should feedback be visible (to see how others voted before you voted)?
-%People go with majority.
-%Better for wisdom of crowds is an independent sample.
-%Maybe we should only allow how other voted to be seen after you voted (but you can still change the vote),
-%but we could then observe that and learn from this and see if this is good or bad.
-%Or we could just store information if you observed existing votes before you voted (you have to click a button to see them).
-%Maybe going with majority is also good when you want more social cohesion in the community and not so much ``optimal''
-%wisdom of crowds result -  you would want that people relax their positions based on how other people care.
-%Cohesion of the group vs. accuracy.
-%Maybe display only number of votes and credibility, maybe credibility direction or graph, but not average votes.
-%
-%A list of potential questions:
-%\begin{itemize}
-%\item Delegation works?
-%\item Statistical quorum works?
-%\item Real-time feedback influences?
-%\item Private vs. public delegation (do you know how member to whom you delegated is voting)?
-%\item When delegating, how to assure that you cannot learn how somebody else voted, but that you can still somehow
-%learn how your vote is used?
-%\item Knowing how much power you have vs. not knowing -- does it change how people behave? (Knowing might corrupt you,
-%but also motivate you to do more thorough work voting because you know other depend on you.)
-%\end{itemize}
-%
-%Performance/theory/scaling in computational terms (how big could effectively community can be to be able to compute it).
-%Some other theoretical properties?
-%Computing statistical quorum is an ``embarrassingly parallel'' problem, and so its
-%use would not suffer in large populations.
-%Sparsity of delegation matrix also enables use of (recently developed?)
-%efficient matrix computation algorithms.
-%
-%Out of scope of this work:
-%
-%\begin{itemize}
-%\item How to assure privacy of the delegation network (it is highly sensitive data, probably even more than votes)?
-%\item How to compute delegation results in a way that it does not require access to the whole network of delegation at once?
-%\item How can we make this system distributed (and still assure privacy)?
-%\end{itemize}
-%
-%\section{Related work}
-%
-%\cite{andersen2008trust, dirnstorfer2010voting, ford2002delegative, rodriguez2007smartocracy, yamakawa2007toward}
+What problem are we solving? Why is it important to solve? What is the story?
+
+Finding a way to lower attention required but increase legitimacy of democratic decisions.
+Traditionally legitimacy is increased by increasing participation/attention required.
+It is an everlasting problem.
+How to outsource governance and at the same time assure it does not get corrupt
+or tyrannic.
+
+Has anyone else tried to solve it or related problems? How did they try
+and what were the results (subjective or objective)?
+
+Described here: \url{http://mitar.tnode.com/post/73983101095/peer-to-peer-voting-scheme}.
+Some tried delegations, but mostly only to one person.
+Getting issues that you have tree-structure with possible cycles you cannot resolve.
+Often there were reported UI problems as well.
+So just technically improving is not enough, you also have to have UI for the technical solution so that people can trust.
+
+How are you solving the problem? What technologies are you using?
+
+Using dynamic quorum, score voting, delegation, and nice integration with UI/UX.
+
+How will you know how well you solved the problem? How will those technologies address the issues?
+
+If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
+So currently we have issues with reaching the quorum and many people do not participate.
+I would like for them to have easier time participating (less time required) and as a consequence more people to
+participate collectively.
+If I observe that I think I will know I solved the problem.
+
+\todo[inline]{This rest of the section is mostly notes. Organize and improve.}
+
+We can often hear that the issue of our democracy is low engagement.
+But this is a fallacy, because we should be instead trying to decrease engagement.
+The best governance system is one where you do not have to be involved yourself, but it is still doing things
+like you would like to do them.
+% Or even, doing things you might not even know that you would like to be done, especially in a long term.
+We want to outsource the job of coordination between people.
+Sometimes you want to coordinate people yourself (in your company, team), you might enjoy it, but in most cases
+you cannot or do not want to, so you outsource this service, especially at a large scale.
+
+One reason why increasing engagement sounds like a good idea is because current democratic systems have two
+questionable assumptions: participants have time to participate, and knowledge to participate.
+If you do not participate it is your fault.
+Moreover, the legitimacy of decisions is based on participation as well.
+This is why those participating in the these democratic systems have motivation to try to increase participation
+of others (especially if they believe they will still be in majority), or claim they are at fault to not participate.
+
+engagement is then increased only to assure that people like the results
+
+so if we try to minimize engagement, but still make sure that people like the results, then this is the perfect system
+
+and only in the case when they are on the loosing side, then we have to make sure that they understand the reasons why
+they lost (e.g.,\ the middle panel in our design) or at least to feel that they participated, in at least some way
+(voting, or just delegating)
+
+but the participation is the last on this list of important factors
+
+the more important thing is that they understand why others were thinking differently
+or that they do not even have to be engaged
+
+we want that people do not have to pay attention to what government is doing
+
+currently, we are in a loop where we have a useless feedback loop, where government in fact would like for people
+to pay some type of attention because they need to be elected, so they have to show like they are doing something,
+but on the other hand they do not want too much engagement, because they do not have things in place to really
+engage with everyone
+
+government should be like infrastructure, you should not have to care about it while it works, and be there to fix
+things if it starts failing, but not even you personally, but have mechanisms and processes in place to find this
+
+Terminology:
+\begin{itemize}
+\item community -- group which is making a decision using PeerMind
+\item member -- user of the app who is participating in decision making
+\end{itemize}
+
+\todo[inline]{Terminology: decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)}
+\todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}
+\todo[inline]{Terminology: discarding a vote if you didn't vote and didn't delegate. Some other term?}
+
+The main question is: can we increase legitimacy and decrease attention required at the same time.
+Many others are trying to increase attention (participation) because they want to increase legitimacy,
+assuming that this is the only way.
+
+A list of potential other points:
+\begin{itemize}
+\item legitimacy (quorum, involvement of more people through delegation)
+\item scaling (delegation helps, less attention helps)
+\item attention (at same or better decisions)
+\item bad assumptions for democracy - time, knowledge, it is rational to know that you do not know, common practice is to ask friends
+\item better decisions (more people consulted, attention directed where it should be)
+\end{itemize}
+
+Another question: should feedback be visible (to see how others voted before you voted)?
+People go with majority.
+Better for wisdom of crowds is an independent sample.
+Maybe we should only allow how other voted to be seen after you voted (but you can still change the vote),
+but we could then observe that and learn from this and see if this is good or bad.
+Or we could just store information if you observed existing votes before you voted (you have to click a button to see them).
+Maybe going with majority is also good when you want more social cohesion in the community and not so much ``optimal''
+wisdom of crowds result -  you would want that people relax their positions based on how other people care.
+Cohesion of the group vs. accuracy.
+Maybe display only number of votes and credibility, maybe credibility direction or graph, but not average votes.
+
+A list of potential questions:
+\begin{itemize}
+\item Delegation works?
+\item Statistical quorum works?
+\item Real-time feedback influences?
+\item Private vs. public delegation (do you know how member to whom you delegated is voting)?
+\item When delegating, how to assure that you cannot learn how somebody else voted, but that you can still somehow
+learn how your vote is used?
+\item Knowing how much power you have vs. not knowing -- does it change how people behave? (Knowing might corrupt you,
+but also motivate you to do more thorough work voting because you know other depend on you.)
+\end{itemize}
+
+Performance/theory/scaling in computational terms (how big could effectively community can be to be able to compute it).
+Some other theoretical properties?
+Computing statistical quorum is an ``embarrassingly parallel'' problem, and so its
+use would not suffer in large populations.
+Sparsity of delegation matrix also enables use of (recently developed?)
+efficient matrix computation algorithms.
+
+Out of scope of this work:
+
+\begin{itemize}
+\item How to assure privacy of the delegation network (it is highly sensitive data, probably even more than votes)?
+\item How to compute delegation results in a way that it does not require access to the whole network of delegation at once?
+\item How can we make this system distributed (and still assure privacy)?
+\end{itemize}
+
+\section{Related work}
+
+\cite{andersen2008trust, dirnstorfer2010voting, ford2002delegative, rodriguez2007smartocracy, yamakawa2007toward}
 
@@
 
-\subsection{Range Voting}
+\subsection{Statistical Quorum}
 
-Range voting (score voting) is voting where members of a community express their positions on open proposals by assigning
-each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
-Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
-positions in a highly informative and precise manner.
-We find it especially suitable for a web application because many web users are used to use it when ranking products
-and giving other ``five-star'' feedback.
-Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
-new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
-each proposal would need reassessing (assuming members use a consistent personal scale when voting).
+\subsubsection*{Background}
+A population of voters needs to collectively decide which item (if any) to take action on out of a set of competing items.
+In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a winning item, while also demonstrating that the winning item is preferred over all other competing items.
+We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that the population does not take undesired actions, which in this case encompasses taking action on an item that is not supported by a majority of the population and/or taking action on an item that is not clearly preferred over all other competing alternatives.
 
-\subsection{Dynamic Quorum}
+\subsubsection*{Problem Statement}
+We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority, $90\%$ for a consensus).
+In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater than the bias towards voting ``yes'' on all other competing items.
 
-\subsubsection{Background}
-A population of voters (a community) needs to collectively decide which item (if any) to take action on out of a set of competing items.
-In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a
-winning item, while also demonstrating that the winning item is preferred over all other competing items.
-We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that
-the population does not take undesired actions, which in this case encompasses taking action on an item that is not
-supported by a majority of the population and/or taking action on an item that is not clearly preferred over all
-other competing alternatives.
-
-\subsubsection{Problem Statement}
-We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some
-fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority, $90\%$ for a consensus).
-In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater
-than the bias towards voting ``yes'' on all other competing items.
-
-\subsubsection{Assumptions}
+\subsubsection*{Assumptions}
 We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
@@
 
-\subsubsection{Voting}
-We allow for score voting where each member of the population may choose any real number in the range $[-1,+1]$ to
-represent their vote, with $-1$ representing opposition and $+1$ representing support.
+\subsubsection*{Voting}
+We allow for score voting where each member of the population may choose any real number in the range $[-1,+1]$ to represent their vote, with $-1$ representing opposition and $+1$ representing support.
 In the theory subsection below, we rescale votes to be in the range $[0,1]$ via the notation $\frac{votes + 1}{2}$.
@@
 We treat abstentions as indicating that a voter is removing themselves from the population of voters for an item.
-It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$
-(indicating indifference).
+It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$ (indicating indifference).
 
-\subsubsection{Bayesian Theory}
-To properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial:
-\url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
+\subsubsection*{Bayesian Theory}
+To properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial: \url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
 
 The above tutorial is very similar to the problem we have at hand.
-The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a
-consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises
-naturally when considering samples from a finite population.
+The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises naturally when considering samples from a finite population.
 
-After extracting the sufficient statistics from raw votes --- namely the average score and the number of
-abstentions --- we can adjust the effective population size and corresponding threshold.
-In addition, we can calculate the effective number of ``yes'' and ``no'' votes, which one can think of as the number
-of Heads and Tails in the aforementioned tutorial.
+After extracting the sufficient statistics from raw votes --- namely the average score and the number of abstentions --- we can adjust the effective population size and corresponding threshold.
+In addition, we can calculate the effective number of ``yes'' and ``no'' votes, which one can think of as the number of Heads and Tails in the aforementioned tutorial.
 
@@
 \item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
-\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
-(e.g. $f=\frac{1}{2}$ for a simple majority).
+\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote (e.g. $f=\frac{1}{2}$ for a simple majority).
 \end{itemize}
 \end{description}
-Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when
-no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
+Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
 \begin{align*}
@@
 \begin{itemize}
-\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that
-$X_i$ would remain below threshold.
+\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that $X_i$ would remain below threshold.
 \end{itemize}
@@
 & \Pr(\alpha^*_1 > T_1 \mid X_1) = \nonumber\\
-& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{K_1-i} {(1-p_1)}^{i} \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
+& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{K_1-i} {(1-p_1)}^{i} \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
 \end{align}
 
-In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to
-consider.
+In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to consider.
 Without loss of generality, assume that $X_1$ is the most popular motion within $C$.
-We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other
-$X_i  \in C, i \neq 1$:
+We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other $X_i  \in C, i \neq 1$:
 
@@
 
-We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.
-In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
+We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.  In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
 
@@
 
-%\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
-%we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
-%also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
+\subsection{Score Voting}
+
+\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
+we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
+also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
 
@@
 
-%\subsection{Three step process}
-%
-%\subsection{Real-time feedback}
-%
-%
-%% Real-time communication in general.
-%
-%\subsection{Integration with offline meetings}
-%
-%Decoupling the information presentation and voting processes from the debating process in the full decision-making
-%pipeline (with the idea being that debating is best done in person, and presenting information/voting is best done online).
-%
-%\section{Evaluation}
-%
-%\subsection{Design}
-%
-%User study/evaluation design.
-%To create a controlled evaluation of our solutions, to not depend on general usage of our system among communities,
-%we can evaluate the design with the following user study:
-%\begin{enumerate}
-%\item we find a tight community of around 100 people willing to participate in the user study
-%\item they should know each other well (to know how to delegate)
-%\item we offer them to pay some Amazon credits or something for participation
-%\item every member of the community should define delegations
-%\item we ask them to split into two groups, group A and group B
-%\item group A (target size is around 20% of the community) will meet together or online and discuss various topics through our system
-%\item discussed topics some can be completely random, to get people to be familiar with they system, and also have some
-%easy topics to decide on, like, what ``should be the color of the sky if we could change it'', and some topics which are relevant to the community
-%\item those discussions would be just for evaluation, but community can use the results in any way they find useful
-%\item after this discussions and decisions were done we ask everyone from group B to evaluate the results, we present them with two results (blinded, them not knowing which was made in which way)
-%\begin{itemize}
-%\item results which would be obtained using traditional methods (majority voting without any our improvements)
-%\item our results using our system
-%\end{itemize}
-%\item we ask them to evaluate which one (or both, or none) of decisions they agree with
-%\item prediction is that members who have not participated in the decision making process will
-%find decisions made through our system more often agreeable
-%\item hypothesis here is that what a decision making system is trying to achieve is to find a decision most of the
-%whole community can agree with, or like it, even when not everyone participates and not all information is available
-%on their positions on a topic
-%\end{enumerate}
-%
-%One additional part of the experimental delegation study:
-%
-%Compute statistical quorum with and without delegations.
-%This semester, do not actually allow delegation to be used for passing statistical quorum, but collect data so that
-%we can see how much earlier delegation reached quorum before official statistical quorum.
-%(Only count how much earlier at the point where delegation passed quorum and never went below thereafter, thus
-%showing the point where the decision truly was clear.)
-%
-%Evaluation of statistical quorum (given that we don't officially decide by including delegation this semester):
-%From email: ``Do an internal Cloyne comparison that is more mathematical, showing how many agenda items were decided
-%using fewer actual votes than old 33\% quorum, as a result of statistical quorum.
-%This could be measured by average \# votes needed to pass statistical quorum vs. \% of those items that get
-%formally passed at council via consent agenda.
-%(If it's 100\%, then you can say that we saved X\% of votes on
-%average by using our online voting).
-%Then you could also show that some items needed more than 33\% of votes, implying that a hard-line 33\% is also bad.'' 
-%
-%Subjective evaluation of entire platform: Council Surveys comparing past two semesters to this semester for other
-%houses vs. Cloyne.
-%``use the previous 2 semesters' worth of council surveys as control, and a council survey for
-%Cloyne at the end of this semester as results.
-%These results would speak to those specific subjective questions
-%and would be one useful thing to look at.''
-%
-%\bibliographystyle{sigchi}
-%\bibliography{refs}
-%
-%More:
-%\begin{itemize}
-%\item https://mason.gmu.edu/~rhanson/futarchy.html
-%\item http://merkle.com/papers/DAOdemocracyDraft.pdf
-%\end{itemize}
+\subsection{Three step process}
+
+\subsection{Real-time feedback}
+
+% Real-time communication in general.
+
+\subsection{Integration with offline meetings}
+
+Decoupling the information presentation and voting processes from the debating process in the full decision-making
+pipeline (with the idea being that debating is best done in person, and presenting information/voting is best done online).
+
+\section{Evaluation}
+
+\subsection{Design}
+
+User study/evaluation design.
+To create a controlled evaluation of our solutions, to not depend on general usage of our system among communities,
+we can evaluate the design with the following user study:
+\begin{enumerate}
+\item we find a tight community of around 100 people willing to participate in the user study
+\item they should know each other well (to know how to delegate)
+\item we offer them to pay some Amazon credits or something for participation
+\item every member of the community should define delegations
+\item we ask them to split into two groups, group A and group B
+\item group A (target size is around 20% of the community) will meet together or online and discuss various topics through our system
+\item discussed topics some can be completely random, to get people to be familiar with they system, and also have some
+easy topics to decide on, like, what ``should be the color of the sky if we could change it'', and some topics which are relevant to the community
+\item those discussions would be just for evaluation, but community can use the results in any way they find useful
+\item after this discussions and decisions were done we ask everyone from group B to evaluate the results, we present them with two results (blinded, them not knowing which was made in which way)
+\begin{itemize}
+\item results which would be obtained using traditional methods (majority voting without any our improvements)
+\item our results using our system
+\end{itemize}
+\item we ask them to evaluate which one (or both, or none) of decisions they agree with
+\item prediction is that members who have not participated in the decision making process will
+find decisions made through our system more often agreeable
+\item hypothesis here is that what a decision making system is trying to achieve is to find a decision most of the
+whole community can agree with, or like it, even when not everyone participates and not all information is available
+on their positions on a topic
+\end{enumerate}
+
+One additional part of the experimental delegation study:
+
+Compute statistical quorum with and without delegations.
+This semester, do not actually allow delegation to be used for passing statistical quorum, but collect data so that
+we can see how much earlier delegation reached quorum before official statistical quorum.
+(Only count how much earlier at the point where delegation passed quorum and never went below thereafter, thus
+showing the point where the decision truly was clear.)
+
+Evaluation of statistical quorum (given that we don't officially decide by including delegation this semester):
+From email: ``Do an internal Cloyne comparison that is more mathematical, showing how many agenda items were decided
+using fewer actual votes than old 33\% quorum, as a result of statistical quorum.
+This could be measured by average \# votes needed to pass statistical quorum vs. \% of those items that get
+formally passed at council via consent agenda.
+(If it's 100\%, then you can say that we saved X\% of votes on
+average by using our online voting).
+Then you could also show that some items needed more than 33\% of votes, implying that a hard-line 33\% is also bad.''
+
+Subjective evaluation of entire platform: Council Surveys comparing past two semesters to this semester for other
+houses vs. Cloyne.
+``use the previous 2 semesters' worth of council surveys as control, and a council survey for
+Cloyne at the end of this semester as results.
+These results would speak to those specific subjective questions
+and would be one useful thing to look at.''
+
+\bibliographystyle{sigchi}
+\bibliography{refs}
+
+More:
+\begin{itemize}
+\item https://mason.gmu.edu/~rhanson/futarchy.html
+\item http://merkle.com/papers/DAOdemocracyDraft.pdf
+\end{itemize}
 

### 2017-04-24 dynamic-quorum.tex
Organizing papers.

\documentclass{sigchi}

% Use this section to set the ACM copyright statement (e.g. for
% preprints).  Consult the conference website for the camera-ready
% copyright statement.

% Copyright
\CopyrightYear{2016}
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}
% DOI
\doi{http://dx.doi.org/10.475/123_4}
% ISBN
\isbn{123-4567-24-567/08/06}
%Conference
\conferenceinfo{CHI'16,}{May 07--12, 2016, San Jose, CA, USA}
%Price
\acmPrice{\$15.00}

% Use this command to override the default ACM copyright statement
% (e.g. for preprints).  Consult the conference website for the
% camera-ready copyright statement.

%% HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP --
%% Please note you need to make sure the copy for your specific
%% license is used here!
% \toappear{
% Permission to make digital or hard copies of all or part of this work
% for personal or classroom use is granted without fee provided that
% copies are not made or distributed for profit or commercial advantage
% and that copies bear this notice and the full citation on the first
% page. Copyrights for components of this work owned by others than ACM
% must be honored. Abstracting with credit is permitted. To copy
% otherwise, or republish, to post on servers or to redistribute to
% lists, requires prior specific permission and/or a fee. Request
% permissions from \href{mailto:Permissions@acm.org}{Permissions@acm.org}. \\
% \emph{CHI '16},  May 07--12, 2016, San Jose, CA, USA \\
% ACM xxx-x-xxxx-xxxx-x/xx/xx\ldots \$15.00 \\
% DOI: \url{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
% }

% Arabic page numbers for submission.  Remove this line to eliminate
% page numbers for the camera ready copy
% \pagenumbering{arabic}

% Load basic packages
\usepackage{balance}       % to better equalize the last page
\usepackage{graphics}      % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}

% Some optional stuff you might like/need.
\usepackage{microtype}        % Improved Tracking and Kerning
% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
\usepackage{ccicons}          % Cite your images correctly!
% \usepackage[utf8]{inputenc} % for a UTF8 editor only

% If you want to use todo notes, marginpars etc. during creation of
% your draft document, you have to enable the "chi_draft" option for
% the document class. To do this, change the very first line to:
% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
% by using the "\todo{...}"  command. Make sure to disable the draft
% option again before submitting your final document.
\usepackage{todonotes}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{SIGCHI Conference Proceedings Format}
\def\plainauthor{First Author, Second Author, Third Author,
  Fourth Author, Fifth Author, Sixth Author}
\def\emptyauthor{}
\def\plainkeywords{Authors' choice; of terms; separated; by
  semicolons; include commas, within terms only; required.}
\def\plaingeneralterms{Documentation, Standardization}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  pdfdisplaydoctitle=true, % For Accessibility
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{\plaintitle}

\numberofauthors{3}
\author{%
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
}

\maketitle

\begin{abstract}
The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.  While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?  We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic.  We developed an online system centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 28 weeks.  We found that the community was able to use the online system to simultaneously improve decision-making legitimacy and efficiency.
\end{abstract}

\category{H.5.m.}{Information Interfaces and Presentation
  (e.g. HCI)}{Miscellaneous} \category{See
  \url{http://acm.org/about/class/1998/} for the full list of ACM
  classifiers. This section is required.}{}{}

\keywords{\plainkeywords}

\section{Introduction}

This format is to be used for submissions that are published in the
conference proceedings. We wish to give this volume a consistent,
high-quality appearance. We therefore ask that authors follow some
simple guidelines. You should format your paper exactly like this
document. The easiest way to do this is to replace the content with
your own material.  This document describes how to prepare your
submissions using \LaTeX.

\section{Page Size and Columns}
On each page your material should fit within a rectangle of 7 $\times$
9.15 inches (18 $\times$ 23.2 cm), centered on a US Letter page (8.5
$\times$ 11 inches), beginning 0.85 inches (1.9 cm) from the top of
the page, with a 0.3 inches (0.85 cm) space between two 3.35 inches
(8.4 cm) columns. Right margins should be justified, not
ragged. Please be sure your document and PDF are US letter and not A4.

\section{Typeset Text}
The styles contained in this document have been modified from the
default styles to reflect ACM formatting conventions. For example,
content paragraphs like this one are formatted using the Normal style.

\LaTeX\ sometimes will create overfull lines that extend into columns.
To attempt to combat this, the \texttt{.cls} file has a command,
\texttt{{\textbackslash}sloppy}, that essentially asks \LaTeX\ to
prefer underfull lines with extra whitespace.  For more details on
this, and info on how to control it more finely, check out
{\url{http://www.economics.utoronto.ca/osborne/latex/PMAKEUP.HTM}}.

\subsection{Title and Authors}

Your paper's title, authors and affiliations should run across the
full width of the page in a single column 17.8 cm (7 in.) wide.  The
title should be in Helvetica or Arial 18-point bold.  Authors' names
should be in Times New Roman or Times Roman 12-point bold, and
affiliations in 12-point regular.

See \texttt{{\textbackslash}author} section of this template for
instructions on how to format the authors. For more than three
authors, you may have to place some address information in a footnote,
or in a named section at the end of your paper. Names may optionally
be placed in a single centered row instead of at the top of each
column. Leave one 10-point line of white space below the last line of
affiliations.

\subsection{Abstract and Keywords}

Every submission should begin with an abstract of about 150 words,
followed by a set of Author Keywords and ACM Classification
Keywords. The abstract and keywords should be placed in the left
column of the first page under the left half of the title. The
abstract should be a concise statement of the problem, approach, and
conclusions of the work described. It should clearly state the paper's
contribution to the field of HCI\@.

\subsection{Normal or Body Text}

Please use a 10-point Times New Roman or Times Roman font or, if this
is unavailable, another proportional font with serifs, as close as
possible in appearance to Times Roman 10-point. Other than Helvetica
or Arial headings, please use sans-serif or non-proportional fonts
only for special purposes, such as source code text.

\subsection{First Page Copyright Notice}
This template include a sample ACM copyright notice at the bottom of
page 1, column 1.  Upon acceptance, you will be provided with the
appropriate copyright statement and unique DOI string for publication.
Accepted papers will be distributed in the conference
publications. They will also be placed in the ACM Digital Library,
where they will remain accessible to thousands of researchers and
practitioners worldwide. See
\url{http://acm.org/publications/policies/copyright_policy} for the
ACM's copyright and permissions policy.

\subsection{Subsequent Pages}

On pages beyond the first, start at the top of the page and continue
in double-column format.  The two columns on the last page should be
of equal length.

\begin{figure}
\centering
  %\includegraphics[width=0.9\columnwidth]{figures/sigchi-logo}
  \caption{Insert a caption below each figure. Do not alter the
    Caption style.  One-line captions should be centered; multi-line
    should be justified. }~\label{fig:figure1}
\end{figure}

\subsection{References and Citations}

Use a numbered list of references at the end of the article, ordered
alphabetically by last name of first author, and referenced by numbers
in
brackets~\cite{acm_categories,ethics,Klemmer:2002:WSC:503376.503378}.
Your references should be published materials accessible to the
public. Internal technical reports may be cited only if they are
easily accessible (i.e., you provide the address for obtaining the
report within your citation) and may be obtained by any reader for a
nominal fee. Proprietary information may not be cited. Private
communications should be acknowledged in the main text, not referenced
(e.g., ``[Borriello, personal communication]'').

References should be in ACM citation format:
\url{http://acm.org/publications/submissions/latex_style}. This
includes citations to internet
resources~\cite{acm_categories,cavender:writing,CHINOSAUR:venue,psy:gangnam}
according to ACM format, although it is often appropriate to include
URLs directly in the text, as above.

% Use a numbered list of references at the end of the article, ordered
% alphabetically by first author, and referenced by numbers in
% brackets~\cite{ethics, Klemmer:2002:WSC:503376.503378,
%   Mather:2000:MUT, Zellweger:2001:FAO:504216.504224}. For papers from
% conference proceedings, include the title of the paper and an
% abbreviated name of the conference (e.g., for Interact 2003
% proceedings, use \textit{Proc. Interact 2003}). Do not include the
% location of the conference or the exact date; do include the page
% numbers if available. See the examples of citations at the end of this
% document. Within this template file, use the \texttt{References} style
% for the text of your citation.

% Your references should be published materials accessible to the
% public.  Internal technical reports may be cited only if they are
% easily accessible (i.e., you provide the address for obtaining the
% report within your citation) and may be obtained by any reader for a
% nominal fee.  Proprietary information may not be cited. Private
% communications should be acknowledged in the main text, not referenced
% (e.g., ``[Robertson, personal communication]'').

\begin{table}
  \centering
  \begin{tabular}{l r r r}
    % \toprule
    & & \multicolumn{2}{c}{\small{\textbf{Test Conditions}}} \\
    \cmidrule(r){3-4}
    {\small\textit{Name}}
    & {\small \textit{First}}
      & {\small \textit{Second}}
    & {\small \textit{Final}} \\
    \midrule
    Marsden & 223.0 & 44 & 432,321 \\
    Nass & 22.2 & 16 & 234,333 \\
    Borriello & 22.9 & 11 & 93,123 \\
    Karat & 34.9 & 2200 & 103,322 \\
    % \bottomrule
  \end{tabular}
  \caption{Table captions should be placed below the table. We
    recommend table lines be 1 point, 25\% black. Minimize use of
    table grid lines.}~\label{tab:table1}
\end{table}

\section{Sections}

The heading of a section should be in Helvetica or Arial 9-point bold,
all in capitals. Sections should \textit{not} be numbered.

\subsection{Subsections}

Headings of subsections should be in Helvetica or Arial 9-point bold
with initial letters capitalized.  For sub-sections and
sub-subsections, a word like \emph{the} or \emph{of} is not
capitalized unless it is the first word of the heading.

\subsubsection{Sub-subsections}

Headings for sub-subsections should be in Helvetica or Arial 9-point
italic with initial letters capitalized.  Standard
\texttt{{\textbackslash}section}, \texttt{{\textbackslash}subsection},
and \texttt{{\textbackslash}subsubsection} commands will work fine in
this template.

\section{Figures/Captions}

Place figures and tables at the top or bottom of the appropriate
column or columns, on the same page as the relevant text (see
Figure~\ref{fig:figure1}). A figure or table may extend across both
columns to a maximum width of 17.78 cm (7 in.).

\begin{figure*}
  \centering
  %\includegraphics[width=1.75\columnwidth]{figures/map}
  \caption{In this image, the map maximizes use of space. You can make
    figures as wide as you need, up to a maximum of the full width of
    both columns. Note that \LaTeX\ tends to render large figures on a
    dedicated page. Image: \ccbynd~ayman on
    Flickr.}~\label{fig:figure2}
\end{figure*}

Captions should be Times New Roman or Times Roman 9-point bold.  They
should be numbered (e.g., ``Table~\ref{tab:table1}'' or
``Figure~\ref{fig:figure1}''), centered and placed beneath the figure
or table.  Please note that the words ``Figure'' and ``Table'' should
be spelled out (e.g., ``Figure'' rather than ``Fig.'') wherever they
occur. Figures, like Figure~\ref{fig:figure2}, may span columns and
all figures should also include alt text for improved accessibility.
Papers and notes may use color figures, which are included in the page
limit; the figures must be usable when printed in black-and-white in
the proceedings.

The paper may be accompanied by a short video figure up to five
minutes in length. However, the paper should stand on its own without
the video figure, as the video may not be available to everyone who
reads the paper.

\subsection{Inserting Images}
When possible, include a vector formatted graphic (i.e. PDF or EPS).
When including bitmaps,  use an image editing tool to resize the image
at the appropriate printing resolution (usually 300 dpi).

\section{Quotations}
Quotations may be italicized when \textit{``placed inline''} (Anab,
23F).

\begin{quote}
Longer quotes, when placed in their own paragraph, need not be
italicized or in quotation marks when indented (Ramon, 39M).  
\end{quote}

\section{Language, Style, and Content}

The written and spoken language of SIGCHI is English. Spelling and
punctuation may use any dialect of English (e.g., British, Canadian,
US, etc.) provided this is done consis- tently. Hyphenation is
optional. To ensure suitability for an international audience, please
pay attention to the following:

\begin{itemize}
\item Write in a straightforward style.
\item Try to avoid long or complex sentence structures.
\item Briefly define or explain all technical terms that may be
  unfamiliar to readers.
\item Explain all acronyms the first time they are used in your
  text---e.g., ``Digital Signal Processing (DSP)''.
\item Explain local references (e.g., not everyone knows all city
  names in a particular country).
\item Explain ``insider'' comments. Ensure that your whole audience
  understands any reference whose meaning you do not describe (e.g.,
  do not assume that everyone has used a Macintosh or a particular
  application).
\item Explain colloquial language and puns. Understanding phrases like
  ``red herring'' may require a local knowledge of English.  Humor and
  irony are difficult to translate.
\item Use unambiguous forms for culturally localized concepts, such as
  times, dates, currencies, and numbers (e.g., ``1--5--97'' or
  ``5/1/97'' may mean 5 January or 1 May, and ``seven o'clock'' may
  mean 7:00 am or 19:00). For currencies, indicate equivalences:
  ``Participants were paid {\fontfamily{txr}\selectfont \textwon}
  25,000, or roughly US \$22.''
\item Be careful with the use of gender-specific pronouns (he, she)
  and other gendered words (chairman, manpower, man-months). Use
  inclusive language that is gender-neutral (e.g., she or he, they,
  s/he, chair, staff, staff-hours, person-years). See the
  \textit{Guidelines for Bias-Free Writing} for further advice and
  examples regarding gender and other personal
  attributes~\cite{Schwartz:1995:GBF}. Be particularly aware of
  considerations around writing about people with disabilities.
\item If possible, use the full (extended) alphabetic character set
  for names of persons, institutions, and places (e.g.,
  Gr{\o}nb{\ae}k, Lafreni\'ere, S\'anchez, Nguy{\~{\^{e}}}n,
  Universit{\"a}t, Wei{\ss}enbach, Z{\"u}llighoven, \r{A}rhus, etc.).
  These characters are already included in most versions and variants
  of Times, Helvetica, and Arial fonts.
\end{itemize}

\section{Accessibility}
The Executive Council of SIGCHI has committed to making SIGCHI
conferences more inclusive for researchers, practitioners, and
educators with disabilities. As a part of this goal, the all authors
are asked to work on improving the accessibility of their
submissions. Specifically, we encourage authors to carry out the
following five steps:
\begin{enumerate}
\item Add alternative text to all figures
\item Mark table headings
\item Add tags to the PDF
\item Verify the default language
\item Set the tab order to ``Use Document Structure''
\end{enumerate}
For more information and links to instructions and resources, please
see: \url{http://chi2016.acm.org/accessibility}.  The
\texttt{{\textbackslash}hyperref} package allows you to create well tagged PDF files,
please see the preamble of this template for an example.

\section{Page Numbering, Headers and Footers}
Your final submission should not contain footer or header information
at the top or bottom of each page. Specifically, your final submission
should not include page numbers. Initial submissions may include page
numbers, but these must be removed for camera-ready. Page numbers will
be added to the PDF when the proceedings are assembled.

\section{Producing and Testing PDF Files}

We recommend that you produce a PDF version of your submission well
before the final deadline.  Your PDF file must be ACM DL
Compliant. The requirements for an ACM Compliant PDF are available at:
{\url{http://www.sheridanprinting.com/typedept/ACM-distilling-settings.htm}}.

Test your PDF file by viewing or printing it with the same software we
will use when we receive it, Adobe Acrobat Reader Version 10. This is
widely available at no cost. Note that most
reviewers will use a North American/European version of Acrobat
reader, so please check your PDF accordingly.

When creating your PDF from Word, ensure that you generate a tagged
PDF from improved accessibility. This can be done by using the Adobe
PDF add-in, also called PDFMaker. Select Acrobat | Preferences from
the ribbon and ensure that ``Enable Accessibility and Reflow with
tagged Adobe PDF'' is selected. You can then generate a tagged PDF by
selecting ``Create PDF'' from the Acrobat ribbon.

\section{Conclusion}

It is important that you write for the SIGCHI audience. Please read
previous years' proceedings to understand the writing style and
conventions that successful authors have used. It is particularly
important that you state clearly what you have done, not merely what
you plan to do, and explain how your work is different from previously
published work, i.e., the unique contribution that your work makes to
the field. Please consider what the reader will learn from your
submission, and how they will find your work useful. If you write with
these questions in mind, your work is more likely to be successful,
both in being accepted into the conference, and in influencing the
work of our field.

\section{Acknowledgments}

Sample text: We thank all the volunteers, and all publications support
and staff, who wrote and provided helpful comments on previous
versions of this document. Authors 1, 2, and 3 gratefully acknowledge
the grant from NSF (\#1234--2012--ABC). \textit{This whole paragraph is
  just an example.}

% Balancing columns in a ref list is a bit of a pain because you
% either use a hack like flushend or balance, or manually insert
% a column break.  http://www.tex.ac.uk/cgi-bin/texfaq2html?label=balance
% multicols doesn't work because we're already in two-column mode,
% and flushend isn't awesome, so I choose balance.  See this
% for more info: http://cs.brown.edu/system/software/latex/doc/balance.pdf
%
% Note that in a perfect world balance wants to be in the first
% column of the last page.
%
% If balance doesn't work for you, you can remove that and
% hard-code a column break into the bbl file right before you
% submit:
%
% http://stackoverflow.com/questions/2149854/how-to-manually-equalize-columns-
% in-an-ieee-paper-if-using-bibtex
%
% Or, just remove \balance and give up on balancing the last page.
%
\balance{}

\section{References Format}
Your references should be published materials accessible to the
public. Internal technical reports may be cited only if they are
easily accessible and may be obtained by any reader for a nominal
fee. Proprietary information may not be cited. Private communications
should be acknowledged in the main text, not referenced (e.g.,
[Golovchinsky, personal communication]). References must be the same
font size as other body text. References should be in alphabetical
order by last name of first author. Use a numbered list of references
at the end of the article, ordered alphabetically by last name of
first author, and referenced by numbers in brackets. For papers from
conference proceedings, include the title of the paper and the name of
the conference. Do not include the location of the conference or the
exact date; do include the page numbers if available.

References should be in ACM citation format:
\url{http://www.acm.org/publications/submissions/latex_style}.  This
includes citations to Internet
resources~\cite{CHINOSAUR:venue,cavender:writing,psy:gangnam}
according to ACM format, although it is often appropriate to include
URLs directly in the text, as above. Example reference formatting for
individual journal articles~\cite{ethics}, articles in conference
proceedings~\cite{Klemmer:2002:WSC:503376.503378},
books~\cite{Schwartz:1995:GBF}, theses~\cite{sutherland:sketchpad},
book chapters~\cite{winner:politics}, an entire journal
issue~\cite{kaye:puc},
websites~\cite{acm_categories,cavender:writing},
tweets~\cite{CHINOSAUR:venue}, patents~\cite{heilig:sensorama}, 
games~\cite{supermetroid:snes}, and
online videos~\cite{psy:gangnam} is given here.  See the examples of
citations at the end of this document and in the accompanying
\texttt{BibTeX} document. This formatting is a edited version of the
format automatically generated by the ACM Digital Library
(\url{http://dl.acm.org}) as ``ACM Ref.'' DOI and/or URL links are
optional but encouraged as are full first names. Note that the
Hyperlink style used throughout this document uses blue links;
however, URLs in the references section may optionally appear in
black.

% BALANCE COLUMNS
\balance{}

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{sample}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-04-24 delegation.tex
Organizing papers.
Syncing templates.

--- 2016-11-29 paper.tex
@@
 
-%% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
-% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is      granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
-Copyrights for components of this work owned by others than ACM must be honored.
-Abstracting with credit is permitted.
-To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee.
-Request permissions from permissions@acm.org.
-\\
-% {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
-% Copyright \copyright~2014 ACM ISBN/14/04...\$15.00. \\
-% DOI string from ACM form confirmation}
-%% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
+% Use this section to set the ACM copyright statement (e.g. for
+% preprints).  Consult the conference website for the camera-ready
+% copyright statement.
+
+% Copyright
+%\CopyrightYear{2016}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+% DOI
+%\doi{http://dx.doi.org/10.475/123_4}
+% ISBN
+%\isbn{123-4567-24-567/08/06}
+%Conference
+%\conferenceinfo{CHI'16,}{May 07--12, 2016, San Jose, CA, USA}
+%Price
+%\acmPrice{\$15.00}
+
+% Use this command to override the default ACM copyright statement
+% (e.g. for preprints).  Consult the conference website for the
+% camera-ready copyright statement.
+
+%% HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP --
+%% Please note you need to make sure the copy for your specific
+%% license is used here!
+% \toappear{
+% Permission to make digital or hard copies of all or part of this work
+% for personal or classroom use is granted without fee provided that
+% copies are not made or distributed for profit or commercial advantage
+% and that copies bear this notice and the full citation on the first
+% page. Copyrights for components of this work owned by others than ACM
+% must be honored. Abstracting with credit is permitted. To copy
+% otherwise, or republish, to post on servers or to redistribute to
+% lists, requires prior specific permission and/or a fee. Request
+% permissions from \href{mailto:Permissions@acm.org}{Permissions@acm.org}. \\
+% \emph{CHI '16},  May 07--12, 2016, San Jose, CA, USA \\
+% ACM xxx-x-xxxx-xxxx-x/xx/xx\ldots \$15.00 \\
+% DOI: \url{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
+% }
 
@@
 % page numbers for the camera ready copy
-
 \pagenumbering{arabic}
@@
 \usepackage{graphics}      % for EPS, load graphicx instead
-\usepackage[T1]{fontenc}
+\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
 \usepackage{txfonts}
-\usepackage{times}    % comment if you want LaTeX's default font
-\usepackage[pdftex]{hyperref}
-\usepackage{url}      % llt: nicely formatted URLs
+\usepackage{mathptmx}
+\usepackage[pdflang={en-US},pdftex]{hyperref}
 \usepackage{color}
+\usepackage{booktabs}
 \usepackage{textcomp}
@@
 \usepackage{ccicons}
-\usepackage{todonotes}
 \usepackage{pgf}
@@
 
+% Some optional stuff you might like/need.
+\usepackage{microtype}        % Improved Tracking and Kerning
+% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
+\usepackage{ccicons}          % Cite your images correctly!
+% \usepackage[utf8]{inputenc} % for a UTF8 editor only
+
+% If you want to use todo notes, marginpars etc. during creation of
+% your draft document, you have to enable the "chi_draft" option for
+% the document class. To do this, change the very first line to:
+% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
+% by using the "\todo{...}"  command. Make sure to disable the draft
+% option again before submitting your final document.
+\usepackage{todonotes}
+
+% Paper metadata (use plain text, for PDF inclusion and later
+% re-using, if desired).  Use \emtpyauthor when submitting for review
+% so you remain anonymous.
+\def\plaintitle{PeerMind: Delegation}
+%\def\plainauthor{First Author, Second Author, Third Author,
+%  Fourth Author, Fifth Author, Sixth Author}
+%\def\emptyauthor{}
+%\def\plainkeywords{Authors' choice; of terms; separated; by
+%  semicolons; include commas, within terms only; required.}
+%\def\plaingeneralterms{Documentation, Standardization}
+
 % llt: Define a global style for URLs, rather that the default one
@@
 \def\url@leostyle{%
-  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
+  \@ifundefined{selectfont}{
+    \def\UrlFont{\sf}
+  }{
+    \def\UrlFont{\small\bf\ttfamily}
+  }}
 \makeatother
@@
 %\hypersetup{%
-%  pdftitle={SIGCHI Conference Proceedings Format},
-%  pdfauthor={LaTeX},
-%  pdfkeywords={SIGCHI, proceedings, archival format},
+%  pdftitle={\plaintitle},
+%% Use \plainauthor for final version.
+%%  pdfauthor={\plainauthor},
+%  pdfauthor={\emptyauthor},
+%  pdfkeywords={\plainkeywords},
+%  pdfdisplaydoctitle=true, % For Accessibility
 %  bookmarksnumbered,
@@
 %  breaklinks=true,
+%  hypertexnames=false
 %}
@@
 
+% End of preamble. Here it comes the document.
 \begin{document}
 
-\title{PeerMind: Towards Collective Decision-Making}
+\title{\plaintitle}
 
@@
 %\author{%
-%  \alignauthor{1st Author Name\\
-%    \affaddr{Affiliation}\\
+%  \alignauthor{Leave Authors Anonymous\\
+%    \affaddr{for Submission}\\
 %    \affaddr{City, Country}\\
 %    \email{e-mail address}}\\
-%  \alignauthor{2nd Author Name\\
-%    \affaddr{Affiliation}\\
+%  \alignauthor{Leave Authors Anonymous\\
+%    \affaddr{for Submission}\\
 %    \affaddr{City, Country}\\
 %    \email{e-mail address}}\\
-%  \alignauthor{3rd Author Name\\
-%    \affaddr{Affiliation}\\
+%  \alignauthor{Leave Authors Anonymous\\
+%    \affaddr{for Submission}\\
 %    \affaddr{City, Country}\\
@@
 
-%\keywords{Authors' choice; of terms; separated; by semi\-colons;
-%  commas, within terms only; this section is required.}
-
 %\category{H.5.m.}{Information Interfaces and Presentation
@@
 %  classifiers. This section is required.}{}{}
+%
+%\keywords{\plainkeywords
 
@@
 \item[$N = $] population size,
-\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the raw votes for motion $X_i$, excluding ``abstentions'',
+\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for motion $X_i$, excluding ``abstentions'',
 \item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on motion $X_i$,
@@
 \Pr(\alpha^*_1 \leq T_1 \mid X_1) &= \int_0^1 \Pr(\alpha^*_1 \leq T_1 \mid p_1)\cdot \Pr(p_1 \mid X_1) \mathrm{d}p_1\\
-&= \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{K-i} {(1-p_1)}^{i} \\
+&= \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \\
 & \qquad\qquad\quad \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1,
@@
 & \Pr(\alpha^*_1 > T_1 \mid X_1) = \nonumber\\
-& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{K_1-i} {(1-p_1)}^{i} \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
+& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
 \end{align}
@@
 %
+%\todo[inline]{Count number of decision being made before using PeerMind, and number of them this semester. How many people participated in voting?}
+%
 %\subsection{Design}
@@
 %
+% REFERENCES FORMAT
+% References must be the same font size as other body text.
 %\bibliographystyle{sigchi}
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2017-04-24 statistical-quorum.tex
Syncing templates.

--- 2016-12-03 paper.tex
@@
 \usepackage{graphics}      % for EPS, load graphicx instead 
-\usepackage[T1]{fontenc}
+\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
 \usepackage{txfonts}
-\usepackage{times}    % comment if you want LaTeX's default font
-\usepackage[pdftex]{hyperref}
-\usepackage{url}      % llt: nicely formatted URLs
+\usepackage{mathptmx}
+\usepackage[pdflang={en-US},pdftex]{hyperref}
 \usepackage{color}
+\usepackage{booktabs}
 \usepackage{textcomp}
-\usepackage{booktabs}
-\usepackage{ccicons}
+
+% Some optional stuff you might like/need.
+\usepackage{microtype}        % Improved Tracking and Kerning
+% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
+\usepackage{ccicons}          % Cite your images correctly!
+% \usepackage[utf8]{inputenc} % for a UTF8 editor only
+
+% If you want to use todo notes, marginpars etc. during creation of
+% your draft document, you have to enable the "chi_draft" option for
+% the document class. To do this, change the very first line to:
+% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
+% by using the "\todo{...}"  command. Make sure to disable the draft
+% option again before submitting your final document.
 \usepackage{todonotes}
-\usepackage{pgf}
-\usepackage{tikz}
-\usepackage{tkz-graph}
-\usepackage{mathtools}
-\usepackage{amsmath}
+
+% Paper metadata (use plain text, for PDF inclusion and later
+% re-using, if desired).  Use \emtpyauthor when submitting for review
+% so you remain anonymous.
+\def\plaintitle{PeerMind: Statistical Quorum}
+%\def\plainauthor{First Author, Second Author, Third Author,
+%  Fourth Author, Fifth Author, Sixth Author}
+%\def\emptyauthor{}
+%\def\plainkeywords{Authors' choice; of terms; separated; by
+%  semicolons; include commas, within terms only; required.}
+%\def\plaingeneralterms{Documentation, Standardization}
 
@@
 
-\GraphInit[vstyle=Dijkstra]
-\tikzset{
-  EdgeStyle/.append style={->,bend left}
-}
-
-\hyphenation{PeerMind}
-
+% End of preamble. Here it comes the document.
 \begin{document}
 
-\title{PeerMind}
+\title{\plaintitle}
 
@@
 \begin{abstract}
-
-%\todo[inline]{TODO}
-
+The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
+While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?
+We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic.
+We developed an online system centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 28 weeks.
+We found that the community was able to use the online system to simultaneously improve decision-making legitimacy and efficiency.
 \end{abstract}
-
-%\keywords{Authors' choice; of terms; separated; by semi\-colons;
-%  commas, within terms only; this section is required.}
 
@@
 %  classifiers. This section is required.}{}{}
+%
+%\keywords{\plainkeywords}
 
@@
 
-The broad idea is that instead of just trying to move the voting process online -- something many others do -- we in
-fact add additional value by voting online.
-Specifically, we can compute results faster and more accurately, we can use sophisticated statistics to estimate
-quantities for the entire voting population, and we can promote clarity and full understanding of the voting items for all.
-
-\todo[inline]{Main questions and answers follow.}
-
-What problem are we solving? Why is it important to solve? What is the story?
-
-Finding a way to lower attention required but increase legitimacy of democratic decisions.
-Traditionally legitimacy is increased by increasing participation/attention required.
-It is an everlasting problem.
-How to outsource governance and at the same time assure it does not get corrupt
-or tyrannic.
-
-Has anyone else tried to solve it or related problems? How did they try
-and what were the results (subjective or objective)?
-
-Described here: \url{http://mitar.tnode.com/post/73983101095/peer-to-peer-voting-scheme}.
-Some tried delegations, but mostly only to one person.
-Getting issues that you have tree-structure with possible cycles you cannot resolve.
-Often there were reported UI problems as well.
-So just technically improving is not enough, you also have to have UI for the technical solution so that people can trust.
-
-How are you solving the problem? What technologies are you using?
-
-Using dynamic quorum, score voting, delegation, and nice integration with UI/UX.
-
-How will you know how well you solved the problem? How will those technologies address the issues?
-
-If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
-So currently we have issues with reaching the quorum and many people do not participate.
-I would like for them to have easier time participating (less time required) and as a consequence more people to
-participate collectively.
-If I observe that I think I will know I solved the problem.
-
-\todo[inline]{This rest of the section is mostly notes. Organize and improve.}
-
-We can often hear that the issue of our democracy is low engagement.
-But this is a fallacy, because we should be instead trying to decrease engagement.
-The best governance system is one where you do not have to be involved yourself, but it is still doing things
-like you would like to do them.
-% Or even, doing things you might not even know that you would like to be done, especially in a long term.
-We want to outsource the job of coordination between people.
-Sometimes you want to coordinate people yourself (in your company, team), you might enjoy it, but in most cases
-you cannot or do not want to, so you outsource this service, especially at a large scale.
-
-One reason why increasing engagement sounds like a good idea is because current democratic systems have two
-questionable assumptions: participants have time to participate, and knowledge to participate.
-If you do not participate it is your fault.
-Moreover, the legitimacy of decisions is based on participation as well.
-This is why those participating in the these democratic systems have motivation to try to increase participation
-of others (especially if they believe they will still be in majority), or claim they are at fault to not participate.
-
-engagement is then increased only to assure that people like the results
-
-so if we try to minimize engagement, but still make sure that people like the results, then this is the perfect system
-
-and only in the case when they are on the loosing side, then we have to make sure that they understand the reasons why
-they lost (e.g.,\ the middle panel in our design) or at least to feel that they participated, in at least some way
-(voting, or just delegating)
-
-but the participation is the last on this list of important factors
-
-the more important thing is that they understand why others were thinking differently
-or that they do not even have to be engaged
-
-we want that people do not have to pay attention to what government is doing
-
-currently, we are in a loop where we have a useless feedback loop, where government in fact would like for people
-to pay some type of attention because they need to be elected, so they have to show like they are doing something,
-but on the other hand they do not want too much engagement, because they do not have things in place to really
-engage with everyone
-
-government should be like infrastructure, you should not have to care about it while it works, and be there to fix
-things if it starts failing, but not even you personally, but have mechanisms and processes in place to find this
-
-Terminology:
-\begin{itemize}
-\item community -- group which is making a decision using PeerMind
-\item member -- user of the app who is participating in decision making
-\end{itemize}
-
-\todo[inline]{Terminology: decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)}
-\todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}
-\todo[inline]{Terminology: discarding a vote if you didn't vote and didn't delegate. Some other term?}
-
-The main question is: can we increase legitimacy and decrease attention required at the same time.
-Many others are trying to increase attention (participation) because they want to increase legitimacy,
-assuming that this is the only way.
-
-A list of potential other points:
-\begin{itemize}
-\item legitimacy (quorum, involvement of more people through delegation)
-\item scaling (delegation helps, less attention helps)
-\item attention (at same or better decisions)
-\item bad assumptions for democracy - time, knowledge, it is rational to know that you do not know, common practice is to ask friends
-\item better decisions (more people consulted, attention directed where it should be)
-\end{itemize}
-
-Another question: should feedback be visible (to see how others voted before you voted)?
-People go with majority.
-Better for wisdom of crowds is an independent sample.
-Maybe we should only allow how other voted to be seen after you voted (but you can still change the vote),
-but we could then observe that and learn from this and see if this is good or bad.
-Or we could just store information if you observed existing votes before you voted (you have to click a button to see them).
-Maybe going with majority is also good when you want more social cohesion in the community and not so much ``optimal''
-wisdom of crowds result -  you would want that people relax their positions based on how other people care.
-Cohesion of the group vs. accuracy.
-Maybe display only number of votes and credibility, maybe credibility direction or graph, but not average votes.
-
-A list of potential questions:
-\begin{itemize}
-\item Delegation works?
-\item Statistical quorum works?
-\item Real-time feedback influences?
-\item Private vs. public delegation (do you know how member to whom you delegated is voting)?
-\item When delegating, how to assure that you cannot learn how somebody else voted, but that you can still somehow
-learn how your vote is used?
-\item Knowing how much power you have vs. not knowing -- does it change how people behave? (Knowing might corrupt you,
-but also motivate you to do more thorough work voting because you know other depend on you.)
-\end{itemize}
-
-Performance/theory/scaling in computational terms (how big could effectively community can be to be able to compute it).
-Some other theoretical properties?
-Computing statistical quorum is an ``embarrassingly parallel'' problem, and so its
-use would not suffer in large populations.
-Sparsity of delegation matrix also enables use of (recently developed?)
-efficient matrix computation algorithms.
-
-Out of scope of this work:
-
-\begin{itemize}
-\item How to assure privacy of the delegation network (it is highly sensitive data, probably even more than votes)?
-\item How to compute delegation results in a way that it does not require access to the whole network of delegation at once?
-\item How can we make this system distributed (and still assure privacy)?
-\end{itemize}
-
-\section{Related work}
-
-\cite{andersen2008trust, dirnstorfer2010voting, ford2002delegative, rodriguez2007smartocracy, yamakawa2007toward}
-
-\section{Features}
-
-\subsection{Statistical Quorum}
-
-\subsubsection*{Background}
-A population of voters needs to collectively decide which item (if any) to take action on out of a set of competing items.
-In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a winning item, while also demonstrating that the winning item is preferred over all other competing items.
-We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that the population does not take undesired actions, which in this case encompasses taking action on an item that is not supported by a majority of the population and/or taking action on an item that is not clearly preferred over all other competing alternatives.
-
-\subsubsection*{Problem Statement}
-We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority, $90\%$ for a consensus).
-In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater than the bias towards voting ``yes'' on all other competing items.
-
-\subsubsection*{Assumptions}
-We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
-
-Without any votes, we assume that this bias may be any value between $0\%$ and $100\%$ with equal probability.
-
-\subsubsection*{Voting}
-We allow for score voting where each member of the population may choose any real number in the range $[-1,+1]$ to represent their vote, with $-1$ representing opposition and $+1$ representing support.
-In the theory subsection below, we rescale votes to be in the range $[0,1]$ via the notation $\frac{votes + 1}{2}$.
-
-We allow voters to ``abstain'' from voting.
-We treat abstentions as indicating that a voter is removing themselves from the population of voters for an item.
-It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$ (indicating indifference).
-
-\subsubsection*{Bayesian Theory}
-To properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial: \url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
-
-The above tutorial is very similar to the problem we have at hand.
-The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises naturally when considering samples from a finite population.
-
-After extracting the sufficient statistics from raw votes --- namely the average score and the number of abstentions --- we can adjust the effective population size and corresponding threshold.
-In addition, we can calculate the effective number of ``yes'' and ``no'' votes, which one can think of as the number of Heads and Tails in the aforementioned tutorial.
-
-Deriving our expression of interest now immediately follows:
-
-Let
-\begin{description}
-\begin{itemize}
-\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing motions,
-\item[$N = $] population size,
-\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the raw votes for motion $X_i$, excluding ``abstentions'',
-\item[$N_i = $] $(N - $ number of abstentions on $X_i)$, the effective population voting on motion $X_i$,
-\item[$\alpha_i = $] $\sum\limits_{v \in X_i} (\frac{v+1}{2})$, the effective number of ``yes'' votes on motion $X_i$,
-\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
-\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
-\item[$p_i = $] true latent population bias towards voting ``yes'' on $X_i$.
-\end{itemize}
-\end{description}
-
-Often, we have a single motion open for voting.  In this scenario, $C=\{X_1\}$ and we simply wish to compute:
-\begin{align}\label{eq1}
-\Pr(\alpha^*_1 > T_1 \mid X_1),
-\end{align}
-where
-\begin{description}
-\begin{itemize}
-\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
-\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote (e.g. $f=\frac{1}{2}$ for a simple majority).
-\end{itemize}
-\end{description}
-Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
-\begin{align*}
-\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \Pr(\alpha^*_1 \leq T_1 \mid X_1),\\
-\Pr(\alpha^*_1 \leq T_1 \mid X_1) &= \int_0^1 \Pr(\alpha^*_1 \leq T_1 \mid p_1)\cdot \Pr(p_1 \mid X_1) \mathrm{d}p_1\\
-&= \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{K-i} {(1-p_1)}^{i} \\
-& \qquad\qquad\quad \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1,
-\end{align*}
-where
-\begin{description}
-\begin{itemize}
-\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that $X_i$ would remain below threshold.
-\end{itemize}
-\end{description}
-
-Hence, we have
-\begin{align}\label{eq2}
-& \Pr(\alpha^*_1 > T_1 \mid X_1) = \nonumber\\
-& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{K_1-i} {(1-p_1)}^{i} \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
-\end{align}
-
-In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to consider.
-Without loss of generality, assume that $X_1$ is the most popular motion within $C$.
-We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other $X_i  \in C, i \neq 1$:
-
-\begin{align}\label{eq3}
-\Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)
-\end{align}
-
-Using a safe assumption that votes on motions are independent across motions, we can decompose \eqref{eq3} as follows:
-\begin{align*}
-& \Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)\\
-=& \prod_{i=2}^n \Pr(\alpha^*_1 > \alpha^*_i \mid C)\\
-=& \prod_{i=2}^n \iint\limits_{p_1, p_i} \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \Pr(p_1 \mid X_1) \cdot \Pr(p_i \mid X_i) \mathrm{d}p_i \mathrm{d}p_1\\
-=& \prod_{i=2}^n \int_0^1 \int_0^1 \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \\
-& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1\\
-\end{align*}
-Deriving an expression for $\Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)$, we get:
-\begin{align*}
-& \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)\\
-=& 1 - \Pr(\alpha^*_1 \leq \alpha^*_i \mid p_1, p_i)\\
-=& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
-\end{align*}
-where
-\begin{description}
-\begin{itemize}
-\item[$v_{i,j} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
-\end{itemize}
-\end{description}
-
-Denoting
-\begin{align}\label{eq4}
-& \Phi_i(p_1,p_i) {\overset {\underset {\mathrm {def} }{}}{=}} \nonumber\\
-& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
-\end{align}
-we can simplify \eqref{eq3} to:
-\begin{align}\label{eq5}
-& \Pr\bigg(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\bigg)\nonumber\\
-&= \prod_{i=2}^n \int_0^1 \int_0^1 \Phi_i(p_1,p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \nonumber\\
-& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1
-\end{align}
-
-We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.  In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/simple_majority}
-\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
-\label{simple_majority}
-\end{figure}
-%Data:
-%Curve 1: {{0,0.000299648},{1,0.00213083},{2,0.00817989},{3,0.0224777},{4,0.049512},{5,0.0928927},{6,0.154136},{7,0.232044},{8,0.322859},{9,0.421057},{10,0.520467},{11,0.615358},{12,0.701241},{13,0.775267},{14,0.836245},{15,0.884385},{16,0.920892},{17,0.947532},{18,0.966265},{19,0.978973},{20,0.987296},{21,0.992561},{22,0.99578},{23,0.997682},{24,0.998767},{25,0.999366}}
-%Curve 2: {x, 0.99} for all x
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/super_majority}
-\caption{Super Majority: $\alpha$ vs. $\Pr(\alpha > \frac{2}{3}N)$ for $N=100$, $\beta = 10$.}
-\label{super_majority}
-\end{figure}
-%Data:
-%Curve 1: {{10,0.0395173},{11,0.0591278},{12,0.0846131},{13,0.116407},{14,0.15463},{15,0.199057},{16,0.24911},{17,0.303899},{18,0.362273},{19,0.422909},{20,0.484399},{21,0.545345},{22,0.604445},{23,0.660559},{24,0.712764},{25,0.760382},{26,0.802985},{27,0.840385},{28,0.872612},{29,0.899875},{30,0.922519},{31,0.940987},{32,0.955777},{33,0.967407},{34,0.976384},{35,0.983186},{36,0.988243},{37,0.99193},{38,0.994567},{39,0.996414},{40,0.997682},{41,0.998533},{42,0.999093},{43,0.999452},{44,0.999677},{45,0.999815},{46,0.999896},{47,0.999944},{48,0.99997},{49,0.999985},{50,0.999993}}
-%Curve 2: {x, 0.99} for all x
-
-\subsection{Score Voting}
-
-\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
-we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
-also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
-
-\subsection{Delegation}
-\todo[inline]{Max: I think delegation graph edges should point in the direction towards the person that the vote is
-being given to (more intuitive).
-\\
-Mitar: We are not giving votes to people, we are delegating a decision to people, but votes are computed for each person
-separately based on how delegates voted themselves. The idea is that you ask a delegate how the voted and then compute
-your vote. The do not get your vote to do anything with it. This is why arrows are like that. Maybe the word
-"delegation" is confusing and we should maybe use surrogate or some other word.}
-Because not every member of the community has time or knowledge to vote all the time, each member can delegate
-decisions to one or more other members of the community.
-When a member does not vote themselves, information how those delegates voted is combined into a vote of
-the non-voting member.
-If a delegate has not voted themselves either, their vote is computed through delegation as well, recursively.
-
-Each member $M$ can declare their delegates and ratios $r$ in which delegates' votes are combined. For example:
-
-\begin{itemize}
-\item Member $M_1$ decides not to delegate at all.
-\item Member $M_2$ delegates to $M_1$ and $M_4$ with ratios $0.75$ ($r_{2,1}$) and $0.25$ ($r_{2,4}$), respectively.
-\item Member $M_3$ delegates to $M_4$ with ratio $1.00$ ($r_{3,4}$).
-\item Member $M_4$ delegates to $M_2$, $M_3$, and $M_5$ with ratios $0.40$ ($r_{4,2}$), $0.40$ ($r_{4,3}$), and $0.20$ ($r_{4,5}$), respectively.
-\item Member $M_5$ decides not to delegate.
-\end{itemize}
-
-The sum of ratios of all delegations of a member has to be $1.0$.
-Members $M_1$ and $M_5$ decided not to participate in delegation and if they do not vote they are
-counted towards non-voting population.
-
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \SetGraphUnit{2.5}
-
-    \Vertex[Math,L=M_1]{M1}
-    \EA[Math,L=M_2](M1){M2}
-    \SO[Math,L=M_4](M2){M4}
-    \WE[Math,L=M_3](M4){M3}
-    \EA[Math,L=M_5](M4){M5}
-
-    \Edge[label=$0.40$](M2)(M4)
-    \Edge[label=$0.75$](M1)(M2)
-    \Edge[label=$0.25$](M4)(M2)
-    \Edge[label=$0.20$](M5)(M4)
-
-    \tikzset{EdgeStyle/.append style={bend right}}
-    \Edge[label=$1.00$](M4)(M3)
-    \Edge[label=$0.40$](M3)(M4)
-  \end{tikzpicture}
-  \caption{Example delegation network. Edges are directed towards a member who is delegating, representing how their
-  vote is computed from delegates' votes. E.g., $V_2 = 0.75 V_1 + 0.25 V_4$}~\label{fig:delegation-network}
-\end{figure}
-
-We can represent these delegations as a graph we call \emph{delegation network} as seen in
-Figure~\ref{fig:delegation-network}.
-We direct edges towards a member who is delegating, to represent how their vote is computed from delegates' votes.
-For example, member $M_2$ delegates the decision to members $M_1$ and $M_4$.
-$M_1$ and $M_4$ become $M_2$'s delegates.
-As a consequence, $M_2$'s vote $V_2$ is computed from $M_1$'s and $M_4$'s computed votes ($V_1$ and $V_4$)
-if $M_2$ does not vote themselves.
-
-With $v_i$ we denote a vote of a member $M_i$ if they voted, otherwise it is $0$.
-$V_i$ denotes their vote, if they have voted, or a computed vote based on their delegations, if they have not voted.
-
-A delegation network can be described as a system of equations.
-
-\todo[inline]{Max: Double check accuracy of this method, Improve explanation here and perhaps use nicer numbers so
-readers can follow along certain parts in their heads easily.}
-
-\begin{displaymath}
-\begin{array}{rcl}
-V_1 & = & \begin{dcases*}
- v_1 & if $M_1$ votes \\
- 0 & otherwise \\
-\end{dcases*} \\
-V_2 & = & \begin{dcases*}
- v_2 & if $M_2$ votes \\
- r_{2,1} V_1 + r_{2,4} V_4 = 0.75 V_1 + 0.25 V_4 & otherwise \\
-\end{dcases*} \\
-V_3 & = & \begin{dcases*}
- v_3 & if $M_3$ votes \\
- r_{3,4} V_4 = 1.00 V_4 & otherwise \\
-\end{dcases*} \\
-V_4 & = & \begin{dcases*}
- v_4 & if $M_4$ votes \\
- r_{4,2} V_2+ r_{4,3} V_3 + r_{4,5} V_5 = & otherwise \\
- = 0.40 V_2 + 0.40 V_3 + 0.20 V_5 & \\
-\end{dcases*} \\
-V_5 & = & \begin{dcases*}
- v_5 & if $M_5$ votes \\
- 0 & otherwise \\
-\end{dcases*} \\
-\end{array}
-\end{displaymath}
-
-When only some members vote we can use a delegation network to compute votes
-of all members.
-We convert a delegation network to a matrix $D$:
-
-\begin{displaymath}
-D = \left[d_{ij}\right] = \begin{dcases*}
- 1 & if $i = j$ \\
- -r_{i,j} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with ratio $r_{i,j}$} \\
- 0 & otherwise \\
-\end{dcases*}
-\end{displaymath}
-
-We solve the matrix equation for $\boldsymbol{V}$:
-
-\begin{displaymath}
-D \boldsymbol{V} = \boldsymbol{v},\quad \boldsymbol{v} = \left[v_{i}\right],\quad \boldsymbol{V} = \left[V_{i}\right]
-\end{displaymath}
-
-To continue with our example, let's say that $M_1$ and $M_3$ voted for a decision with $0.7$ and against with $0.4$,
-respectively.
-We are using score voting on $[-1, 1]$ range, so those votes are $v_1 = V_1 = 0.7$ and
-$v_3 = V_3 = -0.4$.
-
-\begin{displaymath}
-D = \left[ \begin{array}{ccccc}
-1 & 0 & 0 & 0 & 0 \\
--0.75 & 1 & -0.25 & 0 & 0 \\
-0 & 0 & 1 & 0 & 0 \\
-0 & -0.40 & -0.40 & 1 & -0.20 \\
-0 & 0 & 0 & 0 & 1 \\
-\end{array} \right]
-\end{displaymath}
-
-\begin{displaymath}
-\boldsymbol{v} = \left[ \begin{array}{c}
-0.7 \\
-0 \\
--0.4 \\
-0 \\
-0 \\
-\end{array} \right],\quad \boldsymbol{V} = \left[ \begin{array}{c}
-0.70 \\
-0.54 \\
--0.40 \\
-0.056 \\
-0 \\
-\end{array} \right]
-\end{displaymath}
-
-\todo[inline]{Max: FYI, computing who to discard from the population as ``non-voting'' is a separate computation in itself.
-You need to figure out what votes can be calculated partially as a consequence of delegation, and which cannot.
-Use indicator vectors denoting 1 if member has voted, else 0.
-\\
-Mitar: This is exactly what $\boldsymbol{w}$ is later on.}
-
-There are few details we have to look into.
-Despite delegation some votes or parts of votes can still happen to be discarded.
-Members who did not vote might have not declared their delegates.
-Other members could delegate to them as well, potentially propagating lack of a vote further.
-In our example, member $M_5$ has not voted so their vote is discarded.
-As a consequence $M_4$'s vote cannot be computed fully, which influences $M_2$'s vote as well.
-
-Another issue is a question of abstentions.
-In PeerMind members can abstain from voting which is different from simply not participating.
-Abstentions are counted towards quorum, but they do not contribute to the result.
-
-Moreover, from $\boldsymbol{V}$ alone we cannot see which elements represent vote $0$ and which
-represent that it was not possible to compute a vote for a member.
-
-To address all this, we define two vectors:
-
-\begin{displaymath}
-\boldsymbol{w} = \left[w_i\right] = \begin{dcases*}
- 1 & if $M_i$ voted \\
- 0 & otherwise \\
-\end{dcases*},\quad \boldsymbol{a} = \left[a_i\right] = \begin{dcases*}
- 1 & if $M_i$ abstained \\
- 0 & otherwise \\
-\end{dcases*}
-\end{displaymath}
-
-We solve $D \boldsymbol{W} = \boldsymbol{w}$ and $D \boldsymbol{A} = \boldsymbol{a}$ for
-$\boldsymbol{W} = \left[W_{i}\right]$ and $\boldsymbol{A} = \left[A_{i}\right]$, respectively.
-
-Non-zero elements of $\boldsymbol{W}$ correspond to members who voted or for which we can compute
-a vote.
-
-\todo[inline]{If there is a clique/cycle/closed subgroup in delegation, without any vote inside a clique/cycle/closed subgroup, the matrix becomes singular.
-We should find a way to detect such rows and set corresponding elements of $\boldsymbol{W}$ to $0$ (or the row
-to identity row before we try to compute $\boldsymbol{W}$).
-This is exactly what we want in fact, that those in such a clique/cycle/closed subgroup are seen as non-computable.
-Then the rest works.
-We should describe a problem, maybe with an example.}
-
-We define a normalized matrix:
-
-\begin{displaymath}
-D' = \left[d'_{ij}\right] = \begin{dcases*}
- 1 & if $i = j$ \\
- 0 & if $W_j = 0$ ($V_j$ cannot be computed) \\
- -r'_{i,j} & if $M_i$ did not vote \\
- 0 & otherwise \\
-\end{dcases*}
-\end{displaymath}
-
-Where:
-
-\begin{displaymath}
-r'_{i,j} = \frac{r_{i,j}}{R_i},\quad R_i = \sum_{W_k \ne 0} r_{i,k}
-\end{displaymath}
-
-We normalize ratios to only those delegates who voted or for which we can
-compute their votes recursively.
-So $r_{4,2}$ and $r_{4,3}$ are normalized from $0.40$ to $0.50$.
-
-To compute votes with normalized ratios for only those delegates who voted or for which we can compute votes recursively,
-we solve $D' \boldsymbol{V} = \boldsymbol{v}$ for $\boldsymbol{V}$:
-
-\begin{displaymath}
-\mathbf{V} = \left[ \begin{array}{c}
-0.70 \\
-0.54 \\
--0.40 \\
-0.07 \\
-0 \\
-\end{array} \right]
-\end{displaymath}
-
-If an element of $\boldsymbol{W}$ is equal to $0$, then we could not compute a vote for the corresponding
-member and we count it towards the non-voting population.
-We just have to make sure that they are not in fact abstentions.
-For elements $\boldsymbol{W} = 0$ we check their corresponding elements of $\boldsymbol{A}$ and if they are
-non-zero we count them as abstentions, otherwise as non-votes.
-
-\begin{itemize}
-\item Number of directly voting members: $N_v = \left| \{ k: M_k~\mathrm{voted} \} \right|$
-\item Number of voting members: $N_V = \left| \{ k: W_k \ne 0 \} \right|$
-\item Number of abstentions: $N_A = \left| \{ k:  W_k = 0 \land A_k \ne 0 \} \right|$
-\item Number of non-voting members: $\left| \{ k: W_k = 0 \land A_k = 0 \} \right|$
-\end{itemize}
-
-The result (average of scores) without delegation is:
-
-\begin{displaymath}
-\overline{v} = \frac{\sum_{M_k~\mathrm{voted}} v_k}{N_v} = 0.15
-\end{displaymath}
-
-With delegation:
-
-\begin{displaymath}
-\overline{V} = \frac{\sum_{W_k \ne 0} V_k}{N_V} = 0.23
-\end{displaymath}
-
-In our example, delegation slightly pushes decision in favor.
-
-We pass computed votes $V$ and number of abstentions $N_A$ into computation of dynamic quorum to determine
-a quorum and credibility, in the same manner as if members voted themselves.
-
-\todo[inline]{Do we want really to compute quorum using delegation, or should quorum be computed
-using direct votes only?}
-
-% In general, there are various ways to use this delegation scheme.
-To whom to delegate could be done for each vote separately, or decisions could be categorized into various
-categories and then delegations for that category would be used.
-But for simplicity, PeerMind currently supports only defining delegates separately from a decision making process,
-and those delegates are then in effect for all votes where a member does not vote themselves.
-A member is invited to define initial delegations at a registration time, and can change them later on at will,
-but independently from a particular decision being made.
-
-\subsection{Three step process}
-
-\subsection{Real-time feedback}
-
-% Real-time communication in general.
-
-\subsection{Integration with offline meetings}
-
-Decoupling the information presentation and voting processes from the debating process in the full decision-making
-pipeline (with the idea being that debating is best done in person, and presenting information/voting is best done online).
-
-\section{Evaluation}
-
-\subsection{Design}
-
-User study/evaluation design.
-To create a controlled evaluation of our solutions, to not depend on general usage of our system among communities,
-we can evaluate the design with the following user study:
-\begin{enumerate}
-\item we find a tight community of around 100 people willing to participate in the user study
-\item they should know each other well (to know how to delegate)
-\item we offer them to pay some Amazon credits or something for participation
-\item every member of the community should define delegations
-\item we ask them to split into two groups, group A and group B
-\item group A (target size is around 20% of the community) will meet together or online and discuss various topics through our system
-\item discussed topics some can be completely random, to get people to be familiar with they system, and also have some
-easy topics to decide on, like, what ``should be the color of the sky if we could change it'', and some topics which are relevant to the community
-\item those discussions would be just for evaluation, but community can use the results in any way they find useful
-\item after this discussions and decisions were done we ask everyone from group B to evaluate the results, we present them with two results (blinded, them not knowing which was made in which way)
-\begin{itemize}
-\item results which would be obtained using traditional methods (majority voting without any our improvements)
-\item our results using our system
-\end{itemize}
-\item we ask them to evaluate which one (or both, or none) of decisions they agree with
-\item prediction is that members who have not participated in the decision making process will
-find decisions made through our system more often agreeable
-\item hypothesis here is that what a decision making system is trying to achieve is to find a decision most of the
-whole community can agree with, or like it, even when not everyone participates and not all information is available
-on their positions on a topic
-\end{enumerate}
-
-One additional part of the experimental delegation study:
-
-Compute statistical quorum with and without delegations.
-This semester, do not actually allow delegation to be used for passing statistical quorum, but collect data so that
-we can see how much earlier delegation reached quorum before official statistical quorum.
-(Only count how much earlier at the point where delegation passed quorum and never went below thereafter, thus
-showing the point where the decision truly was clear.)
-
-Evaluation of statistical quorum (given that we don't officially decide by including delegation this semester):
-From email: ``Do an internal Cloyne comparison that is more mathematical, showing how many agenda items were decided
-using fewer actual votes than old 33\% quorum, as a result of statistical quorum.
-This could be measured by average \# votes needed to pass statistical quorum vs. \% of those items that get
-formally passed at council via consent agenda.
-(If it's 100\%, then you can say that we saved X\% of votes on
-average by using our online voting).
-Then you could also show that some items needed more than 33\% of votes, implying that a hard-line 33\% is also bad.''
-
-Subjective evaluation of entire platform: Council Surveys comparing past two semesters to this semester for other
-houses vs. Cloyne.
-``use the previous 2 semesters' worth of council surveys as control, and a council survey for
-Cloyne at the end of this semester as results.
-These results would speak to those specific subjective questions
-and would be one useful thing to look at.''
-
-\bibliographystyle{sigchi}
-\bibliography{refs}
-
-More:
-\begin{itemize}
-\item https://mason.gmu.edu/~rhanson/futarchy.html
-\item http://merkle.com/papers/DAOdemocracyDraft.pdf
-\end{itemize}
+% REFERENCES FORMAT
+% References must be the same font size as other body text.
+%\bibliographystyle{sigchi}
+%\bibliography{refs}
 
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2017-04-24 delegation.tex
Splitting papers.

@@
 
-Two of the most shocking democratic votes in recent history occurred in 2016:
-The referendum vote for Britain to leave the European Union (``Brexit'') and the election of Donald Trump as
-President of the United States.
-Both votes were decided by the slimmest of margins, but more significantly, both votes had poor turnout.
-We often hear that the Achilles heel of a democracy is low engagement, but this is a fallacy~---~in fact, our goal
-should be to decrease engagement.
-The ideal governance system is one where citizens do not have to be constantly involved, but the system is still
-acting as the citizens would like it to. Can we increase legitimacy within a democracy while also decreasing the
-attention required at the same time?
-
-In this paper we present our work on improving the collective decision-making through use of a web application
-to facilitate the process.
-In contrast with existing similar solutions, our web application integrates new ideas in how to improve this process:
-use of range voting, dynamic quorum, and vote delegation.
-The broad idea is that instead of just trying to move the voting process online~---~something many others have done
-previously~---~we in fact wish to add value by voting online.
-Specifically, we can compute results faster, more accurately,
-%and securely,
-we can use sophisticated statistics to
-estimate important quantities for the entire voting population, we can allow for members to entrust others who have
-more expertise on topics than they do themselves, and we can promote clarity and full understanding of the voting
-items for all.
+%Two of the most shocking democratic votes in recent history occurred in 2016:
+%The referendum vote for Britain to leave the European Union (``Brexit'') and the election of Donald Trump as
+%President of the United States.
+%Both votes were decided by the slimmest of margins, but more significantly, both votes had poor turnout.
+%We often hear that the Achilles heel of a democracy is low engagement, but this is a fallacy~---~in fact, our goal
+%should be to decrease engagement.
+%The ideal governance system is one where citizens do not have to be constantly involved, but the system is still
+%acting as the citizens would like it to. Can we increase legitimacy within a democracy while also decreasing the
+%attention required at the same time?
+%
+%In this paper we present our work on improving the collective decision-making through use of a web application
+%to facilitate the process.
+%In contrast with existing similar solutions, our web application integrates new ideas in how to improve this process:
+%use of range voting, dynamic quorum, and vote delegation.
+%The broad idea is that instead of just trying to move the voting process online~---~something many others have done
+%previously~---~we in fact wish to add value by voting online.
+%Specifically, we can compute results faster, more accurately,
+%%and securely,
+%we can use sophisticated statistics to
+%estimate important quantities for the entire voting population, we can allow for members to entrust others who have
+%more expertise on topics than they do themselves, and we can promote clarity and full understanding of the voting
+%items for all.
 
@@
 \section{Features}
-
-\subsection{Range Voting}
-
-Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
-each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
-A proposal passes if the average score of all votes on a position is positive.
-When multiple competing proposals are being voted on, the one with highest average score wins.
-
-Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
-positions in a highly informative and precise manner.
-We find it especially suitable for a web application because many web users are used to use it when ranking products
-and giving other ``five-star'' feedback.
-Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
-new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
-each proposal would need reassessing (assuming members use a consistent personal scale when voting).
-
-\subsection{Dynamic Quorum}
-
-%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal, motion -> proposal}
-
-Dynamic quorum uses statistical hypothesis testing to compute the minimum number of required votes needed
-to achieve a desired level of confidence before a community adopts an outcome as being truly representative of
-its members' positions.
-For example, in a community of 100 members, if 20 members vote in full support of a proposal, then we can conclude
-with very high confidence that, had all 100 members actually voted, a majority them would support the proposal,
-despite in actuality 80 members not having voted.
-On the other hand, if 80 members vote but 41 vote in support of the proposal and 39 vote in opposition,
-the community cannot yet conclude anything with high confidence.
-Hence, the suggested action would be for the community to gather more votes and continue discussion around the proposal.
-By utilizing a dynamic quorum, we can direct attention and participation of the community towards contentious proposals,
-while reducing attention towards proposals which already have overwhelming support.
-
-\subsubsection{Background}
-A population of voters (a community) needs to collectively decide which item (if any) to take action on out of a set of competing items.
-In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a
-winning item, while also demonstrating that the winning item is preferred over all other competing items.
-We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that
-the population does not take undesired actions, which in this case encompasses taking action on an item that is not
-supported by a majority of the population and/or taking action on an item that is not clearly preferred over all
-other competing alternatives.
-
-\subsubsection{Problem Statement}
-We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some
-fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority).
-In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater
-than the bias towards voting ``yes'' on all other competing items.
-
-\subsubsection{Assumptions}
-We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
-
-Without any votes, we assume that this bias may be any value between $0\%$ and $100\%$ with equal probability.
-
-\subsubsection{Voting}
-We allow for range voting where each member of the population may choose any real number in the range $[-1,+1]$ to
-represent their vote, with $-1$ representing opposition and $+1$ representing support.
-In the theory subsection below, we rescale votes to be in the range $[0,1]$ via the notation $\frac{votes + 1}{2}$.
-
-We allow voters to ``abstain'' from voting.
-We treat abstentions as indicating that a voter is removing themselves from the population of voters for an item.
-It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$
-(indicating indifference).
-
-\subsubsection{Bayesian Theory}
-
-The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a
-consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises
-naturally when considering samples from a finite population.\footnote{To
-properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial:
-\url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
-The tutorial is very similar to the problem we have at hand.}
-
-After extracting the sufficient statistics from raw votes~---~namely the average score and the number of
-abstentions~---~we can adjust the effective population size and corresponding threshold.
-In addition, we can calculate the effective number of ``yes'' and ``no'' votes.\footnote{Which one can think
-of as the number of Heads and Tails in the aforementioned tutorial.}
-
-Let
-\begin{description}
-\begin{itemize}
-\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing motions,
-\item[$N = $] population size,
-\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for motion $X_i$, excluding ``abstentions'',
-\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on motion $X_i$,
-\item[$\alpha_i = $] $\sum\limits_{v \in X_i} \left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
-\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
-\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
-\item[$p_i = $] true latent population bias towards voting ``yes'' on $X_i$.
-\end{itemize}
-\end{description}
-
-When we have a single motion open for voting, $C=\{X_1\}$, we simply wish to compute:
-\begin{align}\label{eq1}
-\Pr(\alpha^*_1 > T_1 \mid X_1),
-\end{align}
-where
-\begin{description}
-\begin{itemize}
-\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
-\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
-(e.g. $f=\frac{1}{2}$ for a simple majority).
-\end{itemize}
-\end{description}
-Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when
-no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
-\begin{align*}
-\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \Pr(\alpha^*_1 \leq T_1 \mid X_1),\\
-\Pr(\alpha^*_1 \leq T_1 \mid X_1) &= \int_0^1 \Pr(\alpha^*_1 \leq T_1 \mid p_1)\cdot \Pr(p_1 \mid X_1) \mathrm{d}p_1\\
-&= \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \\
-& \qquad\qquad\quad \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1,
-\end{align*}
-where
-\begin{description}
-\begin{itemize}
-\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that
-$X_i$ would remain below threshold.
-\end{itemize}
-\end{description}
-
-Hence, we have
-\begin{align}\label{eq2}
-& \Pr(\alpha^*_1 > T_1 \mid X_1) = \nonumber\\
-& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
-\end{align}
-
-In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to
-consider.
-Without loss of generality, assume that $X_1$ is the most popular motion within $C$.
-We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other
-$X_i  \in C, i \neq 1$:
-
-\begin{align}\label{eq3}
-\Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)
-\end{align}
-
-Using the assumption
-that votes on motions are independent across motions, which, in the common
-scenario where there is a negative covariance between votes on competing
-motions, will provide a lower bound on the probability, we can decompose \eqref{eq3} as follows:
-\begin{align*}
-& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\\
-=& \prod_{i=2}^n \Pr(\alpha^*_1 > \alpha^*_i \mid C)\\
-=& \prod_{i=2}^n \iint\limits_{p_1, p_i} \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \Pr(p_1 \mid X_1) \cdot \Pr(p_i \mid X_i) \mathrm{d}p_i \mathrm{d}p_1\\
-=& \prod_{i=2}^n \int_0^1 \int_0^1 \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \\
-& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1\\
-\end{align*}
-Deriving an expression for $\Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)$, we get:
-\begin{align*}
-& \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)\\
-=& 1 - \Pr(\alpha^*_1 \leq \alpha^*_i \mid p_1, p_i)\\
-=& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
-\end{align*}
-where
-\begin{description}
-\begin{itemize}
-\item[$v_{i,j} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
-\end{itemize}
-\end{description}
-
-Denoting
-\begin{align}\label{eq4}
-& \Phi_i(p_1,p_i) {\overset {\underset {\mathrm {def} }{}}{=}} \nonumber\\
-& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
-\end{align}
-we can simplify \eqref{eq3} to:
-\begin{align}\label{eq5}
-& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\nonumber\\
-&= \prod_{i=2}^n \int_0^1 \int_0^1 \Phi_i(p_1,p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \nonumber\\
-& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1
-\end{align}
-
-We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.
-In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/simple_majority}
-\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
-\label{simple_majority}
-\end{figure}
-%Data:
-%Curve 1: {{0,0.000299648},{1,0.00213083},{2,0.00817989},{3,0.0224777},{4,0.049512},{5,0.0928927},{6,0.154136},{7,0.232044},{8,0.322859},{9,0.421057},{10,0.520467},{11,0.615358},{12,0.701241},{13,0.775267},{14,0.836245},{15,0.884385},{16,0.920892},{17,0.947532},{18,0.966265},{19,0.978973},{20,0.987296},{21,0.992561},{22,0.99578},{23,0.997682},{24,0.998767},{25,0.999366}}
-%Curve 2: {x, 0.99} for all x
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/super_majority}
-\caption{Super Majority: $\alpha$ vs. $\Pr(\alpha > \frac{2}{3}N)$ for $N=100$, $\beta = 10$.}
-\label{super_majority}
-\end{figure}
-%Data:
-%Curve 1: {{10,0.0395173},{11,0.0591278},{12,0.0846131},{13,0.116407},{14,0.15463},{15,0.199057},{16,0.24911},{17,0.303899},{18,0.362273},{19,0.422909},{20,0.484399},{21,0.545345},{22,0.604445},{23,0.660559},{24,0.712764},{25,0.760382},{26,0.802985},{27,0.840385},{28,0.872612},{29,0.899875},{30,0.922519},{31,0.940987},{32,0.955777},{33,0.967407},{34,0.976384},{35,0.983186},{36,0.988243},{37,0.99193},{38,0.994567},{39,0.996414},{40,0.997682},{41,0.998533},{42,0.999093},{43,0.999452},{44,0.999677},{45,0.999815},{46,0.999896},{47,0.999944},{48,0.99997},{49,0.999985},{50,0.999993}}
-%Curve 2: {x, 0.99} for all x
-
-%\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
-%we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
-%also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
 

### 2017-04-24 statistical-quorum.tex
Splitting papers.

@@
 
+\section{Features}
+
+\section{Range Voting}
+
+Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
+each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
+A proposal passes if the average score of all votes on a position is positive.
+When multiple competing proposals are being voted on, the one with highest average score wins.
+
+Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
+positions in a highly informative and precise manner.
+We find it especially suitable for a web application because many web users are used to use it when ranking products
+and giving other ``five-star'' feedback.
+Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
+new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
+each proposal would need reassessing (assuming members use a consistent personal scale when voting).
+
+\subsection{Dynamic Quorum}
+
+%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal, motion -> proposal}
+
+Dynamic quorum uses statistical hypothesis testing to compute the minimum number of required votes needed
+to achieve a desired level of confidence before a community adopts an outcome as being truly representative of
+its members' positions.
+For example, in a community of 100 members, if 20 members vote in full support of a proposal, then we can conclude
+with very high confidence that, had all 100 members actually voted, a majority them would support the proposal,
+despite in actuality 80 members not having voted.
+On the other hand, if 80 members vote but 41 vote in support of the proposal and 39 vote in opposition,
+the community cannot yet conclude anything with high confidence.
+Hence, the suggested action would be for the community to gather more votes and continue discussion around the proposal.
+By utilizing a dynamic quorum, we can direct attention and participation of the community towards contentious proposals,
+while reducing attention towards proposals which already have overwhelming support.
+
+\subsubsection{Background}
+A population of voters (a community) needs to collectively decide which item (if any) to take action on out of a set of competing items.
+In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a
+winning item, while also demonstrating that the winning item is preferred over all other competing items.
+We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that
+the population does not take undesired actions, which in this case encompasses taking action on an item that is not
+supported by a majority of the population and/or taking action on an item that is not clearly preferred over all
+other competing alternatives.
+
+\subsubsection{Problem Statement}
+We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some
+fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority).
+In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater
+than the bias towards voting ``yes'' on all other competing items.
+
+\subsubsection{Assumptions}
+We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
+
+Without any votes, we assume that this bias may be any value between $0\%$ and $100\%$ with equal probability.
+
+\subsubsection{Voting}
+We allow for range voting where each member of the population may choose any real number in the range $[-1,+1]$ to
+represent their vote, with $-1$ representing opposition and $+1$ representing support.
+In the theory subsection below, we rescale votes to be in the range $[0,1]$ via the notation $\frac{votes + 1}{2}$.
+
+We allow voters to ``abstain'' from voting.
+We treat abstentions as indicating that a voter is removing themselves from the population of voters for an item.
+It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$
+(indicating indifference).
+
+\subsubsection{Bayesian Theory}
+
+The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a
+consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises
+naturally when considering samples from a finite population.\footnote{To
+properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial:
+\url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
+The tutorial is very similar to the problem we have at hand.}
+
+After extracting the sufficient statistics from raw votes~---~namely the average score and the number of
+abstentions~---~we can adjust the effective population size and corresponding threshold.
+In addition, we can calculate the effective number of ``yes'' and ``no'' votes.\footnote{Which one can think
+of as the number of Heads and Tails in the aforementioned tutorial.}
+
+Let
+\begin{description}
+\begin{itemize}
+\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing motions,
+\item[$N = $] population size,
+\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for motion $X_i$, excluding ``abstentions'',
+\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on motion $X_i$,
+\item[$\alpha_i = $] $\sum\limits_{v \in X_i} \left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
+\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
+\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
+\item[$p_i = $] true latent population bias towards voting ``yes'' on $X_i$.
+\end{itemize}
+\end{description}
+
+When we have a single motion open for voting, $C=\{X_1\}$, we simply wish to compute:
+\begin{align}\label{eq1}
+\Pr(\alpha^*_1 > T_1 \mid X_1),
+\end{align}
+where
+\begin{description}
+\begin{itemize}
+\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
+\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
+(e.g. $f=\frac{1}{2}$ for a simple majority).
+\end{itemize}
+\end{description}
+Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when
+no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
+\begin{align*}
+\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \Pr(\alpha^*_1 \leq T_1 \mid X_1),\\
+\Pr(\alpha^*_1 \leq T_1 \mid X_1) &= \int_0^1 \Pr(\alpha^*_1 \leq T_1 \mid p_1)\cdot \Pr(p_1 \mid X_1) \mathrm{d}p_1\\
+&= \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \\
+& \qquad\qquad\quad \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1,
+\end{align*}
+where
+\begin{description}
+\begin{itemize}
+\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that
+$X_i$ would remain below threshold.
+\end{itemize}
+\end{description}
+
+Hence, we have
+\begin{align}\label{eq2}
+& \Pr(\alpha^*_1 > T_1 \mid X_1) = \nonumber\\
+& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
+\end{align}
+
+In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to
+consider.
+Without loss of generality, assume that $X_1$ is the most popular motion within $C$.
+We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other
+$X_i  \in C, i \neq 1$:
+
+\begin{align}\label{eq3}
+\Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)
+\end{align}
+
+Using the assumption
+that votes on motions are independent across motions, which, in the common
+scenario where there is a negative covariance between votes on competing
+motions, will provide a lower bound on the probability, we can decompose \eqref{eq3} as follows:
+\begin{align*}
+& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\\
+=& \prod_{i=2}^n \Pr(\alpha^*_1 > \alpha^*_i \mid C)\\
+=& \prod_{i=2}^n \iint\limits_{p_1, p_i} \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \Pr(p_1 \mid X_1) \cdot \Pr(p_i \mid X_i) \mathrm{d}p_i \mathrm{d}p_1\\
+=& \prod_{i=2}^n \int_0^1 \int_0^1 \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \\
+& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1\\
+\end{align*}
+Deriving an expression for $\Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)$, we get:
+\begin{align*}
+& \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)\\
+=& 1 - \Pr(\alpha^*_1 \leq \alpha^*_i \mid p_1, p_i)\\
+=& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
+\end{align*}
+where
+\begin{description}
+\begin{itemize}
+\item[$v_{i,j} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
+\end{itemize}
+\end{description}
+
+Denoting
+\begin{align}\label{eq4}
+& \Phi_i(p_1,p_i) {\overset {\underset {\mathrm {def} }{}}{=}} \nonumber\\
+& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
+\end{align}
+we can simplify \eqref{eq3} to:
+\begin{align}\label{eq5}
+& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\nonumber\\
+&= \prod_{i=2}^n \int_0^1 \int_0^1 \Phi_i(p_1,p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \nonumber\\
+& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1
+\end{align}
+
+We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.
+In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/simple_majority}
+\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
+\label{simple_majority}
+\end{figure}
+%Data:
+%Curve 1: {{0,0.000299648},{1,0.00213083},{2,0.00817989},{3,0.0224777},{4,0.049512},{5,0.0928927},{6,0.154136},{7,0.232044},{8,0.322859},{9,0.421057},{10,0.520467},{11,0.615358},{12,0.701241},{13,0.775267},{14,0.836245},{15,0.884385},{16,0.920892},{17,0.947532},{18,0.966265},{19,0.978973},{20,0.987296},{21,0.992561},{22,0.99578},{23,0.997682},{24,0.998767},{25,0.999366}}
+%Curve 2: {x, 0.99} for all x
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/super_majority}
+\caption{Super Majority: $\alpha$ vs. $\Pr(\alpha > \frac{2}{3}N)$ for $N=100$, $\beta = 10$.}
+\label{super_majority}
+\end{figure}
+%Data:
+%Curve 1: {{10,0.0395173},{11,0.0591278},{12,0.0846131},{13,0.116407},{14,0.15463},{15,0.199057},{16,0.24911},{17,0.303899},{18,0.362273},{19,0.422909},{20,0.484399},{21,0.545345},{22,0.604445},{23,0.660559},{24,0.712764},{25,0.760382},{26,0.802985},{27,0.840385},{28,0.872612},{29,0.899875},{30,0.922519},{31,0.940987},{32,0.955777},{33,0.967407},{34,0.976384},{35,0.983186},{36,0.988243},{37,0.99193},{38,0.994567},{39,0.996414},{40,0.997682},{41,0.998533},{42,0.999093},{43,0.999452},{44,0.999677},{45,0.999815},{46,0.999896},{47,0.999944},{48,0.99997},{49,0.999985},{50,0.999993}}
+%Curve 2: {x, 0.99} for all x
+
+%\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
+%we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
+%also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
+
 % REFERENCES FORMAT

### 2017-04-25 statistical-quorum.tex
New title.
Better title.
Initial introduction.

@@
-\documentclass{sigchi}
+\documentclass[chi_draft]{sigchi}
 
@@
 % so you remain anonymous.
-\def\plaintitle{PeerMind: Statistical Quorum}
+\def\plaintitle{Using statistical quorum to improve legitimacy of decisions and utilization of collective attention in democratic communities}
 %\def\plainauthor{First Author, Second Author, Third Author,
@@
 The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
-While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?
-We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic.
-We developed an online system centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 28 weeks.
-We found that the community was able to use the online system to simultaneously improve decision-making legitimacy and efficiency.
+While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
+different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
+maximize productivity?
+We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
+attention, in turn making the entire decision-making process dynamic.
+We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
+councils over the course of 28 weeks.  We found that the community was able to use the online system to simultaneously
+improve decision-making legitimacy and efficiency.
 \end{abstract}
@@
 \section{Introduction}
+
+Many democratic communities use voting to make collective decisions.
+\todo{Should we support this statement with a reference?}
+In its basic form voting consist of members of a community casting a vote \emph{in favor} or \emph{against} a decision and if
+majority ($> 50\%$) of votes are in favor of the decision, the decision \emph{passes} and is adopted as a decision
+of the whole community.
+
+There have been many improvements proposed to this basic form and various analysis made.
+\todo{Add references.}
+In this paper we focus on the question of \emph{quorum} and how high the quorum should be.
+Quorum is the minimum number of members participating in voting for voting to be seen as representative of the whole
+community, i.e., legitimate.
+If quorum is too low, decisions adopted might not represent the community.
+If it is too high, the community might have issues achieving the quorum.
+Not enough members might decide to participate in voting, or members might on purpose boycott voting to
+prevent decision to be adopted, even if those opposing it are in minority.
+Communities often decide quorum arbitrary\footnote{Robert's Rules of Order states that the quorum
+``should approximate the largest number that can be depended on to attend any meeting except in very bad weather
+or other extremely unfavorable conditions.''~\cite{roberts}}
+and try to increase participation of members in communities' democratic processes.
+\todo{Add references.}
+Moreover, many democratic communities have issues achieving quorum which hinders their operation.
+\todo{Add references.}
 
@@
 % References must be the same font size as other body text.
-%\bibliographystyle{sigchi}
-%\bibliography{refs}
+\bibliographystyle{sigchi}
+\bibliography{refs}
 

### 2017-04-25 statistical-quorum.tex
More on introduction.

@@
 or other extremely unfavorable conditions.''~\cite{roberts}}
-and try to increase participation of members in communities' democratic processes.
-\todo{Add references.}
+and try to increase participation of members in communities' democratic processes through various means
+to address the issue of too low turnout.
+\todo{Add references. Mention forced participation in some countries.}
 Moreover, many democratic communities have issues achieving quorum which hinders their operation.
 \todo{Add references.}
+
+\todo[inline]{Structure of the rest of the paper.}
+
+\section{Related Work}
 

### 2017-04-25 statistical-quorum.tex
Added table for example.

@@
 \todo{Add references.}
+
+Even if voting turnout has reached legally required quorum, it is still a question how to evaluate if
+the turnout is high enough to find results legitimate.
+As an example we can look at two recent votes, Brexit, a 2016 referendum for United Kingdom's withdrawal
+from the European Union, and popular vote in 2016 USA presidential elections, as seen in Table~\ref{tab:example}.
+
+\begin{table*}
+  \centering
+  \begin{tabular}{l r r r r r r r}
+    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Statistical quorum}} \\
+    \midrule
+    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & \\
+    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 250,055,734 & $51.5\%$ & \\
+  \end{tabular}
+  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union,
+  and 2016 USA presidential elections, popular vote. Simplified to only two options and without invalid votes.}\label{tab:example}
+\end{table*}
+
+%for example: trump and brexit
+%
+%we define measure of creditibility, statistical quorm, to evaluate this
+%
+%the easiest is to claim that results are legitimate becaue people had an option to participate, and they decided not to
+%
+%but what if they didn't have time to participate? or determined that they are not informed enough to participate?
+%
+%instead of trying to increase participation, we can look at participation as it is, and try to utilize it better
+%
+%we can use the same statistical quorum to direct collective attention to decisions which require more participation
+%and away from those which do not
+%
+%instead of requiring same arbitirary high static quorum for all decisions, we can require same creditibility, which
+%depends on votes made define if enough members participated
+%
+%Example
 

### 2017-04-26 statistical-quorum.tex
reworded last sentence of abstract
Adding references and results.

@@
 
-Even if voting turnout has reached legally required quorum, it is still a question how to evaluate if
-the turnout is high enough to find results legitimate.
+Even if voting turnout has reached legally required quorum, or when one is not defined,
+it is still a question how to evaluate if the turnout is high enough to find results legitimate.
 As an example we can look at two recent votes, Brexit, a 2016 referendum for United Kingdom's withdrawal
 from the European Union, and popular vote in 2016 USA presidential elections, as seen in Table~\ref{tab:example}.
+Ignoring particularities of USA presidential elections process and using a popular vote as an example,
+how can we reason about legitimacy of results?
 
@@
     \midrule
-    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & \\
-    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 250,055,734 & $51.5\%$ & \\
+    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
+    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
   \end{tabular}
-  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union,
-  and 2016 USA presidential elections, popular vote. Simplified to only two options and without invalid votes.}\label{tab:example}
+  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
+  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without invalid votes.}\label{tab:example}
 \end{table*}
+
+%traditionally, we just claim that everyone had an option
+
+%we can use statistical quroum, defined in section X
 
@@
 
+% threshold and quorum at https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf
+% how does some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate them
+
 \section{Features}
@@
 
-\subsection{Dynamic Quorum}
+\section{Statistical Quorum}
+\label{sec:statistical-quorum}
 

### 2017-04-26 statistical-quorum.tex
Finished first pass of introduction.

@@
 \begin{abstract}
-The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
-While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
-different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
-maximize productivity?
-We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
-attention, in turn making the entire decision-making process dynamic.
-We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
-councils over the course of 28 weeks.  We found that the community was able to use the online system to simultaneously
-improve decision-making legitimacy and efficiency.
+%The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
+%While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
+%different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
+%maximize productivity?
+%We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
+%attention, in turn making the entire decision-making process dynamic.
+%We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
+%councils over the course of 28 weeks.  We found that the community was able to use the online system to simultaneously
+%improve the legitimacy and efficiency of their decision-making process.
 \end{abstract}
@@
   \begin{tabular}{l r r r r r r r}
-    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Statistical quorum}} \\
+    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Decision credibility}} \\
     \midrule
@@
   \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
-  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without invalid votes.}\label{tab:example}
+  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
+  invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:statistical-quorum}. We see that
+  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
+  with more votes any option could end up winning.}\label{tab:example}
 \end{table*}
 
-%traditionally, we just claim that everyone had an option
+To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
+hypothesis testing to evaluate the credibility of the result.
+It uses distribution of votes made to compute the probability of decision passing if everyone voted,
+and not just those who voted.
+Intuitively, if in a community of 100 members 40 members vote for a decision to pass and nobody against,
+we can see that if everyone voted, the probability of majority of population voting for a decision is high.
+On the other hand, if 41 members for in favor and 40 against, we cannot really say much about what would a decision
+be if everyone voted.
+Only two more votes against can completely change the outcome.
+We describe decision credibility in more detail in Section~\ref{sec:statistical-quorum}.
 
-%we can use statistical quroum, defined in section X
+Based on decision credibility we can define a \emph{statistical quorum} as a threshold which a decision's credibility
+has to pass for a decision to be seen as legitimate.
+For example, if we adopt $90\%$ as a reasonable requirement for statistical quorum and we compute decision
+credibility for votes in Table~\ref{tab:example} we see that in both cases we do not reach statistical quorum
+which puts legitimacy of those decisions into a question.
+We discuss a way for a community to determine a reasonable statistical quorum in Section~\ref{sec:discussion}.
 
-%for example: trump and brexit
-%
-%we define measure of creditibility, statistical quorm, to evaluate this
-%
-%the easiest is to claim that results are legitimate becaue people had an option to participate, and they decided not to
-%
-%but what if they didn't have time to participate? or determined that they are not informed enough to participate?
-%
-%instead of trying to increase participation, we can look at participation as it is, and try to utilize it better
-%
-%we can use the same statistical quorum to direct collective attention to decisions which require more participation
-%and away from those which do not
-%
-%instead of requiring same arbitirary high static quorum for all decisions, we can require same creditibility, which
-%depends on votes made define if enough members participated
-%
-%Example
+Traditionally it is argued that legitimacy comes from the fact that members had a chance to participate
+and they decided not to, giving legitimacy to make a decision to those who did participate.
+We argue that this is a problematic view because there are many restrictions on why members could not
+participate and vote.
+For example, they might not had time to participate because of other commitments, or they might determined
+that they are not informed enough to participate.
+Instead of assuming that everyone could participate if they would want, or trying to increase participation,
+we can see amount of participation members are able to do as a limited resource, a \emph{collective attention}
+a community has available to utilize.
+
+We can compute decision credibility during ongoing voting process to give feedback to the community
+on how contentious a proposed decision is how much more votes are necessary for a statistical quorum
+to be reached.
+This allows the community to focus its collective attention towards decisions which require more votes
+to determine a legitimate result.
+Instead of trying to increase participation on all democratic decisions being made, a community can
+do that selectively through an automatic feedback loop.
+
+To evaluate the use of statistical quorum in decision making of democratic communities we have developed
+a web application PeerMind which integrates online discussion with online voting.
+The application exposes computed decision credibility to users in real-time and through that provides
+the feedback loop needed for improving utilization of collective attention.
+We present the app in Section~\ref{sec:peermind}.
 
@@
 
-\section{Features}
-
-\section{Range Voting}
+\section{PeerMind Web Application}
+\label{sec:peermind}
 
@@
 
-\section{Statistical Quorum}
+\section{Decision Credibility and Statistical Quorum}
 \label{sec:statistical-quorum}
@@
 
+\section{Evaluation}
+\label{sec:evaluation}
+
+\section{Discussion}
+\label{sec:discussion}
+
+% describe why 0.90% is a reasnable statistical quorum and how to determine one for your community
+
 % REFERENCES FORMAT

### 2017-04-26 statistical-quorum.tex
Preparing appendix.

@@
 Only two more votes against can completely change the outcome.
-We describe decision credibility in more detail in Section~\ref{sec:statistical-quorum}.
+We describe decision credibility in more detail in Section~\ref{sec:statistical-quorum}\todo{Hm, SIGCGI template seems to not display section numbers. So how useful is to reference a section number?}.
 
@@
 participate and vote.
-For example, they might not had time to participate because of other commitments, or they might determined
+For example, they might not had time to participate because of other commitments, or motivation, or they might determined
 that they are not informed enough to participate.
+Moreover, members are from different reasons in their communities and potentially contribute in other ways
+than by participating in its democratic processes.
 Instead of assuming that everyone could participate if they would want, or trying to increase participation,
@@
 
+\section{Conclusion}
+
+%\section{Acknowledgments}
+
 % REFERENCES FORMAT
@@
 \bibliography{refs}
+
+\appendix
+\section{Derivation of Decision Credibility}
+\label{sec:derivation}
 

### 2017-04-26 statistical-quorum.tex
Moving content to appendix.

@@
 
+\section{Evaluation}
+\label{sec:evaluation}
+
+\section{Discussion}
+\label{sec:discussion}
+
+% describe why 0.90% is a reasnable statistical quorum and how to determine one for your community
+
+\section{Conclusion}
+
+%\section{Acknowledgments}
+
+% REFERENCES FORMAT
+% References must be the same font size as other body text.
+\bibliographystyle{sigchi}
+\bibliography{refs}
+
+\appendix
+\section{Derivation of Decision Credibility}
+\label{sec:derivation}
+
 %\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal, motion -> proposal}
@@
 
-\subsubsection{Background}
+\subsection{Background}
 A population of voters (a community) needs to collectively decide which item (if any) to take action on out of a set of competing items.
@@
 
-\subsubsection{Problem Statement}
+\subsection{Problem Statement}
 We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some
@@
 
-\subsubsection{Assumptions}
+\subsection{Assumptions}
 We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
@@
 
-\subsubsection{Voting}
+\subsection{Voting}
 We allow for range voting where each member of the population may choose any real number in the range $[-1,+1]$ to
@@
 
-\subsubsection{Bayesian Theory}
+\subsection{Bayesian Theory}
 
@@
 
-\section{Evaluation}
-\label{sec:evaluation}
-
-\section{Discussion}
-\label{sec:discussion}
-
-% describe why 0.90% is a reasnable statistical quorum and how to determine one for your community
-
-\section{Conclusion}
-
-%\section{Acknowledgments}
-
-% REFERENCES FORMAT
-% References must be the same font size as other body text.
-\bibliographystyle{sigchi}
-\bibliography{refs}
-
-\appendix
-\section{Derivation of Decision Credibility}
-\label{sec:derivation}
-
 \end{document}

### 2017-04-26 statistical-quorum.tex
Adding a TODO.
Small improvements to introduction.

@@
 we can see amount of participation members are able to do as a limited resource, a \emph{collective attention}
-a community has available to utilize.
+a community has available to utilize for its decision making.
 
@@
 Instead of trying to increase participation on all democratic decisions being made, a community can
-do that selectively through an automatic feedback loop.
+do that selectively through an automatic feedback loop, trying to reach statistical quorum for their decisions.
 
-To evaluate the use of statistical quorum in decision making of democratic communities we have developed
-a web application PeerMind which integrates online discussion with online voting.
+We built an open source web application PeerMind integrating statistical quorum with online voting.
 The application exposes computed decision credibility to users in real-time and through that provides
 the feedback loop needed for improving utilization of collective attention.
-We present the app in Section~\ref{sec:peermind}.
+We present the app in Section~\ref{sec:peermind} and use it for evaluating statistical quorum in
+Section~\ref{sec:evaluation}.
 
-\todo[inline]{Structure of the rest of the paper.}
+\todo[inline]{Should we explicitly list our contributions again?}
+
+\todo[inline]{Repeat the structure of the rest of the paper?}
 
@@
 % how does some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate them
+
+% repeat references from "There have been many improvements proposed to this basic form and various analysis made."?
+
+% voting auditing
+
+% delegation as way to increase legitimacy
 
@@
 
+% theoretical: math, diagrams
+
+% imperial: running it for x weeks in student community which has traditionally weekly meeting
+% issues with quorum
+% participation of everyone
+% they used online app for x weeks
+% we compare their old minutes with performance of new system
+
+% interviews and surveys
+
 \section{Discussion}
@@
 % describe why 0.90% is a reasnable statistical quorum and how to determine one for your community
+
+\section{Future Work}
 

### 2017-04-26 statistical-quorum.tex
Exposing related work TODOs.

@@
 
-% threshold and quorum at https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf
-% how does some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate them
+\todo[inline]{Repeat references from "There have been many improvements proposed to this basic form and various analysis made."?}
 
-% repeat references from "There have been many improvements proposed to this basic form and various analysis made."?
+\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
+How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
 
-% voting auditing
+\todo[inline]{Voting auditing work.}
 
-% delegation as way to increase legitimacy
+\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
 

### 2017-04-26 statistical-quorum.tex
Small changes.

@@
   and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
-  invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:statistical-quorum}. We see that
+  invalid votes. The last column shows computed decision credibility as described in Section \nameref{sec:statistical-quorum}. We see that
   credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
@@
 To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
-hypothesis testing to evaluate the credibility of the result.
-It uses distribution of votes made to compute the probability of decision passing if everyone voted,
+hypothesis testing to evaluate the credibility of the voting result.
+It uses distribution of votes made to evaluate the probability of decision passing if everyone actually voted,
 and not just those who voted.
@@
 Only two more votes against can completely change the outcome.
-We describe decision credibility in more detail in Section~\ref{sec:statistical-quorum}\todo{Hm, SIGCGI template seems to not display section numbers. So how useful is to reference a section number?}.
+Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
+We describe decision credibility in more detail in Section \nameref{sec:statistical-quorum}.
 
@@
 which puts legitimacy of those decisions into a question.
-We discuss a way for a community to determine a reasonable statistical quorum in Section~\ref{sec:discussion}.
+We discuss a way for a community to determine a reasonable statistical quorum in Section \nameref{sec:discussion}.
 
@@
 This allows the community to focus its collective attention towards decisions which require more votes
-to determine a legitimate result.
+to determine a legitimate result, while reducing attention towards proposals which already have overwhelming support.
 Instead of trying to increase participation on all democratic decisions being made, a community can
@@
 the feedback loop needed for improving utilization of collective attention.
-We present the app in Section~\ref{sec:peermind} and use it for evaluating statistical quorum in
-Section~\ref{sec:evaluation}.
+We present the app in Section \nameref{sec:peermind} and use it for evaluating statistical quorum in
+Section \nameref{sec:evaluation}.
 
@@
 
-\todo[inline]{Repeat the structure of the rest of the paper?}
+\todo[inline]{Repeat the structure of the rest of the paper? Is there a better way to reference sections than
+referencing it by name?}
 
@@
 %\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal, motion -> proposal}
-
-Dynamic quorum uses statistical hypothesis testing to compute the minimum number of required votes needed
-to achieve a desired level of confidence before a community adopts an outcome as being truly representative of
-its members' positions.
-For example, in a community of 100 members, if 20 members vote in full support of a proposal, then we can conclude
-with very high confidence that, had all 100 members actually voted, a majority them would support the proposal,
-despite in actuality 80 members not having voted.
-On the other hand, if 80 members vote but 41 vote in support of the proposal and 39 vote in opposition,
-the community cannot yet conclude anything with high confidence.
-Hence, the suggested action would be for the community to gather more votes and continue discussion around the proposal.
-By utilizing a dynamic quorum, we can direct attention and participation of the community towards contentious proposals,
-while reducing attention towards proposals which already have overwhelming support.
 

### 2017-04-26 statistical-quorum.tex
TODOs.

@@
 
+PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
+anonymized dataset of interactions with the app is available at \url{http://redacted.for.blind.review}.
+
 \section{Discussion}
@@
 \section{Future Work}
+
+\todo[inline]{More work on security evaluation of potential attacks when a community is using statistical quorum to force a decision.
+Is there a game theoretical approach to it?}
+
+\todo[inline]{How to surface information about credibility to users better than just a number? Visualizations? How can we help them
+understand what exactly this is and how the math works?}
 

### 2017-04-26 statistical-quorum.tex
Adding TODOs.

@@
 %We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
-%councils over the course of 28 weeks.  We found that the community was able to use the online system to simultaneously
+%councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
 %improve the legitimacy and efficiency of their decision-making process.
@@
 the feedback loop needed for improving utilization of collective attention.
-We present the app in Section \nameref{sec:peermind} and use it for evaluating statistical quorum in
+We present the application in Section \nameref{sec:peermind} and use it for evaluating statistical quorum in
 Section \nameref{sec:evaluation}.
@@
 
-Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
-each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
-A proposal passes if the average score of all votes on a position is positive.
-When multiple competing proposals are being voted on, the one with highest average score wins.
+%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
+%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
+%A proposal passes if the average score of all votes on a position is positive.
+%When multiple competing proposals are being voted on, the one with highest average score wins.
+%
+%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
+%positions in a highly informative and precise manner.
+%We find it especially suitable for a web application because many web users are used to use it when ranking products
+%and giving other ``five-star'' feedback.
+%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
+%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
+%each proposal would need reassessing (assuming members use a consistent personal scale when voting).
 
-Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
-positions in a highly informative and precise manner.
-We find it especially suitable for a web application because many web users are used to use it when ranking products
-and giving other ``five-star'' feedback.
-Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
-new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
-each proposal would need reassessing (assuming members use a consistent personal scale when voting).
+\todo[inline]{The app has features both for use in online and offline setting. This allows one to compute
+statistical quorum also for an in-person meeting because math is now more complicated than just counting hands.
+We focus on online experience in this paper.}
+
+\todo[inline]{Security is out of the scope of this work. We assume here that users and their identities are known and are managed
+by a community's admin. There are many communities where their decision making is public with public minutes and this is who
+the app is currently targeting.}
 
@@
 
+\todo[inline]{Definition and formulas for three types of decision credibility: single option, multiple competing options,
+and dynamic credibility which is increasing through time. In this order.}
+
 \section{Evaluation}
 \label{sec:evaluation}
+
+We evaluate statistical quorum in two ways. First, we evaluate it theoretically and explore its mathematical properties.
+Then we evaluate it through real-world use of the PeerMind application by a 140 members community.
+
+\subsection{Properties of Statistical Quorum}
+
+\todo[inline]{Theoretical evaluation, diagrams. Get readers to understand some of its properties visually.}
+
+\subsection{Use in Existing Democratic Community}
 
@@
 
+\todo[inline]{Make sure that we are supporting both of our claims well: this statistical quorum is a good metric and
+collective attention can be utilized better.}
+
 PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
-anonymized dataset of interactions with the app is available at \url{http://redacted.for.blind.review}.
+anonymized dataset of interactions with the app, including votes, is available at \url{http://redacted.for.blind.review}.
 
@@
 
-% describe why 0.90% is a reasnable statistical quorum and how to determine one for your community
+\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.}
+
+\todo[inline]{Describe interesting example how we had a bug which were showing a very low statistical quorum for a failing
+vote and so the whole community very engaged for a whole month to decide and resolve it. So how they trusted the
+app and the metric and how it was really able to direct the attention of the community. It seems it works even if it is
+buggy or not perfect, because it has worked so well in the past that members got used to the fact that when it was saying
+that there is a consensus or not on an issue that that was also their own read of the community. So for the purpose of
+directing attention the metric does not have to mathematically perfect.}
+
+\todo[inline]{Security. How to prevent a group of people to vote and by trick reaching statistical quorum.
+We use dynamic statistical quorum to address this, but one other approach is also for voting to stay open and
+decisions can then be in the future revoked by people adding more votes, or changing them. But for some decisions
+this might not really be possible, because decisions might not be reversible. You can revoke a law, but if you built
+a building as a consequence of a decision?}
 
@@
 
-%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal, motion -> proposal}
+\todo[inline]{Derivation in order of dynamic credibility which is increasing through time, multiple competing options,
+single option, because this is how they are easier to derive as each following is a special case of the previous one.}
+
+%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> decision, motion -> decision}
 

### 2017-04-26 statistical-quorum.tex
Style section names.

@@
 
+\newcommand\secref[1]{\textit{\nameref{#1}}}
+
 % End of preamble. Here it comes the document.
@@
   and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
-  invalid votes. The last column shows computed decision credibility as described in Section \nameref{sec:statistical-quorum}. We see that
+  invalid votes. The last column shows computed decision credibility as described in Section \secref{sec:statistical-quorum}. We see that
   credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
@@
 Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
-We describe decision credibility in more detail in Section \nameref{sec:statistical-quorum}.
+We describe decision credibility in more detail in Section \secref{sec:statistical-quorum}.
 
@@
 which puts legitimacy of those decisions into a question.
-We discuss a way for a community to determine a reasonable statistical quorum in Section \nameref{sec:discussion}.
+We discuss a way for a community to determine a reasonable statistical quorum in Section \secref{sec:discussion}.
 
@@
 the feedback loop needed for improving utilization of collective attention.
-We present the application in Section \nameref{sec:peermind} and use it for evaluating statistical quorum in
-Section \nameref{sec:evaluation}.
+We present the application in Section \secref{sec:peermind} and use it for evaluating statistical quorum in
+Section \secref{sec:evaluation}.
 
@@
 \subsection{Use in Existing Democratic Community}
-
-% theoretical: math, diagrams
 

### 2017-04-27 statistical-quorum.tex
math

--- 2017-04-26 statistical-quorum.tex
@@
   and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
-  invalid votes. The last column shows computed decision credibility as described in Section \nameref{sec:statistical-quorum}. We see that
+  invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:statistical-quorum}. We see that
   credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
@@
 To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
-hypothesis testing to evaluate the credibility of the voting result.
-It uses distribution of votes made to evaluate the probability of decision passing if everyone actually voted,
+hypothesis testing to evaluate the credibility of the result.
+It uses distribution of votes made to compute the probability of decision passing if everyone voted,
 and not just those who voted.
@@
 Only two more votes against can completely change the outcome.
-Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
-We describe decision credibility in more detail in Section \nameref{sec:statistical-quorum}.
+We describe decision credibility in more detail in Section~\ref{sec:statistical-quorum}\todo{Hm, SIGCGI template seems to not display section numbers. So how useful is to reference a section number?}.
 
@@
 which puts legitimacy of those decisions into a question.
-We discuss a way for a community to determine a reasonable statistical quorum in Section \nameref{sec:discussion}.
+We discuss a way for a community to determine a reasonable statistical quorum in Section~\ref{sec:discussion}.
 
@@
 we can see amount of participation members are able to do as a limited resource, a \emph{collective attention}
-a community has available to utilize for its decision making.
+a community has available to utilize.
 
@@
 This allows the community to focus its collective attention towards decisions which require more votes
-to determine a legitimate result, while reducing attention towards proposals which already have overwhelming support.
+to determine a legitimate result.
 Instead of trying to increase participation on all democratic decisions being made, a community can
-do that selectively through an automatic feedback loop, trying to reach statistical quorum for their decisions.
+do that selectively through an automatic feedback loop.
 
-We built an open source web application PeerMind integrating statistical quorum with online voting.
+To evaluate the use of statistical quorum in decision making of democratic communities we have developed
+a web application PeerMind which integrates online discussion with online voting.
 The application exposes computed decision credibility to users in real-time and through that provides
 the feedback loop needed for improving utilization of collective attention.
-We present the app in Section \nameref{sec:peermind} and use it for evaluating statistical quorum in
-Section \nameref{sec:evaluation}.
+We present the app in Section~\ref{sec:peermind}.
 
@@
 
-\todo[inline]{Repeat the structure of the rest of the paper? Is there a better way to reference sections than
-referencing it by name?}
+\todo[inline]{Structure of the rest of the paper.}
 
@@
 
-\todo[inline]{Repeat references from "There have been many improvements proposed to this basic form and various analysis made."?}
-
-\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
-How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
-
-\todo[inline]{Voting auditing work.}
-
-\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
+% threshold and quorum at https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf
+% how does some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate them
 
@@
 
-\section{Decision Credibility and Statistical Quorum}
+\section{Statistical Quorum}
 \label{sec:statistical-quorum}
+A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
+For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose, then we can conclude with high probability that, had all 100 members taken the time to vote, a majority of them would be in support of the proposal.
+Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote was credible enough to be taken as representative of the entire community's desire.
+On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in the vote.
+Instead, the suggested course of action would be to continue gathering more votes and/or further discussing the proposal until a credible outcome is established.
 
-\section{Evaluation}
-\label{sec:evaluation}
-
-% theoretical: math, diagrams
-
-% imperial: running it for x weeks in student community which has traditionally weekly meeting
-% issues with quorum
-% participation of everyone
-% they used online app for x weeks
-% we compare their old minutes with performance of new system
-
-% interviews and surveys
-
-PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
-anonymized dataset of interactions with the app is available at \url{http://redacted.for.blind.review}.
-
-\section{Discussion}
-\label{sec:discussion}
-
-% describe why 0.90% is a reasnable statistical quorum and how to determine one for your community
-
-\section{Future Work}
-
-\todo[inline]{More work on security evaluation of potential attacks when a community is using statistical quorum to force a decision.
-Is there a game theoretical approach to it?}
-
-\todo[inline]{How to surface information about credibility to users better than just a number? Visualizations? How can we help them
-understand what exactly this is and how the math works?}
-
-\section{Conclusion}
-
-%\section{Acknowledgments}
-
-% REFERENCES FORMAT
-% References must be the same font size as other body text.
-\bibliographystyle{sigchi}
-\bibliography{refs}
-
-\appendix
-\section{Derivation of Decision Credibility}
-\label{sec:derivation}
-
-%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> proposal, motion -> proposal}
+By utilizing a statistical quorum, a community can actively direct attention and participation within the community towards contentious proposals, while avoiding wasting attention on noncontroversial proposals.
 
 \subsection{Background}
-A population of voters (a community) needs to collectively decide which item (if any) to take action on out of a set of competing items.
-In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a
-winning item, while also demonstrating that the winning item is preferred over all other competing items.
-We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that
-the population does not take undesired actions, which in this case encompasses taking action on an item that is not
-supported by a majority of the population and/or taking action on an item that is not clearly preferred over all
-other competing alternatives.
+Consider a community that needs to collectively decide which proposal (if any) to accept out of a set of competing proposals.
+To act in democratic fashion, the community may wish to select a winning proposal that either has the support of a \{simple, super\}-majority of all members, or instead is demonstrably preferred over all other competing proposals.
+(In some cases, the community may want a winning proposal to satisfy both criteria.)
+
+Ideally, we do not want to poll all members in order to officially accept a proposal, but we do want some level of certainty (e.g., $90\%$ certainty) that
+the population is not accidentally accepting an undesired proposal, which in this case encompasses accepting a proposal that is not supported by a majority of members and/or is not clearly preferred over all other competing alternatives.
+
+The primary motivation behind using a statistical quorum is to guarantee a level of credibility before letting a subset of members act on behalf of the entire community.
 
 \subsection{Problem Statement}
-We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some
+We wish to compute the probability that the bias in a community towards supporting a proposal is greater than some
 fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority).
-In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater
-than the bias towards voting ``yes'' on all other competing items.
+
+In addition, we wish to compute the probability that the community bias towards supporting a specific proposal is greater than the community bias towards supporting any competing proposals.
 
 \subsection{Assumptions}
-We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
+We assume that there is a fixed latent bias for the support of a proposal within the community.
+We take each individual vote to be independent and identically distributed (i.i.d.) according to a Bernoulli distribution around this bias.
+It should be noted that the independence assumption may not hold in some settings, such as when votes are accumulated over time (one may observe surges of correlated votes).
+We will partially address this when refining our original formulation of a statistical quorum into a \textit{dynamic} statistical quorum.
 
-Without any votes, we assume that this bias may be any value between $0\%$ and $100\%$ with equal probability.
+Without having collected any votes, we assume that the community bias may be any value between $0\%$ and $100\%$ with equal probability (uniform prior).
 
 \subsection{Voting}
-We allow for range voting where each member of the population may choose any real number in the range $[-1,+1]$ to
-represent their vote, with $-1$ representing opposition and $+1$ representing support.
-In the theory subsection below, we rescale votes to be in the range $[0,1]$ via the notation $\frac{votes + 1}{2}$.
+We allow for range voting where each member of the community may choose any real-valued vote $v \in [-1,+1]$ to
+represent their vote, with $v = -1$ representing full opposition and $v=+1$ representing full support.
+In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning the full range, hence our treatment of $v$ as a discrete random variable.
+Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as an individual's average vote.
+In the theory subsection below, we rescale $v$ to be in the range $[0,1]$ via the notation $\frac{v + 1}{2}$.
 
 We allow voters to ``abstain'' from voting.
-We treat abstentions as indicating that a voter is removing themselves from the population of voters for an item.
-It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$
-(indicating indifference).
+We treat abstentions as indicating that a voter is removing themselves from the community on a given vote.
+It is important to distinguish an ``abstention'' (removing one's self from the voting population) from voting $0$ (indicating indifference).
 
 \subsection{Bayesian Theory}
+After collecting an initial set of votes from members, we want to use that data to construct a prior distribution to help estimate the posterior probability that the proposal is acceptable to the entire community.
+Because the sum of  the collected votes follows a binomial distribution, a commonplace and convenient choice for the form of the prior on the community's bias is the Beta distribution (due to its conjugacy with the binomial distribution).
 
-The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a
-consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises
-naturally when considering samples from a finite population.\footnote{To
-properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial:
-\url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
-The tutorial is very similar to the problem we have at hand.}
-
-After extracting the sufficient statistics from raw votes~---~namely the average score and the number of
-abstentions~---~we can adjust the effective population size and corresponding threshold.
-In addition, we can calculate the effective number of ``yes'' and ``no'' votes.\footnote{Which one can think
-of as the number of Heads and Tails in the aforementioned tutorial.}
+After extracting the sufficient statistics from raw votes --- namely, the average score and the number of abstentions --- we can adjust the effective community size and corresponding threshold.
+In addition, we can calculate the effective number of ``yes'' ($+1$) and ``no'' ($-1$) votes:
 
@@
 \begin{itemize}
-\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing motions,
+\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing proposals,
 \item[$N = $] population size,
-\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for motion $X_i$, excluding ``abstentions'',
-\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on motion $X_i$,
-\item[$\alpha_i = $] $\sum\limits_{v \in X_i} \left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
+\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding abstentions,
+\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on proposal $X_i$,
+\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
 \item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
 \item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
-\item[$p_i = $] true latent population bias towards voting ``yes'' on $X_i$.
+\item[$p_i = $] true latent community bias towards voting ``yes'' on $X_i$.
 \end{itemize}
@@
 
-When we have a single motion open for voting, $C=\{X_1\}$, we simply wish to compute:
+In Eqs. \eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we refer the reader to the \hyperref[sec:derivation]{Appendix} for complete derivation details and justifications.
+
+When considering a single proposal $X_i$ in isolation, we wish to compute:
 \begin{align}\label{eq1}
-\Pr(\alpha^*_1 > T_1 \mid X_1),
+\Pr(\alpha^*_i > T_i \mid X_i),
 \end{align}
@@
 \item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
-(e.g. $f=\frac{1}{2}$ for a simple majority).
+(e.g., $f=\frac{1}{2}$ for a simple majority).
 \end{itemize}
 \end{description}
-Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when
-no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
+Using a desirable prior 
+\begin{align}\label{eq2}
+p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
+\end{align}
+corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can express \eqref{eq1} as:
+\begin{align}\label{eq3}
+\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
+\end{align}
+for
 \begin{align*}
-\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \Pr(\alpha^*_1 \leq T_1 \mid X_1),\\
-\Pr(\alpha^*_1 \leq T_1 \mid X_1) &= \int_0^1 \Pr(\alpha^*_1 \leq T_1 \mid p_1)\cdot \Pr(p_1 \mid X_1) \mathrm{d}p_1\\
-&= \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \\
-& \qquad\qquad\quad \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1,
+\nu_i & \sim \operatorname{Beta-Binomial}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}, K_i),\\
+{CDF}_{\nu_i}(V_i) & = \Pr(\nu_i \leq V_i),
 \end{align*}
@@
 \begin{itemize}
-\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that
-$X_i$ would remain below threshold.
+\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes $X_i$ could receive such that it would remain below threshold.
 \end{itemize}
@@
 
-Hence, we have
-\begin{align}\label{eq2}
-& \Pr(\alpha^*_1 > T_1 \mid X_1) = \nonumber\\
-& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
+When considering a proposal $X_i$ relative to other proposals in $C$, we wish to compute
+\begin{align}\label{eq4}
+\Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right)
 \end{align}
 
-In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to
-consider.
-Without loss of generality, assume that $X_1$ is the most popular motion within $C$.
-We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other
-$X_i  \in C, i \neq 1$:
-
-\begin{align}\label{eq3}
-\Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)
+Under the assumption that the community's votes on motions are independent across motions (which, in many instances, will actually give a lower bound on \eqref{eq3} when a community's support of one proposal is negatively correlated with its support on a competing proposal), we can decompose \eqref{eq3} as follows:
+\begin{align}\label{eq5}
+& \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) \nonumber \\
+= & \prod_{j \neq i}\left(1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \Pr(\nu_j = k) \cdot {CDF}_{\nu_i}(\alpha_j + k) \right)
 \end{align}
-
-Using the assumption
-that votes on motions are independent across motions, which, in the common
-scenario where there is a negative covariance between votes on competing
-motions, will provide a lower bound on the probability, we can decompose \eqref{eq3} as follows:
-\begin{align*}
-& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\\
-=& \prod_{i=2}^n \Pr(\alpha^*_1 > \alpha^*_i \mid C)\\
-=& \prod_{i=2}^n \iint\limits_{p_1, p_i} \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \Pr(p_1 \mid X_1) \cdot \Pr(p_i \mid X_i) \mathrm{d}p_i \mathrm{d}p_1\\
-=& \prod_{i=2}^n \int_0^1 \int_0^1 \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \\
-& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1\\
-\end{align*}
-Deriving an expression for $\Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)$, we get:
-\begin{align*}
-& \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)\\
-=& 1 - \Pr(\alpha^*_1 \leq \alpha^*_i \mid p_1, p_i)\\
-=& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
-\end{align*}
 where
@@
 \begin{itemize}
-\item[$v_{i,j} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
+\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
 \end{itemize}
@@
 
-Denoting
-\begin{align}\label{eq4}
-& \Phi_i(p_1,p_i) {\overset {\underset {\mathrm {def} }{}}{=}} \nonumber\\
-& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
+Using Eqs.
+\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
+For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.
+
+\subsection{Dynamic Statistical Quorum}
+Oftentimes, not all proposals are relevant to the entire community.
+Instead, a niche proposal may only affect a subset of the community, and hence to require much of the community to indicate their indifference and/or decide their opinion is a waste of the their fixed attentional budget.
+While the underlying formulation for a statistical quorum was to establish credibility guarantees before taking action, a natural consequence in common scenarios of the kind just described leads to a reduction in decision-making efficiency.
+How can we maintain credibility guarantees while simultaneously boosting efficiency?
+
+The solution is to infer the \textit{relevant} population size that cares about a given proposal so that we may establish credibility within only the subset population (naturally requiring less global attention).
+By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs.
+\eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).
+
+To infer $N_i$, we turn to a dynamic voting process whereby information embedded within the dynamics is exploited.
+By modeling the arrival rate of new votes as a nonhomogeneous Poisson process, we can extrapolate into the future.
+For a decaying arrival rate of new votes, a nonhomogeneous Poisson process $\{N(t),t\geq 0\}$ probabilistically models the number of future arrivals over a given time interval $[0,t)$ according to:
+\begin{align*}
+\Pr(N(t)=n)={\frac {[\Lambda (t)]^{n}}{n!}}e^{-\Lambda (t)},
+\end{align*}
+where
+\begin{align*}
+\Lambda (t)=\int _{0}^{t}\lambda (\tilde{t})\mathrm{d}\tilde{t},
+\end{align*}
+for some arrival rate process $\lambda (t)$.
+
+For our purposes, we make a natural choice,
+\begin{align}\label{eq6}
+\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}},
 \end{align}
-we can simplify \eqref{eq3} to:
-\begin{align}\label{eq5}
-& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\nonumber\\
-&= \prod_{i=2}^n \int_0^1 \int_0^1 \Phi_i(p_1,p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \nonumber\\
-& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1
+where
+\begin{description}
+\begin{itemize}
+\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that sufficiently smooths the historical rate),
+\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data for the community via a regression analysis.
+\end{itemize}
+\end{description}
+Now, for example, to upper bound with $95\%$ probability the remaining number of votes still pending, we solve for $\hat{K_i}$ to satisfy the following inequality:
+\begin{align}\label{eq7}
+\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}}\\
+& \geq 95\% \nonumber
 \end{align}
+Upon solving for $\hat{K_i}$, we set $N_i = \alpha_i + \beta_i + \hat{K_i}$, and can proceed to check for a statistical quorum.
 
-We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.
-In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=0.5\textwidth]{figures/}
+%\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
+%\label{simple_majority}
+%\end{figure}
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/simple_majority}
-\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
-\label{simple_majority}
-\end{figure}
-%Data:
-%Curve 1: {{0,0.000299648},{1,0.00213083},{2,0.00817989},{3,0.0224777},{4,0.049512},{5,0.0928927},{6,0.154136},{7,0.232044},{8,0.322859},{9,0.421057},{10,0.520467},{11,0.615358},{12,0.701241},{13,0.775267},{14,0.836245},{15,0.884385},{16,0.920892},{17,0.947532},{18,0.966265},{19,0.978973},{20,0.987296},{21,0.992561},{22,0.99578},{23,0.997682},{24,0.998767},{25,0.999366}}
-%Curve 2: {x, 0.99} for all x
+\section{Evaluation}
+\label{sec:evaluation}
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/super_majority}
-\caption{Super Majority: $\alpha$ vs. $\Pr(\alpha > \frac{2}{3}N)$ for $N=100$, $\beta = 10$.}
-\label{super_majority}
-\end{figure}
-%Data:
-%Curve 1: {{10,0.0395173},{11,0.0591278},{12,0.0846131},{13,0.116407},{14,0.15463},{15,0.199057},{16,0.24911},{17,0.303899},{18,0.362273},{19,0.422909},{20,0.484399},{21,0.545345},{22,0.604445},{23,0.660559},{24,0.712764},{25,0.760382},{26,0.802985},{27,0.840385},{28,0.872612},{29,0.899875},{30,0.922519},{31,0.940987},{32,0.955777},{33,0.967407},{34,0.976384},{35,0.983186},{36,0.988243},{37,0.99193},{38,0.994567},{39,0.996414},{40,0.997682},{41,0.998533},{42,0.999093},{43,0.999452},{44,0.999677},{45,0.999815},{46,0.999896},{47,0.999944},{48,0.99997},{49,0.999985},{50,0.999993}}
-%Curve 2: {x, 0.99} for all x
+\section{Discussion}
+\label{sec:discussion}
 
-%\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
-%we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
-%also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
+% describe why 0.90% is a reasonable statistical quorum and how to determine one for your community
 
+\section{Conclusion}
+
+%\section{Acknowledgments}
+
+\bibliographystyle{sigchi}
+\bibliography{refs}
+
+\appendix
+\label{sec:derivation}
+\subsection{Derivation of Eq. \eqref{eq2}}
+In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.
+To update the prior based on collected votes, standard theory suggests updating to $\operatorname{Beta}(1+\alpha_i,1+\beta_i)$; however, there are a few consistency issues with this update that we seek to fix:
+\begin{description}
+\begin{itemize}
+\item[$(1)$] The updated prior should not depend on the absolute number of votes $\alpha_i$, $\beta_i$, as this is not scale-invariant; rather, it should depend on the relative proportions (which are unitless instead of a number of votes),
+\item[$(2)$] as $\alpha_i + \beta_i \rightarrow N_i$, $\mathbb{E}[p_i] \rightarrow \frac{\alpha_i}{\alpha_i+\beta_i}$ (at a rate independent of $N_i$),
+\item[$(3)$] as $\alpha_i + \beta_i \rightarrow N_i$, $\operatorname{Var}[p_i] \rightarrow 0$ (at a rate independent of $N_i$).
+\end{itemize}
+\end{description}
+To satisfy these consistency properties, we seek to linearly update the prior's scale parameters in a more natural way: $\operatorname{Beta}(1+\eta\alpha_i,1+\eta\beta_i)$ for some renormalizing factor $\eta$.
+
+We can now solve for $\eta$:
+\begin{align*}
+\operatorname{Var}[p_i] &= \frac{(1+\eta\alpha_i)(1+\eta\beta_i)}{(2+\eta\alpha_i+\eta\beta_i)^2(3+\eta\alpha_i+\eta\beta_i)}
+\end{align*}
+Clearly, to satisfy $(3)$, we need $\eta \rightarrow \infty$ as $\alpha_i + \beta_i \rightarrow N_i$.
+Consequently, the dominant asymptotic term becomes $\frac{1}{\eta^3 (\alpha_i + \beta_i)^3} = \frac{1}{\eta^3 N_i^3}$.
+Hence, for the variance to asymptotically approach $0$ at a rate independent of $N_i$, we need $\eta \propto \frac{1}{N_i}$.
+
+Combining these two constraints, we arrive at $\eta \propto \frac{1}{N_i}\cdot \frac{1}{K_i/N_i}$, where the latter factor leads to $\eta \rightarrow \infty$ as $\alpha_i + \beta_i \rightarrow N_i$ without changing the units of $\eta$ ($1/\text{votes}$).
+
+Hence we arrive at $\eta = \frac{c}{K_i}$, where $c$ is any constant that can be selected to tune the rate at which the variance approaches $0$.  For simplicity, we use $\eta = \frac{1}{K_i}$.
+
+As a sanity check, one can verify that this form for $\eta$ satisfies properties $(1)-(3)$.
+
+\subsection{Derivation of Eq. \eqref{eq3}}
+\begin{align*}
+\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - \Pr(\alpha^*_i \leq T_i \mid X_i), \text{ where}\\
+\Pr(\alpha^*_i \leq T_i \mid X_i) &= \int_0^1 \Pr(\alpha^*_i \leq T_i \mid p_i)\cdot \Pr(p_i \mid X_i) \mathrm{d}p_i\\
+\end{align*}
+Using the binomial and Beta distributions, we get
+\begin{align*}
+&= \int_0^1 \sum\limits_{v=0}^{V_i} {K_i \choose v} p_i^{v} {(1-p_i)}^{K_i-v} \cdot \operatorname{Beta}(p_i; 1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}) \mathrm{d}p_i,\\
+&= \sum\limits_{v=0}^{V_i} \underbrace{\int_0^1 {K_i \choose v} p_i^{v} {(1-p_i)}^{K_i-v} \cdot \operatorname{Beta}(p_i; 1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}) \mathrm{d}p_i}_{= \Pr(\nu_i = v)\text{ for } \nu_i\ \sim\ \operatorname{Beta-Binomial}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}, K_i)}.
+\end{align*}
+
+Hence, we have
+\begin{align*}
+\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \sum\limits_{v=0}^{V_i} \Pr(\nu_i = v)\\
+&= 1 - \Pr(\nu_i \leq V_i).
+\end{align*}
+
+\subsection{Derivation of Eq. \eqref{eq5}}
+Assuming independence of votes across proposals,
+\begin{align*}
+\Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) &=\prod_{j \neq i} \Pr(\alpha^*_i > \alpha^*_j \mid X_i, X_j).
+\end{align*}
+(If, instead, votes in $X_i$ were negatively correlated with votes in $X_j$, and $X_i$ had more support, then the equality ($=$) above would be replaced by an inequality ($>$).)
+
+Further expanding the expression, we get
+\begin{align*}
+&= \prod_{j \neq i} \left( 1 - \Pr(\alpha^*_i \leq \alpha^*_j \mid X_i, X_j) \right)\\
+&= \prod_{j \neq i} \left( 1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \underbrace{\Pr(\alpha^*_j = \alpha_j + k \mid X_j)}_{= \Pr(\nu_j = k)} \cdot \underbrace{\Pr(\alpha^*_i \leq \alpha_j + k \mid X_i)}_{={CDF}_{\nu_i}(\alpha_j + k)} \right).
+\end{align*}
+
+\subsection{Derivation of Eq. \eqref{eq6}}
+The properties we desire for $\lambda(t)$ are:
+\begin{description}
+\begin{itemize}
+\item[$(1)$] as $K_i \rightarrow 0$, the rate of incoming votes should drop exponentially and should be independent of $N_i$,
+\item[$(2)$] as $t \rightarrow \infty$, the rate of incoming votes should drop exponentially.
+\end{itemize} 
+\end{description}
+Setting $\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}}$ satisfies these constraints while still allowing for the exact rate to be tunable.
+
+\subsection{Derivation of Eq. \eqref{eq7}}
+Using $\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}}$, we have
+\begin{align*}
+\lim_{t \rightarrow \infty} \Lambda (t) &=\int _{0}^{\infty}\lambda (t)\mathrm{d}t\\
+&= \int _{0}^{\infty} \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}} \mathrm{d}t\\
+&= \lambda_0 \frac{K_i/N_i}{\ln{\gamma}}.
+\end{align*}
+Thus, 
+\begin{align*}
+\lim_{t\rightarrow \infty}\Pr\left(N(t) = n\right) = \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}},
+\end{align*}
+and
+\begin{align*}
+\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}}.
+\end{align*}
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2017-04-27 statistical-quorum.tex

@@
   and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
-  invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:statistical-quorum}. We see that
+  invalid votes. The last column shows computed decision credibility as described in Section \secref{sec:statistical-quorum}. We see that
   credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
@@
 Only two more votes against can completely change the outcome.
-We describe decision credibility in more detail in Section~\ref{sec:statistical-quorum}\todo{Hm, SIGCGI template seems to not display section numbers. So how useful is to reference a section number?}.
+Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
+We describe decision credibility in more detail in Section \secref{sec:statistical-quorum}.
 
@@
 which puts legitimacy of those decisions into a question.
-We discuss a way for a community to determine a reasonable statistical quorum in Section~\ref{sec:discussion}.
+We discuss a way for a community to determine a reasonable statistical quorum in Section \secref{sec:discussion}.
 
@@
 This allows the community to focus its collective attention towards decisions which require more votes
-to determine a legitimate result.
+to determine a legitimate result, while reducing attention towards proposals which already have overwhelming support.
 Instead of trying to increase participation on all democratic decisions being made, a community can
-do that selectively through an automatic feedback loop.
+do that selectively through an automatic feedback loop, trying to reach statistical quorum for their decisions.
 
-To evaluate the use of statistical quorum in decision making of democratic communities we have developed
-a web application PeerMind which integrates online discussion with online voting.
+We built an open source web application PeerMind integrating statistical quorum with online voting.
 The application exposes computed decision credibility to users in real-time and through that provides
 the feedback loop needed for improving utilization of collective attention.
-We present the app in Section~\ref{sec:peermind}.
+We present the application in Section \secref{sec:peermind} and use it for evaluating statistical quorum in
+Section \secref{sec:evaluation}.
 
@@
 
-\todo[inline]{Structure of the rest of the paper.}
+\todo[inline]{Repeat the structure of the rest of the paper? Is there a better way to reference sections than
+referencing it by name?}
 
@@
 
-% threshold and quorum at https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf
-% how does some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate them
+\todo[inline]{Repeat references from "There have been many improvements proposed to this basic form and various analysis made."?}
+
+\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
+How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
+
+\todo[inline]{Voting auditing work.}
+
+\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
 
@@
 
+We evaluate statistical quorum in two ways. First, we evaluate it theoretically and explore its mathematical properties.
+Then we evaluate it through real-world use of the PeerMind application by a 140 members community.
+
+\subsection{Properties of Statistical Quorum}
+
+\todo[inline]{Theoretical evaluation, diagrams. Get readers to understand some of its properties visually.}
+
+\subsection{Use in Existing Democratic Community}
+
+% imperial: running it for x weeks in student community which has traditionally weekly meeting
+% issues with quorum
+% participation of everyone
+% they used online app for x weeks
+% we compare their old minutes with performance of new system
+
+% interviews and surveys
+
+\todo[inline]{Make sure that we are supporting both of our claims well: this statistical quorum is a good metric and
+collective attention can be utilized better.}
+
+PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
+anonymized dataset of interactions with the app, including votes, is available at \url{http://redacted.for.blind.review}.
+
 \section{Discussion}
@@
 
-% describe why 0.90% is a reasonable statistical quorum and how to determine one for your community
+\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.}
+
+\todo[inline]{Describe interesting example how we had a bug which were showing a very low statistical quorum for a failing
+vote and so the whole community very engaged for a whole month to decide and resolve it. So how they trusted the
+app and the metric and how it was really able to direct the attention of the community. It seems it works even if it is
+buggy or not perfect, because it has worked so well in the past that members got used to the fact that when it was saying
+that there is a consensus or not on an issue that that was also their own read of the community. So for the purpose of
+directing attention the metric does not have to mathematically perfect.}
+
+\todo[inline]{Security. How to prevent a group of people to vote and by trick reaching statistical quorum.
+We use dynamic statistical quorum to address this, but one other approach is also for voting to stay open and
+decisions can then be in the future revoked by people adding more votes, or changing them. But for some decisions
+this might not really be possible, because decisions might not be reversible. You can revoke a law, but if you built
+a building as a consequence of a decision?}
+
+\section{Future Work}
+
+\todo[inline]{More work on security evaluation of potential attacks when a community is using statistical quorum to force a decision.
+Is there a game theoretical approach to it?}
+
+\todo[inline]{How to surface information about credibility to users better than just a number? Visualizations? How can we help them
+understand what exactly this is and how the math works?}
 
@@
 
+% REFERENCES FORMAT
+% References must be the same font size as other body text.
 \bibliographystyle{sigchi}

### 2017-04-27 statistical-quorum.tex
fixed math typo

@@
 \begin{align}\label{eq6}
-\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}},
+\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}},
 \end{align}
@@
 \begin{align}\label{eq7}
-\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}}\\
+\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}}\\
 & \geq 95\% \nonumber
@@
 \end{description}
-Setting $\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}}$ satisfies these constraints while still allowing for the exact rate to be tunable.
+Setting $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$ for $0 < \gamma < 1$ satisfies these constraints while still allowing for the exact rate to be tunable.
 
 \subsection{Derivation of Eq. \eqref{eq7}}
-Using $\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}}$, we have
+Using $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$, we have
 \begin{align*}
 \lim_{t \rightarrow \infty} \Lambda (t) &=\int _{0}^{\infty}\lambda (t)\mathrm{d}t\\
-&= \int _{0}^{\infty} \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}} \mathrm{d}t\\
-&= \lambda_0 \frac{K_i/N_i}{\ln{\gamma}}.
+&= \int _{0}^{\infty} \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}} \mathrm{d}t\\
+&= \lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}.
 \end{align*}
@@
 \begin{align*}
-\lim_{t\rightarrow \infty}\Pr\left(N(t) = n\right) = \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}},
+\lim_{t\rightarrow \infty}\Pr\left(N(t) = n\right) = \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}},
 \end{align*}
@@
 \begin{align*}
-\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}}.
+\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}}.
 \end{align*}

### 2017-04-27 statistical-quorum.tex
clarified math
More TODOs.

--- 2017-04-27 statistical-quorum.tex
@@
 
+We built a full-fletched real-time web application for collective decision making and voting using Meteor
+framework~\cite{meteor}.
+Users can initiate new discussions which leads them to the main user interface for each discussion as seen
+in Figure~\ref{fig:peermind}.
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=1.75\columnwidth]{figures/peermind.png}
+  \caption{The main PeerMind interface of a discussion. There is initial discussion description on the top and under it there
+  are three columns. A commenting sections for users to discuss it and proposed decisions, a second column where moderators can
+  summarize arguments made during discussion, and a list of proposals to address the discussion.
+  Proposals can be in different states: a proposal being discussed, a proposal being voted on, and with voting concluded.}\label{fig:peermind}
+\end{figure*}
+
+%and discuss them with others online in real-time and with rich-text content.
+%They can attach images and files using drag\&drop. Moderators can summarize discussions into three groups
+
+\todo[inline]{Present how users can see ongoing results during voting. Show an image of active ballot, but with expanded view for current results.}
+
 %Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
@@
 
-\section{Statistical Quorum}
+\section{Decision Credibility and Statistical Quorum}
 \label{sec:statistical-quorum}
-A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
-For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose, then we can conclude with high probability that, had all 100 members taken the time to vote, a majority of them would be in support of the proposal.
-Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote was credible enough to be taken as representative of the entire community's desire.
-On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in the vote.
-Instead, the suggested course of action would be to continue gathering more votes and/or further discussing the proposal until a credible outcome is established.
 
-By utilizing a statistical quorum, a community can actively direct attention and participation within the community towards contentious proposals, while avoiding wasting attention on noncontroversial proposals.
-
-\subsection{Background}
-Consider a community that needs to collectively decide which proposal (if any) to accept out of a set of competing proposals.
-To act in democratic fashion, the community may wish to select a winning proposal that either has the support of a \{simple, super\}-majority of all members, or instead is demonstrably preferred over all other competing proposals.
-(In some cases, the community may want a winning proposal to satisfy both criteria.)
-
-Ideally, we do not want to poll all members in order to officially accept a proposal, but we do want some level of certainty (e.g., $90\%$ certainty) that
-the population is not accidentally accepting an undesired proposal, which in this case encompasses accepting a proposal that is not supported by a majority of members and/or is not clearly preferred over all other competing alternatives.
-
-The primary motivation behind using a statistical quorum is to guarantee a level of credibility before letting a subset of members act on behalf of the entire community.
-
-\subsection{Problem Statement}
-We wish to compute the probability that the bias in a community towards supporting a proposal is greater than some
-fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority).
-
-In addition, we wish to compute the probability that the community bias towards supporting a specific proposal is greater than the community bias towards supporting any competing proposals.
-
-\subsection{Assumptions}
-We assume that there is a fixed latent bias for the support of a proposal within the community.
-We take each individual vote to be independent and identically distributed (i.i.d.) according to a Bernoulli distribution around this bias.
-It should be noted that the independence assumption may not hold in some settings, such as when votes are accumulated over time (one may observe surges of correlated votes).
-We will partially address this when refining our original formulation of a statistical quorum into a \textit{dynamic} statistical quorum.
-
-Without having collected any votes, we assume that the community bias may be any value between $0\%$ and $100\%$ with equal probability (uniform prior).
-
-\subsection{Voting}
-We allow for range voting where each member of the community may choose any real-valued vote $v \in [-1,+1]$ to
-represent their vote, with $v = -1$ representing full opposition and $v=+1$ representing full support.
-In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning the full range, hence our treatment of $v$ as a discrete random variable.
-Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as an individual's average vote.
-In the theory subsection below, we rescale $v$ to be in the range $[0,1]$ via the notation $\frac{v + 1}{2}$.
-
-We allow voters to ``abstain'' from voting.
-We treat abstentions as indicating that a voter is removing themselves from the community on a given vote.
-It is important to distinguish an ``abstention'' (removing one's self from the voting population) from voting $0$ (indicating indifference).
-
-\subsection{Bayesian Theory}
-After collecting an initial set of votes from members, we want to use that data to construct a prior distribution to help estimate the posterior probability that the proposal is acceptable to the entire community.
-Because the sum of  the collected votes follows a binomial distribution, a commonplace and convenient choice for the form of the prior on the community's bias is the Beta distribution (due to its conjugacy with the binomial distribution).
-
-After extracting the sufficient statistics from raw votes --- namely, the average score and the number of abstentions --- we can adjust the effective community size and corresponding threshold.
-In addition, we can calculate the effective number of ``yes'' ($+1$) and ``no'' ($-1$) votes:
-
-Let
-\begin{description}
-\begin{itemize}
-\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing proposals,
-\item[$N = $] population size,
-\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding abstentions,
-\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on proposal $X_i$,
-\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
-\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
-\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
-\item[$p_i = $] true latent community bias towards voting ``yes'' on $X_i$.
-\end{itemize}
-\end{description}
-
-In Eqs. \eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we refer the reader to the \hyperref[sec:derivation]{Appendix} for complete derivation details and justifications.
-
-When considering a single proposal $X_i$ in isolation, we wish to compute:
-\begin{align}\label{eq1}
-\Pr(\alpha^*_i > T_i \mid X_i),
-\end{align}
-where
-\begin{description}
-\begin{itemize}
-\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
-\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
-(e.g., $f=\frac{1}{2}$ for a simple majority).
-\end{itemize}
-\end{description}
-Using a desirable prior 
-\begin{align}\label{eq2}
-p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
-\end{align}
-corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can express \eqref{eq1} as:
-\begin{align}\label{eq3}
-\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
-\end{align}
-for
-\begin{align*}
-\nu_i & \sim \operatorname{Beta-Binomial}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}, K_i),\\
-{CDF}_{\nu_i}(V_i) & = \Pr(\nu_i \leq V_i),
-\end{align*}
-where
-\begin{description}
-\begin{itemize}
-\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes $X_i$ could receive such that it would remain below threshold.
-\end{itemize}
-\end{description}
-
-When considering a proposal $X_i$ relative to other proposals in $C$, we wish to compute
-\begin{align}\label{eq4}
-\Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right)
-\end{align}
-
-Under the assumption that the community's votes on motions are independent across motions (which, in many instances, will actually give a lower bound on \eqref{eq3} when a community's support of one proposal is negatively correlated with its support on a competing proposal), we can decompose \eqref{eq3} as follows:
-\begin{align}\label{eq5}
-& \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) \nonumber \\
-= & \prod_{j \neq i}\left(1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \Pr(\nu_j = k) \cdot {CDF}_{\nu_i}(\alpha_j + k) \right)
-\end{align}
-where
-\begin{description}
-\begin{itemize}
-\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
-\end{itemize}
-\end{description}
-
-Using Eqs.
-\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
-For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.
-
-\subsection{Dynamic Statistical Quorum}
-Oftentimes, not all proposals are relevant to the entire community.
-Instead, a niche proposal may only affect a subset of the community, and hence to require much of the community to indicate their indifference and/or decide their opinion is a waste of the their fixed attentional budget.
-While the underlying formulation for a statistical quorum was to establish credibility guarantees before taking action, a natural consequence in common scenarios of the kind just described leads to a reduction in decision-making efficiency.
-How can we maintain credibility guarantees while simultaneously boosting efficiency?
-
-The solution is to infer the \textit{relevant} population size that cares about a given proposal so that we may establish credibility within only the subset population (naturally requiring less global attention).
-By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs.
-\eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).
-
-To infer $N_i$, we turn to a dynamic voting process whereby information embedded within the dynamics is exploited.
-By modeling the arrival rate of new votes as a nonhomogeneous Poisson process, we can extrapolate into the future.
-For a decaying arrival rate of new votes, a nonhomogeneous Poisson process $\{N(t),t\geq 0\}$ probabilistically models the number of future arrivals over a given time interval $[0,t)$ according to:
-\begin{align*}
-\Pr(N(t)=n)={\frac {[\Lambda (t)]^{n}}{n!}}e^{-\Lambda (t)},
-\end{align*}
-where
-\begin{align*}
-\Lambda (t)=\int _{0}^{t}\lambda (\tilde{t})\mathrm{d}\tilde{t},
-\end{align*}
-for some arrival rate process $\lambda (t)$.
-
-For our purposes, we make a natural choice,
-\begin{align}\label{eq6}
-\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}},
-\end{align}
-where
-\begin{description}
-\begin{itemize}
-\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that sufficiently smooths the historical rate),
-\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data for the community via a regression analysis.
-\end{itemize}
-\end{description}
-Now, for example, to upper bound with $95\%$ probability the remaining number of votes still pending, we solve for $\hat{K_i}$ to satisfy the following inequality:
-\begin{align}\label{eq7}
-\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}}\\
-& \geq 95\% \nonumber
-\end{align}
-Upon solving for $\hat{K_i}$, we set $N_i = \alpha_i + \beta_i + \hat{K_i}$, and can proceed to check for a statistical quorum.
-
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=0.5\textwidth]{figures/}
-%\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
-%\label{simple_majority}
-%\end{figure}
+\todo[inline]{Definition and formulas for three types of decision credibility: single option, multiple competing options,
+and dynamic credibility which is increasing through time. In this order.}
 
@@
 
-% imperial: running it for x weeks in student community which has traditionally weekly meeting
-% issues with quorum
-% participation of everyone
-% they used online app for x weeks
-% we compare their old minutes with performance of new system
+\todo[inline]{Explain how we were using this for 16 weeks in fall 2016 in a community where traditionally they have
+been meeting weekly and now they used it online. How X proposals have been made with Y votes.}
 
-% interviews and surveys
+\todo[inline]{Explain how we compare with history: using minutes from spring 2016 for the same community. Issues with
+quorum often not bein reached or being lost during the meeting. How many proposals were made then.}
+
+\todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}
+
+\todo[inline]{We compare how much time collectively was spend on decision making based on minutes in spring 2016 and
+how much based on interactions with the platform in fall 2016. Claim is that we made more decisions with less time people
+having to be involved in decision making, and with better legitimacy.}
+
+\todo[inline]{We did satisfaction surveys with decision making both in spring and fall 2016. Pull out some interesting
+observation when comparing.}
+
+\todo[inline]{We made 6 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
+to work. But how there are other issues by moving decision making online: meetings also serve community building which
+does not happen online. No fun. So interesting insights about using online platforms for offline communities
+for making decisions which are not really answering our hypothesis, but are useful as insights we got.}
 
@@
 
+\todo[inline]{Discuss how this might or might not lead to a more inclusive decision making: some people prefer it because
+they have time to collect thoughts and can participate when they have time, some people do not because they feel
+they cannot communicate as effectively, especially with lack of emotions being conveyed.}
+
 \section{Future Work}
@@
 \appendix
+\section{Derivation of Decision Credibility}
 \label{sec:derivation}
-\subsection{Derivation of Eq. \eqref{eq2}}
-In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.
-To update the prior based on collected votes, standard theory suggests updating to $\operatorname{Beta}(1+\alpha_i,1+\beta_i)$; however, there are a few consistency issues with this update that we seek to fix:
+
+\todo[inline]{Derivation in order of dynamic credibility which is increasing through time, multiple competing options,
+single option, because this is how they are easier to derive as each following is a special case of the previous one.}
+
+%\todo[inline]{Unify terminology: population -> community, voters -> members, item -> decision, motion -> decision}
+
+\subsection{Background}
+A population of voters (a community) needs to collectively decide which item (if any) to take action on out of a set of competing items.
+In order to democratically take action, the population needs to have a \{simple, super\}-majority of ``yes'' votes on a
+winning item, while also demonstrating that the winning item is preferred over all other competing items.
+We do not want to poll the entire population, but we do want to be certain (specifically, $99\%$ certain) that
+the population does not take undesired actions, which in this case encompasses taking action on an item that is not
+supported by a majority of the population and/or taking action on an item that is not clearly preferred over all
+other competing alternatives.
+
+\subsection{Problem Statement}
+We wish to compute the probability that the bias in a population towards voting ``yes'' is greater than some
+fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority).
+In addition, we wish to compute the probability that the bias towards voting ``yes'' on a specific item is greater
+than the bias towards voting ``yes'' on all other competing items.
+
+\subsection{Assumptions}
+We assume that there is a fixed underlying bias in the population that is to be estimated given raw voting data.
+
+Without any votes, we assume that this bias may be any value between $0\%$ and $100\%$ with equal probability.
+
+\subsection{Voting}
+We allow for range voting where each member of the population may choose any real number in the range $[-1,+1]$ to
+represent their vote, with $-1$ representing opposition and $+1$ representing support.
+In the theory subsection below, we rescale votes to be in the range $[0,1]$ via the notation $\frac{votes + 1}{2}$.
+
+We allow voters to ``abstain'' from voting.
+We treat abstentions as indicating that a voter is removing themselves from the population of voters for an item.
+It is important to distinguish an ``abstention'' (removing yourself from the voting population) from voting $0$
+(indicating indifference).
+
+\subsection{Bayesian Theory}
+
+The commonplace and convenient choice of using the Beta distribution as our prior on the population bias is a
+consequence of the Beta distribution's conjugacy with the binomial distribution, a distribution that arises
+naturally when considering samples from a finite population.\footnote{To
+properly understand the Bayesian approach to parameter estimation, read the following instructive tutorial:
+\url{http://www.behind-the-enemy-lines.com/2008/01/are-you-bayesian-or-frequentist-or.html}
+The tutorial is very similar to the problem we have at hand.}
+
+After extracting the sufficient statistics from raw votes~---~namely the average score and the number of
+abstentions~---~we can adjust the effective population size and corresponding threshold.
+In addition, we can calculate the effective number of ``yes'' and ``no'' votes.\footnote{Which one can think
+of as the number of Heads and Tails in the aforementioned tutorial.}
+
+Let
 \begin{description}
 \begin{itemize}
-\item[$(1)$] The updated prior should not depend on the absolute number of votes $\alpha_i$, $\beta_i$, as this is not scale-invariant; rather, it should depend on the relative proportions (which are unitless instead of a number of votes),
-\item[$(2)$] as $\alpha_i + \beta_i \rightarrow N_i$, $\mathbb{E}[p_i] \rightarrow \frac{\alpha_i}{\alpha_i+\beta_i}$ (at a rate independent of $N_i$),
-\item[$(3)$] as $\alpha_i + \beta_i \rightarrow N_i$, $\operatorname{Var}[p_i] \rightarrow 0$ (at a rate independent of $N_i$).
+\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing motions,
+\item[$N = $] population size,
+\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for motion $X_i$, excluding ``abstentions'',
+\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on motion $X_i$,
+\item[$\alpha_i = $] $\sum\limits_{v \in X_i} \left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
+\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
+\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
+\item[$p_i = $] true latent population bias towards voting ``yes'' on $X_i$.
 \end{itemize}
 \end{description}
-To satisfy these consistency properties, we seek to linearly update the prior's scale parameters in a more natural way: $\operatorname{Beta}(1+\eta\alpha_i,1+\eta\beta_i)$ for some renormalizing factor $\eta$.
 
-We can now solve for $\eta$:
+When we have a single motion open for voting, $C=\{X_1\}$, we simply wish to compute:
+\begin{align}\label{eq1}
+\Pr(\alpha^*_1 > T_1 \mid X_1),
+\end{align}
+where
+\begin{description}
+\begin{itemize}
+\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
+\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
+(e.g. $f=\frac{1}{2}$ for a simple majority).
+\end{itemize}
+\end{description}
+Using the prior $p_1 \mid X_1 \sim \operatorname{Beta}(1+\alpha_1,1+\beta_1)$, corresponding to a uniform prior when
+no votes have been cast ($\alpha_1 = \beta_1 = 0$), we can express \eqref{eq1} as:
 \begin{align*}
-\operatorname{Var}[p_i] &= \frac{(1+\eta\alpha_i)(1+\eta\beta_i)}{(2+\eta\alpha_i+\eta\beta_i)^2(3+\eta\alpha_i+\eta\beta_i)}
+\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \Pr(\alpha^*_1 \leq T_1 \mid X_1),\\
+\Pr(\alpha^*_1 \leq T_1 \mid X_1) &= \int_0^1 \Pr(\alpha^*_1 \leq T_1 \mid p_1)\cdot \Pr(p_1 \mid X_1) \mathrm{d}p_1\\
+&= \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \\
+& \qquad\qquad\quad \times \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1,
 \end{align*}
-Clearly, to satisfy $(3)$, we need $\eta \rightarrow \infty$ as $\alpha_i + \beta_i \rightarrow N_i$.
-Consequently, the dominant asymptotic term becomes $\frac{1}{\eta^3 (\alpha_i + \beta_i)^3} = \frac{1}{\eta^3 N_i^3}$.
-Hence, for the variance to asymptotically approach $0$ at a rate independent of $N_i$, we need $\eta \propto \frac{1}{N_i}$.
-
-Combining these two constraints, we arrive at $\eta \propto \frac{1}{N_i}\cdot \frac{1}{K_i/N_i}$, where the latter factor leads to $\eta \rightarrow \infty$ as $\alpha_i + \beta_i \rightarrow N_i$ without changing the units of $\eta$ ($1/\text{votes}$).
-
-Hence we arrive at $\eta = \frac{c}{K_i}$, where $c$ is any constant that can be selected to tune the rate at which the variance approaches $0$.  For simplicity, we use $\eta = \frac{1}{K_i}$.
-
-As a sanity check, one can verify that this form for $\eta$ satisfies properties $(1)-(3)$.
-
-\subsection{Derivation of Eq. \eqref{eq3}}
-\begin{align*}
-\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - \Pr(\alpha^*_i \leq T_i \mid X_i), \text{ where}\\
-\Pr(\alpha^*_i \leq T_i \mid X_i) &= \int_0^1 \Pr(\alpha^*_i \leq T_i \mid p_i)\cdot \Pr(p_i \mid X_i) \mathrm{d}p_i\\
-\end{align*}
-Using the binomial and Beta distributions, we get
-\begin{align*}
-&= \int_0^1 \sum\limits_{v=0}^{V_i} {K_i \choose v} p_i^{v} {(1-p_i)}^{K_i-v} \cdot \operatorname{Beta}(p_i; 1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}) \mathrm{d}p_i,\\
-&= \sum\limits_{v=0}^{V_i} \underbrace{\int_0^1 {K_i \choose v} p_i^{v} {(1-p_i)}^{K_i-v} \cdot \operatorname{Beta}(p_i; 1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}) \mathrm{d}p_i}_{= \Pr(\nu_i = v)\text{ for } \nu_i\ \sim\ \operatorname{Beta-Binomial}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}, K_i)}.
-\end{align*}
+where
+\begin{description}
+\begin{itemize}
+\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes allowed such that
+$X_i$ would remain below threshold.
+\end{itemize}
+\end{description}
 
 Hence, we have
+\begin{align}\label{eq2}
+& \Pr(\alpha^*_1 > T_1 \mid X_1) = \nonumber\\
+& 1 - \int_0^1 \sum\limits_{i=0}^{V_1} {K_1 \choose i} p_1^{i} {(1-p_1)}^{K_1-i} \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \mathrm{d}p_1
+\end{align}
+
+In the scenario where $\mathbf{card}(C) = n > 1$, i.e., there are competing motions, there is additional complexity to
+consider.
+Without loss of generality, assume that $X_1$ is the most popular motion within $C$.
+We now wish to not only compute \eqref{eq1}, but also the probability that $X_1$ is superior to all other
+$X_i  \in C, i \neq 1$:
+
+\begin{align}\label{eq3}
+\Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)
+\end{align}
+
+Using the assumption
+that votes on motions are independent across motions, which, in the common
+scenario where there is a negative covariance between votes on competing
+motions, will provide a lower bound on the probability, we can decompose \eqref{eq3} as follows:
 \begin{align*}
-\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \sum\limits_{v=0}^{V_i} \Pr(\nu_i = v)\\
-&= 1 - \Pr(\nu_i \leq V_i).
+& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\\
+=& \prod_{i=2}^n \Pr(\alpha^*_1 > \alpha^*_i \mid C)\\
+=& \prod_{i=2}^n \iint\limits_{p_1, p_i} \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \Pr(p_1 \mid X_1) \cdot \Pr(p_i \mid X_i) \mathrm{d}p_i \mathrm{d}p_1\\
+=& \prod_{i=2}^n \int_0^1 \int_0^1 \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \\
+& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1\\
 \end{align*}
-
-\subsection{Derivation of Eq. \eqref{eq5}}
-Assuming independence of votes across proposals,
+Deriving an expression for $\Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)$, we get:
 \begin{align*}
-\Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) &=\prod_{j \neq i} \Pr(\alpha^*_i > \alpha^*_j \mid X_i, X_j).
+& \Pr(\alpha^*_1 > \alpha^*_i \mid p_1, p_i)\\
+=& 1 - \Pr(\alpha^*_1 \leq \alpha^*_i \mid p_1, p_i)\\
+=& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
 \end{align*}
-(If, instead, votes in $X_i$ were negatively correlated with votes in $X_j$, and $X_i$ had more support, then the equality ($=$) above would be replaced by an inequality ($>$).)
-
-Further expanding the expression, we get
-\begin{align*}
-&= \prod_{j \neq i} \left( 1 - \Pr(\alpha^*_i \leq \alpha^*_j \mid X_i, X_j) \right)\\
-&= \prod_{j \neq i} \left( 1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \underbrace{\Pr(\alpha^*_j = \alpha_j + k \mid X_j)}_{= \Pr(\nu_j = k)} \cdot \underbrace{\Pr(\alpha^*_i \leq \alpha_j + k \mid X_i)}_{={CDF}_{\nu_i}(\alpha_j + k)} \right).
-\end{align*}
-
-\subsection{Derivation of Eq. \eqref{eq6}}
-The properties we desire for $\lambda(t)$ are:
+where
 \begin{description}
 \begin{itemize}
-\item[$(1)$] as $K_i \rightarrow 0$, the rate of incoming votes should drop exponentially and should be independent of $N_i$,
-\item[$(2)$] as $t \rightarrow \infty$, the rate of incoming votes should drop exponentially.
+\item[$v_{i,j} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
 \end{itemize}
 \end{description}
-Setting $\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}}$ satisfies these constraints while still allowing for the exact rate to be tunable.
 
-\subsection{Derivation of Eq. \eqref{eq7}}
-Using $\lambda (t) = \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}}$, we have
-\begin{align*}
-\lim_{t \rightarrow \infty} \Lambda (t) &=\int _{0}^{\infty}\lambda (t)\mathrm{d}t\\
-&= \int _{0}^{\infty} \lambda_0 {(\gamma)}^{-\frac{t}{K_i/N_i}} \mathrm{d}t\\
-&= \lambda_0 \frac{K_i/N_i}{\ln{\gamma}}.
-\end{align*}
-Thus, 
-\begin{align*}
-\lim_{t\rightarrow \infty}\Pr\left(N(t) = n\right) = \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}},
-\end{align*}
-and
-\begin{align*}
-\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{\gamma}}}.
-\end{align*}
+Denoting
+\begin{align}\label{eq4}
+& \Phi_i(p_1,p_i) {\overset {\underset {\mathrm {def} }{}}{=}} \nonumber\\
+& 1 - \sum\limits_{m=0}^{K_1} \sum\limits_{n=\lceil{v_{1,i}+m}\rceil}^{K_i} {K_1 \choose m} p_1^{m} {(1-p_1)}^{K_1 - m} \cdot {K_i \choose n} p_i^{n} {(1-p_i)}^{K_i-n},
+\end{align}
+we can simplify \eqref{eq3} to:
+\begin{align}\label{eq5}
+& \Pr\left(\bigwedge_{i=2}^n (\alpha^*_1 > \alpha^*_i) \mid C\right)\nonumber\\
+&= \prod_{i=2}^n \int_0^1 \int_0^1 \Phi_i(p_1,p_i) \cdot \operatorname{Beta}(p_1; 1+\alpha_1,1+\beta_1) \nonumber\\
+& \qquad\qquad\qquad \times \operatorname{Beta}(p_i; 1+\alpha_i,1+\beta_i) \mathrm{d}p_i \mathrm{d}p_1
+\end{align}
+
+We can numerically compute \eqref{eq2} and \eqref{eq5} given raw voting data.
+In order to be certain before taking any actions, we would like for the computed credibilities to be $\geq 99\%$.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/simple_majority}
+\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
+\label{simple_majority}
+\end{figure}
+%Data:
+%Curve 1: {{0,0.000299648},{1,0.00213083},{2,0.00817989},{3,0.0224777},{4,0.049512},{5,0.0928927},{6,0.154136},{7,0.232044},{8,0.322859},{9,0.421057},{10,0.520467},{11,0.615358},{12,0.701241},{13,0.775267},{14,0.836245},{15,0.884385},{16,0.920892},{17,0.947532},{18,0.966265},{19,0.978973},{20,0.987296},{21,0.992561},{22,0.99578},{23,0.997682},{24,0.998767},{25,0.999366}}
+%Curve 2: {x, 0.99} for all x
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/super_majority}
+\caption{Super Majority: $\alpha$ vs. $\Pr(\alpha > \frac{2}{3}N)$ for $N=100$, $\beta = 10$.}
+\label{super_majority}
+\end{figure}
+%Data:
+%Curve 1: {{10,0.0395173},{11,0.0591278},{12,0.0846131},{13,0.116407},{14,0.15463},{15,0.199057},{16,0.24911},{17,0.303899},{18,0.362273},{19,0.422909},{20,0.484399},{21,0.545345},{22,0.604445},{23,0.660559},{24,0.712764},{25,0.760382},{26,0.802985},{27,0.840385},{28,0.872612},{29,0.899875},{30,0.922519},{31,0.940987},{32,0.955777},{33,0.967407},{34,0.976384},{35,0.983186},{36,0.988243},{37,0.99193},{38,0.994567},{39,0.996414},{40,0.997682},{41,0.998533},{42,0.999093},{43,0.999452},{44,0.999677},{45,0.999815},{46,0.999896},{47,0.999944},{48,0.99997},{49,0.999985},{50,0.999993}}
+%Curve 2: {x, 0.99} for all x
+
+%\todo[inline]{Do we take into the account the distribution of votes? So if you have a very large standard deviation
+%we might want to prefer a decision less than some other decision with smaller standard deviation, but maybe
+%also smaller (but still positive) average. Is distribution of votes influencing dynamic quorum?}
+
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2017-04-27 statistical-quorum.tex

--- 2017-04-27 statistical-quorum.tex
@@
 \begin{align}\label{eq7}
-\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}}\\
+\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda}\\
 & \geq 95\% \nonumber
 \end{align}
+where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i)}{\ln{1/\gamma}}$.
 Upon solving for $\hat{K_i}$, we set $N_i = \alpha_i + \beta_i + \hat{K_i}$, and can proceed to check for a statistical quorum.
@@
 a building as a consequence of a decision?}
+
+\todo[inline]{Discuss how this might or might not lead to a more inclusive decision making: some people prefer it because
+they have time to collect thoughts and can participate when they have time, some people do not because they feel
+they cannot communicate as effectively, especially with lack of emotions being conveyed.}
 

### 2017-04-27 statistical-quorum.tex
some figures

--- 2017-04-27 statistical-quorum.tex
@@
 Because the sum of  the collected votes follows a binomial distribution, a commonplace and convenient choice for the form of the prior on the community's bias is the Beta distribution (due to its conjugacy with the binomial distribution).
+Together, they form the Beta-binomial distribution (related to the P\'olya urn model).
 
@@
 \begin{align}\label{eq7}
-\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}}\\
+\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda}\\
 & \geq 95\% \nonumber
 \end{align}
-Upon solving for $\hat{K_i}$, we set $N_i = \alpha_i + \beta_i + \hat{K_i}$, and can proceed to check for a statistical quorum.
+where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i)}{\ln{1/\gamma}}$.
+Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$, and can proceed to check for a statistical quorum.
 

### 2017-04-27 statistical-quorum.tex

--- 2017-04-27 statistical-quorum.tex
@@
 where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i)}{\ln{1/\gamma}}$.
-Upon solving for $\hat{K_i}$, we set $N_i = \alpha_i + \beta_i + \hat{K_i}$, and can proceed to check for a statistical quorum.
+Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$, and can proceed to check for a statistical quorum.
 

### 2017-04-28 statistical-quorum.tex
Improving secref macro.

@@
 
-\newcommand\secref[1]{\textit{\nameref{#1}}}
+\newcommand\secref[1]{section \textit{\nameref{#1}}}
 
@@
   and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
-  invalid votes. The last column shows computed decision credibility as described in Section \secref{sec:statistical-quorum}. We see that
+  invalid votes. The last column shows computed decision credibility as described in \secref{sec:statistical-quorum}. We see that
   credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
@@
 Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
-We describe decision credibility in more detail in Section \secref{sec:statistical-quorum}.
+We describe decision credibility in more detail in \secref{sec:statistical-quorum}.
 
@@
 which puts legitimacy of those decisions into a question.
-We discuss a way for a community to determine a reasonable statistical quorum in Section \secref{sec:discussion}.
+We discuss a way for a community to determine a reasonable statistical quorum in \secref{sec:discussion}.
 
@@
 the feedback loop needed for improving utilization of collective attention.
-We present the application in Section \secref{sec:peermind} and use it for evaluating statistical quorum in
-Section \secref{sec:evaluation}.
+We present the application in \secref{sec:peermind} and use it for evaluating statistical quorum in
+\secref{sec:evaluation}.
 

### 2017-04-28 statistical-quorum.tex
Small improvements.

@@
 % so you remain anonymous.
-\def\plaintitle{Using statistical quorum to improve legitimacy of decisions and utilization of collective attention in democratic communities}
+\def\plaintitle{Using decision credibility to improve legitimacy of decisions and utilization of collective attention in democratic communities}
 %\def\plainauthor{First Author, Second Author, Third Author,
@@
 
+%%% GLOSSARY %%%
+%
+% community = group which votes
+% proposal = what you vote on
+% decision = result of a vote, accepted or rejected proposal
+%
+%%% END GLOSSARY %%%
+
 % End of preamble. Here it comes the document.
@@
 
-Many democratic communities use voting to make collective decisions.
+Many communities use voting to make collective decisions.
 \todo{Should we support this statement with a reference?}
-In its basic form voting consist of members of a community casting a vote \emph{in favor} or \emph{against} a decision and if
-majority ($> 50\%$) of votes are in favor of the decision, the decision \emph{passes} and is adopted as a decision
+In its basic form, voting consists of members of a community casting a vote \emph{in favor} or \emph{against} a decision.
+If the majority ($> 50\%$) of votes are in favor of the decision, the decision \emph{passes} and is adopted as a decision
 of the whole community.
@@
 prevent decision to be adopted, even if those opposing it are in minority.
-Communities often decide quorum arbitrary\footnote{Robert's Rules of Order states that the quorum
+Communities often decide quorum arbitrarily\footnote{Robert's Rules of Order states that the quorum
 ``should approximate the largest number that can be depended on to attend any meeting except in very bad weather
@@
 hypothesis testing to evaluate the credibility of the voting result.
-It uses distribution of votes made to evaluate the probability of decision passing if everyone actually voted,
-and not just those who voted.
-Intuitively, if in a community of 100 members 40 members vote for a decision to pass and nobody against,
+It uses the distribution of votes made to evaluate the probability of a decision passing if everyone actually voted,
+and not just those who did vote.
+Intuitively, if, in a community of 100 members, 40 members vote for a decision to pass and nobody against,
 we can see that if everyone voted, the probability of majority of population voting for a decision is high.

### 2017-04-28 statistical-quorum.tex
small notes

--- 2017-04-27 statistical-quorum.tex
@@
 
+\todo[inline]{Alternative titles:\\
+A statistical quorum to (improve/guarantee) credibility and efficiency of decision-making within democratic communities}
+
 \begin{abstract}
-%The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
-%While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
-%different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
-%maximize productivity?
-%We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
-%attention, in turn making the entire decision-making process dynamic.
-%We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
-%councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
-%improve the legitimacy and efficiency of their decision-making process.
+The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
+While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
+different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
+maximize productivity?
+We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
+attention, in turn making the entire decision-making process dynamic.
+We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
+councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
+improve the legitimacy and efficiency of their decision-making process.
 \end{abstract}
@@
 
-\subsection{Properties of Statistical Quorum}
+\subsection{Theory}
 
-\todo[inline]{Theoretical evaluation, diagrams. Get readers to understand some of its properties visually.}
+\todo[inline]{Theoretical evaluation, diagrams. Get readers to understand some of its properties visually:\\
+4 Figures and 1 Table Pending}
 
@@
 \label{sec:derivation}
+
+\todo[inline]{Focus on only including bare minimum math, (no ``decorative math'').
+Emphasis on justifications and properties, minimal arithmetic.
+Good heuristic: Every line of math should have something said about it, otherwise it is unnecessary.}
+
 \subsection{Derivation of Eq. \eqref{eq2}}

### 2017-04-29 statistical-quorum.tex
TODOs

@@
 
+\todo[inline]{Something we need to think about is the level of detail we want to go into regarding the implementation/features of PeerMind beyond statistical quorum and range voting.
+I think we might want to focus on only those two features for this paper.\\
+I think PeerMind as a comprehensive platform should have its own paper where all of the data is analyzed across the board.
+This would be a much more typical HCI paper.
+(If I can become president of Ridge in a future semester, we'll have another chance to run a clean, well-designed experiment with a much more developed version of PeerMind.)\\
+The message in the current paper might be clearest if we keep the scope focused on the theory + evaluation/verification.  We should make sure we're not trying to pack too much into one paper.
+\\---Max}
+
 We built a full-fletched real-time web application for collective decision making and voting using Meteor
@@
 \subsection{Bayesian Theory}
+
+\todo[inline]{Weave in a single concrete example to ground all the abstract theory so that the entire derivation and motivation has a clear and cohesive thread from start to finish.}
+
 After collecting an initial set of votes from members, we want to use that data to construct a prior distribution to help estimate the posterior probability that the proposal is acceptable to the entire community.
@@
 
-\todo[inline]{Theoretical evaluation, diagrams. Get readers to understand some of its properties visually:\\
+\todo[inline]{Visualize the credibility and dynamics of the voting process under this framework:\\
 4 Figures and 1 Table Pending}
@@
 \subsection{Use in Existing Democratic Community}
+
+\todo[inline]{See my comment at the beginning of the PeerMind Web App section.
+I think the evaluation you note below about the interviews is out of the scope of this paper, which should be focused on statistical quorum and less online vs.
+offline platforms.\\
+---Max}
 
@@
 
-\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.}
+\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.\\
+Part of the answer here is honestly just that the admin should tune the threshold (and a few of the other hyperparameters we introduce) to their intuition.
+Once you set the parameters, the real benefit is that you have consistency, rather than having to always use your intuition as an admin about evaluating credibility.}
 
@@
 this might not really be possible, because decisions might not be reversible. You can revoke a law, but if you built
-a building as a consequence of a decision?}
+a building as a consequence of a decision?\\
+(See near/post-quorum validation phase idea in GitHub issues as one suggested solution.)}
 

### 2017-05-26 statistical-quorum.tex
improved wording in math section

@@
 \end{description}
-Now, for example, to upper bound with $95\%$ probability the remaining number of votes still pending, we solve for $\hat{K_i}$ to satisfy the following inequality:
+Now, for example, to upper bound with $95\%$ probability the remaining number of votes still pending, we wish to solve for $\hat{K_i}$ such that the following inequality is satisfied:
 \begin{align}\label{eq7}
-\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda}\\
-& \geq 95\% \nonumber
+\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda} \geq 95\%,
 \end{align}
 where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i)}{\ln{1/\gamma}}$.
-Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$, and can proceed to check for a statistical quorum.
+Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$, and can proceed to check for a statistical quorum using Eq. \ref{eq5}.
 

### 2017-09-03 statistical-quorum.tex
modifying constant used in model
Introduction.

@@
 \usepackage{microtype}      % Improved Tracking and Kerning
-% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
+\usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
 \usepackage{ccicons}        % Cite your images correctly!
-% \usepackage[utf8]{inputenc} % for a UTF8 editor only
+\usepackage[utf8]{inputenc} % for a UTF8 editor only
 
@@
 % so you remain anonymous.
-\def\plaintitle{Using decision credibility to improve legitimacy of decisions and utilization of collective attention in democratic communities}
+\def\plaintitle{Improving legitimacy and efficiency of decisions in democratic communities}
 %\def\plainauthor{First Author, Second Author, Third Author,
@@
 
-\title{\plaintitle}
+\title{Improving legitimacy and efficiency \\ of decisions in democratic communities}
 
@@
 
-\todo[inline]{Alternative titles:\\
-A statistical quorum to (improve/guarantee) credibility and efficiency of decision-making within democratic communities}
-
 \begin{abstract}
-The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
-While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
-different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
-maximize productivity?
-We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
-attention, in turn making the entire decision-making process dynamic.
-We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
-councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
-improve the legitimacy and efficiency of their decision-making process.
+%The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
+%While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
+%different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
+%maximize productivity?
+%We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
+%attention, in turn making the entire decision-making process dynamic.
+%We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
+%councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
+%improve the legitimacy and efficiency of their decision-making process.
+\todo[inline]{TODO}
 \end{abstract}
@@
 
-Many communities use voting to make collective decisions.
-\todo{Should we support this statement with a reference?}
-In its basic form, voting consists of members of a community casting a vote \emph{in favor} or \emph{against} a decision.
-If the majority ($> 50\%$) of votes are in favor of the decision, the decision \emph{passes} and is adopted as a decision
-of the whole community.
+In many historical and recent elections it is not infrequent that we get close results.
+Especially when this is combined with low voter turnout, how can we have confidence in results?
+What if elections would be a day longer? What if they would be on a different day? What if
+more people would participate? Would results still be the same?
 
-There have been many improvements proposed to this basic form and various analysis made.
-\todo{Add references.}
-In this paper we focus on the question of \emph{quorum} and how high the quorum should be.
-Quorum is the minimum number of members participating in voting for voting to be seen as representative of the whole
-community, i.e., legitimate.
-If quorum is too low, decisions adopted might not represent the community.
-If it is too high, the community might have issues achieving the quorum.
-Not enough members might decide to participate in voting, or members might on purpose boycott voting to
-prevent decision to be adopted, even if those opposing it are in minority.
-Communities often decide quorum arbitrarily\footnote{Robert's Rules of Order states that the quorum
-``should approximate the largest number that can be depended on to attend any meeting except in very bad weather
-or other extremely unfavorable conditions.''~\cite{roberts}}
-and try to increase participation of members in communities' democratic processes through various means
-to address the issue of too low turnout.
-\todo{Add references. Mention forced participation in some countries.}
-Moreover, many democratic communities have issues achieving quorum which hinders their operation.
-\todo{Add references.}
+An underlying question is what should we do when not everyone in the population votes.
 
-Even if voting turnout has reached legally required quorum, or when one is not defined,
-it is still a question how to evaluate if the turnout is high enough to find results legitimate.
-As an example we can look at two recent votes, Brexit, a 2016 referendum for United Kingdom's withdrawal
-from the European Union, and popular vote in 2016 USA presidential elections, as seen in Table~\ref{tab:example}.
-Ignoring particularities of USA presidential elections process and using a popular vote as an example,
-how can we reason about legitimacy of results?
+Traditionally, we can have a stance that everyone had an opportunity to vote and express
+their preference, so whatever are results based those who voted are legitimate results.
+Research shows, though, that there are many reasons why people do not vote and that there
+are many issues related to how elections are organized, i.e., are they held on working
+or work-free day.
 
-\begin{table*}
-  \centering
-  \begin{tabular}{l r r r r r r r}
-    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Decision credibility}} \\
-    \midrule
-    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
-    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
-  \end{tabular}
-  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
-  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
-  invalid votes. The last column shows computed decision credibility as described in \secref{sec:statistical-quorum}. We see that
-  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
-  with more votes any option could end up winning.}\label{tab:example}
-\end{table*}
+Alternatively, we can see voting as a survey and cast votes as a sample, although imperfect
+and non-randomly selected, from a larger population, with a goal to infer the preference
+of the whole population. In this setting we can use statistical methods to evaluate the
+confidence we can have in the results.
 
-To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
-hypothesis testing to evaluate the credibility of the voting result.
-It uses the distribution of votes made to evaluate the probability of a decision passing if everyone actually voted,
-and not just those who did vote.
-Intuitively, if, in a community of 100 members, 40 members vote for a decision to pass and nobody against,
-we can see that if everyone voted, the probability of majority of population voting for a decision is high.
-On the other hand, if 41 members for in favor and 40 against, we cannot really say much about what would a decision
-be if everyone voted.
-Only two more votes against can completely change the outcome.
-Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
-We describe decision credibility in more detail in \secref{sec:statistical-quorum}.
+Our contributions are as follows:
 
-Based on decision credibility we can define a \emph{statistical quorum} as a threshold which a decision's credibility
-has to pass for a decision to be seen as legitimate.
-For example, if we adopt $90\%$ as a reasonable requirement for statistical quorum and we compute decision
-credibility for votes in Table~\ref{tab:example} we see that in both cases we do not reach statistical quorum
-which puts legitimacy of those decisions into a question.
-We discuss a way for a community to determine a reasonable statistical quorum in \secref{sec:discussion}.
+\begin{itemize}
+\item We developed a framework of computing statistical credibility of results
+based on sample size, votes cast, and voting duration.
+\item Based on this framework we developed a concept of \emph{statistical quorum}
+as a threshold of votes required to deem results legitimate.
+\item We implemented an online application \emph{PeerMind},
+integrating credibility of results and statistical quorum, and designed
+a workflow of its use in a decision making processes.
+\item We deployed the application in a democratically run community of 140
+members for 5 months.
+\item \todo[inline]{We did this analysis and found these results. Legitimacy. Efficiency. Attention.}
+\end{itemize}
 
-Traditionally it is argued that legitimacy comes from the fact that members had a chance to participate
-and they decided not to, giving legitimacy to make a decision to those who did participate.
-We argue that this is a problematic view because there are many restrictions on why members could not
-participate and vote.
-For example, they might not had time to participate because of other commitments, or motivation, or they might determined
-that they are not informed enough to participate.
-Moreover, members are from different reasons in their communities and potentially contribute in other ways
-than by participating in its democratic processes.
-Instead of assuming that everyone could participate if they would want, or trying to increase participation,
-we can see amount of participation members are able to do as a limited resource, a \emph{collective attention}
-a community has available to utilize for its decision making.
+\todo[inline]{Structure of the paper.}
 
-We can compute decision credibility during ongoing voting process to give feedback to the community
-on how contentious a proposed decision is how much more votes are necessary for a statistical quorum
-to be reached.
-This allows the community to focus its collective attention towards decisions which require more votes
-to determine a legitimate result, while reducing attention towards proposals which already have overwhelming support.
-Instead of trying to increase participation on all democratic decisions being made, a community can
-do that selectively through an automatic feedback loop, trying to reach statistical quorum for their decisions.
-
-We built an open source web application PeerMind integrating statistical quorum with online voting.
-The application exposes computed decision credibility to users in real-time and through that provides
-the feedback loop needed for improving utilization of collective attention.
-We present the application in \secref{sec:peermind} and use it for evaluating statistical quorum in
-\secref{sec:evaluation}.
-
-\todo[inline]{Should we explicitly list our contributions again?}
-
-\todo[inline]{Repeat the structure of the rest of the paper? Is there a better way to reference sections than
-referencing it by name?}
+%
+%
+%Many communities use voting to make collective decisions.
+%\todo{Should we support this statement with a reference?}
+%In its basic form, voting consists of members of a community casting a vote \emph{in favor} or \emph{against} a decision.
+%If the majority ($> 50\%$) of votes are in favor of the decision, the decision \emph{passes} and is adopted as a decision
+%of the whole community.
+%
+%There have been many improvements proposed to this basic form and various analysis made.
+%\todo{Add references.}
+%In this paper we focus on the question of \emph{quorum} and how high the quorum should be.
+%Quorum is the minimum number of members participating in voting for voting to be seen as representative of the whole
+%community, i.e., legitimate.
+%If quorum is too low, decisions adopted might not represent the community.
+%If it is too high, the community might have issues achieving the quorum.
+%Not enough members might decide to participate in voting, or members might on purpose boycott voting to
+%prevent decision to be adopted, even if those opposing it are in minority.
+%Communities often decide quorum arbitrarily\footnote{Robert's Rules of Order states that the quorum
+%``should approximate the largest number that can be depended on to attend any meeting except in very bad weather
+%or other extremely unfavorable conditions.''~\cite{roberts}}
+%and try to increase participation of members in communities' democratic processes through various means
+%to address the issue of too low turnout.
+%\todo{Add references. Mention forced participation in some countries.}
+%Moreover, many democratic communities have issues achieving quorum which hinders their operation.
+%\todo{Add references.}
+%
+%Even if voting turnout has reached legally required quorum, or when one is not defined,
+%it is still a question how to evaluate if the turnout is high enough to find results legitimate.
+%As an example we can look at two recent votes, Brexit, a 2016 referendum for United Kingdom's withdrawal
+%from the European Union, and popular vote in 2016 USA presidential elections, as seen in Table~\ref{tab:example}.
+%Ignoring particularities of USA presidential elections process and using a popular vote as an example,
+%how can we reason about legitimacy of results?
+%
+%\begin{table*}
+%  \centering
+%  \begin{tabular}{l r r r r r r r}
+%    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Decision credibility}} \\
+%    \midrule
+%    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
+%    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
+%  \end{tabular}
+%  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
+%  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
+%  invalid votes. The last column shows computed decision credibility as described in \secref{sec:decision-credibility}. We see that
+%  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
+%  with more votes any option could end up winning.}\label{tab:example}
+%\end{table*}
+%
+%To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
+%hypothesis testing to evaluate the credibility of the voting result.
+%It uses the distribution of votes made to evaluate the probability of a decision passing if everyone actually voted,
+%and not just those who did vote.
+%Intuitively, if, in a community of 100 members, 40 members vote for a decision to pass and nobody against,
+%we can see that if everyone voted, the probability of majority of population voting for a decision is high.
+%On the other hand, if 41 members for in favor and 40 against, we cannot really say much about what would a decision
+%be if everyone voted.
+%Only two more votes against can completely change the outcome.
+%Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
+%We describe decision credibility in more detail in \secref{sec:decision-credibility}.
+%
+%Based on decision credibility we can define a \emph{statistical quorum} as a threshold which a decision's credibility
+%has to pass for a decision to be seen as legitimate.
+%For example, if we adopt $90\%$ as a reasonable requirement for statistical quorum and we compute decision
+%credibility for votes in Table~\ref{tab:example} we see that in both cases we do not reach statistical quorum
+%which puts legitimacy of those decisions into a question.
+%We discuss a way for a community to determine a reasonable statistical quorum in \secref{sec:discussion}.
+%
+%Traditionally it is argued that legitimacy comes from the fact that members had a chance to participate
+%and they decided not to, giving legitimacy to make a decision to those who did participate.
+%We argue that this is a problematic view because there are many restrictions on why members could not
+%participate and vote.
+%For example, they might not had time to participate because of other commitments, or motivation, or they might determined
+%that they are not informed enough to participate.
+%Moreover, members are from different reasons in their communities and potentially contribute in other ways
+%than by participating in its democratic processes.
+%Instead of assuming that everyone could participate if they would want, or trying to increase participation,
+%we can see amount of participation members are able to do as a limited resource, a \emph{collective attention}
+%a community has available to utilize for its decision making.
+%
+%We can compute decision credibility during ongoing voting process to give feedback to the community
+%on how contentious a proposed decision is how much more votes are necessary for a statistical quorum
+%to be reached.
+%This allows the community to focus its collective attention towards decisions which require more votes
+%to determine a legitimate result, while reducing attention towards proposals which already have overwhelming support.
+%Instead of trying to increase participation on all democratic decisions being made, a community can
+%do that selectively through an automatic feedback loop, trying to reach statistical quorum for their decisions.
+%
+%We built an open source web application PeerMind integrating statistical quorum with online voting.
+%The application exposes computed decision credibility to users in real-time and through that provides
+%the feedback loop needed for improving utilization of collective attention.
+%We present the application in \secref{sec:peermind} and use it for evaluating statistical quorum in
+%\secref{sec:evaluation}.
+%
+%\todo[inline]{Should we explicitly list our contributions again?}
+%
+%\todo[inline]{Repeat the structure of the rest of the paper? Is there a better way to reference sections than
+%referencing it by name?}
 
@@
 
-\todo[inline]{Repeat references from "There have been many improvements proposed to this basic form and various analysis made."?}
-
-\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
-How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
-
-\todo[inline]{Voting auditing work.}
-
-\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
-
-\section{PeerMind Web Application}
-\label{sec:peermind}
-
-\todo[inline]{Something we need to think about is the level of detail we want to go into regarding the implementation/features of PeerMind beyond statistical quorum and range voting.
+%\todo[inline]{Repeat references from "There have been many improvements proposed to this basic form and various analysis made."?}
+%
+%\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
+%How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
+%
+%\todo[inline]{Voting auditing work.}
+%
+%\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
+%
+%\section{PeerMind Web Application}
+%\label{sec:peermind}
+%
+%\todo[inline]{Something we need to think about is the level of detail we want to go into regarding the implementation/features of PeerMind beyond statistical quorum and range voting.
 I think we might want to focus on only those two features for this paper.\\
-I think PeerMind as a comprehensive platform should have its own paper where all of the data is analyzed across the board.
+%I think PeerMind as a comprehensive platform should have its own paper where all of the data is analyzed across the board.
 This would be a much more typical HCI paper.
 (If I can become president of Ridge in a future semester, we'll have another chance to run a clean, well-designed experiment with a much more developed version of PeerMind.)\\
-The message in the current paper might be clearest if we keep the scope focused on the theory + evaluation/verification.  We should make sure we're not trying to pack too much into one paper.
-\\---Max}
+%The message in the current paper might be clearest if we keep the scope focused on the theory + evaluation/verification.  We should make sure we're not trying to pack too much into one paper.
+%\\---Max}
+%
+%We built a full-fletched real-time web application for collective decision making and voting using Meteor
+%framework~\cite{meteor}.
+%Users can initiate new discussions which leads them to the main user interface for each discussion as seen
+%in Figure~\ref{fig:peermind}.
+%
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=1.75\columnwidth]{figures/peermind.png}
+%  \caption{The main PeerMind interface of a discussion. There is initial discussion description on the top and under it there
+%  are three columns. A commenting sections for users to discuss it and proposed decisions, a second column where moderators can
+%  summarize arguments made during discussion, and a list of proposals to address the discussion.
+%  Proposals can be in different states: a proposal being discussed, a proposal being voted on, and with voting concluded.}\label{fig:peermind}
+%\end{figure*}
+%
+%%and discuss them with others online in real-time and with rich-text content.
+%%They can attach images and files using drag\&drop. Moderators can summarize discussions into three groups
+%
+%\todo[inline]{Present how users can see ongoing results during voting. Show an image of active ballot, but with expanded view for current results.}
+%
+%%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
+%%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
+%%A proposal passes if the average score of all votes on a position is positive.
+%%When multiple competing proposals are being voted on, the one with highest average score wins.
+%%
+%%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
+%%positions in a highly informative and precise manner.
+%%We find it especially suitable for a web application because many web users are used to use it when ranking products
+%%and giving other ``five-star'' feedback.
+%%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
+%%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
+%%each proposal would need reassessing (assuming members use a consistent personal scale when voting).
+%
+%\todo[inline]{The app has features both for use in online and offline setting. This allows one to compute
+%statistical quorum also for an in-person meeting because math is now more complicated than just counting hands.
+%We focus on online experience in this paper.}
+%
+%\todo[inline]{Security is out of the scope of this work. We assume here that users and their identities are known and are managed
+%by a community's admin. There are many communities where their decision making is public with public minutes and this is who
+%the app is currently targeting.}
 
-We built a full-fletched real-time web application for collective decision making and voting using Meteor
-framework~\cite{meteor}.
-Users can initiate new discussions which leads them to the main user interface for each discussion as seen
-in Figure~\ref{fig:peermind}.
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=1.75\columnwidth]{figures/peermind.png}
-  \caption{The main PeerMind interface of a discussion. There is initial discussion description on the top and under it there
-  are three columns. A commenting sections for users to discuss it and proposed decisions, a second column where moderators can
-  summarize arguments made during discussion, and a list of proposals to address the discussion.
-  Proposals can be in different states: a proposal being discussed, a proposal being voted on, and with voting concluded.}\label{fig:peermind}
-\end{figure*}
-
-%and discuss them with others online in real-time and with rich-text content.
-%They can attach images and files using drag\&drop. Moderators can summarize discussions into three groups
-
-\todo[inline]{Present how users can see ongoing results during voting. Show an image of active ballot, but with expanded view for current results.}
-
-%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
-%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
-%A proposal passes if the average score of all votes on a position is positive.
-%When multiple competing proposals are being voted on, the one with highest average score wins.
-%
-%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
-%positions in a highly informative and precise manner.
-%We find it especially suitable for a web application because many web users are used to use it when ranking products
-%and giving other ``five-star'' feedback.
-%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
-%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
-%each proposal would need reassessing (assuming members use a consistent personal scale when voting).
-
-\todo[inline]{The app has features both for use in online and offline setting. This allows one to compute
-statistical quorum also for an in-person meeting because math is now more complicated than just counting hands.
-We focus on online experience in this paper.}
-
-\todo[inline]{Security is out of the scope of this work. We assume here that users and their identities are known and are managed
-by a community's admin. There are many communities where their decision making is public with public minutes and this is who
-the app is currently targeting.}
-
-\section{Statistical Quorum}
-\label{sec:statistical-quorum}
+\section{Decision Credibility}
+\label{sec:decision-credibility}
 A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
@@
 \begin{align*}
-\nu_i & \sim \operatorname{Beta-Binomial}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}, K_i),\\
-{CDF}_{\nu_i}(V_i) & = \Pr(\nu_i \leq V_i),
+&\nu_i \sim \operatorname{Beta-Binomial}(1 + \frac{\alpha_i}{K_i}c, 1 + \frac{\beta_i}{K_i}c, K_i),\\
+&{CDF}_{\nu_i}(V_i)  = \Pr(\nu_i \leq V_i),
 \end{align*}
+\todo[inline]{Note that we choose $c=10$ because it leads to intuitive results.}
 where
@@
 
-Hence we arrive at $\eta = \frac{c}{K_i}$, where $c$ is any constant that can be selected to tune the rate at which the variance approaches $0$.  For simplicity, we use $\eta = \frac{1}{K_i}$.
+Hence we arrive at $\eta = \frac{c}{K_i}$, where $c$ is any constant that can be selected to tune the rate at which the variance approaches $0$.  In this study, we prefer $\eta = \frac{10}{K_i}$.
 

### 2017-09-04 statistical-quorum.tex
Write it out.
Even more things.

@@
 members for 5 months.
-\item \todo[inline]{We did this analysis and found these results. Legitimacy. Efficiency. Attention.}
+\item We show how use of statistical quorum increases legitimacy of decisions,
+improves utilization of collective attention, increases overall participation,
+and increases number of decisions made while not sacrificing their legitimacy.
 \end{itemize}

### 2017-09-19 statistical-quorum.tex
Adding TODOs.
Front page.

--- 2017-09-03 statistical-quorum.tex
@@
 %  Fourth Author, Fifth Author, Sixth Author}
-%\def\emptyauthor{}
-%\def\plainkeywords{Authors' choice; of terms; separated; by
-%  semicolons; include commas, within terms only; required.}
+\def\emptyauthor{}
+\def\plainkeywords{Decision making; voting; democracy; quorum; legitimacy; statistical quorum.}
 %\def\plaingeneralterms{Documentation, Standardization}
@@
 \begin{abstract}
-%The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
-%While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
-%different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
-%maximize productivity?
-%We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
-%attention, in turn making the entire decision-making process dynamic.
-%We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
-%councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
-%improve the legitimacy and efficiency of their decision-making process.
-\todo[inline]{TODO}
+The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
+While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
+different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
+maximize productivity?
+We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to focus
+attention, in turn making the entire decision-making process dynamic.
+We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
+councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
+improve the legitimacy and efficiency of their decision-making process.
 \end{abstract}
 
-%\category{H.5.m.}{Information Interfaces and Presentation
-%  (e.g. HCI)}{Miscellaneous} \category{See
-%  \url{http://acm.org/about/class/1998/} for the full list of ACM
-%  classifiers. This section is required.}{}{}
-%
-%\keywords{\plainkeywords}
+\category{H.5.3.}{Information Interfaces and Presentation
+ (e.g. HCI)}{Computer-supported cooperative work}
+
+\keywords{\plainkeywords}
 
@@
 Traditionally, we can have a stance that everyone had an opportunity to vote and express
-their preference, so whatever are results based those who voted are legitimate results.
+their preference, so whatever are results based on those who voted are legitimate results.
 Research shows, though, that there are many reasons why people do not vote and that there
-are many issues related to how elections are organized, i.e., are they held on working
-or work-free day.
+are many issues related to how elections are organized~\cite{CANCELA2016264, GIMPEL2003471}.
 
@@
 
+%Moreover, we can see overall participation by everyone in the population as a
+%limited resource, a collective attention, we have at our disposal to make collective decisions.
+%Traditionally, a common solution to the low voter turnout is to focus on
+%increasing the turnout and through that increasing the collective attention available
+%for decision making.
+%But increasing collective attention for decision making is at odds with other
+%demands for the collective and individual attention.
+%But once we have a statistical measure of results, we can utilize it to
+%optimize the allocation of the collective attention by directing participation towards
+%contentious issues.
+
 Our contributions are as follows:
@@
 based on sample size, votes cast, and voting duration.
-\item Based on this framework we developed a concept of \emph{statistical quorum}
+\item Based on this framework we developed a novel concept of \emph{statistical quorum}
 as a threshold of votes required to deem results legitimate.
@@
 \item We deployed the application in a democratically run community of 140
-members for 5 months.
-\item \todo[inline]{We did this analysis and found these results. Legitimacy. Efficiency. Attention.}
+members for 16 weeks.
+\item Using the analysis of decision making process before and after introduction
+of the application, using surveys, and using semi-structured interviews, we show
+how use of statistical quorum increases legitimacy of decisions,
+improves utilization of collective attention,
+and increases number of decisions made while not sacrificing their legitimacy.
 \end{itemize}
 
-\todo[inline]{Structure of the paper.}
+The rest of the paper is structured as follows. In Section~\ref{sec:related-work} we
+present related work. We continue by describing a framework of computing statistical
+credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
+We present our methodology in Section~\ref{sec:methodology} after which we do evaluation in
+Section~\ref{sec:evaluation}. We discuss results in Section~\ref{sec:discussion} and conclude
+in Section~\ref{sec:conclusion}. At the end we have Appendix where we
+derive formulas presented in Section~\ref{sec:decision-credibility}.
 
@@
 \section{Related Work}
+\label{sec:related-work}
 
@@
 
-\section{Decision Credibility}
+\section{Decision Credibility and Statistical Quorum}
 \label{sec:decision-credibility}
@@
 
+\section{Methodology}
+\label{sec:methodology}
+
 \section{Evaluation}
@@
 
-\section{Future Work}
-
-\todo[inline]{More work on security evaluation of potential attacks when a community is using statistical quorum to force a decision.
-Is there a game theoretical approach to it?}
-
-\todo[inline]{How to surface information about credibility to users better than just a number? Visualizations? How can we help them
-understand what exactly this is and how the math works?}
+%\section{Future Work}
+%
+%\todo[inline]{More work on security evaluation of potential attacks when a community is using statistical quorum to force a decision.
+%Is there a game theoretical approach to it?}
+%
+%\todo[inline]{How to surface information about credibility to users better than just a number? Visualizations? How can we help them
+%understand what exactly this is and how the math works?}
 
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2017-09-19 statistical-quorum.tex
Survey data.

@@
 how use of statistical quorum increases legitimacy of decisions,
-improves utilization of collective attention,
+improves utilization of collective attention, increases overall participation,
 and increases number of decisions made while not sacrificing their legitimacy.
@@
 
+To evaluate statistical quorum in practice, we developed an online web application PeerMind and
+used it in a democratically run community of 140 members for 16 weeks, replacing the traditional
+decision making through in-person meetings. The community is a house of 140
+students living together and making collective decisions about the governance and operation of
+the house, use of resources like common spaces and budgets, and changes to the house itself.
+
+Before the introduction of the application, community had a practice of meeting weekly, with all
+members being invited to participate. For decision making it used variation of Robert's rules~\cite{roberts}
+with majority and super-majority voting to adopt decisions, generally using hand voting and voice voting.
+A fixed quorum of the third of the membership was used. Meetings generally lasted between one to three hours.
+Public minutes were taken of those meetings.
+
+We introduced the application in fall 2016 semester for 16 weeks. We use spring 2016 as a baseline for comparison,
+16 weeks as well, when traditional decision making was used.
+
+To evaluate general perception of members before and after the use of the application for decision making,
+we did a general survey at the end of each semester, asking members to answer general question about how
+they felt the decision making in the house was that semester.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
+\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making in the house.
+For each question, a participant was asked to rank on a scale between 1 (completely disagree) and 7 (completely agree). All members
+of the community were invited to participate in the survey.}
+\label{fig:survey-comparison}
+\end{figure}
+
 \section{Evaluation}

### 2017-09-19 statistical-quorum.tex
Cleaning the language about the survey.

@@
 
-Before the introduction of the application, community had a practice of meeting weekly, with all
+Before the introduction of the application, community had a practice of weekly meetings, councils, with all
 members being invited to participate. For decision making it used variation of Robert's rules~\cite{roberts}
 with majority and super-majority voting to adopt decisions, generally using hand voting and voice voting.
-A fixed quorum of the third of the membership was used. Meetings generally lasted between one to three hours.
-Public minutes were taken of those meetings.
+A fixed quorum of $30\%$ of the membership was used (42 members). Councils generally lasted between one to three hours.
+Public minutes were taken.
 
-We introduced the application in fall 2016 semester for 16 weeks. We use spring 2016 as a baseline for comparison,
-16 weeks as well, when traditional decision making was used.
+We introduced the application in fall 2016 semester which lasted 16 weeks. We use spring 2016 semester as
+a baseline for comparison, which lasted 16 weeks as well and traditional decision making was used.
 
@@
 \includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
-\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making in the house.
+\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
 For each question, a participant was asked to rank on a scale between 1 (completely disagree) and 7 (completely agree). All members

### 2017-09-19 statistical-quorum.tex
Methodology section done.

@@
 attention, in turn making the entire decision-making process dynamic.
-We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
-councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
+We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
+councils over the course of 16 weeks.  We found that the community was able to use the web application to simultaneously
 improve the legitimacy and efficiency of their decision-making process.
@@
 as a threshold of votes required to deem results legitimate.
-\item We implemented an online application \emph{PeerMind},
+\item We implemented a web application \emph{PeerMind},
 integrating credibility of results and statistical quorum, and designed
@@
 
-To evaluate statistical quorum in practice, we developed an online web application PeerMind and
+To evaluate statistical quorum in practice, we developed a web application PeerMind and
 used it in a democratically run community of 140 members for 16 weeks, replacing the traditional
@@
 To evaluate general perception of members before and after the use of the application for decision making,
-we did a general survey at the end of each semester, asking members to answer general question about how
-they felt the decision making in the house was that semester.
+we did a general survey at the end of each semester, asking members to answer general questions about the
+decision making in the house that semester.
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
-\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
-For each question, a participant was asked to rank on a scale between 1 (completely disagree) and 7 (completely agree). All members
-of the community were invited to participate in the survey.}
-\label{fig:survey-comparison}
-\end{figure}
+We instrumented the application to collect information on how much time users spent using the application,
+measuring the time the application was focused. Other information we used for evaluation are available
+through regular features of the application: how many and who made discussion items, comments, motions
+(proposed decisions on discussion items), and vote tallies per motion.
+
+For the baseline, we extracted similar information from minutes: how long did the councils last, how many
+discussions were opened, how many comments by which members were made, how many motions were proposed,
+and vote tallies for them.
+
+Finally, we did semi-structured interviews at the end of the fall semester with 7 members who were participating
+both in spring and fall semesters and ask them to compare their experience between those two semesters and
+evaluate the application and the use of statistical quorum.
 
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
+\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
+For each question, a participant was asked to rank on a scale between 1 (completely disagree) and 7 (completely agree). All members
+of the community were invited to participate in the survey each time.}
+\label{fig:survey-comparison}
+\end{figure}
+
 \subsection{Theory}
@@
 \subsection{Use in Existing Democratic Community}
-
-\todo[inline]{See my comment at the beginning of the PeerMind Web App section.
-I think the evaluation you note below about the interviews is out of the scope of this paper, which should be focused on statistical quorum and less online vs.
-offline platforms.\\
----Max}
 

### 2017-09-19 statistical-quorum.tex
PeerMind.

@@
 
-Our contributions are as follows:
+Our contributions are:
 
@@
 
-For the baseline, we extracted similar information from minutes: how long did the councils last, how many
+For the baseline, we extracted similar information from council minutes: how long did the councils last, how many
 discussions were opened, how many comments by which members were made, how many motions were proposed,
@@
 
-Finally, we did semi-structured interviews at the end of the fall semester with 7 members who were participating
+Finally, we did semi-structured interviews after the fall semester with 7 members who were participating
 both in spring and fall semesters and ask them to compare their experience between those two semesters and
 evaluate the application and the use of statistical quorum.
+
+\subsection{PeerMind}
+\label{sec:peermind}
+
+Web application was developed using Meteor web framework~\cite{meteor} as a mobile-friendly web application.
+Meteor framework supports reactivity which makes the application update user interface based on actions
+of other users in real time. This helps users see comments by other users as they happen, but can also see
+progress during voting in real time.
+
+Each user can open and propose new discussion items. Each discussion item has then three basic sections,
+as seen in Figure~\ref{fig:peer-mind}: commenting section on the left, summary points in the middle,
+and proposed motions with voting on the right. All content inputs fields accept rich text content in WYSWYG manner,
+so users can easily add links, images, and other files.
+
+A general workflow after creation of a discussion item is that users discuss it through comments,
+which moderators summarize into summary points. At some point an user proposes a motion which can be
+further discussed through comments. Once a moderator deems that motion is ready to be voted upon
+can open voting. Users vote through a ballot seen in Figure~\ref{fig:ballot}.
+
+An important design decision in the scope of this paper is that during voting, users can see the current
+results of the vote in real time. Results are shown on a chart depicting average score through time
+as seen in Figure~\ref{fig:results}. Users can see both the current average score of all votes (score
+larger than 0 means that a motion is passing) and the credibility of current result. A community can decide
+how high credibility it requires to deem the result passing statistical quorum. During our evaluation
+we used 0.90 as statistical quorum.
+
+The rationale for displaying the current results during voting is to give users feedback how the vote
+is progressing and allow them to engage more with the discussion. Moreover, displaying the results
+also serves as additional safeguard against potential manipulations of statistical quorum through malicious
+coordination.
+
+After the vote passes the statistical quorum, a moderator can close voting and a discussion item.
+
+Web application contains other features which makes it easier to use as a decision making and discusion
+platform, like notifications, and additional features which allow the application to be used also to
+help run in-person meetings to discuss discussion items made
+online. Facilitators of in-person meetings can use it to prepare agendas for meetings, take notes,
+and display summary points of current discussion and what motions are being voted on. Through that the
+application can also serve as way to integrate both online and in-person decision making. Having all of
+decision making in one platform serves as an important resource for institutional memory.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/peermind.png}
+\caption{User interface of PeerMind web application. One can see three column structure, with commenting
+section on the left, summary points in the middle, and proposed motions with voting on the right.}
+\label{fig:peer-mind}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/ballot.png}
+\caption{Ballot used for voting in PeerMind web application. Users can chose a value between -1 and 1 on a slider,
+where -1 means opposition, 1 means in favor, and 0 indifference. They can also chose abstention (they
+remove themselves from the population of the community for the vote in question), or to not vote (default), which
+means that their vote does not contribute to the quorum.}
+\label{fig:ballot}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/results.png}
+\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of the vote.
+A user can hover over the chart to see details at that time. Users can see these results in real time during voting.}
+\label{fig:results}
+\end{figure}
 

### 2017-09-19 statistical-quorum.tex
Surveys evaluation.

@@
 
+PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
+anonymized dataset of interactions with the app, including votes, is available at \url{http://redacted.for.blind.review}.
+
 \begin{figure}[ht]
@@
 
-We evaluate statistical quorum in two ways. First, we evaluate it theoretically and explore its mathematical properties.
+We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
 Then we evaluate it through real-world use of the PeerMind application by a 140 members community.
+
+\subsection{Analysis}
+
+\todo[inline]{Visualize the credibility and dynamics of the voting process under this framework:\\
+4 Figures and 1 Table Pending}
+
+\subsection{Survey}
+
+In Figure~\ref{fig:survey-comparison} we see questions and results of the survey we made at the end of each semester.
+
+We can see that in fall members were more satisfied with the decision making in the community and only how understandable
+is the whole process received a decrease. We attribute this to the fact that the use of the web application is
+a novelty and requires more effort in educating members than a traditional decision making. Moreover, decision
+making in person allows one to observe others how they are participating and learn from them.
 
@@
 
-\subsection{Theory}
-
-\todo[inline]{Visualize the credibility and dynamics of the voting process under this framework:\\
-4 Figures and 1 Table Pending}
-
-\subsection{Use in Existing Democratic Community}
+\subsection{Usage comparison}
 
@@
 
-PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
-anonymized dataset of interactions with the app, including votes, is available at \url{http://redacted.for.blind.review}.
+\subsection{Interviews}
 

### 2017-09-19 statistical-quorum.tex
Updating the screenshot.

@@
 a novelty and requires more effort in educating members than a traditional decision making. Moreover, decision
-making in person allows one to observe others how they are participating and learn from them.
+making in person allows one to observe others how they are participating and learn from them, instead of
+being alone using the application.
 
@@
 \caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
-For each question, a participant was asked to rank on a scale between 1 (completely disagree) and 7 (completely agree). All members
+For each question, a participant was asked to rank the answer on a scale between 1 (completely disagree) and 7 (completely agree). All members
 of the community were invited to participate in the survey each time.}

### 2017-09-19 statistical-quorum.tex
Moving things around.

@@
 
-%
-%
-%Many communities use voting to make collective decisions.
-%\todo{Should we support this statement with a reference?}
-%In its basic form, voting consists of members of a community casting a vote \emph{in favor} or \emph{against} a decision.
-%If the majority ($> 50\%$) of votes are in favor of the decision, the decision \emph{passes} and is adopted as a decision
-%of the whole community.
-%
-%There have been many improvements proposed to this basic form and various analysis made.
-%\todo{Add references.}
-%In this paper we focus on the question of \emph{quorum} and how high the quorum should be.
-%Quorum is the minimum number of members participating in voting for voting to be seen as representative of the whole
-%community, i.e., legitimate.
-%If quorum is too low, decisions adopted might not represent the community.
-%If it is too high, the community might have issues achieving the quorum.
-%Not enough members might decide to participate in voting, or members might on purpose boycott voting to
-%prevent decision to be adopted, even if those opposing it are in minority.
-%Communities often decide quorum arbitrarily\footnote{Robert's Rules of Order states that the quorum
-%``should approximate the largest number that can be depended on to attend any meeting except in very bad weather
-%or other extremely unfavorable conditions.''~\cite{roberts}}
-%and try to increase participation of members in communities' democratic processes through various means
-%to address the issue of too low turnout.
-%\todo{Add references. Mention forced participation in some countries.}
-%Moreover, many democratic communities have issues achieving quorum which hinders their operation.
-%\todo{Add references.}
-%
-%Even if voting turnout has reached legally required quorum, or when one is not defined,
-%it is still a question how to evaluate if the turnout is high enough to find results legitimate.
-%As an example we can look at two recent votes, Brexit, a 2016 referendum for United Kingdom's withdrawal
-%from the European Union, and popular vote in 2016 USA presidential elections, as seen in Table~\ref{tab:example}.
-%Ignoring particularities of USA presidential elections process and using a popular vote as an example,
-%how can we reason about legitimacy of results?
-%
+% Security is out of the scope of this work. We assume here that users and their identities are known and are managed
+% by a community's admin. There are many communities where their decision making is public with public minutes and this
+% is who the app is currently targeting.
+
 %\begin{table*}
@@
 %\end{table*}
-%
-%To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
-%hypothesis testing to evaluate the credibility of the voting result.
-%It uses the distribution of votes made to evaluate the probability of a decision passing if everyone actually voted,
-%and not just those who did vote.
-%Intuitively, if, in a community of 100 members, 40 members vote for a decision to pass and nobody against,
-%we can see that if everyone voted, the probability of majority of population voting for a decision is high.
-%On the other hand, if 41 members for in favor and 40 against, we cannot really say much about what would a decision
-%be if everyone voted.
-%Only two more votes against can completely change the outcome.
-%Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
-%We describe decision credibility in more detail in \secref{sec:decision-credibility}.
-%
-%Based on decision credibility we can define a \emph{statistical quorum} as a threshold which a decision's credibility
-%has to pass for a decision to be seen as legitimate.
-%For example, if we adopt $90\%$ as a reasonable requirement for statistical quorum and we compute decision
-%credibility for votes in Table~\ref{tab:example} we see that in both cases we do not reach statistical quorum
-%which puts legitimacy of those decisions into a question.
-%We discuss a way for a community to determine a reasonable statistical quorum in \secref{sec:discussion}.
-%
-%Traditionally it is argued that legitimacy comes from the fact that members had a chance to participate
-%and they decided not to, giving legitimacy to make a decision to those who did participate.
-%We argue that this is a problematic view because there are many restrictions on why members could not
-%participate and vote.
-%For example, they might not had time to participate because of other commitments, or motivation, or they might determined
-%that they are not informed enough to participate.
-%Moreover, members are from different reasons in their communities and potentially contribute in other ways
-%than by participating in its democratic processes.
-%Instead of assuming that everyone could participate if they would want, or trying to increase participation,
-%we can see amount of participation members are able to do as a limited resource, a \emph{collective attention}
-%a community has available to utilize for its decision making.
-%
-%We can compute decision credibility during ongoing voting process to give feedback to the community
-%on how contentious a proposed decision is how much more votes are necessary for a statistical quorum
-%to be reached.
-%This allows the community to focus its collective attention towards decisions which require more votes
-%to determine a legitimate result, while reducing attention towards proposals which already have overwhelming support.
-%Instead of trying to increase participation on all democratic decisions being made, a community can
-%do that selectively through an automatic feedback loop, trying to reach statistical quorum for their decisions.
-%
-%We built an open source web application PeerMind integrating statistical quorum with online voting.
-%The application exposes computed decision credibility to users in real-time and through that provides
-%the feedback loop needed for improving utilization of collective attention.
-%We present the application in \secref{sec:peermind} and use it for evaluating statistical quorum in
-%\secref{sec:evaluation}.
-%
-%\todo[inline]{Should we explicitly list our contributions again?}
-%
-%\todo[inline]{Repeat the structure of the rest of the paper? Is there a better way to reference sections than
-%referencing it by name?}
 
@@
 
-%\todo[inline]{Repeat references from "There have been many improvements proposed to this basic form and various analysis made."?}
-%
 %\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
@@
 %\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
-%
-%\section{PeerMind Web Application}
-%\label{sec:peermind}
-%
-%\todo[inline]{Something we need to think about is the level of detail we want to go into regarding the implementation/features of PeerMind beyond statistical quorum and range voting.
-I think we might want to focus on only those two features for this paper.\\
-%I think PeerMind as a comprehensive platform should have its own paper where all of the data is analyzed across the board.
-This would be a much more typical HCI paper.
-(If I can become president of Ridge in a future semester, we'll have another chance to run a clean, well-designed experiment with a much more developed version of PeerMind.)\\
-%The message in the current paper might be clearest if we keep the scope focused on the theory + evaluation/verification.  We should make sure we're not trying to pack too much into one paper.
-%\\---Max}
-%
-%We built a full-fletched real-time web application for collective decision making and voting using Meteor
-%framework~\cite{meteor}.
-%Users can initiate new discussions which leads them to the main user interface for each discussion as seen
-%in Figure~\ref{fig:peermind}.
-%
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=1.75\columnwidth]{figures/peermind.png}
-%  \caption{The main PeerMind interface of a discussion. There is initial discussion description on the top and under it there
-%  are three columns. A commenting sections for users to discuss it and proposed decisions, a second column where moderators can
-%  summarize arguments made during discussion, and a list of proposals to address the discussion.
-%  Proposals can be in different states: a proposal being discussed, a proposal being voted on, and with voting concluded.}\label{fig:peermind}
-%\end{figure*}
-%
-%%and discuss them with others online in real-time and with rich-text content.
-%%They can attach images and files using drag\&drop. Moderators can summarize discussions into three groups
-%
-%\todo[inline]{Present how users can see ongoing results during voting. Show an image of active ballot, but with expanded view for current results.}
-%
-%%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
-%%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
-%%A proposal passes if the average score of all votes on a position is positive.
-%%When multiple competing proposals are being voted on, the one with highest average score wins.
-%%
-%%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
-%%positions in a highly informative and precise manner.
-%%We find it especially suitable for a web application because many web users are used to use it when ranking products
-%%and giving other ``five-star'' feedback.
-%%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
-%%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
-%%each proposal would need reassessing (assuming members use a consistent personal scale when voting).
-%
-%\todo[inline]{The app has features both for use in online and offline setting. This allows one to compute
-%statistical quorum also for an in-person meeting because math is now more complicated than just counting hands.
-%We focus on online experience in this paper.}
-%
-%\todo[inline]{Security is out of the scope of this work. We assume here that users and their identities are known and are managed
-%by a community's admin. There are many communities where their decision making is public with public minutes and this is who
-%the app is currently targeting.}
 
@@
 \label{sec:decision-credibility}
+
+%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
+%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
+%A proposal passes if the average score of all votes on a position is positive.
+%When multiple competing proposals are being voted on, the one with highest average score wins.
+%
+%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
+%positions in a highly informative and precise manner.
+%We find it especially suitable for a web application because many web users are used to use it when ranking products
+%and giving other ``five-star'' feedback.
+%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
+%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
+%each proposal would need reassessing (assuming members use a consistent personal scale when voting).
+
 A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
@@
 
-Web application was developed using Meteor web framework~\cite{meteor} as a mobile-friendly web application.
+We built a full-fletched mobile-friendly web applicationusing Meteor web framework~\cite{meteor}.
 Meteor framework supports reactivity which makes the application update user interface based on actions
@@
 
-Each user can open and propose new discussion items. Each discussion item has then three basic sections,
-as seen in Figure~\ref{fig:peer-mind}: commenting section on the left, summary points in the middle,
+Each user can open and propose new discussion items, which leads them to the main user interface for each discussion as seen in Figure~\ref{fig:peer-mind}, having three sections: commenting section on the left, summary points in the middle,
 and proposed motions with voting on the right. All content inputs fields accept rich text content in WYSWYG manner,
@@
 further discussed through comments. Once a moderator deems that motion is ready to be voted upon
-can open voting. Users vote through a ballot seen in Figure~\ref{fig:ballot}.
+can open it for voting. Users vote through a ballot seen in Figure~\ref{fig:ballot}.
 

### 2018-04-19 statistical-quorum.tex
Clean.
Updated abstract.

@@
 \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
-\usepackage{ccicons}        % Cite your images correctly!
+%\usepackage{ccicons}        % Cite your images correctly!
 \usepackage[utf8]{inputenc} % for a UTF8 editor only
@@
 % so you remain anonymous.
-\def\plaintitle{Improving legitimacy and efficiency of decisions in democratic communities}
+\def\plaintitle{Improving credibility and efficiency of decision-making within democratic communities via statistical quorum}
 %\def\plainauthor{First Author, Second Author, Third Author,
@@
 \def\emptyauthor{}
-\def\plainkeywords{Decision making; voting; democracy; quorum; legitimacy; statistical quorum.}
+\def\plainkeywords{Decision making; voting; democracy; quorum; credibility; statistical quorum.}
 %\def\plaingeneralterms{Documentation, Standardization}
@@
 
-\title{Improving legitimacy and efficiency \\ of decisions in democratic communities}
+\title{Improving credibility and efficiency of decision-making \\ within democratic communities via statistical quorum}
 
@@
 maximize productivity?
-We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to focus
-attention, in turn making the entire decision-making process dynamic.
-We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
-councils over the course of 16 weeks.  We found that the community was able to use the web application to simultaneously
-improve the legitimacy and efficiency of their decision-making process.
+We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to
+focus attention, in turn making the entire decision-making process dynamic.
+We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
+councils over the course of 16 weeks.
+We found that the community was able to use the online system to simultaneously improve the credibility and efficiency
+of their decision-making process.
 \end{abstract}
 
-\category{H.5.3.}{Information Interfaces and Presentation
- (e.g. HCI)}{Computer-supported cooperative work}
+\category{H.5.3.}{Information Interfaces and Presentation (e.g. HCI)}{Computer-supported cooperative work}
 
@@
 
-\subsection{Survey}
-
-In Figure~\ref{fig:survey-comparison} we see questions and results of the survey we made at the end of each semester.
-
-We can see that in fall members were more satisfied with the decision making in the community and only how understandable
-is the whole process received a decrease. We attribute this to the fact that the use of the web application is
-a novelty and requires more effort in educating members than a traditional decision making. Moreover, decision
-making in person allows one to observe others how they are participating and learn from them, instead of
-being alone using the application.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
-\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
-For each question, a participant was asked to rank the answer on a scale between 1 (completely disagree) and 7 (completely agree). All members
-of the community were invited to participate in the survey each time.}
-\label{fig:survey-comparison}
-\end{figure}
-
-\subsection{Usage comparison}
+\subsection{Use in Practice}
 
@@
 collective attention can be utilized better.}
+
+\subsection{Survey}
+
+In Figure~\ref{fig:survey-comparison} we see questions and results of the survey we made at the end of each semester.
+
+We can see that in fall members were more satisfied with the decision making in the community and only how understandable
+is the whole process received a decrease. We attribute this to the fact that the use of the web application is
+a novelty and requires more effort in educating members than a traditional decision making. Moreover, decision
+making in person allows one to observe others how they are participating and learn from them, instead of
+being alone using the application.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
+\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
+For each question, a participant was asked to rank the answer on a scale between 1 (completely disagree) and 7 (completely agree). All members
+of the community were invited to participate in the survey each time.}
+\label{fig:survey-comparison}
+\end{figure}
 

### 2018-04-19 statistical-quorum.tex
Moving around.

@@
 
-\todo[inline]{We made 6 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
-to work. But how there are other issues by moving decision making online: meetings also serve community building which
-does not happen online. No fun. So interesting insights about using online platforms for offline communities
-for making decisions which are not really answering our hypothesis, but are useful as insights we got.}
-
 \todo[inline]{Make sure that we are supporting both of our claims well: this statistical quorum is a good metric and
@@
 \subsection{Interviews}
+
+\todo[inline]{We made 7 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
+to work. But how there are other issues by moving decision making online: meetings also serve community building which
+does not happen online. No fun. So interesting insights about using online platforms for offline communities
+for making decisions which are not really answering our hypothesis, but are useful as insights we got. How some
+people prefer online and offline meetings.}
 

### 2018-04-19 statistical-quorum.tex
Updating to new format.

@@
-\documentclass[chi_draft]{sigchi}
+\documentclass[format=acmsmall, review=false, screen=true, anonymous=false]{acmart}
 
-% Use this section to set the ACM copyright statement (e.g. for
-% preprints).  Consult the conference website for the camera-ready
-% copyright statement.
+\usepackage{booktabs} % For formal tables
+
+\usepackage[ruled]{algorithm2e} % For algorithms
+\renewcommand{\algorithmcfname}{ALGORITHM}
+\SetAlFnt{\small}
+\SetAlCapFnt{\small}
+\SetAlCapNameFnt{\small}
+\SetAlCapHSkip{0pt}
+\IncMargin{-\parindent}
+
+\usepackage{todonotes}
+\usepackage{bookmark}
+
+% Metadata Information
+%\acmJournal{TWEB}
+%\acmVolume{9}
+%\acmNumber{4}
+%\acmArticle{39}
+%\acmYear{2010}
+%\acmMonth{3}
+%\copyrightyear{2009}
+%\acmArticleSeq{9}
 
 % Copyright
-%\CopyrightYear{2016}
 %\setcopyright{acmcopyright}
@@
 %\setcopyright{cagovmixed}
+
 % DOI
-%\doi{http://dx.doi.org/10.475/123_4}
-% ISBN
-%\isbn{123-4567-24-567/08/06}
-%Conference
-%\conferenceinfo{CHI'16,}{May 07--12, 2016, San Jose, CA, USA}
-%Price
-%\acmPrice{\$15.00}
+%\acmDOI{0000001.0000001}
 
-% Use this command to override the default ACM copyright statement
-% (e.g. for preprints).  Consult the conference website for the
-% camera-ready copyright statement.
+% Paper history
+%\received{February 2007}
+%\received[revised]{March 2009}
+%\received[accepted]{June 2009}
 
-%% HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP --
-%% Please note you need to make sure the copy for your specific
-%% license is used here!
-% \toappear{
-% Permission to make digital or hard copies of all or part of this work
-% for personal or classroom use is granted without fee provided that
-% copies are not made or distributed for profit or commercial advantage
-% and that copies bear this notice and the full citation on the first
-% page. Copyrights for components of this work owned by others than ACM
-% must be honored. Abstracting with credit is permitted. To copy
-% otherwise, or republish, to post on servers or to redistribute to
-% lists, requires prior specific permission and/or a fee. Request
-% permissions from \href{mailto:Permissions@acm.org}{Permissions@acm.org}. \\
-% \emph{CHI '16},  May 07--12, 2016, San Jose, CA, USA \\
-% ACM xxx-x-xxxx-xxxx-x/xx/xx\ldots \$15.00 \\
-% DOI: \url{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
-% }
+% Document starts
+\begin{document}
+% Title portion. Note the short title for running heads
+\title[Improving credibility and efficiency of decision-making]{Improving credibility and efficiency of decision-making \\ within democratic communities via statistical quorum}
 
-\toappear{}
-
-% Arabic page numbers for submission.  Remove this line to eliminate
-% page numbers for the camera ready copy
-\pagenumbering{arabic}
-
-% Load basic packages
-\usepackage{balance}       % to better equalize the last page
-\usepackage{graphics}      % for EPS, load graphicx instead
-\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
-\usepackage{txfonts}
-\usepackage{mathptmx}
-\usepackage[pdflang={en-US},pdftex]{hyperref}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{textcomp}
-
-% Some optional stuff you might like/need.
-\usepackage{microtype}      % Improved Tracking and Kerning
-\usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
-%\usepackage{ccicons}        % Cite your images correctly!
-\usepackage[utf8]{inputenc} % for a UTF8 editor only
-
-% If you want to use todo notes, marginpars etc. during creation of
-% your draft document, you have to enable the "chi_draft" option for
-% the document class. To do this, change the very first line to:
-% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
-% by using the "\todo{...}"  command. Make sure to disable the draft
-% option again before submitting your final document.
-\usepackage{todonotes}
-
-% Paper metadata (use plain text, for PDF inclusion and later
-% re-using, if desired).  Use \emtpyauthor when submitting for review
-% so you remain anonymous.
-\def\plaintitle{Improving credibility and efficiency of decision-making within democratic communities via statistical quorum}
-%\def\plainauthor{First Author, Second Author, Third Author,
-%  Fourth Author, Fifth Author, Sixth Author}
-\def\emptyauthor{}
-\def\plainkeywords{Decision making; voting; democracy; quorum; credibility; statistical quorum.}
-%\def\plaingeneralterms{Documentation, Standardization}
-
-% llt: Define a global style for URLs, rather that the default one
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{
-    \def\UrlFont{\sf}
-  }{
-    \def\UrlFont{\small\bf\ttfamily}
-  }}
-\makeatother
-\urlstyle{leo}
-
-% To make various LaTeX processors do the right thing with page size.
-\def\pprw{8.5in}
-\def\pprh{11in}
-\special{papersize=\pprw,\pprh}
-\setlength{\paperwidth}{\pprw}
-\setlength{\paperheight}{\pprh}
-\setlength{\pdfpagewidth}{\pprw}
-\setlength{\pdfpageheight}{\pprh}
-
-% Make sure hyperref comes last of your loaded packages, to give it a
-% fighting chance of not being over-written, since its job is to
-% redefine many LaTeX commands.
-\definecolor{linkColor}{RGB}{6,125,233}
-%\hypersetup{%
-%  pdftitle={\plaintitle},
-%% Use \plainauthor for final version.
-%%  pdfauthor={\plainauthor},
-%  pdfauthor={\emptyauthor},
-%  pdfkeywords={\plainkeywords},
-%  pdfdisplaydoctitle=true, % For Accessibility
-%  bookmarksnumbered,
-%  pdfstartview={FitH},
-%  colorlinks,
-%  citecolor=black,
-%  filecolor=black,
-%  linkcolor=black,
-%  urlcolor=linkColor,
-%  breaklinks=true,
-%  hypertexnames=false
-%}
-
-% create a shortcut to typeset table headings
-% \newcommand\tabhead[1]{\small\textbf{#1}}
-
-\newcommand\secref[1]{section \textit{\nameref{#1}}}
-
-%%% GLOSSARY %%%
-%
-% community = group which votes
-% proposal = what you vote on
-% decision = result of a vote, accepted or rejected proposal
-%
-%%% END GLOSSARY %%%
-
-% End of preamble. Here it comes the document.
-\begin{document}
-
-\title{Improving credibility and efficiency of decision-making \\ within democratic communities via statistical quorum}
-
-%\numberofauthors{3}
-%\author{%
-%  \alignauthor{Leave Authors Anonymous\\
-%    \affaddr{for Submission}\\
-%    \affaddr{City, Country}\\
-%    \email{e-mail address}}\\
-%  \alignauthor{Leave Authors Anonymous\\
-%    \affaddr{for Submission}\\
-%    \affaddr{City, Country}\\
-%    \email{e-mail address}}\\
-%  \alignauthor{Leave Authors Anonymous\\
-%    \affaddr{for Submission}\\
-%    \affaddr{City, Country}\\
-%    \email{e-mail address}}\\
-%}
-
-\maketitle
+\author{Gang Zhou}
+\orcid{1234-5678-9012-3456}
+\affiliation{%
+  \institution{College of William and Mary}
+  \streetaddress{104 Jamestown Rd}
+  \city{Williamsburg}
+  \state{VA}
+  \postcode{23185}
+  \country{USA}}
+\email{gang_zhou@wm.edu}
+\author{Valerie B\'eranger}
+\affiliation{%
+  \institution{Inria Paris-Rocquencourt}
+  \city{Rocquencourt}
+  \country{France}
+}
+\email{beranger@inria.fr}
+\author{Aparna Patel}
+\affiliation{%
+ \institution{Rajiv Gandhi University}
+ \streetaddress{Rono-Hills}
+ \city{Doimukh}
+ \state{Arunachal Pradesh}
+ \country{India}}
+\email{aprna_patel@rguhs.ac.in}
+\author{Huifen Chan}
+\affiliation{%
+  \institution{Tsinghua University}
+  \streetaddress{30 Shuangqing Rd}
+  \city{Haidian Qu}
+  \state{Beijing Shi}
+  \country{China}
+}
+\email{chan0345@tsinghua.edu.cn}
+\author{Ting Yan}
+\affiliation{%
+  \institution{Eaton Innovation Center}
+  \city{Prague}
+  \country{Czech Republic}}
+\email{yanting02@gmail.com}
+\author{Tian He}
+\affiliation{%
+  \institution{University of Virginia}
+  \department{School of Engineering}
+  \city{Charlottesville}
+  \state{VA}
+  \postcode{22903}
+  \country{USA}
+}
+\affiliation{%
+  \institution{University of Minnesota}
+  \country{USA}}
+\email{tinghe@uva.edu}
+\author{Chengdu Huang}
+\author{John A. Stankovic}
+\author{Tarek F. Abdelzaher}
+\affiliation{%
+  \institution{University of Virginia}
+  \department{School of Engineering}
+  \city{Charlottesville}
+  \state{VA}
+  \postcode{22903}
+  \country{USA}
+}
 
@@
 
-\category{H.5.3.}{Information Interfaces and Presentation (e.g. HCI)}{Computer-supported cooperative work}
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below.
+%
+\begin{CCSXML}
+<ccs2012>
+ <concept>
+  <concept_id>10010520.10010553.10010562</concept_id>
+  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+  <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010575.10010755</concept_id>
+  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010553.10010554</concept_id>
+  <concept_desc>Computer systems organization~Robotics</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10003033.10003083.10003095</concept_id>
+  <concept_desc>Networks~Network reliability</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+</ccs2012>
+\end{CCSXML}
 
-\keywords{\plainkeywords}
+\ccsdesc[500]{Computer systems organization~Embedded systems}
+\ccsdesc[300]{Computer systems organization~Redundancy}
+\ccsdesc{Computer systems organization~Robotics}
+\ccsdesc[100]{Networks~Network reliability}
+
+%
+% End generated code
+%
+
+\keywords{Decision making; voting; democracy; quorum; credibility; statistical quorum}
+
+\maketitle
+
+% The default list of authors is too long for headers.
+\renewcommand{\shortauthors}{M. Milutinovic et al.}
 
@@
 
+designed integration with system
+
+minutes and issues with how to get data out
+
+workflow we created -> over one in-person meetings, but because of all decision making happining online, this
+didn't happen
+
+issues with quorum
+unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after the vote
+
+X people choose not to participate
+
 \todo[inline]{Explain how we were using this for 16 weeks in fall 2016 in a community where traditionally they have
@@
 
-%\section{Acknowledgments}
-
-% REFERENCES FORMAT
-% References must be the same font size as other body text.
-\bibliographystyle{sigchi}
-\bibliography{refs}
-
+% Appendix
 \appendix
@@
 
-\subsection{Derivation of Eq. \eqref{eq2}}
+\section{Derivation of Eq. \eqref{eq2}}
 In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.
@@
 
-\subsection{Derivation of Eq. \eqref{eq3}}
+\section{Derivation of Eq. \eqref{eq3}}
 \begin{align*}
@@
 
-\subsection{Derivation of Eq. \eqref{eq5}}
+\section{Derivation of Eq. \eqref{eq5}}
 Assuming independence of votes across proposals,
@@
 
-\subsection{Derivation of Eq. \eqref{eq6}}
+\section{Derivation of Eq. \eqref{eq6}}
 The properties we desire for $\lambda(t)$ are:
@@
 
-\subsection{Derivation of Eq. \eqref{eq7}}
+\section{Derivation of Eq. \eqref{eq7}}
 Using $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$, we have
@@
 \end{document}
+
+%\begin{acks}
+%\end{acks}
+
+% Bibliography
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{refs}

### 2018-04-19 statistical-quorum.tex
Fixing.
Updated authors.
Addresses.
Enable review mode.
Set language.
Added CSS concepts.

@@
-\documentclass[format=acmsmall, review=false, screen=true, anonymous=false]{acmart}
+\documentclass[format=acmsmall, review=true, screen=true, anonymous=true]{acmart}
+
+\usepackage[english]{babel}
 
@@
 
-\author{Gang Zhou}
-\orcid{1234-5678-9012-3456}
+\author{Mitar Milutinovic}
 \affiliation{%
-  \institution{College of William and Mary}
-  \streetaddress{104 Jamestown Rd}
-  \city{Williamsburg}
-  \state{VA}
-  \postcode{23185}
-  \country{USA}}
-\email{gang_zhou@wm.edu}
-\author{Valerie B\'eranger}
+  \institution{UC Berkeley}}
+\email{mitar@cs.berkeley.edu}
+\author{Maxinder S. Kanwal}
 \affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
-\email{beranger@inria.fr}
-\author{Aparna Patel}
+  \institution{UC Berkeley}}
+\email{maxowaxo2013@gmail.com}
+\author{David Marn}
 \affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
-\email{aprna_patel@rguhs.ac.in}
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}
-}
-\email{chan0345@tsinghua.edu.cn}
-\author{Ting Yan}
-\affiliation{%
-  \institution{Eaton Innovation Center}
-  \city{Prague}
-  \country{Czech Republic}}
-\email{yanting02@gmail.com}
-\author{Tian He}
-\affiliation{%
-  \institution{University of Virginia}
-  \department{School of Engineering}
-  \city{Charlottesville}
-  \state{VA}
-  \postcode{22903}
-  \country{USA}
-}
-\affiliation{%
-  \institution{University of Minnesota}
-  \country{USA}}
-\email{tinghe@uva.edu}
-\author{Chengdu Huang}
-\author{John A. Stankovic}
-\author{Tarek F. Abdelzaher}
-\affiliation{%
-  \institution{University of Virginia}
-  \department{School of Engineering}
-  \city{Charlottesville}
-  \state{VA}
-  \postcode{22903}
-  \country{USA}
-}
+  \institution{UC Berkeley}}
+\email{david.marn@gmail.com}
 
@@
 <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+<concept_id>10003120.10003130.10003131.10003570</concept_id>
+<concept_desc>Human-centered computing~Computer supported cooperative work</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
+<concept_id>10003120.10003130.10011764</concept_id>
+<concept_desc>Human-centered computing~Collaborative and social computing devices</concept_desc>
+<concept_significance>500</concept_significance>
 </concept>
 <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
+<concept_id>10010405.10010476.10010936.10003590</concept_id>
+<concept_desc>Applied computing~Voting / election technologies</concept_desc>
+<concept_significance>500</concept_significance>
 </concept>
@@
 
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+\ccsdesc[500]{Human-centered computing~Computer supported cooperative work}
+\ccsdesc[500]{Human-centered computing~Collaborative and social computing devices}
+\ccsdesc[500]{Applied computing~Voting / election technologies}
 
@@
 \begin{description}
-\begin{itemize}
 \item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing proposals,
@@
 \item[$p_i = $] true latent community bias towards voting ``yes'' on $X_i$.
-\end{itemize}
 \end{description}
@@
 \begin{description}
-\begin{itemize}
 \item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
@@
 (e.g., $f=\frac{1}{2}$ for a simple majority).
-\end{itemize}
 \end{description}
@@
 \begin{description}
-\begin{itemize}
 \item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes $X_i$ could receive such that it would remain below threshold.
-\end{itemize}
 \end{description}
@@
 \begin{description}
-\begin{itemize}
 \item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
-\end{itemize}
 \end{description}
@@
 \begin{description}
-\begin{itemize}
 \item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that sufficiently smooths the historical rate),
 \item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data for the community via a regression analysis.
-\end{itemize}
 \end{description}
@@
 \begin{description}
-\begin{itemize}
 \item[$(1)$] The updated prior should not depend on the absolute number of votes $\alpha_i$, $\beta_i$, as this is not scale-invariant; rather, it should depend on the relative proportions (which are unitless instead of a number of votes),
@@
 \item[$(3)$] as $\alpha_i + \beta_i \rightarrow N_i$, $\operatorname{Var}[p_i] \rightarrow 0$ (at a rate independent of $N_i$).
-\end{itemize}
 \end{description}
@@
 \begin{description}
-\begin{itemize}
 \item[$(1)$] as $K_i \rightarrow 0$, the rate of incoming votes should drop exponentially and should be independent of $N_i$,
 \item[$(2)$] as $t \rightarrow \infty$, the rate of incoming votes should drop exponentially.
-\end{itemize}
 \end{description}
@@
 \end{align*}
-\end{document}
 
@@
 \bibliography{refs}
+
+\end{document}

### 2018-04-19 statistical-quorum.tex
Introduction and example.

@@
 \usepackage{bookmark}
+\usepackage[T1]{fontenc}
+\usepackage{microtype}
+\usepackage[utf8]{inputenc}
 
@@
 focus attention, in turn making the entire decision-making process dynamic.
-We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
+We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
 councils over the course of 16 weeks.
-We found that the community was able to use the online system to simultaneously improve the credibility and efficiency
+We found that the community was able to use the web application to simultaneously improve the credibility and efficiency
 of their decision-making process.
@@
 
-In many historical and recent elections it is not infrequent that we get close results.
+In many historical and recent elections it is not infrequent that we get close results (see Table~\ref{tab:example}).
 Especially when this is combined with low voter turnout, how can we have confidence in results?
@@
 more people would participate? Would results still be the same?
+
+\begin{table}
+  \centering
+  \begin{small}
+  \begin{tabular}{l r r r r r r r}
+    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible}} & {\small \textit{Turnout}} & {\small \textit{Decision}} \\
+    & & & & & {\small \textit{voters}} & & {\small \textit{credibility}} \\
+    \midrule
+    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
+    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
+  \end{tabular}
+  \end{small}
+  \bigskip\caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
+  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
+  invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:decision-credibility}. We see that
+  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
+  with more votes any option could end up winning.}\label{tab:example}
+  \vspace{-0.3cm}
+\end{table}
 
@@
 \item Based on this framework we developed a novel concept of \emph{statistical quorum}
-as a threshold of votes required to deem results legitimate.
-\item We implemented a web application \emph{PeerMind},
+as a threshold of votes required to deem results credible and legitimate.
+\item We implemented web application \emph{PeerMind},
 integrating credibility of results and statistical quorum, and designed
@@
 of the application, using surveys, and using semi-structured interviews, we show
-how use of statistical quorum increases legitimacy of decisions,
-improves utilization of collective attention, increases overall participation,
-and increases number of decisions made while not sacrificing their legitimacy.
+how use of statistical quorum increases credibility of decisions,
+improves utilization of collective attention, increases and diversifies overall participation,
+and increases number of decisions made.
 \end{itemize}
@@
 credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
-We present our methodology in Section~\ref{sec:methodology} after which we do evaluation in
-Section~\ref{sec:evaluation}. We discuss results in Section~\ref{sec:discussion} and conclude
-in Section~\ref{sec:conclusion}. At the end we have Appendix where we
-derive formulas presented in Section~\ref{sec:decision-credibility}.
-
-% Security is out of the scope of this work. We assume here that users and their identities are known and are managed
-% by a community's admin. There are many communities where their decision making is public with public minutes and this
-% is who the app is currently targeting.
-
-%\begin{table*}
-%  \centering
-%  \begin{tabular}{l r r r r r r r}
-%    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Decision credibility}} \\
-%    \midrule
-%    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
-%    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
-%  \end{tabular}
-%  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
-%  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
-%  invalid votes. The last column shows computed decision credibility as described in \secref{sec:decision-credibility}. We see that
-%  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
-%  with more votes any option could end up winning.}\label{tab:example}
-%\end{table*}
+We present our methodology in Section~\ref{sec:methodology} after which we do qualitative and
+quantitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
+Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
+have Appendix where we derive formulas presented in Section~\ref{sec:decision-credibility}.
 

### 2018-04-19 statistical-quorum.tex
Motions into proposals.

@@
 
-A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
-For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose, then we can conclude with high probability that, had all 100 members taken the time to vote, a majority of them would be in support of the proposal.
-Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote was credible enough to be taken as representative of the entire community's desire.
-On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in the vote.
-Instead, the suggested course of action would be to continue gathering more votes and/or further discussing the proposal until a credible outcome is established.
+A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome
+of an ongoing voting before the community adopts the outcome as being truly representative of all
+members' positions.
+For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose,
+then we can conclude with high probability that, had all 100 members taken the time to vote, a
+majority of them would be in support of the proposal.
+Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote is
+credible enough to be taken as representative of the entire community's desire.
+On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we
+cannot yet conclude that the decision to accept the proposal is credible enough for the entire
+community to officially adopt, despite the significant participation in the vote.
+Instead, the suggested course of action would be to continue gathering more votes and/or further
+discussing the proposal until a credible outcome is established.
 
-By utilizing a statistical quorum, a community can actively direct attention and participation within the community towards contentious proposals, while avoiding wasting attention on noncontroversial proposals.
+By utilizing a statistical quorum, a community can actively direct attention and participation
+within the community towards contentious proposals, while avoiding wasting attention on
+noncontroversial proposals.
 
 \subsection{Background}
-Consider a community that needs to collectively decide which proposal (if any) to accept out of a set of competing proposals.
-To act in democratic fashion, the community may wish to select a winning proposal that either has the support of a \{simple, super\}-majority of all members, or instead is demonstrably preferred over all other competing proposals.
+Consider a community that needs to collectively decide which proposal (if any) to accept out of
+a set of competing proposals.
+To act in democratic fashion, the community may wish to select a winning proposal that either has
+the support of a \{simple, super\}-majority of all members, or instead is demonstrably preferred
+over all other competing proposals.
 (In some cases, the community may want a winning proposal to satisfy both criteria.)
 
-Ideally, we do not want to poll all members in order to officially accept a proposal, but we do want some level of certainty (e.g., $90\%$ certainty) that
-the population is not accidentally accepting an undesired proposal, which in this case encompasses accepting a proposal that is not supported by a majority of members and/or is not clearly preferred over all other competing alternatives.
+Ideally, we do not want to poll all members in order to officially accept a proposal, but we do
+want some level of certainty (e.g., $90\%$ certainty) that the population is not accidentally
+accepting an undesired proposal, which in this case encompasses accepting a proposal that is
+not supported by a majority of members and/or is not clearly preferred over all other competing
+alternatives.
 
-The primary motivation behind using a statistical quorum is to guarantee a level of credibility before letting a subset of members act on behalf of the entire community.
+The primary motivation behind using a statistical quorum is to guarantee a level of credibility
+before making a decision on behalf of the entire community.
 
 \subsection{Problem Statement}
-We wish to compute the probability that the bias in a community towards supporting a proposal is greater than some
-fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority).
+We wish to compute the probability that the bias in a community towards supporting a
+proposal is greater than some fraction of the population (e.g., $50\%$ for a simple majority,
+$67\%$ for a super majority).
 
-In addition, we wish to compute the probability that the community bias towards supporting a specific proposal is greater than the community bias towards supporting any competing proposals.
+In addition, we wish to compute the probability that the community bias towards supporting
+a specific proposal is greater than the community bias towards supporting any competing proposals.
 
@@
 We assume that there is a fixed latent bias for the support of a proposal within the community.
-We take each individual vote to be independent and identically distributed (i.i.d.) according to a Bernoulli distribution around this bias.
-It should be noted that the independence assumption may not hold in some settings, such as when votes are accumulated over time (one may observe surges of correlated votes).
-We will partially address this when refining our original formulation of a statistical quorum into a \textit{dynamic} statistical quorum.
+We take each individual vote to be independent and identically distributed (i.i.d.) according
+to a Bernoulli distribution around this bias.
+It should be noted that the independence assumption may not hold in some settings, such as when
+votes are accumulated over time (one may observe surges of correlated votes).
+We will partially address this when refining our original formulation of a statistical quorum
+into a \textit{dynamic} statistical quorum.
 
-Without having collected any votes, we assume that the community bias may be any value between $0\%$ and $100\%$ with equal probability (uniform prior).
+Without having collected any votes, we assume that the community bias may be any value between
+$0\%$ and $100\%$ with equal probability (uniform prior).
 
 \subsection{Voting}
-We allow for range voting where each member of the community may choose any real-valued vote $v \in [-1,+1]$ to
-represent their vote, with $v = -1$ representing full opposition and $v=+1$ representing full support.
-In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning the full range, hence our treatment of $v$ as a discrete random variable.
-Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as an individual's average vote.
-In the theory subsection below, we rescale $v$ to be in the range $[0,1]$ via the notation $\frac{v + 1}{2}$.
+We allow for range voting where each member of the community may choose any real-valued vote
+$v \in [-1,+1]$ to represent their vote, with $v = -1$ representing full opposition and $v=+1$
+representing full support.
+In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning
+the full range, hence our treatment of $v$ as a discrete random variable.
+Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as
+an individual's average vote.
+In the following subsection, we rescale $v$ to be in the range $[0,1]$ via the
+notation $\frac{v + 1}{2}$.
 
 We allow voters to ``abstain'' from voting.
-We treat abstentions as indicating that a voter is removing themselves from the community on a given vote.
-It is important to distinguish an ``abstention'' (removing one's self from the voting population) from voting $0$ (indicating indifference).
+We treat abstentions as indicating that a voter is removing themselves from the community on a
+given vote.
+It is important to distinguish an ``abstention'' (removing one's self from the voting population)
+from voting $0$ (indicating indifference).
 
@@
 
-\todo[inline]{Weave in a single concrete example to ground all the abstract theory so that the entire derivation and motivation has a clear and cohesive thread from start to finish.}
+\todo[inline]{Weave in a single concrete example to ground all the abstract theory so that the
+entire derivation and motivation has a clear and cohesive thread from start to finish.}
 
-After collecting an initial set of votes from members, we want to use that data to construct a prior distribution to help estimate the posterior probability that the proposal is acceptable to the entire community.
-Because the sum of  the collected votes follows a binomial distribution, a commonplace and convenient choice for the form of the prior on the community's bias is the Beta distribution (due to its conjugacy with the binomial distribution).
+After collecting an initial set of votes from members, we want to use that data to construct
+a prior distribution to help estimate the posterior probability that the proposal is acceptable
+to the entire community.
+Because the sum of the collected votes follows a binomial distribution, a commonplace and
+convenient choice for the form of the prior on the community's bias is the Beta distribution
+(due to its conjugacy with the binomial distribution).
 Together, they form the Beta-binomial distribution (related to the P\'olya urn model).
 
-After extracting the sufficient statistics from raw votes --- namely, the average score and the number of abstentions --- we can adjust the effective community size and corresponding threshold.
+After extracting the sufficient statistics from raw votes --- namely, the average score
+and the number of abstentions --- we can adjust the effective community size and corresponding
+threshold.
 In addition, we can calculate the effective number of ``yes'' ($+1$) and ``no'' ($-1$) votes:
@@
 \item[$N = $] population size,
-\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding abstentions,
-\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on proposal $X_i$,
-\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
-\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
+\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding
+abstentions,
+\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting
+on proposal $X_i$,
+\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number
+of ``yes'' votes on proposal $X_i$,
+\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on proposal $X_i$,
 \item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
@@
 
-In Eqs. \eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we refer the reader to the \hyperref[sec:derivation]{Appendix} for complete derivation details and justifications.
+In Eqs.~\eqref{EQ2}, \eqref{EQ3}, \eqref{EQ5}, \eqref{EQ6}, \eqref{EQ7} that follow, we refer the
+reader to the \hyperref[sec:derivation]{Appendix} for complete derivation details and justifications.
 
 When considering a single proposal $X_i$ in isolation, we wish to compute:
-\begin{align}\label{eq1}
+\begin{align}\label{EQ1}
 \Pr(\alpha^*_i > T_i \mid X_i),
@@
 \begin{description}
-\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
-\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
+\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire
+population,
+\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed
+for the vote
 (e.g., $f=\frac{1}{2}$ for a simple majority).
@@
 Using a desirable prior
-\begin{align}\label{eq2}
+\begin{align}\label{EQ2}
 p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
 \end{align}
-corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can express \eqref{eq1} as:
-\begin{align}\label{eq3}
+corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can
+express \eqref{EQ1} as:
+\begin{align}\label{EQ3}
 \Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
@@
 \begin{description}
-\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes $X_i$ could receive such that it would remain below threshold.
+\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes''
+votes $X_i$ could receive such that it would remain below threshold.
 \end{description}
@@
 When considering a proposal $X_i$ relative to other proposals in $C$, we wish to compute
-\begin{align}\label{eq4}
+\begin{align}\label{EQ4}
 \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right)
@@
 
-Under the assumption that the community's votes on motions are independent across motions (which, in many instances, will actually give a lower bound on \eqref{eq3} when a community's support of one proposal is negatively correlated with its support on a competing proposal), we can decompose \eqref{eq3} as follows:
-\begin{align}\label{eq5}
+Under the assumption that the community's votes on proposals are independent across proposals
+(which, in many instances, will actually give a lower bound on \eqref{EQ3} when a community's
+support of one proposal is negatively correlated with its support on a competing proposal),
+we can decompose \eqref{EQ3} as follows:
+\begin{align}\label{EQ5}
 & \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) \nonumber \\
@@
 \begin{description}
-\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
+\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$
+leads $X_j$ by.
 \end{description}
 
-Using Eqs.
-\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
-For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.
+Using Eqs.~\eqref{EQ2} and \eqref{EQ4}, we can numerically compute the credibility of a
+vote's outcome in two key scenarios relevant for democratic communities.
+For example, a community may only consider accepting a proposal when a vote's credibility
+is above $90\%$.
 
 \subsection{Dynamic Statistical Quorum}
-Oftentimes, not all proposals are relevant to the entire community.
-Instead, a niche proposal may only affect a subset of the community, and hence to require much of the community to indicate their indifference and/or decide their opinion is a waste of the their fixed attentional budget.
-While the underlying formulation for a statistical quorum was to establish credibility guarantees before taking action, a natural consequence in common scenarios of the kind just described leads to a reduction in decision-making efficiency.
+Oftentimes not all proposals are relevant to the entire community.
+Instead, a niche proposal may only affect a subset of the community, and hence to require much
+of the community to indicate their indifference and/or decide their opinion is a waste of their
+attentional budget.
+While the underlying formulation for a statistical quorum was to establish credibility guarantees
+before making a decision, a natural consequence in common scenarios of the kind just described
+leads to a reduction in decision-making efficiency.
 How can we maintain credibility guarantees while simultaneously boosting efficiency?
 
-The solution is to infer the \textit{relevant} population size that cares about a given proposal so that we may establish credibility within only the subset population (naturally requiring less global attention).
-By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs.
-\eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).
+The solution is to infer the \textit{relevant} population size that cares about a given proposal
+so that we may establish credibility within only the subset population (naturally requiring less
+global attention).
+By replacing the absolute upper bound on this relevant population size ($N_i$) with a high
+probability upper bound, we can continue to use the same formula presented in Eqs.~\eqref{EQ2}
+and \eqref{EQ4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn
+modifies $T_i$, $K_i$, and $V_i$).
 
-To infer $N_i$, we turn to a dynamic voting process whereby information embedded within the dynamics is exploited.
-By modeling the arrival rate of new votes as a nonhomogeneous Poisson process, we can extrapolate into the future.
-For a decaying arrival rate of new votes, a nonhomogeneous Poisson process $\{N(t),t\geq 0\}$ probabilistically models the number of future arrivals over a given time interval $[0,t)$ according to:
+To infer $N_i$, we turn to a dynamic voting process whereby information embedded within the
+dynamics is exploited.
+By modeling the arrival rate of new votes as a nonhomogeneous Poisson process, we can extrapolate
+into the future.
+For a decaying arrival rate of new votes, a nonhomogeneous Poisson process
+$\{N(t),t\geq 0\}$ probabilistically models the number of future arrivals over a given time
+interval $[0,t)$ according to:
 \begin{align*}
@@
 For our purposes, we make a natural choice,
-\begin{align}\label{eq6}
+\begin{align}\label{EQ6}
 \lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}},
@@
 \begin{description}
-\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that sufficiently smooths the historical rate),
-\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data for the community via a regression analysis.
+\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that
+sufficiently smooths the historical rate),
+\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data
+for the community via a regression analysis.
 \end{description}
-Now, for example, to upper bound with $95\%$ probability the remaining number of votes still pending, we wish to solve for $\hat{K_i}$ such that the following inequality is satisfied:
-\begin{align}\label{eq7}
+Now, for example, to upper bound with $95\%$ probability the remaining number of votes still
+pending, we wish to solve for $\hat{K_i}$ such that the following inequality is satisfied:
+\begin{align}\label{EQ7}
 \lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda} \geq 95\%,
@@
 where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i)}{\ln{1/\gamma}}$.
-Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$, and can proceed to check for a statistical quorum using Eq. \ref{eq5}.
+Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$,
+and can proceed to check for a statistical quorum using Eq.~\ref{EQ5}.
 
@@
 measuring the time the application was focused. Other information we used for evaluation are available
-through regular features of the application: how many and who made discussion items, comments, motions
-(proposed decisions on discussion items), and vote tallies per motion.
+through regular features of the application: how many and who made discussion items, comments, proposals
+(proposed decisions on discussion items), and vote tallies per proposal.
 
 For the baseline, we extracted similar information from council minutes: how long did the councils last, how many
-discussions were opened, how many comments by which members were made, how many motions were proposed,
+discussions were opened, how many comments by which members were made, how many proposals were proposed,
 and vote tallies for them.
@@
 Each user can open and propose new discussion items, which leads them to the main user interface for each discussion as seen in Figure~\ref{fig:peer-mind}, having three sections: commenting section on the left, summary points in the middle,
-and proposed motions with voting on the right. All content inputs fields accept rich text content in WYSWYG manner,
+and proposed proposals with voting on the right. All content inputs fields accept rich text content in WYSWYG manner,
 so users can easily add links, images, and other files.
@@
 A general workflow after creation of a discussion item is that users discuss it through comments,
-which moderators summarize into summary points. At some point an user proposes a motion which can be
-further discussed through comments. Once a moderator deems that motion is ready to be voted upon
+which moderators summarize into summary points. At some point an user proposes a proposal which can be
+further discussed through comments. Once a moderator deems that proposal is ready to be voted upon
 can open it for voting. Users vote through a ballot seen in Figure~\ref{fig:ballot}.
@@
 as seen in Figure~\ref{fig:results}. Users can see both the current average score of all votes (score
-larger than 0 means that a motion is passing) and the credibility of current result. A community can decide
+larger than 0 means that a proposal is passing) and the credibility of current result. A community can decide
 how high credibility it requires to deem the result passing statistical quorum. During our evaluation
@@
 online. Facilitators of in-person meetings can use it to prepare agendas for meetings, take notes,
-and display summary points of current discussion and what motions are being voted on. Through that the
+and display summary points of current discussion and what proposals are being voted on. Through that the
 application can also serve as way to integrate both online and in-person decision making. Having all of
@@
 \caption{User interface of PeerMind web application. One can see three column structure, with commenting
-section on the left, summary points in the middle, and proposed motions with voting on the right.}
+section on the left, summary points in the middle, and proposed proposals with voting on the right.}
 \label{fig:peer-mind}
@@
 
-\section{Derivation of Eq. \eqref{eq2}}
+\section{Derivation of Eq.~\eqref{EQ2}}
 In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.
@@
 
-\section{Derivation of Eq. \eqref{eq3}}
+\section{Derivation of Eq.~\eqref{EQ3}}
 \begin{align*}
@@
 
-\section{Derivation of Eq. \eqref{eq5}}
+\section{Derivation of Eq.~\eqref{EQ5}}
 Assuming independence of votes across proposals,
@@
 
-\section{Derivation of Eq. \eqref{eq6}}
+\section{Derivation of Eq.~\eqref{EQ6}}
 The properties we desire for $\lambda(t)$ are:
@@
 
-\section{Derivation of Eq. \eqref{eq7}}
+\section{Derivation of Eq.~\eqref{EQ7}}
 Using $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$, we have

### 2018-04-19 statistical-quorum.tex
Updating image sizes.

@@
 
-We built a full-fletched mobile-friendly web applicationusing Meteor web framework~\cite{meteor}.
+We built a full-fletched mobile-friendly web application using Meteor web framework~\cite{meteor}.
 Meteor framework supports reactivity which makes the application update user interface based on actions
@@
 
-PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
-anonymized dataset of interactions with the app, including votes, is available at \url{http://redacted.for.blind.review}.
+PeerMind source code is available under open source licence at \url{https://redacted.for.blind.review} and
+anonymized dataset of interactions with the app, including votes, is available at \url{https://redacted.for.blind.review}.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/peermind.png}
+\includegraphics[width=0.9\textwidth]{figures/peermind.png}
 \caption{User interface of PeerMind web application. One can see three column structure, with commenting
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/results.png}
+\includegraphics[width=0.6\textwidth]{figures/results.png}
 \caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of the vote.

### 2018-04-19 statistical-quorum.tex
The vote -> voting.

@@
 cannot yet conclude that the decision to accept the proposal is credible enough for the entire
-community to officially adopt, despite the significant participation in the vote.
+community to officially adopt, despite the significant participation in voting.
 Instead, the suggested course of action would be to continue gathering more votes and/or further
@@
 \item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed
-for the vote
 (e.g., $f=\frac{1}{2}$ for a simple majority).
@@
 An important design decision in the scope of this paper is that during voting, users can see the current
-results of the vote in real time. Results are shown on a chart depicting average score through time
+results of voting in real time. Results are shown on a chart depicting average score through time
 as seen in Figure~\ref{fig:results}. Users can see both the current average score of all votes (score
@@
 
-The rationale for displaying the current results during voting is to give users feedback how the vote
+The rationale for displaying the current results during voting is to give users feedback how voting
 is progressing and allow them to engage more with the discussion. Moreover, displaying the results
@@
 
-After the vote passes the statistical quorum, a moderator can close voting and a discussion item.
+After voting passes the statistical quorum, a moderator can close voting and a discussion item.
 
@@
 where -1 means opposition, 1 means in favor, and 0 indifference. They can also chose abstention (they
-remove themselves from the population of the community for the vote in question), or to not vote (default), which
+remove themselves from the population of the community for voting in question), or to not vote (default), which
 means that their vote does not contribute to the quorum.}
@@
 \includegraphics[width=0.6\textwidth]{figures/results.png}
-\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of the vote.
+\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of voting.
 A user can hover over the chart to see details at that time. Users can see these results in real time during voting.}
@@
 issues with quorum
-unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after the vote
+unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after voting
 

### 2018-04-19 statistical-quorum.tex
Updating methodology.

@@
 
-For the baseline, we extracted similar information from council minutes: how long did the councils last, how many
-discussions were opened, how many comments by which members were made, how many proposals were proposed,
-and vote tallies for them.
+For the baseline, we extracted similar information from council minutes: how long did the councils last,
+how many members participated, how many proposals were proposed and how many passed, and vote tallies for them.
 
@@
 We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
-Then we evaluate it through real-world use of the PeerMind application by a 140 members community.
+Then we do qualitative and quantitative evaluation through real-world use of the PeerMind application by a 140 members
+community.
 

### 2018-04-19 statistical-quorum.tex
Things were less clear.

@@
 
-In Figure~\ref{fig:survey-comparison} we see questions and results of the survey we made at the end of each semester.
+In Figure~\ref{fig:survey-comparison} we see questions and results of the survey we made at the end of both semesters.
 
 We can see that in fall members were more satisfied with the decision making in the community and only how understandable
-is the whole process received a decrease. We attribute this to the fact that the use of the web application is
-a novelty and requires more effort in educating members than a traditional decision making. Moreover, decision
-making in person allows one to observe others how they are participating and learn from them, instead of
-being alone using the application.
+is the whole process decreased.
+We attribute this to the fact that the use of the web application is
+a novelty, including using a novel concept like statistical quroum, and requires more effort
+in educating members than a traditional decision making.
+Moreover, decision making in person allows one to observe others how they are participating and
+learn from them, instead of being alone using the application.
+We explore this question also through interviews in Section~\ref{sec:interviews}.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
-\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
+\includegraphics[width=0.9\textwidth]{figures/survey-comparison.pdf}
+\caption{Results of a survey at the end of spring (baseline) and fall (using PeerMind) semesters about general perception of the decision making (councils) in the community.
 For each question, a participant was asked to rank the answer on a scale between 1 (completely disagree) and 7 (completely agree). All members
@@
 \subsection{Interviews}
+\label{sec:interviews}
 

### 2018-04-19 statistical-quorum.tex
Updating title.

@@
 % Title portion. Note the short title for running heads
-\title[Improving credibility and efficiency of decision-making]{Improving credibility and efficiency of decision-making \\ within democratic communities via statistical quorum}
+\title[Improving credibility and efficiency of decision-making]{Improving credibility and efficiency of decision-making \\ within democratic communities via a statistical quorum}
 

### 2018-04-19 statistical-quorum.tex
theory updates, plus figures

--- 2018-04-19 statistical-quorum.tex
@@
 % Title portion. Note the short title for running heads
-\title[Improving credibility and efficiency of decision-making]{Improving credibility and efficiency of decision-making \\ within democratic communities via statistical quorum}
+\title{Improving credibility and efficiency of decision-making within democratic communities via a statistical quorum}
 
@@
   \institution{UC Berkeley}}
-\email{maxowaxo2013@gmail.com}
+\email{mkanwal@berkeley.edu}
 \author{David Marn}
@@
 \begin{abstract}
-The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
-While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
-different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
-maximize productivity?
-We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to
-focus attention, in turn making the entire decision-making process dynamic.
-We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
-councils over the course of 16 weeks.
-We found that the community was able to use the web application to simultaneously improve the credibility and efficiency
-of their decision-making process.
+The Achilles heel of many major democracies is the perpetual need for sufficient engagement and attention from the community.
+While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?
+We propose a novel \textit{statistical quorum}, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic.
+We developed a web application centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 16 weeks.
+We found that the community was able to use the web application to simultaneously improve the credibility and efficiency of their decision-making process.
 \end{abstract}
@@
 
+Two of the most shocking democratic votes in recent history occurred in 2016: The referendum vote for Britain to leave the European Union (“Brexit”) and the election of Donald Trump as President of the United States.
+Both votes were decided by the slimmest of margins, but more significantly, both votes also had poor turnout.
+The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
+Hence, there have been major efforts towards increasing these attributes within the community.
+We believe this is a fallacy.
+Democracy need not suffer under low engagement — in fact, our goal should be to make democracy work with minimal engagement.
+The ideal governance system is one where citizens do not have to be constantly involved, but the system is still acting as the citizens would like it to.
+We explore a fundamentally different approach to making democracy work: \textit{For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?} Can we increase the credibility of decisions made within a democracy while simultaneously skirting the need for an increase in attention?
+We formulated a novel ‘statistical quorum’, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic.
+We have developed a web application centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 16 weeks.
+We found that the community was able to use the web application to simultaneously improve the legitimacy and efficiency of their decision-making process.
+Our prototyped web application for collective decision-making contrasts with existing similar solutions.
+Our web application integrates new ideas in how to improve this process: use of \textbf{range voting}, a \textbf{dynamic statistical quorum}, and \textbf{voter delegation}.
+The broad idea is that, instead of just trying to move the voting process online — something many others have done previously — we in fact wish to add value by voting online.
+Specifically, we can compute results faster, more accurately, and securely, we can use powerful statistical methods to estimate important quantities for the entire voting population, we can allow for members to entrust their vote to others who have more expertise on specific topics than they themselves do, and we can promote clarity of thought and complete understanding among all members for all voting items.
+\textit{Range voting} is voting where members of a community express their positions on open proposals by assigning each a score between –1 and +1.
+Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their positions in a highly informative and precise manner.
+Moreover, if a new proposal is added to a set of existing proposals on which voting was already open, only the new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of each proposal would need reassessing.
+A \textit{dynamic statistical quorum} uses statistical hypothesis testing to compute the minimum number of required votes needed to achieve a desired level of confidence before a community adopts an outcome as being truly representative of its members’ positions.
+For example, in a community of 100 members, if 20 members all vote in full support of a proposal, then we can conclude with very high confidence that, had all 100 members actually voted, a majority of them would support the proposal (despite in actuality 80 members having not yet voted).
+On the other hand, if 80 members vote with 41 voting in favor of the proposal and 39 voting in opposition, then the community cannot yet conclude anything with high confidence (despite the high voter turnout).
+Hence, the suggested course of action would be for the community to gather more votes and continue discussing the proposal.
+By utilizing a dynamic statistical quorum, we can direct attention and participation within the community towards contentious proposals, while reducing wasted attention directed towards proposals that already have overwhelming support.
+\textit{Voter delegation} addresses two questionable assumptions about current democratic systems: (1) participants have the time to participate and (2) participants have the knowledge to participate.
+Not all members of a community can participate at all times and not all members have the time to educate themselves about each and every decision being made.
+By allowing members to define delegates to whom they can give their own vote to — likely those who share their values — their proxy vote can be inferred from their delegates’ votes.
+In this way, their values are still voiced even if they do not vote directly.
+If at any point a member decides to vote themselves, then their direct vote is used instead.
+All of the above features are combined into a guided process to help community members participate in decision-making.
+The process starts with a new discussion item where members can comment and discuss a proposal.
+Moderators then summarize the discussion into three types of points: “in favor,” “against,” and “other”.
+Members can then use those summaries to help form an opinion before they vote, without having to read every discussion comment made in full detail, thus lowering the barrier of entry into any discussion.
+As members vote, the app shows the community the average score and confidence in real-time.
+When the confidence passes a preset threshold, the proposal can be firmly accepted by the community.
+
 In many historical and recent elections it is not infrequent that we get close results (see Table~\ref{tab:example}).
@@
   invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:decision-credibility}. We see that
-  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
-  with more votes any option could end up winning.}\label{tab:example}
+  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so with a larger turnout, any option could have ended up winning.}\label{tab:example}
   \vspace{-0.3cm}
@@
 
-Our contributions are:
+Our main contributions are:
 
-\begin{itemize}
+\begin{enumerate}
 \item We developed a framework of computing statistical credibility of results
 based on sample size, votes cast, and voting duration.
-\item Based on this framework we developed a novel concept of \emph{statistical quorum}
-as a threshold of votes required to deem results credible and legitimate.
-\item We implemented web application \emph{PeerMind},
+\item Based on this framework we developed the novel concept of a \emph{statistical quorum} as a threshold of votes required to deem results credible and legitimate.
+\item We implement a web application \emph{PeerMind},
 integrating credibility of results and statistical quorum, and designed
-a workflow of its use in a decision making processes.
+a workflow of its use in decision making processes.
 \item We deployed the application in a democratically run community of 140
@@
 and increases number of decisions made.
-\end{itemize}
+\end{enumerate}
 
@@
 %\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
+\section{Related Work}
+\label{sec:related-work}
+
+%\todo[inline]{Repeat references from "There have been many improvements proposed to this basic form and various analysis made."?}
+%
+%\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
+%How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
+%
+%\todo[inline]{Voting auditing work.}
+%
+%\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
+%
+%\section{PeerMind Web Application}
+%\label{sec:peermind}
+%
+%\todo[inline]{Something we need to think about is the level of detail we want to go into regarding the implementation/features of PeerMind beyond statistical quorum and range voting.
+I think we might want to focus on only those two features for this paper.\\
+%I think PeerMind as a comprehensive platform should have its own paper where all of the data is analyzed across the board.
+This would be a much more typical HCI paper.
+(If I can become president of Ridge in a future semester, we'll have another chance to run a clean, well-designed experiment with a much more developed version of PeerMind.)\\
+%The message in the current paper might be clearest if we keep the scope focused on the theory + evaluation/verification.  We should make sure we're not trying to pack too much into one paper.
+%\\---Max}
+%
+%We built a full-fletched real-time web application for collective decision making and voting using Meteor
+%framework~\cite{meteor}.
+%Users can initiate new discussions which leads them to the main user interface for each discussion as seen
+%in Figure~\ref{fig:peermind}.
+%
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=1.75\columnwidth]{figures/peermind.png}
+%  \caption{The main PeerMind interface of a discussion. There is initial discussion description on the top and under it there
+%  are three columns. A commenting sections for users to discuss it and proposed decisions, a second column where moderators can
+%  summarize arguments made during discussion, and a list of proposals to address the discussion.
+%  Proposals can be in different states: a proposal being discussed, a proposal being voted on, and with voting concluded.}\label{fig:peermind}
+%\end{figure*}
+%
+%%and discuss them with others online in real-time and with rich-text content.
+%%They can attach images and files using drag\&drop. Moderators can summarize discussions into three groups
+%
+%\todo[inline]{Present how users can see ongoing results during voting. Show an image of active ballot, but with expanded view for current results.}
+%
+%%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
+%%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
+%%A proposal passes if the average score of all votes on a position is positive.
+%%When multiple competing proposals are being voted on, the one with highest average score wins.
+%%
+%%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
+%%positions in a highly informative and precise manner.
+%%We find it especially suitable for a web application because many web users are used to use it when ranking products
+%%and giving other ``five-star'' feedback.
+%%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
+%%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
+%%each proposal would need reassessing (assuming members use a consistent personal scale when voting).
+%
+%\todo[inline]{The app has features both for use in online and offline setting. This allows one to compute
+%statistical quorum also for an in-person meeting because math is now more complicated than just counting hands.
+%We focus on online experience in this paper.}
+%
+%\todo[inline]{Security is out of the scope of this work. We assume here that users and their identities are known and are managed
+%by a community's admin. There are many communities where their decision making is public with public minutes and this is who
+%the app is currently targeting.}
 
@@
 \label{sec:decision-credibility}
+A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
+For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose, then we can conclude with high probability that, had all 100 members taken the time to vote, a majority of them would be in support of the proposal.
+Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote was credible enough to be taken as representative of the entire community's desire.
+On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in the vote.
+Instead, the suggested course of action would be to continue gathering more votes and/or further discussing the proposal until a credible outcome is established.
 
-%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
-%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
-%A proposal passes if the average score of all votes on a position is positive.
-%When multiple competing proposals are being voted on, the one with highest average score wins.
-%
-%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
-%positions in a highly informative and precise manner.
-%We find it especially suitable for a web application because many web users are used to use it when ranking products
-%and giving other ``five-star'' feedback.
-%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
-%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
-%each proposal would need reassessing (assuming members use a consistent personal scale when voting).
-
-A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome
-of an ongoing voting before the community adopts the outcome as being truly representative of all
-members' positions.
-For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose,
-then we can conclude with high probability that, had all 100 members taken the time to vote, a
-majority of them would be in support of the proposal.
-Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote is
-credible enough to be taken as representative of the entire community's desire.
-On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we
-cannot yet conclude that the decision to accept the proposal is credible enough for the entire
-community to officially adopt, despite the significant participation in voting.
-Instead, the suggested course of action would be to continue gathering more votes and/or further
-discussing the proposal until a credible outcome is established.
-
-By utilizing a statistical quorum, a community can actively direct attention and participation
-within the community towards contentious proposals, while avoiding wasting attention on
-noncontroversial proposals.
+By utilizing a statistical quorum, a community can actively direct attention and participation within the community towards contentious proposals, while avoiding wasting attention on noncontroversial proposals.
 
 \subsection{Background}
-Consider a community that needs to collectively decide which proposal (if any) to accept out of
-a set of competing proposals.
-To act in democratic fashion, the community may wish to select a winning proposal that either has
-the support of a \{simple, super\}-majority of all members, or instead is demonstrably preferred
-over all other competing proposals.
+Consider a community that needs to collectively decide which proposal (if any) to accept out of a set of competing proposals.
+To act in democratic fashion, the community may wish to select a winning proposal that either has the support of a \{simple, super\}-majority of all members, or instead is demonstrably preferred over all other competing proposals.
 (In some cases, the community may want a winning proposal to satisfy both criteria.)
 
-Ideally, we do not want to poll all members in order to officially accept a proposal, but we do
-want some level of certainty (e.g., $90\%$ certainty) that the population is not accidentally
-accepting an undesired proposal, which in this case encompasses accepting a proposal that is
-not supported by a majority of members and/or is not clearly preferred over all other competing
-alternatives.
+Ideally, we do not want to poll all members in order to officially accept a proposal, but we do want some level of certainty (e.g., $90\%$ certainty) that
+the population is not accidentally accepting an undesired proposal, which in this case encompasses accepting a proposal that is not supported by a majority of members and/or is not clearly preferred over all other competing alternatives.
 
-The primary motivation behind using a statistical quorum is to guarantee a level of credibility
-before making a decision on behalf of the entire community.
+The primary motivation behind using a statistical quorum is to guarantee a level of credibility before letting a subset of members act on behalf of the entire community.
 
 \subsection{Problem Statement}
-We wish to compute the probability that the bias in a community towards supporting a
-proposal is greater than some fraction of the population (e.g., $50\%$ for a simple majority,
-$67\%$ for a super majority).
+We wish to compute the probability that the bias in a community towards supporting a proposal is greater than some
+fraction of the population (e.g., $50\%$ for a simple majority, $67\%$ for a super majority).
 
-In addition, we wish to compute the probability that the community bias towards supporting
-a specific proposal is greater than the community bias towards supporting any competing proposals.
+In addition, we wish to compute the probability that the community bias towards supporting a specific proposal is greater than the community bias towards supporting any competing proposals.
 
@@
 We assume that there is a fixed latent bias for the support of a proposal within the community.
-We take each individual vote to be independent and identically distributed (i.i.d.) according
-to a Bernoulli distribution around this bias.
-It should be noted that the independence assumption may not hold in some settings, such as when
-votes are accumulated over time (one may observe surges of correlated votes).
-We will partially address this when refining our original formulation of a statistical quorum
-into a \textit{dynamic} statistical quorum.
+We take each individual vote to be independent and identically distributed (i.i.d.) according to a Bernoulli distribution around this bias.
+It should be noted that the independence assumption may not hold in some settings, such as when votes are accumulated over time (one may observe surges of correlated votes).
+We will attempt to address this when refining our original formulation of a statistical quorum into a \textit{dynamic} statistical quorum.
 
-Without having collected any votes, we assume that the community bias may be any value between
-$0\%$ and $100\%$ with equal probability (uniform prior).
+Without having collected any votes, we assume that the community's overall bias may be any value between $0\%$ and $100\%$ with equal probability (uniform prior).
 
 \subsection{Voting}
-We allow for range voting where each member of the community may choose any real-valued vote
-$v \in [-1,+1]$ to represent their vote, with $v = -1$ representing full opposition and $v=+1$
-representing full support.
-In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning
-the full range, hence our treatment of $v$ as a discrete random variable.
-Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as
-an individual's average vote.
-In the following subsection, we rescale $v$ to be in the range $[0,1]$ via the
-notation $\frac{v + 1}{2}$.
+
+%\url{https://youtu.be/Q60ZXoXP6Hg}
+
+We allow for range voting where each member of the community may choose any real-valued vote $v \in [-1,+1]$ to
+represent their vote, with $v = -1$ representing full opposition and $v=+1$ representing full support.
+In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning the full range, hence our treatment of $v$ as a discrete random variable.
+Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as an individual's average vote.
+In the theory subsection below, we rescale $v$ to be in the range $[0,1]$ via the notation $\frac{v + 1}{2}$.
 
 We allow voters to ``abstain'' from voting.
-We treat abstentions as indicating that a voter is removing themselves from the community on a
-given vote.
-It is important to distinguish an ``abstention'' (removing one's self from the voting population)
-from voting $0$ (indicating indifference).
+We treat abstentions as indicating that a voter is removing themselves from the community on a given vote.
+It is important to distinguish an ``abstention'' (removing one's self from the voting population) from voting $0$ (indicating indifference).
 
@@
 
-\todo[inline]{Weave in a single concrete example to ground all the abstract theory so that the
-entire derivation and motivation has a clear and cohesive thread from start to finish.}
+\todo[inline]{Weave in a single concrete example to ground all the abstract theory so that the entire derivation and motivation has a clear and cohesive thread from start to finish.}
 
-After collecting an initial set of votes from members, we want to use that data to construct
-a prior distribution to help estimate the posterior probability that the proposal is acceptable
-to the entire community.
-Because the sum of the collected votes follows a binomial distribution, a commonplace and
-convenient choice for the form of the prior on the community's bias is the Beta distribution
-(due to its conjugacy with the binomial distribution).
-Together, they form the Beta-binomial distribution (related to the P\'olya urn model).
+\todo[inline]{Hyperparameters: Explain not only what you did, but also why you did it, so that readers (including reviewers) can be convinced that you made appropriate choices.}
 
-After extracting the sufficient statistics from raw votes --- namely, the average score
-and the number of abstentions --- we can adjust the effective community size and corresponding
-threshold.
+After collecting an initial set of votes from members, we want to use that data to construct a prior distribution to help estimate the posterior probability that the proposal is acceptable to the entire community.
+Because the sum of  the collected votes follows a binomial distribution, a commonplace and convenient choice for the form of the prior on the community's bias is the Beta distribution (due to its conjugacy with the binomial distribution).
+Together, they form the Beta-binomial distribution (also known as the P\'olya urn model).
+
+After extracting the sufficient statistics from raw votes --- namely, the average score and the number of abstentions --- we can adjust the effective community size and corresponding threshold.
 In addition, we can calculate the effective number of ``yes'' ($+1$) and ``no'' ($-1$) votes:
@@
 Let
-\begin{description}
+
+\begin{itemize}
 \item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing proposals,
 \item[$N = $] population size,
-\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding
-abstentions,
-\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting
-on proposal $X_i$,
-\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number
-of ``yes'' votes on proposal $X_i$,
-\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on proposal $X_i$,
+\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding abstentions,
+\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on proposal $X_i$,
+\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
+\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
 \item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
 \item[$p_i = $] true latent community bias towards voting ``yes'' on $X_i$.
-\end{description}
+\end{itemize}
 
-In Eqs.~\eqref{EQ2}, \eqref{EQ3}, \eqref{EQ5}, \eqref{EQ6}, \eqref{EQ7} that follow, we refer the
-reader to the \hyperref[sec:derivation]{Appendix} for complete derivation details and justifications.
+In Eqs.
+\eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we urge the reader to refer to the \hyperref[sec:derivation]{Appendix} for complete derivation details, including motivations and justifications.
 
 When considering a single proposal $X_i$ in isolation, we wish to compute:
-\begin{align}\label{EQ1}
+\begin{align}\label{eq1}
 \Pr(\alpha^*_i > T_i \mid X_i),
@@
 where
-\begin{description}
-\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire
-population,
-\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed
+\begin{itemize}
+\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
+\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
 (e.g., $f=\frac{1}{2}$ for a simple majority).
-\end{description}
+\end{itemize}
 Using a desirable prior 
-\begin{align}\label{EQ2}
+\begin{align}\label{eq2}
 p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
 \end{align}
-corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can
-express \eqref{EQ1} as:
-\begin{align}\label{EQ3}
+corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can express \eqref{eq1} as:
+\begin{align}\label{eq3}
 \Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
@@
 \end{align*}
-\todo[inline]{Note that we choose $c=10$ because it leads to intuitive results.}
 where
-\begin{description}
-\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes''
-votes $X_i$ could receive such that it would remain below threshold.
-\end{description}
+\begin{itemize}
+\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes $X_i$ could receive such that it would remain below threshold.,
+\item[$c = $] $10$, a hyperparameter chosen by the community based on their relative concern about false alarms versus (temporarily) missed detection.
+\end{itemize}
 
 When considering a proposal $X_i$ relative to other proposals in $C$, we wish to compute
-\begin{align}\label{EQ4}
+\begin{align}\label{eq4}
 \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right)
@@
 
-Under the assumption that the community's votes on proposals are independent across proposals
-(which, in many instances, will actually give a lower bound on \eqref{EQ3} when a community's
-support of one proposal is negatively correlated with its support on a competing proposal),
-we can decompose \eqref{EQ3} as follows:
-\begin{align}\label{EQ5}
-& \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) \nonumber \\
-= & \prod_{j \neq i}\left(1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \Pr(\nu_j = k) \cdot {CDF}_{\nu_i}(\alpha_j + k) \right)
+Under the assumption that the community's votes on motions are independent across motions (which, in many instances, will actually give a lower bound on \eqref{eq3} when a community's support of one proposal is negatively correlated with its support on a competing proposal), we can decompose \eqref{eq3} as follows:
+\begin{align}\label{eq5}
+& \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) = \prod_{j \neq i}\left(1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \Pr(\nu_j = k) \cdot {CDF}_{\nu_i}(k-\delta_{ij}) \right)
 \end{align}
 where
-\begin{description}
-\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$
-leads $X_j$ by.
-\end{description}
+\begin{itemize}
+\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
+\end{itemize}
 
-Using Eqs.~\eqref{EQ2} and \eqref{EQ4}, we can numerically compute the credibility of a
-vote's outcome in two key scenarios relevant for democratic communities.
-For example, a community may only consider accepting a proposal when a vote's credibility
-is above $90\%$.
+Using Eqs.
+\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
+For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.
 
 \subsection{Dynamic Statistical Quorum}
-Oftentimes not all proposals are relevant to the entire community.
-Instead, a niche proposal may only affect a subset of the community, and hence to require much
-of the community to indicate their indifference and/or decide their opinion is a waste of their
-attentional budget.
-While the underlying formulation for a statistical quorum was to establish credibility guarantees
-before making a decision, a natural consequence in common scenarios of the kind just described
-leads to a reduction in decision-making efficiency.
+Oftentimes, not all proposals are relevant to the entire community.
+Instead, a niche proposal may only affect a subset of the community, and hence to require much of the community to indicate their indifference and/or decide their opinion is a waste of the their fixed attentional budget.
+While the underlying formulation for a statistical quorum was to establish credibility guarantees before taking action, a natural consequence in common scenarios of the kind just described leads to a reduction in decision-making efficiency.
 How can we maintain credibility guarantees while simultaneously boosting efficiency?
 
-The solution is to infer the \textit{relevant} population size that cares about a given proposal
-so that we may establish credibility within only the subset population (naturally requiring less
-global attention).
-By replacing the absolute upper bound on this relevant population size ($N_i$) with a high
-probability upper bound, we can continue to use the same formula presented in Eqs.~\eqref{EQ2}
-and \eqref{EQ4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn
-modifies $T_i$, $K_i$, and $V_i$).
+Our proposed solution is to infer the \textit{relevant} population size that cares about a given proposal so that we may establish credibility within only the subset population (naturally requiring less global attention).
+By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs.
+\eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).
 
-To infer $N_i$, we turn to a dynamic voting process whereby information embedded within the
-dynamics is exploited.
-By modeling the arrival rate of new votes as a nonhomogeneous Poisson process, we can extrapolate
-into the future.
-For a decaying arrival rate of new votes, a nonhomogeneous Poisson process
-$\{N(t),t\geq 0\}$ probabilistically models the number of future arrivals over a given time
-interval $[0,t)$ according to:
+To infer $N_i$, we turn to a dynamic voting process whereby information embedded within the dynamics is exploited.
+By modeling the arrival rate of new votes as a nonhomogeneous Poisson process, we can extrapolate into the future.
+For a decaying arrival rate of new votes, a nonhomogeneous Poisson process $\{N(t),t\geq 0\}$ probabilistically models the number of future arrivals over a given time interval $[0,t)$ according to:
 \begin{align*}
@@
 For our purposes, we make a natural choice,
-\begin{align}\label{EQ6}
+\begin{align}\label{eq6}
 \lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}},
@@
 where
-\begin{description}
-\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that
-sufficiently smooths the historical rate),
-\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data
-for the community via a regression analysis.
-\end{description}
-Now, for example, to upper bound with $95\%$ probability the remaining number of votes still
-pending, we wish to solve for $\hat{K_i}$ such that the following inequality is satisfied:
-\begin{align}\label{EQ7}
+\begin{itemize}
+\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that sufficiently smooths the historical rate),
+\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data for the community via a regression analysis.
+\end{itemize}
+Now, for example, to upper bound with $95\%$ probability the remaining number of votes still pending, we wish to solve for $\hat{K_i}$ such that the following inequality is satisfied:
+\begin{align}\label{eq7}
 \lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda} \geq 95\%,
 \end{align}
-where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i)}{\ln{1/\gamma}}$.
-Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$,
-and can proceed to check for a statistical quorum using Eq.~\ref{EQ5}.
+where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i}{\ln{(1/\gamma)}}$.
+Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$, and can proceed to check for a statistical quorum using Eq. \ref{eq5}.
 
@@
 
-For the baseline, we extracted similar information from council minutes: how long did the councils last,
-how many members participated, how many proposals were proposed and how many passed, and vote tallies for them.
+For the baseline, we extracted similar information from council minutes: how long did the councils last, how many
+discussions were opened, how many comments by which members were made, how many proposals were proposed,
+and vote tallies for them.
 
@@
 
-We built a full-fletched mobile-friendly web application using Meteor web framework~\cite{meteor}.
+We built a full-fletched mobile-friendly web applicationusing Meteor web framework~\cite{meteor}.
 Meteor framework supports reactivity which makes the application update user interface based on actions
@@
 An important design decision in the scope of this paper is that during voting, users can see the current
-results of voting in real time. Results are shown on a chart depicting average score through time
+results of the vote in real time. Results are shown on a chart depicting average score through time
 as seen in Figure~\ref{fig:results}. Users can see both the current average score of all votes (score
@@
 
-The rationale for displaying the current results during voting is to give users feedback how voting
+The rationale for displaying the current results during voting is to give users feedback how the vote
 is progressing and allow them to engage more with the discussion. Moreover, displaying the results
@@
 
-After voting passes the statistical quorum, a moderator can close voting and a discussion item.
+After the vote passes the statistical quorum, a moderator can close voting and a discussion item.
 
@@
 
-PeerMind source code is available under open source licence at \url{https://redacted.for.blind.review} and
-anonymized dataset of interactions with the app, including votes, is available at \url{https://redacted.for.blind.review}.
+PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
+anonymized dataset of interactions with the app, including votes, is available at \url{http://redacted.for.blind.review}.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/peermind.png}
+\includegraphics[width=0.5\textwidth]{figures/peermind.png}
 \caption{User interface of PeerMind web application. One can see three column structure, with commenting
@@
 where -1 means opposition, 1 means in favor, and 0 indifference. They can also chose abstention (they
-remove themselves from the population of the community for voting in question), or to not vote (default), which
+remove themselves from the population of the community for the vote in question), or to not vote (default), which
 means that their vote does not contribute to the quorum.}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/results.png}
-\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of voting.
+\includegraphics[width=0.5\textwidth]{figures/results.png}
+\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of the vote.
 A user can hover over the chart to see details at that time. Users can see these results in real time during voting.}
@@
 We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
-Then we do qualitative and quantitative evaluation through real-world use of the PeerMind application by a 140 members
-community.
+Then we evaluate it through real-world use of the PeerMind application by a 140 members community.
 
@@
 4 Figures and 1 Table Pending}
+
+To gain an intuitive grasp of the decision credibility used in a statistical quorum, we can examine two toy scenarios.
+
+In Figure \ref{fig:credible_interval}, we examine how the $90\%$ credible interval converges as more votes are collected out of a population of $100$.
+A statistical quorum is achieved when the credible interval is contained entirely above the x-axis.
+In this example, the latent bias of the community is that $75\%$ of the community supports the proposal while $25\%$ opposes it.
+
+In 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.45\textwidth]{figures/credible_interval.pdf}
+\caption{.}
+\label{fig:credible_interval}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=.45\textwidth]{figures/simple_majority.pdf}
+\caption{.}
+\label{fig:simple_majority}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=.45\textwidth]{figures/competing_proposals.pdf}
+\caption{.}
+\label{fig:simple_majority}
+\end{figure}
 
@@
 issues with quorum
-unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after voting
+unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after the vote
 
@@
 \appendix
+\section*{Appendix}
 \label{sec:derivation}
@@
 
-\section{Derivation of Eq.~\eqref{EQ2}}
+\subsection*{Derivation of Eq. \eqref{eq2}}
 In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.
 To update the prior based on collected votes, standard theory suggests updating to $\operatorname{Beta}(1+\alpha_i,1+\beta_i)$; however, there are a few consistency issues with this update that we seek to fix:
-\begin{description}
+\begin{itemize}
 \item[$(1)$] The updated prior should not depend on the absolute number of votes $\alpha_i$, $\beta_i$, as this is not scale-invariant; rather, it should depend on the relative proportions (which are unitless instead of a number of votes),
@@
 \item[$(3)$] as $\alpha_i + \beta_i \rightarrow N_i$, $\operatorname{Var}[p_i] \rightarrow 0$ (at a rate independent of $N_i$).
-\end{description}
+\end{itemize}
 To satisfy these consistency properties, we seek to linearly update the prior's scale parameters in a more natural way: $\operatorname{Beta}(1+\eta\alpha_i,1+\eta\beta_i)$ for some renormalizing factor $\eta$.
@@
 
-Hence we arrive at $\eta = \frac{c}{K_i}$, where $c$ is any constant that can be selected to tune the rate at which the variance approaches $0$.  In this study, we prefer $\eta = \frac{10}{K_i}$.
+Hence we arrive at $\eta = \frac{c}{K_i}$, where $c$ is any constant that can be selected to tune the rate at which the variance approaches $0$.
+In this study, we chose $\eta = \frac{10}{K_i}$ based on the preference of the community.
 
@@
 
-\section{Derivation of Eq.~\eqref{EQ3}}
+\subsection*{Derivation of Eq. \eqref{eq3}}
 \begin{align*}
@@
 
-\section{Derivation of Eq.~\eqref{EQ5}}
+\subsection*{Derivation of Eq. \eqref{eq5}}
 Assuming independence of votes across proposals,
@@
 &= \prod_{j \neq i} \left( 1 - \Pr(\alpha^*_i \leq \alpha^*_j \mid X_i, X_j) \right)\\
-&= \prod_{j \neq i} \left( 1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \underbrace{\Pr(\alpha^*_j = \alpha_j + k \mid X_j)}_{= \Pr(\nu_j = k)} \cdot \underbrace{\Pr(\alpha^*_i \leq \alpha_j + k \mid X_i)}_{={CDF}_{\nu_i}(\alpha_j + k)} \right).
+&= \prod_{j \neq i} \left( 1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \underbrace{\Pr(\alpha^*_j = \alpha_j + k \mid X_j)}_{= \Pr(\nu_j = k)} \cdot \underbrace{\Pr(\alpha^*_i \leq \alpha_j + k \mid X_i)}_{={CDF}_{\nu_i}(k - \delta_{ij})} \right).
 \end{align*}
 
-\section{Derivation of Eq.~\eqref{EQ6}}
+\subsection*{Derivation of Eq. \eqref{eq6}}
 The properties we desire for $\lambda(t)$ are:
-\begin{description}
+\begin{itemize}
 \item[$(1)$] as $K_i \rightarrow 0$, the rate of incoming votes should drop exponentially and should be independent of $N_i$,
 \item[$(2)$] as $t \rightarrow \infty$, the rate of incoming votes should drop exponentially.
-\end{description}
+\end{itemize}
 Setting $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$ for $0 < \gamma < 1$ satisfies these constraints while still allowing for the exact rate to be tunable.
 
-\section{Derivation of Eq.~\eqref{EQ7}}
+\subsection*{Derivation of Eq. \eqref{eq7}}
 Using $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$, we have
@@
 &= \int _{0}^{\infty} \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}} \mathrm{d}t\\
-&= \lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}.
+&= \lambda_0 \frac{K_i/N_i}{\ln{(1/\gamma)}}.
 \end{align*}
@@
 \begin{align*}
-\lim_{t\rightarrow \infty}\Pr\left(N(t) = n\right) = \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}},
+\lim_{t\rightarrow \infty}\Pr\left(N(t) = n\right) = \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{(1/\gamma)})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{(1/\gamma)}}},
 \end{align*}
@@
 \begin{align*}
-\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}}.
+\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{(1/\gamma)})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{(1/\gamma)}}}.
 \end{align*}

### 2018-04-19 statistical-quorum.tex

@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/results.png}
-\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of the vote.
+\includegraphics[width=0.6\textwidth]{figures/results.png}
+\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of voting.
 A user can hover over the chart to see details at that time. Users can see these results in real time during voting.}

### 2018-04-19 statistical-quorum.tex
Adding data.

--- 2018-04-19 statistical-quorum.tex
@@
 credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
-We present our methodology in Section~\ref{sec:methodology} after which we do qualitative and
-quantitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
+We present our methodology in Section~\ref{sec:methodology} after which we do
+quantitative and qualitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
 Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
@@
 anonymized dataset of interactions with the app, including votes, is available at \url{https://redacted.for.blind.review}.
+The research protocol with human subjects was approved by \emph{redacted for blind review} under
+ID \emph{redacted for blind review}.
 
@@
 We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
-Then we do qualitative and quantitative evaluation through real-world use of the PeerMind application by a 140 members
+Then we do quantitative and qualitative evaluation through real-world use of the PeerMind application by a 140 members
 community.
@@
 \subsection{Use in Practice}
+
+We tracked web application activity of 132 PeerMind users who consented to the experiment.
+We looked at time they had the web application focused in their browsers to determine the
+time they used the web application.
+In total, they spent 924.886 seconds using the web application (which translates to roughly 2 hours
+per member on average) during the fall 2016 semester.
+
+In comparison, we analyzed spring 2016 council minutes when community was meeting in-person and
+estimated that members spent approximately 2.600.000 seconds total doing decision making
+(around 5 hours per member on average).
+We estimated this based on information about beginning and end of councils multiplied
+with an estimation of the number of people attending each council.
+We estimated the number of people attending based on number of votes made for different proposals
+and the quorum size, taking the maximum of both.
+We find this a good and conservative proxy because councils have a rule that they have to adjurn
+if the number of members present fall under the quorum (42 members) for more than 10 minutes
+which in practice members discover soon through the number of votes being made or when count
+is explictly requested.
+
+Users on PeerMind made 139 proposals over the span of a semester (this figure
+excludes proposals that were withdrawn before voting began).
+Out of those 139 proposals, 98 proposals passed. In comparison, 70 out of 73 proposals passed
+in the semester prior.
+
+PeerMind users made 3958 votes, whereas 1396 votes were made during spring semester.
+Even after adjusting for the larger number of proposals made in the fall, participation
+increased significantly.
+From around 19 votes per proposal to 28 votes per proposal, on average.
+
+Plotting time spent per user on PeerMind gives further insights into the democratization of decision-making.
+While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours.
+We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes).
+Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.
 
@@
 
-We can see that in fall members were more satisfied with the decision making in the community and only how understandable
+We can see that in fall (using PeerMind) members were more satisfied with the decision making in the community and only how understandable
 is the whole process decreased.
-We attribute this to the fact that the use of the web application is
-a novelty, including using a novel concept like statistical quroum, and requires more effort
-in educating members than a traditional decision making.
-Moreover, decision making in person allows one to observe others how they are participating and
-learn from them, instead of being alone using the application.
-We explore this question also through interviews in Section~\ref{sec:interviews}.
 
@@
 \label{sec:discussion}
+
+decline in understanding of the process:
+
+We attribute this to the fact that the use of the web application is
+a novelty, including using a novel concept like statistical quorum, and requires more effort
+in educating members than a traditional decision making.
+Moreover, decision making in person allows one to observe others how they are participating and
+learn from them, instead of being alone using the application.
+We explore this question also through interviews in Section~\ref{sec:interviews}.
+
+This not only means that PeerMind users were more productive in terms of decisions made, they also spent significantly less time per motion than the control group.
 

### 2018-04-19 statistical-quorum.tex

--- 2018-04-19 statistical-quorum.tex
@@
 
+We tracked web application activity of 132 PeerMind users who consented to the experiment.
+We looked at time they had the web application focused in their browsers to determine the
+time they used the web application.
+In total, they spent 924.886 seconds using the web application (which translates to roughly 2 hours
+per member on average) during the fall 2016 semester.
+
+In comparison, we analyzed spring 2016 council minutes when community was meeting in-person and
+estimated that members spent approximately 2.600.000 seconds total doing decision making
+(around 5 hours per member on average).
+We estimated this based on information about beginning and end of councils multiplied
+with an estimation of the number of people attending each council.
+We estimated the number of people attending based on number of votes made for different proposals
+and the quorum size, taking the maximum of both.
+We find this a good and conservative proxy because councils have a rule that they have to adjurn
+if the number of members present fall under the quorum (42 members) for more than 10 minutes
+which in practice members discover soon through the number of votes being made or when count
+is explictly requested.
+
+Users on PeerMind made 139 proposals over the span of a semester (this figure
+excludes proposals that were withdrawn before voting began).
+Out of those 139 proposals, 98 proposals passed. In comparison, 70 out of 73 proposals passed
+in the semester prior.
+
+PeerMind users made 3958 votes, whereas 1396 votes were made during spring semester.
+Even after adjusting for the larger number of proposals made in the fall, participation
+increased significantly.
+From around 19 votes per proposal to 28 votes per proposal, on average.
+
+Plotting time spent per user on PeerMind gives further insights into the democratization of decision-making.
+While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours.
+We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes).
+Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.
+
 designed integration with system

### 2018-04-19 statistical-quorum.tex
Trailing spaces.
updated brexit/POTUS table with last column; needs formatting

@@
   \begin{small}
-  \begin{tabular}{l r r r r r r r}
+  \begin{tabular}{l r r r r r r r r}
     {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible}} & {\small \textit{Turnout}} & {\small \textit{Decision}} \\
-    & & & & & {\small \textit{voters}} & & {\small \textit{credibility}} \\
+    & & & & & {\small \textit{voters}} & & {\small \textit{credibility}} & {\small \textit{>$90\%$}}\\
     \midrule
-    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
-    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
+    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ & \textbf{No}\\
+    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ & \textbf{No}\\
   \end{tabular}
@@
 %\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
-\section{Related Work}
-\label{sec:related-work}
 

### 2018-04-19 statistical-quorum.tex
Title and abstract.

--- 2018-04-19 statistical-quorum.tex
@@
 % Title portion. Note the short title for running heads
-\title{Improving credibility and efficiency of decision-making within democratic communities via a statistical quorum}
+\title[Improving credibility and efficiency of decision-making]{Improving credibility and efficiency of decision-making \\ within democratic communities via a statistical quorum}
 
@@
 \begin{abstract}
-The Achilles heel of many major democracies is the perpetual need for sufficient engagement and attention from the community.
-While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?
-We propose a novel \textit{statistical quorum}, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic.
-We developed a web application centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 16 weeks.
-We found that the community was able to use the web application to simultaneously improve the credibility and efficiency of their decision-making process.
+The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
+While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
+different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
+maximize productivity?
+We propose a novel \textit{statistical quorum}, derived using first principles, as one metric the community can use to
+focus attention, in turn making the entire decision-making process dynamic.
+We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
+councils over the course of 16 weeks.
+We found that the community was able to use the web application to simultaneously improve the credibility and efficiency
+of their decision-making process.
 \end{abstract}

### 2018-04-19 statistical-quorum.tex
Data.

@@
 
-Users on PeerMind made 139 proposals over the span of a semester (this figure
+Members on PeerMind made 139 proposals over the span of a semester (this figure
 excludes proposals that were withdrawn before voting began).
@@
 PeerMind users made 3958 votes, whereas 1396 votes were made during spring semester.
-Even after adjusting for the larger number of proposals made in the fall, participation
+Even after adjusting for the larger number of proposals made in the fall, voting participation
 increased significantly.
-From around 19 votes per proposal to 28 votes per proposal, on average.
-
-Plotting time spent per user on PeerMind gives further insights into the democratization of decision-making.
-While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours.
-We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes).
-Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.
+From around 19 votes per proposal in the spring to 28 votes per proposal in the fall, on average.
+(Not everyone present at a council has to vote.)
 
@@
 \label{sec:discussion}
+
+Plotting time spent per community member on PeerMind gives insights into the
+democratization of decision-making.
+While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours.
+We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes).
+Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.
 

### 2018-04-19 statistical-quorum.tex
Use ~.

@@
 
-In Eqs.
-\eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we urge the reader to refer to the \hyperref[sec:derivation]{Appendix} for complete derivation details, including motivations and justifications.
+In Eqs.~\eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we urge the reader to refer to the \hyperref[sec:derivation]{Appendix} for complete derivation details, including motivations and justifications.
 
@@
 
-Using Eqs.
-\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
+Using Eqs.~\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
 For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.
@@
 Our proposed solution is to infer the \textit{relevant} population size that cares about a given proposal so that we may establish credibility within only the subset population (naturally requiring less global attention).
-By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs.
-\eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).
+By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs.~\eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).
 
@@
 
-\subsection*{Derivation of Eq. \eqref{eq2}}
+\subsection*{Derivation of Eq.~\eqref{eq2}}
 In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.
@@
 
-\subsection*{Derivation of Eq. \eqref{eq3}}
+\subsection*{Derivation of Eq.~\eqref{eq3}}
 \begin{align*}
@@
 
-\subsection*{Derivation of Eq. \eqref{eq5}}
+\subsection*{Derivation of Eq.~\eqref{eq5}}
 Assuming independence of votes across proposals,
@@
 
-\subsection*{Derivation of Eq. \eqref{eq6}}
+\subsection*{Derivation of Eq.~\eqref{eq6}}
 The properties we desire for $\lambda(t)$ are:
@@
 
-\subsection*{Derivation of Eq. \eqref{eq7}}
+\subsection*{Derivation of Eq.~\eqref{eq7}}
 Using $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$, we have

### 2018-04-19 statistical-quorum.tex
Cleaning up introduction. This belongs to later sections, if anywhere.

@@
 
-Two of the most shocking democratic votes in recent history occurred in 2016: The referendum vote for Britain to leave the European Union (“Brexit”) and the election of Donald Trump as President of the United States.
+%Two of the most shocking democratic votes in recent history occurred in 2016: The referendum vote for Britain to leave the European Union (“Brexit”) and the election of Donald Trump as President of the United States.
 Both votes were decided by the slimmest of margins, but more significantly, both votes also had poor turnout.
@@
 The ideal governance system is one where citizens do not have to be constantly involved, but the system is still acting as the citizens would like it to.
-We explore a fundamentally different approach to making democracy work: \textit{For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?} Can we increase the credibility of decisions made within a democracy while simultaneously skirting the need for an increase in attention?
+%We explore a fundamentally different approach to making democracy work: \textit{For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?} Can we increase the credibility of decisions made within a democracy while simultaneously skirting the need for an increase in attention?
 We formulated a novel ‘statistical quorum’, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic.
@@
 We found that the community was able to use the web application to simultaneously improve the legitimacy and efficiency of their decision-making process.
-Our prototyped web application for collective decision-making contrasts with existing similar solutions.
+%Our prototyped web application for collective decision-making contrasts with existing similar solutions.
 Our web application integrates new ideas in how to improve this process: use of \textbf{range voting}, a \textbf{dynamic statistical quorum}, and \textbf{voter delegation}.
@@
 Specifically, we can compute results faster, more accurately, and securely, we can use powerful statistical methods to estimate important quantities for the entire voting population, we can allow for members to entrust their vote to others who have more expertise on specific topics than they themselves do, and we can promote clarity of thought and complete understanding among all members for all voting items.
-\textit{Range voting} is voting where members of a community express their positions on open proposals by assigning each a score between –1 and +1.
+%\textit{Range voting} is voting where members of a community express their positions on open proposals by assigning each a score between –1 and +1.
 Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their positions in a highly informative and precise manner.
 Moreover, if a new proposal is added to a set of existing proposals on which voting was already open, only the new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of each proposal would need reassessing.
-A \textit{dynamic statistical quorum} uses statistical hypothesis testing to compute the minimum number of required votes needed to achieve a desired level of confidence before a community adopts an outcome as being truly representative of its members’ positions.
+%A \textit{dynamic statistical quorum} uses statistical hypothesis testing to compute the minimum number of required votes needed to achieve a desired level of confidence before a community adopts an outcome as being truly representative of its members’ positions.
 For example, in a community of 100 members, if 20 members all vote in full support of a proposal, then we can conclude with very high confidence that, had all 100 members actually voted, a majority of them would support the proposal (despite in actuality 80 members having not yet voted).
@@
 By utilizing a dynamic statistical quorum, we can direct attention and participation within the community towards contentious proposals, while reducing wasted attention directed towards proposals that already have overwhelming support.
-\textit{Voter delegation} addresses two questionable assumptions about current democratic systems: (1) participants have the time to participate and (2) participants have the knowledge to participate.
+%\textit{Voter delegation} addresses two questionable assumptions about current democratic systems: (1) participants have the time to participate and (2) participants have the knowledge to participate.
 Not all members of a community can participate at all times and not all members have the time to educate themselves about each and every decision being made.
@@
 If at any point a member decides to vote themselves, then their direct vote is used instead.
-All of the above features are combined into a guided process to help community members participate in decision-making.
+%All of the above features are combined into a guided process to help community members participate in decision-making.
 The process starts with a new discussion item where members can comment and discuss a proposal.

### 2018-04-19 statistical-quorum.tex
subfigures

--- 2018-04-19 statistical-quorum.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage{subcaption}
 
@@
 % Title portion. Note the short title for running heads
-\title[Improving credibility and efficiency of decision-making]{Improving credibility and efficiency of decision-making \\ within democratic communities via a statistical quorum}
+\title{Improving credibility and efficiency of decision-making within democratic communities via a statistical quorum}
 
@@
 \begin{abstract}
-The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
-While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
-different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
-maximize productivity?
-We propose a novel \textit{statistical quorum}, derived using first principles, as one metric the community can use to
-focus attention, in turn making the entire decision-making process dynamic.
-We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
-councils over the course of 16 weeks.
-We found that the community was able to use the web application to simultaneously improve the credibility and efficiency
-of their decision-making process.
+The Achilles heel of many major democracies is the perpetual need for sufficient engagement and attention from the community.
+While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?
+We propose a novel \textit{statistical quorum}, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic.
+We developed a web application centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 16 weeks.
+We found that the community was able to use the web application to simultaneously improve the credibility and efficiency of their decision-making process.
 \end{abstract}
@@
 
-In Eqs.~\eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we urge the reader to refer to the \hyperref[sec:derivation]{Appendix} for complete derivation details, including motivations and justifications.
+In Eqs.
+\eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we urge the reader to refer to the \hyperref[sec:derivation]{Appendix} for complete derivation details, including motivations and justifications.
 
@@
 
-Using Eqs.~\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
+Using Eqs.
+\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
 For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.
@@
 Our proposed solution is to infer the \textit{relevant} population size that cares about a given proposal so that we may establish credibility within only the subset population (naturally requiring less global attention).
-By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs.~\eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).
+By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs.
+\eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).
 
@@
 
+\newpage
+
 \section{Evaluation}
@@
 \centering
-\includegraphics[width=0.45\textwidth]{figures/credible_interval.pdf}
+\includegraphics[width=0.6\textwidth]{figures/credible_interval.pdf}
 \caption{.}
@@
 \centering
-\includegraphics[width=.45\textwidth]{figures/simple_majority.pdf}
+\begin{subfigure}{.5\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/simple_majority.pdf}
 \caption{.}
 \label{fig:simple_majority}
-\end{figure}
-
-\begin{figure}[ht]
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
 \centering
-\includegraphics[width=.45\textwidth]{figures/competing_proposals.pdf}
+\includegraphics[width=\textwidth]{figures/competing_proposals.pdf}
 \caption{.}
 \label{fig:simple_majority}
+\end{subfigure}
 \end{figure}
@@
 
-Members on PeerMind made 139 proposals over the span of a semester (this figure
+Users on PeerMind made 139 proposals over the span of a semester (this figure
 excludes proposals that were withdrawn before voting began).
@@
 PeerMind users made 3958 votes, whereas 1396 votes were made during spring semester.
-Even after adjusting for the larger number of proposals made in the fall, voting participation
+Even after adjusting for the larger number of proposals made in the fall, participation
 increased significantly.
-From around 19 votes per proposal in the spring to 28 votes per proposal in the fall, on average.
-(Not everyone present at a council has to vote.)
+From around 19 votes per proposal to 28 votes per proposal, on average.
+
+Plotting time spent per user on PeerMind gives further insights into the democratization of decision-making.
+While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours.
+We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes).
+Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.
 
@@
 
-Plotting time spent per community member on PeerMind gives insights into the
-democratization of decision-making.
-While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours.
-We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes).
-Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.
-
 decline in understanding of the process:
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq2}}
+\subsection*{Derivation of Eq. \eqref{eq2}}
 In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq3}}
+\subsection*{Derivation of Eq. \eqref{eq3}}
 \begin{align*}
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq5}}
+\subsection*{Derivation of Eq. \eqref{eq5}}
 Assuming independence of votes across proposals,
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq6}}
+\subsection*{Derivation of Eq. \eqref{eq6}}
 The properties we desire for $\lambda(t)$ are:
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq7}}
+\subsection*{Derivation of Eq. \eqref{eq7}}
 Using $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$, we have

### 2018-04-19 statistical-quorum.tex
90% is just a question of community, not universal constant. This column does not add anything.
You can add this to caption, if you want. Discussion that they are below
a resonable number like 90%.

--- 2018-04-19 statistical-quorum.tex
@@
   \begin{small}
-  \begin{tabular}{l r r r r r r r r}
+  \begin{tabular}{l r r r r r r r}
     {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible}} & {\small \textit{Turnout}} & {\small \textit{Decision}} \\
-    & & & & & {\small \textit{voters}} & & {\small \textit{credibility}} & {\small \textit{>$90\%$}}\\
+    & & & & & {\small \textit{voters}} & & {\small \textit{credibility}} \\
     \midrule
-    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ & \textbf{No}\\
-    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ & \textbf{No}\\
+    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
+    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
   \end{tabular}

### 2018-04-19 statistical-quorum.tex
Style they had in the template.

@@
 \appendix
-\section*{Appendix}
+\section{Derivations}
 \label{sec:derivation}
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq2}}
+\subsection{Derivation of Eq.~\eqref{eq2}}
 In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq3}}
+\subsection{Derivation of Eq.~\eqref{eq3}}
 \begin{align*}
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq5}}
+\subsection{Derivation of Eq.~\eqref{eq5}}
 Assuming independence of votes across proposals,
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq6}}
+\subsection{Derivation of Eq.~\eqref{eq6}}
 The properties we desire for $\lambda(t)$ are:
@@
 
-\subsection*{Derivation of Eq.~\eqref{eq7}}
+\subsection{Derivation of Eq.~\eqref{eq7}}
 Using $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$, we have

### 2018-04-19 statistical-quorum.tex
Move figures.
Removing done TODOs.

@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/peermind.png}
+\caption{User interface of PeerMind web application. One can see three column structure, with commenting
+section on the left, summary points in the middle, and proposed proposals with voting on the right.}
+\label{fig:peer-mind}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/ballot.png}
+\caption{Ballot used for voting in PeerMind web application. Users can chose a value between -1 and 1 on a slider,
+where -1 means opposition, 1 means in favor, and 0 indifference. They can also chose abstention (they
+remove themselves from the population of the community for voting in question), or to not vote (default), which
+means that their vote does not contribute to the quorum.}
+\label{fig:ballot}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/results.png}
+\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of voting.
+A user can hover over the chart to see details at that time. Users can see these results in real time during voting.}
+\label{fig:results}
+\end{figure}
+
 Each user can open and propose new discussion items, which leads them to the main user interface for each discussion as seen in Figure~\ref{fig:peer-mind}, having three sections: commenting section on the left, summary points in the middle,
@@
 ID \emph{redacted for blind review}.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.9\textwidth]{figures/peermind.png}
-\caption{User interface of PeerMind web application. One can see three column structure, with commenting
-section on the left, summary points in the middle, and proposed proposals with voting on the right.}
-\label{fig:peer-mind}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/ballot.png}
-\caption{Ballot used for voting in PeerMind web application. Users can chose a value between -1 and 1 on a slider,
-where -1 means opposition, 1 means in favor, and 0 indifference. They can also chose abstention (they
-remove themselves from the population of the community for voting in question), or to not vote (default), which
-means that their vote does not contribute to the quorum.}
-\label{fig:ballot}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.6\textwidth]{figures/results.png}
-\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of voting.
-A user can hover over the chart to see details at that time. Users can see these results in real time during voting.}
-\label{fig:results}
-\end{figure}
 
@@
 
-designed integration with system
-
-minutes and issues with how to get data out
-
-workflow we created -> over one in-person meetings, but because of all decision making happining online, this
-didn't happen
-
-issues with quorum
-unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after voting
-
-X people choose not to participate
-
-\todo[inline]{Explain how we were using this for 16 weeks in fall 2016 in a community where traditionally they have
-been meeting weekly and now they used it online. How X proposals have been made with Y votes.}
-
-\todo[inline]{Explain how we compare with history: using minutes from spring 2016 for the same community. Issues with
-quorum often not bein reached or being lost during the meeting. How many proposals were made then.}
-
 \todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}
-
-\todo[inline]{We compare how much time collectively was spend on decision making based on minutes in spring 2016 and
-how much based on interactions with the platform in fall 2016. Claim is that we made more decisions with less time people
-having to be involved in decision making, and with better legitimacy.}
-
-\todo[inline]{We did satisfaction surveys with decision making both in spring and fall 2016. Pull out some interesting
-observation when comparing.}
-
-\todo[inline]{Make sure that we are supporting both of our claims well: this statistical quorum is a good metric and
-collective attention can be utilized better.}
 
@@
 \label{sec:discussion}
+
+issues with quorum
+unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after voting
+Loosing a quorum often during meetings in spring
 

### 2018-04-19 statistical-quorum.tex
done with Section 5.1 Analysis

--- 2018-04-19 statistical-quorum.tex
@@
 \end{itemize}
-Using a desirable prior
+Using the desirable prior
 \begin{align}\label{eq2}
@@
 
-Using Eqs.~\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
+Using Eqs.~\eqref{eq3} and \eqref{eq5}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.
 For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.
@@
 
-To gain an intuitive grasp of the decision credibility used in a statistical quorum, we can examine two toy scenarios.
+To gain an intuitive grasp of the decision credibility used in a statistical quorum, we can examine a few toy scenarios:
 
 In Figure \ref{fig:credible_interval}, we examine how the $90\%$ credible interval converges as more votes are collected out of a population of $100$.
-A statistical quorum is achieved when the credible interval is contained entirely above the x-axis.
+A statistical quorum is achieved when the credible interval is contained entirely above zero, corresponding to a positive bias toward the proposal.
 In this example, the latent bias of the community is that $75\%$ of the community supports the proposal while $25\%$ opposes it.
+After approximately $35$ members have voted, a statistical quorum is achieved.
 
-In
+In Figure \ref{fig:simple_majority}, we examine how many affirmative votes are required from a community of size $100$ to reach a statistical quorum \textit{after} the proposal starts off with $10$ votes against it.
+Initially, the $0-10$ vote tally in favor of rejecting the proposal gives the outcome over an $80\%$ credibility; however, as affirmative votes begin arriving, the credibility decreases until the `yes' votes outnumber the `no' votes.
+After $21$ `yes' votes have been cast against the $10$ `no' votes, a statistical quorum of $90\%$ credibility is achieved.
+
+In Figure \ref{fig:competing_proposals}, we examine how a set of three competing proposals gets resolved.
+We simulate three proposals competing against each other in a community of size $100$, where the community has a latent bias of $75\%$, $50\%$, $25\%$ toward each of the proposals.
+This scenario models what happens when a community slightly prefers the first proposal, is relatively indifferent or equivalently split on the second proposal, and slightly disproves of the third proposal.
+Each member of the community casts a vote of $+0.5$, $0$, $-0.5$ for proposal 1, 2, 3, respectively.
+(We have removed the stochasticity of the votes cast according to the community's latent biases in order to simplify the simulation.)  After approximately $35$ votes, we see that a statistical quorum of $90\%$ is achieved, signifying that the community's preference for proposal 1 over proposals 2, 3 is clear.
+
 \begin{figure}[ht]
 \centering
-\includegraphics[width=0.45\textwidth]{figures/credible_interval.pdf}
-\caption{.}
+\includegraphics[width=0.6\textwidth]{figures/credible_interval.pdf}
+\caption{Convergence of the $90\%$ credible interval as more votes are cast in the population of size $100$.  A statistical quorum is achieved when the credible interval is strictly above $0$.}
 \label{fig:credible_interval}
@@
 \centering
-\includegraphics[width=.45\textwidth]{figures/simple_majority.pdf}
-\caption{.}
+\begin{subfigure}[t]{.5\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/simple_majority.pdf}
+\caption{Credibility of a proposal requiring a simple majority to pass: After $10$ `No' votes are initially cast, `Yes' votes arrive until a statistical quorum is achieved.}
 \label{fig:simple_majority}
+\end{subfigure}%
+\begin{subfigure}[t]{.5\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/competing_proposals.pdf}
+\caption{Credibility of a proposal competing against two other proposals: As more votes are cast according to the population's bias towards the three proposals ($75\%$, $50\%$, $25\%$), the first proposal eventually achieves a statistical quorum.}
+\label{fig:competing_proposals}
+\end{subfigure}
+\caption{}
 \end{figure}
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=.45\textwidth]{figures/competing_proposals.pdf}
-\caption{.}
-\label{fig:simple_majority}
-\end{figure}
+\newpage
 

### 2018-04-19 statistical-quorum.tex

--- 2018-04-19 statistical-quorum.tex
@@
 
-To gain an intuitive grasp of the decision credibility used in a statistical quorum, we can examine two toy scenarios.
+To gain an intuitive grasp of the decision credibility used in a statistical quorum, we can examine a few toy scenarios:
 
 In Figure \ref{fig:credible_interval}, we examine how the $90\%$ credible interval converges as more votes are collected out of a population of $100$.
-A statistical quorum is achieved when the credible interval is contained entirely above the x-axis.
+A statistical quorum is achieved when the credible interval is contained entirely above zero, corresponding to a positive bias toward the proposal.
 In this example, the latent bias of the community is that $75\%$ of the community supports the proposal while $25\%$ opposes it.
+After approximately $35$ members have voted, a statistical quorum is achieved.
 
-In
+In Figure \ref{fig:simple_majority}, we examine how many affirmative votes are required from a community of size $100$ to reach a statistical quorum \textit{after} the proposal starts off with $10$ votes against it.
+Initially, the $0-10$ vote tally in favor of rejecting the proposal gives the outcome over an $80\%$ credibility; however, as affirmative votes begin arriving, the credibility decreases until the `yes' votes outnumber the `no' votes.
+After $21$ `yes' votes have been cast against the $10$ `no' votes, a statistical quorum of $90\%$ credibility is achieved.
+
+In Figure \ref{fig:competing_proposals}, we examine how a set of three competing proposals gets resolved.
+We simulate three proposals competing against each other in a community of size $100$, where the community has a latent bias of $75\%$, $50\%$, $25\%$ toward each of the proposals.
+This scenario models what happens when a community slightly prefers the first proposal, is relatively indifferent or equivalently split on the second proposal, and slightly disproves of the third proposal.
+Each member of the community casts a vote of $+0.5$, $0$, $-0.5$ for proposal 1, 2, 3, respectively.
+(We have removed the stochasticity of the votes cast according to the community's latent biases in order to simplify the simulation.)  After approximately $35$ votes, we see that a statistical quorum of $90\%$ is achieved, signifying that the community's preference for proposal 1 over proposals 2, 3 is clear.
+
 \begin{figure}[ht]
@@
 \includegraphics[width=0.6\textwidth]{figures/credible_interval.pdf}
-\caption{.}
+\caption{Convergence of the $90\%$ credible interval as more votes are cast in the population of size $100$.  A statistical quorum is achieved when the credible interval is strictly above $0$.}
 \label{fig:credible_interval}
@@
 \centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}[t]{.5\textwidth}
 \centering
 \includegraphics[width=\textwidth]{figures/simple_majority.pdf}
-\caption{.}
+\caption{Credibility of a proposal requiring a simple majority to pass: After $10$ `No' votes are initially cast, `Yes' votes arrive until a statistical quorum is achieved.}
 \label{fig:simple_majority}
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}[t]{.5\textwidth}
 \centering
 \includegraphics[width=\textwidth]{figures/competing_proposals.pdf}
-\caption{.}
-\label{fig:simple_majority}
+\caption{Credibility of a proposal competing against two other proposals: As more votes are cast according to the population's bias towards the three proposals ($75\%$, $50\%$, $25\%$), the first proposal eventually achieves a statistical quorum.}
+\label{fig:competing_proposals}
 \end{subfigure}
+\caption{}
 \end{figure}
+
+\newpage
 

### 2018-04-19 statistical-quorum.tex
Added chart by time.

--- 2018-04-19 statistical-quorum.tex
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/people-sorted-time-chart.png}
+\caption{A chart showing time spent per member on PeerMind, sorted by time.}
+\label{fig:time}
+\end{figure}
+
+Figure~\ref{fig:time} shows time spent per member on PeerMind, sorted by time. While on average, members
+spent 7002 seconds (around 2 hours) using the application, the distribution of users is fairly
+long-tailed with many members only spending 30 minutes and some members spending upwards of 20 hours.
+
 \todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}

### 2018-04-19 statistical-quorum.tex
Layout not important for draft. No time to check as we are changing other stuff.
The vote to voting.

@@
 Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote was credible enough to be taken as representative of the entire community's desire.
-On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in the vote.
+On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in voting.
 Instead, the suggested course of action would be to continue gathering more votes and/or further discussing the proposal until a credible outcome is established.
@@
 \item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
-\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
+\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed
 (e.g., $f=\frac{1}{2}$ for a simple majority).
@@
 ID \emph{redacted for blind review}.
-
-\newpage
 

### 2018-04-19 statistical-quorum.tex
related work in progress

--- 2018-04-19 statistical-quorum.tex
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.6\textwidth]{figures/people-sorted-time-chart.png}
-\caption{A chart showing time spent per member on PeerMind, sorted by time.}
-\label{fig:time}
-\end{figure}
-
-Figure~\ref{fig:time} shows time spent per member on PeerMind, sorted by time. While on average, members
-spent 7002 seconds (around 2 hours) using the application, the distribution of users is fairly
-long-tailed with many members only spending 30 minutes and some members spending upwards of 20 hours.
-
 \todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}

### 2018-04-19 statistical-quorum.tex
Layout not important for draft.
Dominated.
Hiding interviews. No time.

--- 2018-04-19 statistical-quorum.tex
@@
 \item Using the analysis of decision making process before and after introduction
-of the application, using surveys, and using semi-structured interviews, we show
+of the application, and using surveys,
+%and using semi-structured interviews,
+we show
 how use of statistical quorum increases credibility of decisions,
@@
 We present our methodology in Section~\ref{sec:methodology} after which we do
-quantitative and qualitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
+quantitative
+%and qualitative
+evaluation in Section~\ref{sec:evaluation}. We discuss results in
 Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
@@
 
-Finally, we did semi-structured interviews after the fall semester with 7 members who were participating
-both in spring and fall semesters and ask them to compare their experience between those two semesters and
-evaluate the application and the use of statistical quorum.
+%Finally, we did semi-structured interviews after the fall semester with 7 members who were participating
+%both in spring and fall semesters and ask them to compare their experience between those two semesters and
+%evaluate the application and the use of statistical quorum.
 
@@
 We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
-Then we do quantitative and qualitative evaluation through real-world use of the PeerMind application by a 140 members
+Then we do quantitative
+%and qualitative
+evaluation through real-world use of the PeerMind application by a 140 members
 community.
@@
 
-\newpage
-
 \subsection{Use in Practice}
@@
 long-tailed with many members only spending 30 minutes and some members spending upwards of 20 hours.
+On the other hand, based on minutes from spring semester we can observe that around 5 same members
+regularly dominated conversation.
 
-\todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}
+%\todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}
 
@@
 
-\subsection{Interviews}
-\label{sec:interviews}
+%\subsection{Interviews}
+%\label{sec:interviews}
 
-\todo[inline]{We made 7 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
-to work. But how there are other issues by moving decision making online: meetings also serve community building which
-does not happen online. No fun. So interesting insights about using online platforms for offline communities
-for making decisions which are not really answering our hypothesis, but are useful as insights we got. How some
-people prefer online and offline meetings.}
+%\todo[inline]{We made 7 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
+%to work. But how there are other issues by moving decision making online: meetings also serve community building which
+%does not happen online. No fun. So interesting insights about using online platforms for offline communities
+%for making decisions which are not really answering our hypothesis, but are useful as insights we got. How some
+%people prefer online and offline meetings.}
 

### 2018-04-19 statistical-quorum.tex
More discussion.

@@
 
-issues with quorum
-unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after voting
-Loosing a quorum often during meetings in spring
+Based on quantitative data we can observe that members using PeerMind were more productive in
+terms of decisions made.
+On the other hand, they also rejected more proposals which shows more involvement in decisions
+than just rubber stamping proposals.
+They also spent significantly less time per proposal than members in spring semester.
+While at the same time they made significantly more votes.
+Combined with data from the survey where members observed that in the fall their time
+is better spend making decisions, we can find support for our hypothesis that directing
+attention of members towards contentious issues using a statistical quorum increases the
+participation where it is needed (which can lead to a proposal being rejected) and
+increases participation overall while keeping the time (or attention) needed low.
 
-Plotting time spent per community member on PeerMind gives insights into the
-democratization of decision-making.
-While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours.
-We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes).
-Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.
+Moreover, Figure~\ref{fig:time} shows that participation is diversified and more different
+members participate over what we observed through analysis of minutes from the semester prior.
+We believe many members who had spent only 30 minutes using PeerMind would not participate
+in the in-person council at all (which is confirmed by much lower attendance and fewer votes).
+Furthermore, people who need more time to discuss are given the freedom to do so without wasting
+time of those who are not interested in the pertinent discussion.
+We have observed that councils in the spring had often issues retaining or even achieving a
+quorum which we attribute to this mismatch in how people want to spend their time during
+decision making and their time availability.
 
-decline in understanding of the process:
+In the case of the in-person decision making it is also unclear how much time to spend on each
+dicussion item: how contentious the discussion really is can be seen only after voting.
+Using PeerMind members can observe the voting process in real-time and continue discussing
+and even changing their votes if they get convinced to change their poisition.
+At the same time using feedback provided by credibility value they can also gauge how
+much discussion is needed.
 
+% People changed vote online X times after discussion.
+
+The decline in understanding of the process we see in the survey is expected and understandable.
 We attribute this to the fact that the use of the web application is
@@
 learn from them, instead of being alone using the application.
-We explore this question also through interviews in Section~\ref{sec:interviews}.
 
-This not only means that PeerMind users were more productive in terms of decisions made, they also spent significantly less time per motion than the control group.
-
-\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.\\
-Part of the answer here is honestly just that the admin should tune the threshold (and a few of the other hyperparameters we introduce) to their intuition.
+%\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.\\
+%Part of the answer here is honestly just that the admin should tune the threshold (and a few of the other hyperparameters we introduce) to their intuition.
 Once you set the parameters, the real benefit is that you have consistency, rather than having to always use your intuition as an admin about evaluating credibility.}

### 2018-04-19 statistical-quorum.tex
Removing todos.

@@
 
-\todo[inline]{Security. How to prevent a group of people to vote and by trick reaching statistical quorum.
-We use dynamic statistical quorum to address this, but one other approach is also for voting to stay open and
-decisions can then be in the future revoked by people adding more votes, or changing them. But for some decisions
-this might not really be possible, because decisions might not be reversible. You can revoke a law, but if you built
-a building as a consequence of a decision?\\
-(See near/post-quorum validation phase idea in GitHub issues as one suggested solution.)}
+%\todo[inline]{Security. How to prevent a group of people to vote and by trick reaching statistical quorum.
+%We use dynamic statistical quorum to address this, but one other approach is also for voting to stay open and
+%decisions can then be in the future revoked by people adding more votes, or changing them. But for some decisions
+%this might not really be possible, because decisions might not be reversible. You can revoke a law, but if you built
+%a building as a consequence of a decision?\\
+%(See near/post-quorum validation phase idea in GitHub issues as one suggested solution.)}
 
-\todo[inline]{Discuss how this might or might not lead to a more inclusive decision making: some people prefer it because
-they have time to collect thoughts and can participate when they have time, some people do not because they feel
-they cannot communicate as effectively, especially with lack of emotions being conveyed.}
+%\todo[inline]{From interviews: Discuss how this might or might not lead to a more inclusive decision making: some people prefer it because
+%they have time to collect thoughts and can participate when they have time, some people do not because they feel
+%they cannot communicate as effectively, especially with lack of emotions being conveyed.}
 

### 2018-04-19 statistical-quorum.tex
crude related  work section done

--- 2018-04-19 statistical-quorum.tex
@@
 
-\usepackage{todonotes}
+\usepackage[disable]{todonotes}
 \usepackage{bookmark}
@@
 \item Using the analysis of decision making process before and after introduction
-of the application, and using surveys,
-%and using semi-structured interviews,
-we show
+of the application, using surveys, and using semi-structured interviews, we show
 how use of statistical quorum increases credibility of decisions,
@@
 We present our methodology in Section~\ref{sec:methodology} after which we do
-quantitative
-%and qualitative
-evaluation in Section~\ref{sec:evaluation}. We discuss results in
+quantitative and qualitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
 Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
@@
 \label{sec:related-work}
+
+Past approaches exist in using Bayesian statistics in the context of voting theory.
+In particular, Bayesian methods have been proposed for auditing election outcomes \citep{rivest2012bayesian}.
+Many further methods have been developed for efficient ballot auditing \cite{lindeman2012bravo}, \citep{rivest2017clipaudit}.
+The problem of verifying an election's outcome is mathematically similar in structure to the notion of a statistical quorum defined in this work; however, a critical difference lies in the role the metric plays in the decision-making process itself.
+Post hoc verification of results does not influence voting.
+A statistical quorum, on the other hand, directly influences the allocation of a community's attention by essentially performing voting audits in real-time.
+To our knowledge, no methods have been formalized for coupling the verification of vote outcomes with the voting process itself.
+Furthermore, the dynamic statistical quorum presented here takes the previous ideas one step further.
+Statistical method for auditing vote outcomes have no way to infer what we call the ``effective population size'', i.e., the portion of the population that actually cares about a vote's outcome.
+We believe this information is crucial and have developed new statistical methods to infer this population and subsequently couple it to the statistical quorum, further influencing the influence on the decision-making process itself.
+Additional work exists for selectively sampling a subset of the voting population to estimate the population's bias \cite{chaum2016random}.
+In random-sample voting, the emphasis is on security and efficiency of voting.
+The approach taken to optimally allocate a community's attention is fundamentally different, however, in that the goal is to selectively distribute votes to random subsets of the population with the intent of reducing the amount of attention required by each individual.
+This does indeed reduce the amount of attention required of individual voters, but at the same time is restricts the freedom of voters to participate only in the votes they have been delegated to participate in.
 
@@
 
-%Finally, we did semi-structured interviews after the fall semester with 7 members who were participating
-%both in spring and fall semesters and ask them to compare their experience between those two semesters and
-%evaluate the application and the use of statistical quorum.
+Finally, we did semi-structured interviews after the fall semester with 7 members who were participating
+both in spring and fall semesters and ask them to compare their experience between those two semesters and
+evaluate the application and the use of statistical quorum.
 
@@
 We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
-Then we do quantitative
-%and qualitative
-evaluation through real-world use of the PeerMind application by a 140 members
+Then we do quantitative and qualitative evaluation through real-world use of the PeerMind application by a 140 members
 community.
@@
 
+\newpage
+
 \subsection{Use in Practice}
@@
 long-tailed with many members only spending 30 minutes and some members spending upwards of 20 hours.
-On the other hand, based on minutes from spring semester we can observe that around 5 same members
-regularly dominated conversation.
 
-%\todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}
+\todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}
 
@@
 
-%\subsection{Interviews}
-%\label{sec:interviews}
+\subsection{Interviews}
+\label{sec:interviews}
 
-%\todo[inline]{We made 7 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
-%to work. But how there are other issues by moving decision making online: meetings also serve community building which
-%does not happen online. No fun. So interesting insights about using online platforms for offline communities
-%for making decisions which are not really answering our hypothesis, but are useful as insights we got. How some
-%people prefer online and offline meetings.}
+\todo[inline]{We made 7 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
+to work. But how there are other issues by moving decision making online: meetings also serve community building which
+does not happen online. No fun. So interesting insights about using online platforms for offline communities
+for making decisions which are not really answering our hypothesis, but are useful as insights we got. How some
+people prefer online and offline meetings.}
 

### 2018-04-19 statistical-quorum.tex
Conclusion.

--- 2018-04-19 statistical-quorum.tex
@@
 
+In this work we have developed a theorethical framework of computing statistical credibility
+of decisions being made through a voting process.
+We applied this framework to design a statistical quroum which we integrated with a web application
+PeerMind.
+We deployed the application in a democratic community to compare its effect in comparison with
+prior time before the introduction of the web application.
+Analyzing the effects of this introduction, we conclude that our hypothesis that using
+crediblity as a feedback mechanism and statistical quorum as a threshold really improves
+the partication of members in decision making, making it more inclusive, more engaging, and
+better spend collective attention available for decision making.
+
 % Appendix

### 2018-04-19 statistical-quorum.tex
spell check

--- 2018-04-19 statistical-quorum.tex
@@
 
-We built a full-fletched mobile-friendly web application using Meteor web framework~\cite{meteor}.
+We built a full-fledged mobile-friendly web application using Meteor web framework~\cite{meteor}.
 Meteor framework supports reactivity which makes the application update user interface based on actions
@@
 
-Web application contains other features which makes it easier to use as a decision making and discusion
+Web application contains other features which makes it easier to use as a decision making and discussion
 platform, like notifications, and additional features which allow the application to be used also to
@@
 
-PeerMind source code is available under open source licence at \url{https://redacted.for.blind.review} and
+PeerMind source code is available under open source license at \url{https://redacted.for.blind.review} and
 anonymized dataset of interactions with the app, including votes, is available at \url{https://redacted.for.blind.review}.
@@
 and the quorum size, taking the maximum of both.
-We find this a good and conservative proxy because councils have a rule that they have to adjurn
+We find this a good and conservative proxy because councils have a rule that they have to adjourn
 if the number of members present fall under the quorum (42 members) for more than 10 minutes
 which in practice members discover soon through the number of votes being made or when count
-is explictly requested.
+is explicitly requested.
 
@@
 In the case of the in-person decision making it is also unclear how much time to spend on each
-dicussion item: how contentious the discussion really is can be seen only after voting.
+discussion item: how contentious the discussion really is can be seen only after voting.
 Using PeerMind members can observe the voting process in real-time and continue discussing
-and even changing their votes if they get convinced to change their poisition.
+and even changing their votes if they get convinced to change their position.
 At the same time using feedback provided by credibility value they can also gauge how

### 2018-04-19 statistical-quorum.tex
Changes to voting.
submission v1

--- 2018-04-19 statistical-quorum.tex
@@
 To our knowledge, no methods have been formalized for coupling the verification of vote outcomes with the voting process itself.
+
 Furthermore, the dynamic statistical quorum presented here takes the previous ideas one step further.
@@
 We believe this information is crucial and have developed new statistical methods to infer this population and subsequently couple it to the statistical quorum, further influencing the influence on the decision-making process itself.
+
 Additional work exists for selectively sampling a subset of the voting population to estimate the population's bias \cite{chaum2016random}.
@@
 and even changing their votes if they get convinced to change their position.
+69 members changed a total of 279 votes 60 minutes or later after their first vote when
+using PeerMind.
 At the same time using feedback provided by credibility value they can also gauge how
 much discussion is needed.
-
-% People changed vote online X times after discussion.
 
@@
 \label{sec:conclusion}
+
+In this work we have developed a theoretical framework of computing statistical credibility
+of decisions being made through a voting process.
+We applied this framework to design a statistical quorum which we integrated with a web application
+PeerMind.
+We deployed the application in a democratic community to compare its effect in comparison with
+prior time before the introduction of the web application.
+Analyzing the effects of this introduction, we conclude that our hypothesis that using
+credibility as a feedback mechanism and statistical quorum as a threshold really improves
+the participation of members in decision making, making it more inclusive, more engaging, and
+better spend collective attention available for decision making.
 

### 2018-04-19 statistical-quorum.tex
Story.

--- 2018-04-19 statistical-quorum.tex
@@
 
-\todo[inline]{Describe interesting example how we had a bug which were showing a very low statistical quorum for a failing
-vote and so the whole community very engaged for a whole month to decide and resolve it. So how they trusted the
-app and the metric and how it was really able to direct the attention of the community. It seems it works even if it is
-buggy or not perfect, because it has worked so well in the past that members got used to the fact that when it was saying
-that there is a consensus or not on an issue that that was also their own read of the community. So for the purpose of
-directing attention the metric does not have to mathematically perfect.}
+A case for how well can use of statistical quorum influence how community directs their
+attention has been also shown through a bug we had at one point in PeerMind web application
+in the fall semester.
+The web application was for a particular decision showing a very low credibility and inviting
+more votes for decision to reach statistical quorum.
+Community engaged and started discussing proposal more, which lasted more than three weeks.
+Almost whole community got involved and many votes were made.
+At that time we realized that a bug is showing that more votes are needed to pass the proposal
+when in fact enough votes to reject the proposal were already made, long ago.
+This example showed us how powerful can such tool at direction attention be and how trusting
+a community has became in it after weeks of successful use.
 

### 2018-04-20 statistical-quorum.tex

@@
 
-In this work we have developed a theorethical framework of computing statistical credibility
+In this work we have developed a theoretical framework of computing statistical credibility
 of decisions being made through a voting process.
-We applied this framework to design a statistical quroum which we integrated with a web application
+We applied this framework to design a statistical quorum which we integrated with a web application
 PeerMind.
@@
 Analyzing the effects of this introduction, we conclude that our hypothesis that using
-crediblity as a feedback mechanism and statistical quorum as a threshold really improves
-the partication of members in decision making, making it more inclusive, more engaging, and
+credibility as a feedback mechanism and statistical quorum as a threshold really improves
+the participation of members in decision making, making it more inclusive, more engaging, and
 better spend collective attention available for decision making.
